### 2023-06-07 texor.Rmd
initial commit

---
title: "Converting 'LaTeX' 'R Journal' Articles into 'RJ-web-articles'"
date: "2022-11-09"
abstract: >
  An abstract of less than 150 words.
draft: true
author:  
  # see ?rjournal_article for more information
  - name: Abhishek Ulayil
    affiliation: Student, Institute of Actuaries of India
    address:
    - Mumbai, India
    orcid: 0009-0000-6935-8690
    email: perricoq@outlook.com
type: package
output: 
  rjtools::rjournal_web_article:
    web_only: yes
    self_contained: yes
    toc: no
bibliography: RJreferences.bib

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(plotly)
library(ggplot2)
library(palmerpenguins)
```

# Introduction

Explain the purpose of the work, some background on the rjournal/rjtools[DC],

# Background

LaTeX, Regex, pandoc
building blocks of the packages, methods used in them [MOW]

# texor package converting LaTeX to Rmarkdown

## Also mention the hard parts
## Lua Filters with pandoc
A lot of details about different components of texor

# rebib package working out bibliography

# getting started

 Examples/ getting started / usage and so on

# Summary

# Acknowledgments
pandoc
 "TODO : Later"

# supplementary Materials
 github repos for the packages, rjtools

### 2023-06-07 texor.tex
initial commit

% !TeX root = RJwrapper.tex
\title{Converting `LaTeX' `R Journal' Articles into `RJ-web-articles'}
\author{by Abhishek Ulayil}

\maketitle

\abstract{%
An abstract of less than 150 words.
}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Explain the purpose of the work, some background on the rjournal/rjtools{[}DC{]},

\hypertarget{background}{%
\section{Background}\label{background}}

LaTeX, Regex, pandoc
building blocks of the packages, methods used in them {[}MOW{]}

\hypertarget{texor-package-converting-latex-to-rmarkdown}{%
\section{texor package converting LaTeX to Rmarkdown}\label{texor-package-converting-latex-to-rmarkdown}}

\hypertarget{also-mention-the-hard-parts}{%
\subsection{Also mention the hard parts}\label{also-mention-the-hard-parts}}

\hypertarget{lua-filters-with-pandoc}{%
\subsection{Lua Filters with pandoc}\label{lua-filters-with-pandoc}}

A lot of details about different components of texor

\hypertarget{rebib-package-working-out-bibliography}{%
\section{rebib package working out bibliography}\label{rebib-package-working-out-bibliography}}

\hypertarget{getting-started}{%
\section{getting started}\label{getting-started}}

Examples/ getting started / usage and so on

\hypertarget{summary}{%
\section{Summary}\label{summary}}

\hypertarget{acknowledgments}{%
\section{Acknowledgments}\label{acknowledgments}}

pandoc
``TODO : Later''

\hypertarget{supplementary-materials}{%
\section{supplementary Materials}\label{supplementary-materials}}

github repos for the packages, rjtools

\bibliography{RJreferences.bib}

\address{%
Abhishek Ulayil\\
Student, Institute of Actuaries of India\\%
Mumbai, India\\
%
%
\textit{ORCiD: \href{https://orcid.org/0009-0000-6935-8690}{0009-0000-6935-8690}}\\%
\href{mailto:perricoq@outlook.com}{\nolinkurl{perricoq@outlook.com}}%
}

### 2023-06-07 texor.R

@@ -1,0 +1,9 @@
+# Generated by `rjournal_pdf_article()` using `knitr::purl()`: do not edit by hand
+# Please edit texor.Rmd to modify this file
+
+## ----setup, include=FALSE-----------------------------------------------------
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+library(plotly)
+library(ggplot2)
+library(palmerpenguins)
+

### 2023-06-18 texor.R
Removed files which can be generated from Rmd plus a portion of article about pandoc

@@ -1,9 +1,0 @@
-# Generated by `rjournal_pdf_article()` using `knitr::purl()`: do not edit by hand
-# Please edit texor.Rmd to modify this file
-
-## ----setup, include=FALSE-----------------------------------------------------
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
-library(plotly)
-library(ggplot2)
-library(palmerpenguins)
-

### 2023-06-18 texor.Rmd
A short para on pandoc usage

@@
 
+### Pandoc Usage
+
+Pandoc is a versatile document conversion program written in Haskell that is core
+to numerous documentation systems, including Rmarkdown and Quarto. Pandoc first 
+converts a document into an abstract syntax tree. From this, it can convert to a
+different format, including custom ones. Although creating bespoke writers/readers
+in Lua can be challenging, pandoc includes filters that allow users to customise 
+the output generated based on matching patterns plus some logic to modify them.
+
+The Texor package uses these filters heavily, even for simple tasks like choosing
+which supporting figures to copy. A lua filter constructs a list of image paths, 
+writes it to an intermediate file and feeds them to the Texor package.
+
+This was one of the earliest and most basic Lua filters, and it was part of the 
+texor package:
+
+```Lua
+function Div(el)
+    if el.classes[1] == 'thebibliography' then
+        return { }
+    end
+end
+```
+
+Above filter reads the abstract syntax tree and filters out all the Div elements.
+Then it looks for the class "thebibliography." It turns out that this Div element 
+contains the LaTeX bibliographic records, which usually appear at the very end 
+of papers. We don't need this portion in the article with the "RJ-web-article" 
+layout,  given its inclusion as meta-data in the footer.
+
 LaTeX, Regex, pandoc

### 2023-06-20 texor.Rmd
added authors and sketched in some details in intro

@@
 ---
-title: "Converting 'LaTeX' 'R Journal' Articles into 'RJ-web-articles'"
+title: "Converting LaTeX R Journal Articles into RJ-web-articles"
 date: "2022-11-09"
@@
   - name: Abhishek Ulayil
-    affiliation: Student, Institute of Actuaries of India
+    affiliation: Institute of Actuaries of India
     address:
@@
     email: perricoq@outlook.com
+  - name: Mitch O'Hara-Wild
+    affiliation: Monash University
+    address:
+    - Melbourne, Australia
+    orcid: 0000-0001-6729-7695
+    email: mail@mitchelloharawild.com
+  - name: Christophe Dervieux
+    affiliation: Posit PBC
+    address:
+    - Paris, France
+    orcid: 0000-0003-4474-2498
+    email: christophe.dervieux@gmail.com
+  - name: Heather Turner
+    affiliation: University of Warwick
+    address:
+    - Newport, United Kingdom
+    orcid: 0000-0002-1256-3375
+    email: ht@heatherturner.net
+  - name: Dianne Cook
+    affiliation: Monash University
+    address:
+    - Melbourne, Australia
+    orcid: 0000-0002-3813-7155
+    email: dicook@monash.edu
 type: package
@@
 library(ggplot2)
-library(palmerpenguins)
 ```
@@
 
-Explain the purpose of the work, some background on the rjournal/rjtools[DC],
+The R Journal is the primary open-access outlet for the R community.
+In 2019, with the help of funding from the R Consortium it was decided that it was time to update operations.
+One aspect of this was to change from LaTeX paper submissions to a more reproducible format, where code was embedded in the document, and the output could be both HTML and pdf.
+
+There are numerous benefits of HTML format:
+
+1. Articles can include interactive graphics and tables.
+2. The format is more accessible to screen readers making the work more accessible to vision impaired researchers.
+
+<!-- traditional HTML is the most universally accessible format - Australian Human Rights Commission https://humanrights.gov.au/our-work/disability-rights/world-wide-web-access-disability-discrimination-act-advisory-notes-ver and https://printdisability.org -->
+
+Describe changes in templates, and web site. Desirability to convert the legacy articles. Decided to convert latex input to Rmd, rather than directly html and why.
 
@@
 
+Where this might be useful in the future, other applications.
+
 # Acknowledgments

### 2023-06-21 texor.Rmd
Pre-Processing and Post-Processing

@@
 
-### Pandoc Usage
+### Pandoc does the heavy-lifting
 
@@
 writes it to an intermediate file and feeds them to the Texor package.
+
+### Pre-Processing : Regular Expressions
+
+LaTeX is very descriptive and allows unfettered levels of customization to authors,
+which poses an issue during conversions. RMarkdown is quite a bit restrictive by 
+syntax though you can always embed HTML to expand the feature set. Although 
+pandoc does most of the heavy lifting, there are some instances where even 
+filters cannot assist. A good demonstration of the issue in hand is code 
+environments.
+
+A limitation of pandoc is that it only works with a "verbatim" environment. 
+However, in "RJournal",  we have multiple custom code environments which are not
+supported or defined due to lower-level tex limitations in pandoc. Hence the code 
+environments used to be eliminated during the conversion process, including a Lua
+filter to find these elements and redefine them as code environments would also 
+fail because pandoc would have already tried to process the code inside an unknown 
+code environment, leading to missing code elements lost in the conversion.
+
+### Post-Processing : Lua Filters
 

### 2023-06-24 texor.Rmd
notes on more to include in intro
Jotted down a part about rebib

@@
   - name: Abhishek Ulayil
-    affiliation: Institute of Actuaries of India
+    affiliation: Student, Institute of Actuaries of India
     address:
@@
 
+- R Journal latex template and formatting
+- Quirks of the legacy articles
+
 # Background
@@
 The Texor package uses these filters heavily, even for simple tasks like choosing
-which supporting figures to copy. A lua filter constructs a list of image paths, 
-writes it to an intermediate file and feeds them to the Texor package.
+which supporting figures to copy. A Lua filter constructs a list of image paths, 
+stores them momentarily and the package uses this data to copy images.
 
@@
 LaTeX is very descriptive and allows unfettered levels of customization to authors,
-which poses an issue during conversions. RMarkdown is quite a bit restrictive by 
+which poses an issue during conversions. RMarkdown is quite restrictive by 
 syntax though you can always embed HTML to expand the feature set. Although 
@@
 However, in "RJournal",  we have multiple custom code environments which are not
-supported or defined due to lower-level tex limitations in pandoc. Hence the code 
-environments used to be eliminated during the conversion process, including a Lua
-filter to find these elements and redefine them as code environments would also 
-fail because pandoc would have already tried to process the code inside an unknown 
-code environment, leading to missing code elements lost in the conversion.
+supported or defined due to lower-level tex limitations in pandoc. So the code 
+environments are eliminated during the conversion process. Including a Lua filter
+to find these elements and redefine them as verbatim environments would also be 
+ineffective. If the code environment is not verbatim, then pandoc will also try 
+to process the actual code content as LaTeX commands and will lose out on 
+details in some cases.
 
@@
 This was one of the earliest and most basic Lua filters, and it was part of the 
-texor package:
+\CRANpkg{texor} package:
 
@@
 
-# texor package converting LaTeX to Rmarkdown
+# Texor package converting LaTeX to Rmarkdown
 
@@
 ## Lua Filters with pandoc
-A lot of details about different components of texor
+A lot of details about different components of \CRANpkg{texor}
 
-# rebib package working out bibliography
+# Rebib is in charge of the bibliography
+
+The \CRANpkg{rebib} package addresses the issue with LaTeX articles using 
+built-in bibliography options with or without BibTeX files. While this works 
+well with LaTeX, it won't work with Rmarkdown. Initially the goal was to use 
+external software like Biber to convert the embedded bibliography to BibTeX. 
+However the integration of external software was not viable, hence the 
+experimental idea of a bibliography parser gained momentum.
+
+It was initially a part of the \CRANpkg{texor} package, as those functions grew 
+in features and became more involved; at that point, it made sense to move those
+functions as a separate package. Initially, there were some reservations about 
+the usage of rebib and its stability with various formats. However,
+the package has improved over time and proven to be a good performer.
+
+Rebib's best feature is its ability to merge bibliographies from both the 
+LaTeX and BibTeX files. \CRANpkg{rebib} saves authors from manually 
+converting to BibTeX format.
 

### 2023-07-02 texor.Rmd
Updated title and `\CRANpkg{}` usage

@@
 ---
-title: "Converting LaTeX R Journal Articles into RJ-web-articles"
+title: "Converting LaTeX R Journal Articles into Rmarkdown using texor and rebib"
 date: "2022-11-09"
@@
 different format, including custom ones. Although creating bespoke writers/readers
-in Lua can be challenging, pandoc includes filters that allow users to customise 
+in Lua can be challenging, pandoc includes filters that allow users to customize 
 the output generated based on matching patterns plus some logic to modify them.
 
-The Texor package uses these filters heavily, even for simple tasks like choosing
+The \CRANpkg{texor} package uses these filters heavily, even for simple tasks like choosing
 which supporting figures to copy. A Lua filter constructs a list of image paths, 
@@
 
-# Texor package converting LaTeX to Rmarkdown
+# Converting LaTeX to Markdown
 
@@
 
-# Rebib is in charge of the bibliography
+# Managing bibliography using \CRANpkg{rebib}
 
@@
 functions as a separate package. Initially, there were some reservations about 
-the usage of rebib and its stability with various formats. However,
+the usage of \CRANpkg{rebib} and its stability with various formats. However,
 the package has improved over time and proven to be a good performer.
 
-Rebib's best feature is its ability to merge bibliographies from both the 
-LaTeX and BibTeX files. \CRANpkg{rebib} saves authors from manually 
-converting to BibTeX format.
-
-# getting started
+# Getting started
 
@@
 
-# supplementary Materials
+# Supplementary materials
  github repos for the packages, rjtools

### 2023-07-13 texor.Rmd
revised intro

@@
 
-The R Journal is the primary open-access outlet for the R community.
+The R Journal is the primary open-access outlet for publications produced by the R community.
+It was born in 2008, evolving from a newsletter, that ran from 2001, into a more formal article publication to encourage documenting statistical computing research.
+
+The format is constantly evolving.
+Early articles were typeset using LateX [@latex], from a specific, but changing, template.
+This requires that code is separated from the documentation, and there is a chance that code chunks in the paper don't reproduce the results reported.
+With the emergence of dynamic document systems like R Markdown [@rmarkdown] a tight-coupling of code and documentation is possible.
+Code chunks are dynamically executed when the document is typeset using a system like \CRANpkg{knitr} [@knitr], making reporting of computing research more reproducible.
+
 In 2019, with the help of funding from the R Consortium it was decided that it was time to update operations.
 One aspect of this was to change from LaTeX paper submissions to a more reproducible format, where code was embedded in the document, and the output could be both HTML and pdf.
-
 There are numerous benefits of HTML format:
@@
 1. Articles can include interactive graphics and tables.
-2. The format is more accessible to screen readers making the work more accessible to vision impaired researchers.
+2. The format is more accessible to screen readers making the work more accessible to vision-impaired researchers.
 
@@
 
+This latter point is a reason to consider converting all of the legacy articles into HTML.
+
+A key decision for creating conversion software was to decide to directly convert LaTeX to HTML, or PDF to HTML, or LaTeX to R Markdown, and then use the current journal tools to create the HTML.
+The latter approach was decided to be the most versatile and useful.
+If an article can be converted from LaTeX to R Markdown, it would help authors make the transition to reproducible publishing, beyond what the R Journal needed.
+Once an article is in R Markdown format it can be adapted to include the code for dynamic execution.
+
+In addition to article format, changes to the web site structure were important for delivering the publication.
+Web site architectures are also constantly evolving, and the emergence of \CRANpkg{distill} [@distill] allows for the journal web site to optimally deliver R Markdown articles.
+
+The \CRANpkg{rjtools} was developed to create articles using R Markdown for the R Journal, and to embed them into the journal web site.
+The packages described here, \CRANpkg{texor} and \CRANpkg{rebib} describes software to convert legacy LaTeX format articles into Rmarkdown, so that they can be rendered in HTML in the new web site.
+
+The paper is organised as follows.
+Section XXX gives an overview of the conversion process.
+Section XXX describes examples of pre-processing using regular expressions, and section XXX provides examples of post-processing using lua filters.
+Section XXX describes tools to do special handling of bibliography files.
+Section XXX runs through an example conversion process.
+
+<!--
 Describe changes in templates, and web site. Desirability to convert the legacy articles. Decided to convert latex input to Rmd, rather than directly html and why.
@@
 - Quirks of the legacy articles
+-->
 

### 2023-07-13 texor.Rmd
tried to make the paper organised, so it is easier to fill in

@@
 
-# Background
+# Converting from LaTeX to R Markdown
 
-### Pandoc does the heavy-lifting
+The decision to convert to R Markdown format means that the output to pdf and HTML will depend on Pandoc [@pandoc]. 
+Pandoc is a versatile document conversion program written in Haskell that is core to numerous documentation systems, including Rmarkdown and Quarto.
+Pandoc first converts a document into an abstract syntax tree.
+From this, it can convert to a different format, including custom ones.
+Although creating bespoke writers/readers in `Lua` can be challenging, pandoc includes filters that allow users to customize the output generated based on matching patterns plus some logic to modify them.
 
-Pandoc is a versatile document conversion program written in Haskell that is core
-to numerous documentation systems, including Rmarkdown and Quarto. Pandoc first 
-converts a document into an abstract syntax tree. From this, it can convert to a
-different format, including custom ones. Although creating bespoke writers/readers
-in Lua can be challenging, pandoc includes filters that allow users to customize 
-the output generated based on matching patterns plus some logic to modify them.
+**DC: Can you explain what a Lua filter is?**
 
-The \CRANpkg{texor} package uses these filters heavily, even for simple tasks like choosing
-which supporting figures to copy. A Lua filter constructs a list of image paths, 
-stores them momentarily and the package uses this data to copy images.
+The \CRANpkg{texor} package uses these filters heavily, even for simple tasks like choosing which supporting figures to copy.
+A `Lua` filter constructs a list of image paths, stores them momentarily and the package uses this data to copy images.
+However, \CRANpkg{texor} also includes many pre-processing functions that employ regular expressions.
+This is needed to convert some text and LaTeX commands that are not recognized or handled by Pandoc.
 
-### Pre-Processing : Regular Expressions
+**DC: Add a diagram here showing flow of work, just pre-process .tex -> pandoc -> lua ???**
 
-LaTeX is very descriptive and allows unfettered levels of customization to authors,
-which poses an issue during conversions. RMarkdown is quite restrictive by 
-syntax though you can always embed HTML to expand the feature set. Although 
-pandoc does most of the heavy lifting, there are some instances where even 
-filters cannot assist. A good demonstration of the issue in hand is code 
-environments.
+**Maybe more details on the supporting packages here: LaTeX, Regex, pandoc building blocks of the packages, methods used in them [MOW]**
+
+## Pre-processing using regular expressions
+
+LaTeX is very descriptive and allows unfettered levels of customization to authors, which poses an issue during conversions.
+RMarkdown is quite restrictive by  syntax though you can always embed HTML to expand the feature set.
+Although
+pandoc does most of the heavy lifting, there are some instances where even  filters cannot assist. A good demonstration of the issue in hand is code environments.
 
 A limitation of pandoc is that it only works with a "verbatim" environment.
-However, in "RJournal",  we have multiple custom code environments which are not
-supported or defined due to lower-level tex limitations in pandoc. So the code 
-environments are eliminated during the conversion process. Including a Lua filter
-to find these elements and redefine them as verbatim environments would also be 
-ineffective. If the code environment is not verbatim, then pandoc will also try 
-to process the actual code content as LaTeX commands and will lose out on 
-details in some cases.
+However, in "RJournal",  we have multiple custom code environments which are not supported or defined due to lower-level tex limitations in pandoc.
+So the code  environments are eliminated during the conversion process.
+Including a Lua filter to find these elements and redefine them as verbatim environments would also be  ineffective.
+If the code environment is not verbatim, then pandoc will also try to process the actual code content as LaTeX commands and will lose out on details in some cases.
 
-### Post-Processing : Lua Filters
+**Be sure to mention the hard parts**
 
-This was one of the earliest and most basic Lua filters, and it was part of the 
-\CRANpkg{texor} package:
+## Post-Processing : Lua Filters
+
+This was one of the earliest and most basic Lua filters, and it was part of the \CRANpkg{texor} package:
 
@@
 Above filter reads the abstract syntax tree and filters out all the Div elements.
-Then it looks for the class "thebibliography." It turns out that this Div element 
-contains the LaTeX bibliographic records, which usually appear at the very end 
-of papers. We don't need this portion in the article with the "RJ-web-article" 
-layout,  given its inclusion as meta-data in the footer.
+Then it looks for the class "thebibliography." It turns out that this Div element  contains the LaTeX bibliographic records, which usually appear at the very end
+of papers. We don't need this portion in the article with the "RJ-web-article" layout,  given its inclusion as meta-data in the footer.
 
-LaTeX, Regex, pandoc
-building blocks of the packages, methods used in them [MOW]
+# Structure of the \CRANpkg{texor} package
 
-# Converting LaTeX to Markdown
+**DC: List the main functions in an organised way, could be the pre-processing functions, the post-processing functions, utility functions.
+This can be also be summarized in a table, but there needs to be text explanations of the main functionality.**
 
-## Also mention the hard parts
-## Lua Filters with pandoc
-A lot of details about different components of \CRANpkg{texor}
+# Using \CRANpkg{texor}
 
-# Managing bibliography using \CRANpkg{rebib}
+**Install instructions**
 
-The \CRANpkg{rebib} package addresses the issue with LaTeX articles using 
-built-in bibliography options with or without BibTeX files. While this works 
-well with LaTeX, it won't work with Rmarkdown. Initially the goal was to use 
-external software like Biber to convert the embedded bibliography to BibTeX. 
-However the integration of external software was not viable, hence the 
-experimental idea of a bibliography parser gained momentum.
+**Examples / getting started / usage and so on**
 
-It was initially a part of the \CRANpkg{texor} package, as those functions grew 
-in features and became more involved; at that point, it made sense to move those
-functions as a separate package. Initially, there were some reservations about 
-the usage of \CRANpkg{rebib} and its stability with various formats. However,
-the package has improved over time and proven to be a good performer.
+# Managing the bibliography using \CRANpkg{rebib}
 
-# Getting started
+## Overview
 
- Examples/ getting started / usage and so on
+The \CRANpkg{rebib} package addresses the issue with LaTeX articles using  built-in bibliography options with or without BibTeX files.
+While this works  well with LaTeX, it won't work with Rmarkdown.
+Initially the goal was to use external software like Biber to convert the embedded bibliography to BibTeX.
+However the integration of external software was not viable, hence the experimental idea of a bibliography parser gained momentum.
+
+It was initially a part of the \CRANpkg{texor} package, as those functions grew in features and became more involved; at that point, it made sense to move those functions as a separate package.
+Initially, there were some reservations about
+the usage of \CRANpkg{rebib} and its stability with various formats. However, the package has improved over time and proven to be a good performer.
+
+## Using \CRANpkg{rebib}
+
+**Install instructions**
+
+**Examples / getting started / usage and so on**
 
@@
 # Supplementary materials
- github repos for the packages, rjtools
+
+ github repos for the packages, texor, rebib, rjtools

### 2023-07-15 supplement/figure-env/figure-example.tex
First supplementary material

% !TeX root = RJwrapper.tex
\title{Working with figure environments}
\author{by Abhishek Ulayil}

\maketitle

\abstract{
This is a small sample article to demonstrate usage of texor to convert figure environments.
}

\section{Introduction}

Images are an essential component in any article, However due to the differences in the support for various graphic formats between LaTeX and markdown/HTML we need to fallback on raster graphics. It is well summarized in the table \ref{tab:graphics}

\begin{table}[t!]
\begin{center}
\begin{tabular}{l | llll }
 \hline
 Graphics Format & LaTeX & Markdown & Rmarkdown & HTML \\
 \hline
 PNG       & ✓ & ✓ & ✓ & ✓ \\
 JPG       & ✓ & ✓ & ✓ & ✓ \\
 PDF       & ✓ & ✗ & ✗ & ✗ \\
 SVG       & ✗ & ✓ & ✓ & ✓ \\
 Tikz      & ✓ & ✗ & ✓ & ✗ \\
 Algorithm & ✓ & ✗ & ✗ & ✗ \\
\hline
\end{tabular}
\label{tab:graphics} \caption{Image Format support in various Markup/Typesetting Languages}
\end{center}
\end{table}

\section{Image without width parameters}

This section may contain a figure such as Figure~\ref{figure:rlogo}.

\section{Image with width parameters}

\begin{verbatim}
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.35\textwidth]{Rlogo-5.png}
  \caption{The logo of R.}
  \label{figure:rlogo}
\end{figure}
\end{verbatim}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.35\textwidth]{Rlogo-5.png}
  \caption{The logo of R.}
  \label{figure:rlogo}
\end{figure}

\section {Images in PDF format}

Image \ref{fig:normal} is a graphical representation of normal distribution.

\begin{verbatim}
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5\textwidth]{normal}
  \caption{PDF of a normal distribution}
  \label{fig:normal}
\end{figure}
\end{verbatim}
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5\textwidth]{normal}
  \caption{PDF of a normal distribution}
  \label{fig:normal}
\end{figure}

\subsection{Multiple images}
 Pandoc v3 and above now supports a new Figure object which supports multiple 
 images side by side or in a grid format.
\subsubsection{Two or more Images side by side}
\begin{verbatim}
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{Rlogo-5.png}\includegraphics[width=0.45\textwidth]{normal}
  \caption{Images side by side}
  \label{fig:twoimages}
\end{figure}
\end{verbatim}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{Rlogo-5.png}\includegraphics[width=0.45\textwidth]{normal}
  \caption{Images side by side}
  \label{fig:twoimages}
\end{figure}
\subsubsection{Four Images in a grid}

\begin{verbatim}
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{Rlogo-5.png}\includegraphics[width=0.45\textwidth]{normal}

  \includegraphics[width=0.45\textwidth]{normal}\includegraphics[width=0.45\textwidth]{Rlogo-5.png}
  \caption{Multiple images in a grid}
  \label{fig:fourimages}
\end{figure}

\end{verbatim}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{Rlogo-5.png}\includegraphics[width=0.45\textwidth]{normal}

  \includegraphics[width=0.45\textwidth]{normal}\includegraphics[width=0.45\textwidth]{Rlogo-5.png}
  \caption{Multiple images in a grid}
  \label{fig:fourimages}
\end{figure}

\section{Tikz images}

Here is a tikz image example in fig \ref{fig:tikz} adapted from \citet{tikzflow}.

Another interesting aspect of including tikz image here is that you can modify the
source and re-convert without making any other changes. Figure will get updated
in the generated html article.

\subsubsection{Tikz Code:}
\begin{verbatim}
\begin{figure}
\includegraphics{tikz/figtikz.png}
%% Generated Image will included as a PNG above automatically
  \centering
\tikzstyle{process} = [rectangle, rounded corners,
minimum width=3cm, 
minimum height=1cm,
text centered, 
draw=black]
\tikzstyle{arrow} = [thick,->,>=stealth]
\begin{tikzpicture}[node distance=4cm]
%Nodes
\node (start) [process] {LaTeX Source with Tikz Image};
\node (isolate) [process, below right of=start] {Isolated tikz env};
\node (pro1) [process, below left of=isolate] {Standalone PDF file};
\node (pro2) [process, below right of=pro1, yshift=-0.5cm] {PNG file};
\node (link) [process, below left of=pro2] {Include PNG in article};
\node (stop) [process, below right of=link] {Convert To RJ-web-article};
% arrows
\draw [arrow] (start) -- node[anchor=east] {Extract Figure Block} (isolate);
\draw [arrow] (isolate) -- node[anchor=east] {Compile} (pro1);
\draw [arrow] (pro1) -- node[anchor=east] {Convert to raster} (pro2);
\draw [arrow] (pro2) -- node[anchor=east] {Include back to source as image} (link);
\draw [arrow] (link) -- node[anchor=east] {Using Pandoc} (stop);
\end{tikzpicture}
\caption{Tikz Image example}
  \label{fig:tikz}
\end{figure}

\end{verbatim}

\subsubsection{Resultant Figure :}
\begin{figure}
\includegraphics{tikz/figtikz.png}

%% Generated Image will included as a PNG above
  \centering
\tikzstyle{process} = [rectangle, rounded corners,
minimum width=3cm, 
minimum height=1cm,
text centered, 
draw=black]
\tikzstyle{arrow} = [thick,->,>=stealth]
\begin{tikzpicture}[node distance=4cm]
%Nodes
\node (start) [process] {LaTeX Source with Tikz Image};
\node (isolate) [process, below right of=start] {Isolated tikz env};
\node (pro1) [process, below left of=isolate] {Standalone PDF file};
\node (pro2) [process, below right of=pro1, yshift=-0.5cm] {PNG file};
\node (link) [process, below left of=pro2] {Include PNG in article};
\node (stop) [process, below right of=link] {Convert To RJ-web-article};
% arrows
\draw [arrow] (start) -- node[anchor=east] {Extract Figure Block} (isolate);
\draw [arrow] (isolate) -- node[anchor=east] {Compile} (pro1);
\draw [arrow] (pro1) -- node[anchor=east] {Convert to raster} (pro2);
\draw [arrow] (pro2) -- node[anchor=east] {Include back to source as image} (link);
\draw [arrow] (link) -- node[anchor=east] {Using Pandoc} (stop);
\end{tikzpicture}
\caption{Tikz Image example}
  \label{fig:tikz}
\end{figure}

\section{Algorithm2e diagrams}

we do support algorithm2e diagrams and images, these will be numbered differently
and we strongly suggest to use "alg:" in labels for best results. 
Here is an example of algorithm \ref{alg:how}, referenced from \citet{algoexample}
\begin{figure}[H]
\includegraphics{alg/alghow.png}

\SetAlgoLined
\KwData{this text}
\KwResult{how to write algorithm with \LaTeX2e }
initialization\;
\While{not at end of this document}{
read current\;
\eIf{understand}{
go to next section\;
current section becomes this one\;
}{
go back to the beginning of current section\;
}
}
\caption{How to write algorithms}
  \label{alg:how}
\end{figure}

\section{Summary}

In summary the \CRANpkg{texor} package supports:
\begin{itemize}
\item Almost all image formats in LaTeX.
\item Algorithm and tikz as well in some capacity.
\item Multiple images in grid,side-by-side configuration.
\item Image Captions with Numbering and Labelling.
\end{itemize}

\begin{thebibliography}{20}
    \providecommand{\natexlab}[1]{#1}
    \providecommand{\url}[1]{\texttt{#1}}
    \expandafter\ifx\csname urlstyle\endcsname\relax
      \providecommand{\doi}[1]{doi: #1}\else
      \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi
\bibitem[Ihaka, Ross and Gentleman, Robert]{ihaka:1996}
Ihaka, Ross and Gentleman, Robert
\newblock R: A Language for Data Analysis and Graphics
\newblock \emph{Journal of Computational and Graphical Statistics}, 3:\penalty0
299--314, 1996.
\newblock URL : \url{https://doi.org/10.1080/10618600.1996.10474713}

\bibitem[R Core Team]{R}
R Core Team
\newblock R: A Language and Environment for Statistical Computing
\newblock \emph{R Foundation for Statistical Computing}, Vienna, Austria \penalty0 2016.
\newblock URL : \url{https://www.R-project.org/}, ISBN 3-900051-07-0

\bibitem[Cassidy]{tikzflow}
Josh Cassidy
\newblock LaTeX Graphics using TikZ: A Tutorial for Beginners (Part 3)—Creating Flowcharts
\newblock \emph{Overleaf tutorials} 2013
\newblock URL \url{https://www.overleaf.com/learn/latex/LaTeX_Graphics_using_TikZ%3A_A_Tutorial_for_Beginners_(Part_3)%E2%80%94Creating_Flowcharts}

\bibitem[Fiorio]{algoexample}
Christophe Fiorio
\newblock algorithm2e.sty — package for algorithms, release 5.2
\newblock \emph{CTAN}, 2017
\newblock URL \url{https://mirror.kku.ac.th/CTAN/macros/latex/contrib/algorithm2e/doc/algorithm2e.pdf}
\end{thebibliography}

\address{%
Abhishek Ulayil\\
Student, Institute of Actuaries of India\\%
Mumbai, India\\
ORCiD: 0009-0000-6935-8690\\
}

### 2023-07-18 supplement/figure-env/figure-example.tex
Fixed algorithm image environment
Made supplementary material compilable to both PDF and HTML

@@
 % !TeX root = RJwrapper.tex
-\title{Working with figure environments}
+\title{Working with figure environments in texor}
 \author{by Abhishek Ulayil}
@@
 Images are an essential component in any article, However due to the differences in the support for various graphic formats between LaTeX and markdown/HTML we need to fallback on raster graphics. 
-It is well summarized in the table \ref{tab:graphics}
+It is well summarized in the table \ref{table:1}
 
-\begin{table}[t!]
-\begin{center}
+\begin{table}[htbp]
+\centering
 \begin{tabular}{l | llll }
@@
  \hline
- PNG       & ✓ & ✓ & ✓ & ✓ \\
- JPG       & ✓ & ✓ & ✓ & ✓ \\
- PDF       & ✓ & ✗ & ✗ & ✗ \\
- SVG       & ✗ & ✓ & ✓ & ✓ \\
- Tikz      & ✓ & ✗ & ✓ & ✗ \\
- Algorithm & ✓ & ✗ & ✗ & ✗ \\
+ PNG       & Yes & Yes & Yes & Yes \\
+ JPG       & Yes & Yes & Yes & Yes \\
+ PDF       & Yes & No & No & No \\
+ SVG       & No & Yes & Yes & Yes \\
+ Tikz      & Yes & No & Yes & No \\
+ Algorithm & Yes & No & No & No \\
 \hline
 \end{tabular}
-\label{tab:graphics} \caption{Image Format support in various Markup/Typesetting Languages}
-\end{center}
+\caption{Image Format support in various Markup/Typesetting Languages}
+\label{table:1}
 \end{table}
 
-\section{Image without width parameters}
-
+\section{Image with width parameters}
 This section may contain a figure such as Figure~\ref{figure:rlogo}.
-
-\section{Image with width parameters}
+This is the most basic example of figure.
 
@@
 
-\section {Images in PDF format}
+\section{Images in PDF format}
 
@@
 
-\subsection{Multiple images}
- Pandoc v3 and above now supports a new Figure object which supports multiple 
+\section{Multiple images}
+ Pandoc v3 and above now supports a new Figure object \citep{pandoc} which supports multiple 
  images side by side or in a grid format.
+%% Two images side by side
 \subsubsection{Two or more Images side by side}
@@
 
-\begin{figure}[htbp]
+\begin{figure*}[htbp]
   \centering
@@
   \label{fig:twoimages}
-\end{figure}
+\end{figure*}
+
+%% Four images in a grid
 \subsubsection{Four Images in a grid}
@@
   \includegraphics[width=0.45\textwidth]{Rlogo-5.png}\includegraphics[width=0.45\textwidth]{normal}
-
   \includegraphics[width=0.45\textwidth]{normal}\includegraphics[width=0.45\textwidth]{Rlogo-5.png}
@@
 
-Here is a tikz image example in fig \ref{fig:tikz} adapted from \citet{tikzflow}.
+Here is a tikz image example in fig \ref{fig:tikz} adapted from \citep{casflow}.
 
@@
 \begin{figure}
-\includegraphics{tikz/figtikz.png}
+
 %% Generated Image will included as a PNG above automatically
@@
 \subsubsection{Resultant Figure :}
-\begin{figure}
-\includegraphics{tikz/figtikz.png}
+\begin{figure*}
 
@@
   \label{fig:tikz}
-\end{figure}
+\end{figure*}
 
@@
 and we strongly suggest to use "alg:" in labels for best results. 
-Here is an example of algorithm \ref{alg:how}, referenced from \citet{algoexample}
-\begin{figure}[H]
-\includegraphics{alg/alghow.png}
+Here is an example of algorithm \ref{alg:how}, referenced from \cite{algoexample}
 
+\begin{verbatim}
+\begin{algorithm}[htbp]
 \SetAlgoLined
@@
   \label{alg:how}
-\end{figure}
+\end{algorithm}
+\end{verbatim}
+
+\begin{algorithm}[htbp]
+\SetAlgoLined
+\KwData{this text}
+\KwResult{how to write algorithm with \LaTeX2e }
+initialization\;
+\While{not at end of this document}{
+read current\;
+\eIf{understand}{
+go to next section\;
+current section becomes this one\;
+}{
+go back to the beginning of current section\;
+}
+}
+\caption{How to write algorithms}
+  \label{alg:how}
+\end{algorithm}
 
@@
 
-\begin{thebibliography}{20}
+\begin{thebibliography}{4}
     \providecommand{\natexlab}[1]{#1}
@@
       \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi
-\bibitem[Ihaka, Ross and Gentleman, Robert]{ihaka:1996}
-Ihaka, Ross and Gentleman, Robert
-\newblock R: A Language for Data Analysis and Graphics
-\newblock \emph{Journal of Computational and Graphical Statistics}, 3:\penalty0
-299--314, 1996.
-\newblock URL : \url{https://doi.org/10.1080/10618600.1996.10474713}
-
-\bibitem[R Core Team]{R}
-R Core Team
-\newblock R: A Language and Environment for Statistical Computing
-\newblock \emph{R Foundation for Statistical Computing}, Vienna, Austria \penalty0 2016.
-\newblock URL : \url{https://www.R-project.org/}, ISBN 3-900051-07-0
-
-\bibitem[Cassidy]{tikzflow}
+\bibitem[Cassidy(2013)]{casflow}
 Josh Cassidy
 \newblock LaTeX Graphics using TikZ: A Tutorial for Beginners (Part 3)—Creating Flowcharts
-\newblock \emph{Overleaf tutorials} 2013
-\newblock URL \url{https://www.overleaf.com/learn/latex/LaTeX_Graphics_using_TikZ%3A_A_Tutorial_for_Beginners_(Part_3)%E2%80%94Creating_Flowcharts}
+\newblock \emph{Overleaf tutorials} \penalty0 2013
+\newblock URL \url{https://www.overleaf.com/learn/latex/}
 
-\bibitem[Fiorio]{algoexample}
+\bibitem[Fiorio (2017)]{algoexample}
 Christophe Fiorio
 \newblock algorithm2e.sty — package for algorithms, release 5.2
-\newblock \emph{CTAN}, 2017
+\newblock \emph{CTAN},\penalty0 2017
 \newblock URL \url{https://mirror.kku.ac.th/CTAN/macros/latex/contrib/algorithm2e/doc/algorithm2e.pdf}
+
+\bibitem[Krewinkel, Lucero (2023)]{pandoc}
+Albert Krewinkel and Aner Lucero
+\newblock pandoc 3.0 Release notes
+\newblock \emph{pandoc} \penalty0 2023
+\newblock URL \url{https://pandoc.org/releases.html#pandoc-3.0-2023-01-18}
+
 \end{thebibliography}

### 2023-07-19 texor.tex
Included rendered html,pdf versions of the article

@@
 % !TeX root = RJwrapper.tex
-\title{Converting `LaTeX' `R Journal' Articles into `RJ-web-articles'}
-\author{by Abhishek Ulayil}
+\title{Converting LaTeX R Journal Articles into Rmarkdown using texor and rebib}
+
+\author{by Abhishek Ulayil, Mitch O'Hara-Wild, Christophe Dervieux, Heather Turner, and Dianne Cook}
 
@@
 
-Explain the purpose of the work, some background on the rjournal/rjtools{[}DC{]},
+The R Journal is the primary open-access outlet for publications produced by the R community.
+It was born in 2008, evolving from a newsletter, that ran from 2001, into a more formal article publication to encourage documenting statistical computing research.
 
-\hypertarget{background}{%
-\section{Background}\label{background}}
+The format is constantly evolving.
+Early articles were typeset using LateX (The LaTeX Project 2023), from a specific, but changing, template.
+This requires that code is separated from the documentation, and there is a chance that code chunks in the paper don't reproduce the results reported.
+With the emergence of dynamic document systems like R Markdown (Xie, Allaire, and Grolemund 2018) a tight-coupling of code and documentation is possible.
+Code chunks are dynamically executed when the document is typeset using a system like \CRANpkg{knitr} (Xie 2015), making reporting of computing research more reproducible.
 
-LaTeX, Regex, pandoc
-building blocks of the packages, methods used in them {[}MOW{]}
+In 2019, with the help of funding from the R Consortium it was decided that it was time to update operations.
+One aspect of this was to change from LaTeX paper submissions to a more reproducible format, where code was embedded in the document, and the output could be both HTML and pdf.
+There are numerous benefits of HTML format:
 
-\hypertarget{texor-package-converting-latex-to-rmarkdown}{%
-\section{texor package converting LaTeX to Rmarkdown}\label{texor-package-converting-latex-to-rmarkdown}}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Articles can include interactive graphics and tables.
+\item
+  The format is more accessible to screen readers making the work more accessible to vision-impaired researchers.
+\end{enumerate}
 
-\hypertarget{also-mention-the-hard-parts}{%
-\subsection{Also mention the hard parts}\label{also-mention-the-hard-parts}}
+This latter point is a reason to consider converting all of the legacy articles into HTML.
 
-\hypertarget{lua-filters-with-pandoc}{%
-\subsection{Lua Filters with pandoc}\label{lua-filters-with-pandoc}}
+A key decision for creating conversion software was to decide to directly convert LaTeX to HTML, or PDF to HTML, or LaTeX to R Markdown, and then use the current journal tools to create the HTML.
+The latter approach was decided to be the most versatile and useful.
+If an article can be converted from LaTeX to R Markdown, it would help authors make the transition to reproducible publishing, beyond what the R Journal needed.
+Once an article is in R Markdown format it can be adapted to include the code for dynamic execution.
 
-A lot of details about different components of texor
+In addition to article format, changes to the web site structure were important for delivering the publication.
+Web site architectures are also constantly evolving, and the emergence of \CRANpkg{distill} (Dervieux et al.
+2022) allows for the journal web site to optimally deliver R Markdown articles.
 
-\hypertarget{rebib-package-working-out-bibliography}{%
-\section{rebib package working out bibliography}\label{rebib-package-working-out-bibliography}}
+The \CRANpkg{rjtools} was developed to create articles using R Markdown for the R Journal, and to embed them into the journal web site.
+The packages described here, \CRANpkg{texor} and \CRANpkg{rebib} describes software to convert legacy LaTeX format articles into Rmarkdown, so that they can be rendered in HTML in the new web site.
 
-\hypertarget{getting-started}{%
-\section{getting started}\label{getting-started}}
+The paper is organised as follows.
+Section XXX gives an overview of the conversion process.
+Section XXX describes examples of pre-processing using regular expressions, and section XXX provides examples of post-processing using lua filters.
+Section XXX describes tools to do special handling of bibliography files.
+Section XXX runs through an example conversion process.
 
-Examples/ getting started / usage and so on
+\hypertarget{converting-from-latex-to-r-markdown}{%
+\section{Converting from LaTeX to R Markdown}\label{converting-from-latex-to-r-markdown}}
+
+The decision to convert to R Markdown format means that the output to pdf and HTML will depend on Pandoc (MacFarlane 2023).
+Pandoc is a versatile document conversion program written in Haskell that is core to numerous documentation systems, including Rmarkdown and Quarto.
+Pandoc first converts a document into an abstract syntax tree.
+From this, it can convert to a different format, including custom ones.
+Although creating bespoke writers/readers in \texttt{Lua} can be challenging, pandoc includes filters that allow users to customize the output generated based on matching patterns plus some logic to modify them.
+
+\textbf{DC: Can you explain what a Lua filter is?}
+
+The \CRANpkg{texor} package uses these filters heavily, even for simple tasks like choosing which supporting figures to copy.
+A \texttt{Lua} filter constructs a list of image paths, stores them momentarily and the package uses this data to copy images.
+However, \CRANpkg{texor} also includes many pre-processing functions that employ regular expressions.
+This is needed to convert some text and LaTeX commands that are not recognized or handled by Pandoc.
+
+\textbf{DC: Add a diagram here showing flow of work, just pre-process .tex -\textgreater{} pandoc -\textgreater{} lua ???}
+
+\textbf{Maybe more details on the supporting packages here: LaTeX, Regex, pandoc building blocks of the packages, methods used in them {[}MOW{]}}
+
+\hypertarget{pre-processing-using-regular-expressions}{%
+\subsection{Pre-processing using regular expressions}\label{pre-processing-using-regular-expressions}}
+
+LaTeX is very descriptive and allows unfettered levels of customization to authors, which poses an issue during conversions.
+RMarkdown is quite restrictive by syntax though you can always embed HTML to expand the feature set.
+Although
+pandoc does most of the heavy lifting, there are some instances where even filters cannot assist. A good demonstration of the issue in hand is code environments.
+
+A limitation of pandoc is that it only works with a ``verbatim'' environment.
+However, in ``RJournal'', we have multiple custom code environments which are not supported or defined due to lower-level tex limitations in pandoc.
+So the code environments are eliminated during the conversion process.
+Including a Lua filter to find these elements and redefine them as verbatim environments would also be ineffective.
+If the code environment is not verbatim, then pandoc will also try to process the actual code content as LaTeX commands and will lose out on details in some cases.
+
+\textbf{Be sure to mention the hard parts}
+
+\hypertarget{post-processing-lua-filters}{%
+\subsection{Post-Processing : Lua Filters}\label{post-processing-lua-filters}}
+
+This was one of the earliest and most basic Lua filters, and it was part of the \CRANpkg{texor} package:
+
+\begin{verbatim}
+function Div(el)
+    if el.classes[1] == 'thebibliography' then
+        return { }
+    end
+end
+\end{verbatim}
+
+Above filter reads the abstract syntax tree and filters out all the Div elements.
+Then it looks for the class ``thebibliography.'' It turns out that this Div element contains the LaTeX bibliographic records, which usually appear at the very end
+of papers. We don't need this portion in the article with the ``RJ-web-article'' layout, given its inclusion as meta-data in the footer.
+
+\hypertarget{structure-of-the-package}{%
+\section{\texorpdfstring{Structure of the \CRANpkg{texor} package}{Structure of the  package}}\label{structure-of-the-package}}
+
+\textbf{DC: List the main functions in an organised way, could be the pre-processing functions, the post-processing functions, utility functions.
+This can be also be summarized in a table, but there needs to be text explanations of the main functionality.}
+
+\hypertarget{using}{%
+\section{\texorpdfstring{Using \CRANpkg{texor}}{Using }}\label{using}}
+
+\textbf{Install instructions}
+
+\textbf{Examples / getting started / usage and so on}
+
+\hypertarget{managing-the-bibliography-using}{%
+\section{\texorpdfstring{Managing the bibliography using \CRANpkg{rebib}}{Managing the bibliography using }}\label{managing-the-bibliography-using}}
+
+\hypertarget{overview}{%
+\subsection{Overview}\label{overview}}
+
+The \CRANpkg{rebib} package addresses the issue with LaTeX articles using built-in bibliography options with or without BibTeX files.
+While this works well with LaTeX, it won't work with Rmarkdown.
+Initially the goal was to use external software like Biber to convert the embedded bibliography to BibTeX.
+However the integration of external software was not viable, hence the experimental idea of a bibliography parser gained momentum.
+
+It was initially a part of the \CRANpkg{texor} package, as those functions grew in features and became more involved; at that point, it made sense to move those functions as a separate package.
+Initially, there were some reservations about
+the usage of \CRANpkg{rebib} and its stability with various formats. However, the package has improved over time and proven to be a good performer.
+
+\hypertarget{using-1}{%
+\subsection{\texorpdfstring{Using \CRANpkg{rebib}}{Using }}\label{using-1}}
+
+\textbf{Install instructions}
+
+\textbf{Examples / getting started / usage and so on}
 
@@
 \section{Summary}\label{summary}}
+
+Where this might be useful in the future, other applications.
 
@@
 \hypertarget{supplementary-materials}{%
-\section{supplementary Materials}\label{supplementary-materials}}
+\section{Supplementary materials}\label{supplementary-materials}}
 
-github repos for the packages, rjtools
+github repos for the packages, texor, rebib, rjtools
+
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-distill}{}}%
+Dervieux, Christophe, JJ Allaire, Rich Iannone, Alison Presmanes Hill, and Yihui Xie.
+2022.
+\emph{Distill: 'R Markdown' Format for Scientific and Technical Writing}.
+\url{https://CRAN.R-project.org/package=distill}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-pandoc}{}}%
+MacFarlane, John. 2023. {``Pandoc: A Universal Document Converter.''} https://pandoc.org.
+
+\leavevmode\vadjust pre{\hypertarget{ref-latex}{}}%
+The LaTeX Project. 2023. {``LaTeX -- a Document Preparation System.''} https://www.latex-project.org.
+
+\leavevmode\vadjust pre{\hypertarget{ref-knitr}{}}%
+Xie, Yihui. 2015. \emph{Dynamic Documents with {R} and Knitr}. 2nd ed. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://yihui.org/knitr/}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-rmarkdown}{}}%
+Xie, Yihui, J. J. Allaire, and Garrett Grolemund. 2018. \emph{R Markdown: The Definitive Guide}. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://bookdown.org/yihui/rmarkdown}.
+
+\end{CSLReferences}
 
@@
 }
+
+\address{%
+Mitch O'Hara-Wild\\
+Monash University\\%
+Melbourne, Australia\\
+%
+%
+\textit{ORCiD: \href{https://orcid.org/0000-0001-6729-7695}{0000-0001-6729-7695}}\\%
+\href{mailto:mail@mitchelloharawild.com}{\nolinkurl{mail@mitchelloharawild.com}}%
+}
+
+\address{%
+Christophe Dervieux\\
+Posit PBC\\%
+Paris, France\\
+%
+%
+\textit{ORCiD: \href{https://orcid.org/0000-0003-4474-2498}{0000-0003-4474-2498}}\\%
+\href{mailto:christophe.dervieux@gmail.com}{\nolinkurl{christophe.dervieux@gmail.com}}%
+}
+
+\address{%
+Heather Turner\\
+University of Warwick\\%
+Newport, United Kingdom\\
+%
+%
+\textit{ORCiD: \href{https://orcid.org/0000-0002-1256-3375}{0000-0002-1256-3375}}\\%
+\href{mailto:ht@heatherturner.net}{\nolinkurl{ht@heatherturner.net}}%
+}
+
+\address{%
+Dianne Cook\\
+Monash University\\%
+Melbourne, Australia\\
+%
+%
+\textit{ORCiD: \href{https://orcid.org/0000-0002-3813-7155}{0000-0002-3813-7155}}\\%
+\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}%
+}

### 2023-07-19 texor.R

@@ -1,0 +1,8 @@
+# Generated by `rjournal_pdf_article()` using `knitr::purl()`: do not edit by hand
+# Please edit texor.Rmd to modify this file
+
+## ----setup, include=FALSE-----------------------------------------------------
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+library(plotly)
+library(ggplot2)
+

### 2023-07-19 supplement/figure-env/RJwrapper.Rmd
updated supplementary material for figures

--- 2023-07-18 supplement/figure-env/figure-example.tex
@@
-% !TeX root = RJwrapper.tex
-\title{Working with figure environments in texor}
-\author{by Abhishek Ulayil}
+---
+title: Working with figure environments in texor
+abstract: This is a small sample article to demonstrate usage of texor to convert
+  figure environments.
+author:
+- name: Abhishek Ulayil
+  affiliation: Student, Institute of Actuaries of India
+  orcid: 0009-0000-6935-8690
+  address: Mumbai, India
+date: '2009-06-01'
+date_received: ~
+journal:
+  firstpage: ~
+  lastpage: ~
+volume: 1
+issue: 1
+slug: figure-env
+packages:
+  cran: texor
+  bioc: []
+preview: preview.png
+bibliography: figure-example.bib
+CTV: ~
+output:
+  rjtools::rjournal_web_article:
+    self_contained: yes
+    toc: no
+    legacy_pdf: yes
+    web_only: yes
+    mathjax: https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-chtml-full.js
+    md_extension: -tex_math_single_backslash
 
-\maketitle
+---
 
-\abstract{
-This is a small sample article to demonstrate usage of texor to convert figure environments.
-}
+::: article
+# Introduction
 
-\section{Introduction}
+Images are an essential component in any article, However due to the
+differences in the support for various graphic formats between LaTeX and
+markdown/HTML we need to fallback on raster graphics. It is well
+summarized in the table [1](#table:1)
 
-Images are an essential component in any article, However due to the differences in the support for various graphic formats between LaTeX and markdown/HTML we need to fallback on raster graphics. 
-It is well summarized in the table \ref{table:1}
+::: {#table:1}
+  -------------------------------------------------------
+  Graphics Format   LaTeX   Markdown   Rmarkdown   HTML
+  ----------------- ------- ---------- ----------- ------
+  PNG               Yes     Yes        Yes         Yes
 
-\begin{table}[htbp]
-\centering
-\begin{tabular}{l | llll }
- \hline
- Graphics Format & LaTeX & Markdown & Rmarkdown & HTML \\
- \hline
- PNG       & Yes & Yes & Yes & Yes \\
- JPG       & Yes & Yes & Yes & Yes \\
- PDF       & Yes & No & No & No \\
- SVG       & No & Yes & Yes & Yes \\
- Tikz      & Yes & No & Yes & No \\
- Algorithm & Yes & No & No & No \\
-\hline
-\end{tabular}
-\caption{Image Format support in various Markup/Typesetting Languages}
-\label{table:1}
-\end{table}
+  JPG               Yes     Yes        Yes         Yes
 
-\section{Image with width parameters}
-This section may contain a figure such as Figure~\ref{figure:rlogo}.
+  PDF               Yes     No         No          No
+
+  SVG               No      Yes        Yes         Yes
+
+  Tikz              Yes     No         Yes         No
+
+  Algorithm         Yes     No         No          No
+  -------------------------------------------------------
+
+  : Table 1: Image Format support in various Markup/Typesetting
+  Languages
+:::
+
+# Image with width parameters
+
+This section may contain a figure such as Figure [1](#figure:rlogo).
 This is the most basic example of figure.
 
-\begin{verbatim}
+``` r
 \begin{figure}[htbp]
@@
 \end{figure}
-\end{verbatim}
+```
 
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.35\textwidth]{Rlogo-5.png}
-  \caption{The logo of R.}
-  \label{figure:rlogo}
-\end{figure}
+![Figure 1: The logo of R.](Rlogo-5.png){#figure:rlogo width="35.0%"}
 
-\section{Images in PDF format}
+# Images in PDF format
 
-Image \ref{fig:normal} is a graphical representation of normal distribution.
+Image [2](#fig:normal) is a graphical representation of normal
+distribution.
 
-\begin{verbatim}
+``` r
 \begin{figure}[htbp]
@@
 \end{figure}
-\end{verbatim}
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5\textwidth]{normal}
-  \caption{PDF of a normal distribution}
-  \label{fig:normal}
-\end{figure}
+```
 
-\section{Multiple images}
- Pandoc v3 and above now supports a new Figure object \citep{pandoc} which supports multiple 
- images side by side or in a grid format.
-%% Two images side by side
-\subsubsection{Two or more Images side by side}
-\begin{verbatim}
+![Figure 2: PDF of a normal distribution](normal.png){#fig:normal
+width="50.0%"}
+
+# Multiple images
+
+Pandoc v3 and above now supports a new Figure object [@pandoc] which
+supports multiple images side by side or in a grid format.
+
+### Two or more Images side by side
+
+``` r
 \begin{figure}[htbp]
@@
 \end{figure}
-\end{verbatim}
+```
 
-\begin{figure*}[htbp]
-  \centering
-  \includegraphics[width=0.45\textwidth]{Rlogo-5.png}\includegraphics[width=0.45\textwidth]{normal}
-  \caption{Images side by side}
-  \label{fig:twoimages}
-\end{figure*}
+<figure id="fig:twoimages">
+<p><img src="Rlogo-5.png" style="width:45.0%" alt="image" /><img
+src="normal.png" style="width:45.0%" alt="image" /></p>
+<figcaption>Figure 3: Images side by side</figcaption>
+</figure>
 
-%% Four images in a grid
-\subsubsection{Four Images in a grid}
+### Four Images in a grid
 
-\begin{verbatim}
+``` r
 \begin{figure}[htbp]
@@
   \includegraphics[width=0.45\textwidth]{Rlogo-5.png}\includegraphics[width=0.45\textwidth]{normal}
+
   \includegraphics[width=0.45\textwidth]{normal}\includegraphics[width=0.45\textwidth]{Rlogo-5.png}
@@
 \end{figure}
+```
 
-\end{verbatim}
+<figure id="fig:fourimages">
+<p><img src="Rlogo-5.png" style="width:45.0%" alt="image" /><img
+src="normal.png" style="width:45.0%" alt="image" /></p>
+<p><img src="normal.png" style="width:45.0%" alt="image" /><img
+src="Rlogo-5.png" style="width:45.0%" alt="image" /></p>
+<figcaption>Figure 4: Multiple images in a grid</figcaption>
+</figure>
 
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.45\textwidth]{Rlogo-5.png}\includegraphics[width=0.45\textwidth]{normal}
-  \includegraphics[width=0.45\textwidth]{normal}\includegraphics[width=0.45\textwidth]{Rlogo-5.png}
-  \caption{Multiple images in a grid}
-  \label{fig:fourimages}
-\end{figure}
+# Tikz images
 
-\section{Tikz images}
+Here is a tikz image example in fig [5](#fig:tikz) adapted from
+[@casflow].
 
-Here is a tikz image example in fig \ref{fig:tikz} adapted from \citep{casflow}.
+Another interesting aspect of including tikz image here is that you can
+modify the source and re-convert without making any other changes.
+Figure will get updated in the generated html article.
 
-Another interesting aspect of including tikz image here is that you can modify the
-source and re-convert without making any other changes. Figure will get updated
-in the generated html article.
+### Tikz Code:
 
-\subsubsection{Tikz Code:}
-\begin{verbatim}
+``` r
 \begin{figure}
+\includegraphics{tikz/figtikz.png}
 
@@
 \end{figure}
+```
 
-\end{verbatim}
+### Resultant Figure :
 
-\subsubsection{Resultant Figure :}
-\begin{figure*}
+<figure id="fig:tikz">
+<img src="tikz/figtikz.png" style="width:100.0%" />
 
-%% Generated Image will included as a PNG above
-  \centering
-\tikzstyle{process} = [rectangle, rounded corners,
-minimum width=3cm, 
-minimum height=1cm,
-text centered, 
-draw=black]
-\tikzstyle{arrow} = [thick,->,>=stealth]
-\begin{tikzpicture}[node distance=4cm]
-%Nodes
-\node (start) [process] {LaTeX Source with Tikz Image};
-\node (isolate) [process, below right of=start] {Isolated tikz env};
-\node (pro1) [process, below left of=isolate] {Standalone PDF file};
-\node (pro2) [process, below right of=pro1, yshift=-0.5cm] {PNG file};
-\node (link) [process, below left of=pro2] {Include PNG in article};
-\node (stop) [process, below right of=link] {Convert To RJ-web-article};
-% arrows
-\draw [arrow] (start) -- node[anchor=east] {Extract Figure Block} (isolate);
-\draw [arrow] (isolate) -- node[anchor=east] {Compile} (pro1);
-\draw [arrow] (pro1) -- node[anchor=east] {Convert to raster} (pro2);
-\draw [arrow] (pro2) -- node[anchor=east] {Include back to source as image} (link);
-\draw [arrow] (link) -- node[anchor=east] {Using Pandoc} (stop);
-\end{tikzpicture}
-\caption{Tikz Image example}
-  \label{fig:tikz}
-\end{figure*}
+<figcaption>Figure 5: Tikz Image example</figcaption>
+</figure>
 
-\section{Algorithm2e diagrams}
+# Algorithm2e diagrams
 
-we do support algorithm2e diagrams and images, these will be numbered differently
-and we strongly suggest to use "alg:" in labels for best results. 
-Here is an example of algorithm \ref{alg:how}, referenced from \cite{algoexample}
+we do support algorithm2e diagrams and images, these will be numbered
+differently and we strongly suggest to use \"alg:\" in labels for best
+results. Here is an example of algorithm [6](#alg:how), referenced from
+[@algoexample]
 
-\begin{verbatim}
-\begin{algorithm}[htbp]
+``` r
+\begin{figure}[htbp]
+\includegraphics{alg/alghow.png}
+
 \SetAlgoLined
@@
   \label{alg:how}
-\end{algorithm}
-\end{verbatim}
+\end{figure}
+```
 
-\begin{algorithm}[htbp]
-\SetAlgoLined
-\KwData{this text}
-\KwResult{how to write algorithm with \LaTeX2e }
-initialization\;
-\While{not at end of this document}{
-read current\;
-\eIf{understand}{
-go to next section\;
-current section becomes this one\;
-}{
-go back to the beginning of current section\;
-}
-}
-\caption{How to write algorithms}
-  \label{alg:how}
-\end{algorithm}
+<figure id="alg:how">
+<img src="alg/alghow.png" style="width:100.0%" />
+<p>initialization</p>
+<figcaption>Algorithm 1: How to write algorithms</figcaption>
+</figure>
 
-\section{Summary}
+# Summary
 
-In summary the \CRANpkg{texor} package supports:
-\begin{itemize}
-\item Almost all image formats in LaTeX.
-\item Algorithm and tikz as well in some capacity.
-\item Multiple images in grid,side-by-side configuration.
-\item Image Captions with Numbering and Labelling.
-\end{itemize}
+In summary the [*texor*](https://CRAN.R-project.org/package=texor)
+package supports:
 
-\begin{thebibliography}{4}
-    \providecommand{\natexlab}[1]{#1}
-    \providecommand{\url}[1]{\texttt{#1}}
-    \expandafter\ifx\csname urlstyle\endcsname\relax
-      \providecommand{\doi}[1]{doi: #1}\else
-      \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi
-\bibitem[Cassidy(2013)]{casflow}
-Josh Cassidy
-\newblock LaTeX Graphics using TikZ: A Tutorial for Beginners (Part 3)—Creating Flowcharts
-\newblock \emph{Overleaf tutorials} \penalty0 2013
-\newblock URL \url{https://www.overleaf.com/learn/latex/}
+-   Almost all image formats in LaTeX.
 
-\bibitem[Fiorio (2017)]{algoexample}
-Christophe Fiorio
-\newblock algorithm2e.sty — package for algorithms, release 5.2
-\newblock \emph{CTAN},\penalty0 2017
-\newblock URL \url{https://mirror.kku.ac.th/CTAN/macros/latex/contrib/algorithm2e/doc/algorithm2e.pdf}
+-   Algorithm and tikz as well in some capacity.
 
-\bibitem[Krewinkel, Lucero (2023)]{pandoc}
-Albert Krewinkel and Aner Lucero
-\newblock pandoc 3.0 Release notes
-\newblock \emph{pandoc} \penalty0 2023
-\newblock URL \url{https://pandoc.org/releases.html#pandoc-3.0-2023-01-18}
+-   Multiple images in grid,side-by-side configuration.
 
-\end{thebibliography}
-
-\address{%
-Abhishek Ulayil\\
-Student, Institute of Actuaries of India\\%
-Mumbai, India\\
-ORCiD: 0009-0000-6935-8690\\
-}
+-   Image Captions with Numbering and Labelling.
+:::

### 2023-07-19 supplement/code-env/code-example.tex
new supplementary material for code-env (incomplete)

% !TeX root = RJwrapper.tex
\title{Working with Code environments in texor}
\author{by Abhishek Ulayil}

\maketitle

\abstract{
This is a small sample article to demonstrate usage of texor to convert code environments.
}

\section{Introduction}

Pandoc naturally converts verbatim environment easily, however the redefination of other commands such as example, example*, Sinput etc to verbatim does not work well in pandoc.

Hence Texor package uses the stream editor to search find and replace matching code environments to verbatim before pandoc touches it.

This way the the code is not lost in conversion, also a pandoc extension is used to add attributes to the markdown code using 
'fenced\_code\_attributes'

Code Environment types are well summarized in the table \ref{table:1}

\begin{table}[htbp]
\centering
\begin{tabular}{l | lllll }
 \hline
 Code Environment Type &  &  &  & & \\
 \hline
 Example          & example & example* & smallexample & & \\
 S.series         & Sin & Sout & Sinput & Soutput & Scode \\
 Special Verbatim & smallverbatim & boxedverbatim &  & & \\
\hline
\end{tabular}
\caption{Code Environment support in texor}
\label{table:1}
\end{table}

\section{Environments}

\subsection{Verbatim Series}
While verbatim is naturally supported in pandoc,other extensions of verbatim environment
like boxedverbatim from moreverb package \citep{moreverb}.
% verbatim

1. verbatim :

\begin{verbatim}
print("Hello world")
\end{verbatim}

2. smallverbatim :

\begin{smallverbatim}
print("Hello world")
\end{smallverbatim}

3. boxedverbatim :

\begin{boxedverbatim}
print("Hello world")
\end{boxedverbatim}

\subsection{S series}

S series code environement is defined in Rjournal.sty file.

1. Sinput :

\begin{Sinput}
print("Hello world")
\end{Sinput}

2. Soutput :

\begin{Soutput}
[1] "hello world"
\end{Soutput}

3. Sin :

\begin{Sin}
print("Hello world")
\end{Sin}

4. Sout :
\begin{Sout}
[1] "hello world"
\end{Sout}

\subsection{Example series}
Example series of code environment is defined in Rjournal.sty file.

1. example :

\begin{example}
print("Hello world")
\end{example}

2. example* :

\begin{example*}
print("Hello world")
\end{example*}

3. smallexample :

\begin{smallexample}
print("Hello world")
\end{smallexample}

\section{Code in Figure Environments}
A small example of this is visble in \ref{code:example}

\begin{figure*}[htbp]
\begin{Sinput}
code_in_figure <- function() {
  if (pandoc_version >= 3) {
    print("Code in Figure Supported")
  }
  else {
    print("code in Figure not supported")
  }
}
\end{Sinput}
\caption{ Example Code inside Figure environment}
\label{code:example}
\end{figure*}
Pandoc v3 or greater \citep{pandoc} has a Figure object which allows non-image
figures to be treated like one.

\section{Code in Table Environments}

\section{Inline Code usage}

\section{Code chunks using Schunk}

\begin{Schunk}
Input :
\begin{Sinput}
print("Hello world")
\end{Sinput}
Output :
\begin{Soutput}
[1] "hello world"
\end{Soutput}

\end{Schunk}
\section{Summary}

In summary the \CRANpkg{texor} package supports:
\begin{itemize}
\item Almost all code environments in RJournal.
\item Code Highlight for R language.
\item Code blocks inside Figure environment with captions.
\item Inline Code.
\end{itemize}

\begin{thebibliography}{2}
    \providecommand{\natexlab}[1]{#1}
    \providecommand{\url}[1]{\texttt{#1}}
    \expandafter\ifx\csname urlstyle\endcsname\relax
      \providecommand{\doi}[1]{doi: #1}\else
      \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi

\bibitem[Krewinkel, Lucero (2023)]{pandoc}
A.~ Krewinkel and A.~ Lucero
\newblock pandoc 3.0 Release notes
\newblock \emph{pandoc} \penalty0 2023
\newblock URL \url{https://pandoc.org/releases.html#pandoc-3.0-2023-01-18}

\bibitem[Fairbairns, Duggan, Schöpf, Eijkhout (2011)]{moreverb}
R.~ Fairbairns, A.~ Duggan, R.~ Schöpf and V.~ Eijkhout
\newblock The moreverb package documentation
\newblock \emph{CTAN} \penalty0 2011
\newblock URL \url{https://mirror.niser.ac.in/ctan/macros/latex/contrib/moreverb/moreverb.pdf}
\end{thebibliography}

\address{%
Abhishek Ulayil\\
Student, Institute of Actuaries of India\\%
Mumbai, India\\
ORCiD: 0009-0000-6935-8690\\
}

### 2023-07-22 supplement/code-env/code-example.tex
Update Code environment supplementary material

@@
 \section{Code in Table Environments}
+We can use code environments in a table using minipage environments.
+
+Table \ref{table:2} is an example of table with code environments within a table.
+
+\begin{table}[htbp]
+\centering
+  \begin{tabular}{|c | c |}
+    \hline
+    Language & Function Defination Syntax \\
+    \hline
+    R & \begin{minipage}{0.75\textwidth}
+\vspace{1mm}
+\begin{example}
+fun <- function(){
+  print("A function in R")
+  return(0)
+}
+
+\end{example}
+\end{minipage}\\
+\hline
+Python & \begin{minipage}{0.75\textwidth}
+\vspace{1mm}
+\begin{example}
+def fun():
+  print("A function in Python")
+
+\end{example}
+\end{minipage}\\
+\hline
+
+Lua & \begin{minipage}{0.75\textwidth}
+\vspace{1mm}
+\begin{example}
+function fun()
+  print("A function in Lua")
+end
+
+\end{example}
+\end{minipage}\\
+\hline
+
+C & \begin{minipage}{0.75\textwidth}
+\vspace{1mm}
+\begin{example}
+int fun() {
+  printf("A function in C");
+  return 0;
+}
+
+\end{example}
+\end{minipage}\\
+\hline
+\end{tabular}
+\caption{Code in a table}
+\label{table:2}
+\end{table}
+
+A similar arrangement can be had for figures/plots besides code environment.
+Table \ref{table:3} demonstrates a table with code and figure.
+\begin{table}[htbp]
+  \centering
+  \begin{tabular}{| c | c |}
+  \hline
+  Code & Plot\\
+    \hline
+    \begin{minipage}{0.45\textwidth}
+\vspace{1mm}
+\begin{example}
+x <- 1:100
+y <- dbinom(x,100,prob = 0.5)
+plot(x,y)
+\end{example}
+    \end{minipage} &
+    \begin{minipage}{0.45\textwidth}
+    \centering
+    \includegraphics[width=1\textwidth]{binom}
+    \end{minipage}\\
+    \hline
+  \end{tabular}
+  \caption{Code and Plot side by side}
+  \label{table:3}
+\end{table}
 
@@
 
+Using inline code in LaTeX is possible using \verb|\verb| command.
+It would be repoduced similarly, as a Inline code element.
+\begin{verbatim}
+\verb|x <- 1:100|
+\end{verbatim}
+will be represented as \verb|x <- 1:100| in Inline format.
+
 \section{Code chunks using Schunk}
+Code chunks within an \verb|Schunk| environment to demonstrate Input/Output
 
@@
 \end{Soutput}
+\end{Schunk}
 
+Similar arrangement can be had for Plots as well using figure environnment.
+
+\begin{Schunk}
+Input :
+\begin{Sinput}
+x <- 1:100
+y <- dbinom(x,100,prob = 0.5)
+plot(x,y)
+\end{Sinput}
+Output :\\
+\includegraphics[width=0.5\textwidth]{binom}
 \end{Schunk}
+
 \section{Summary}
@@
 \newblock URL \url{https://mirror.niser.ac.in/ctan/macros/latex/contrib/moreverb/moreverb.pdf}
+
 \end{thebibliography}

### 2023-07-22 supplement/code-env/code-example.tex
Updated code env supplementary material

@@
 
-Hence Texor package uses the stream editor to search find and replace matching code environments to verbatim before pandoc touches it.
+Hence \CRANpkg{texor} package uses the stream editor to search find and replace matching code environments to verbatim before pandoc touches it.
 
-This way the the code is not lost in conversion, also a pandoc extension is used to add attributes to the markdown code using 
-'fenced\_code\_attributes'
+This way the the code is not lost in conversion, also a pandoc extension is used to add attributes to the markdown code using  \verb|fenced_code_attributes|
 
@@
  \hline
- Example          & example & example* & smallexample & & \\
+ Example          & example & example* &  & & \\
  S.series         & Sin & Sout & Sinput & Soutput & Scode \\
- Special Verbatim & smallverbatim & boxedverbatim &  & & \\
+ Special Verbatim & boxedverbatim & &  & & \\
 \hline
@@
 
-2. smallverbatim :
-
-\begin{smallverbatim}
-print("Hello world")
-\end{smallverbatim}
-
-3. boxedverbatim :
+2. boxedverbatim :
 
@@
 
-3. smallexample :
-
-\begin{smallexample}
-print("Hello world")
-\end{smallexample}
-
+% Code in Figures
 \section{Code in Figure Environments}
@@
 
+% Code in Tables
 \section{Code in Table Environments}
@@
 
-Table \ref{table:2} is an example of table with code environments within a table.
+Table \ref{table:2} is an example of code environments within a table.
 
@@
 
+% Inline Code
 \section{Inline Code usage}
@@
 \item Code Highlight for R language.
-\item Code blocks inside Figure environment with captions.
 \item Inline Code.
+\item Code in different environments like tables/figures.
 \end{itemize}

### 2023-07-23 supplement/code-env/code-example.tex
Edits to code example supplement

@@
 
-Hence \CRANpkg{texor} package uses the stream editor to search find and replace matching code environments to verbatim before pandoc touches it.
+Hence, the \CRANpkg{texor} package uses a stream editor to search find and replace matching code environments to verbatim before pandoc touches it.
 
@@
 While verbatim is naturally supported in pandoc,other extensions of verbatim environment
-like boxedverbatim from moreverb package \citep{moreverb}.
+like boxedverbatim from moreverb package \citep{moreverb} falls back to normal verbatim.
 % verbatim
@@
 
-S series code environement is defined in Rjournal.sty file.
+S series code environement is defined in Rjournal.sty file. Most of these are
+extensions of verbatim environment and are represented as vanilla verbatim in HTML.
 
@@
 \subsection{Example series}
-Example series of code environment is defined in Rjournal.sty file.
+Example series of code environment is defined in Rjournal.sty file. Examples are
+extensions of verbatim environment and are represented as vanilla verbatim in HTML.
 
@@
 \section{Code in Figure Environments}
-A small example of this is visble in \ref{code:example}
+A small example of this is visble in \ref{code:example}. This is a common practice
+in Rnews articles as it used to add a boxed border around the code which looks 
+attractive. However, in web articles there isn't much advantage to it.
 
-\begin{figure*}[htbp]
-\begin{Sinput}
+\begin{figure}[htbp]
+\begin{center}
+\begin{verbatim}
 code_in_figure <- function() {
@@
 }
-\end{Sinput}
+\end{verbatim}
 \caption{ Example Code inside Figure environment}
 \label{code:example}
-\end{figure*}
+\end{center}
+\end{figure}
+
 Pandoc v3 or greater \citep{pandoc} has a Figure object which allows non-image
-figures to be treated like one.
+figures to be treated like one. This is why \CRANpkg{texor} package requires atleast version 3
+of pandoc.
 
@@
 \section{Code in Table Environments}
-We can use code environments in a table using minipage environments.
+We can use code environments in a table using minipage environments. This is not
+a common practice among LaTeX article authors, but a few articles had such complex
+structures. So, as a example to demonstrate pandoc and \CRANpkg{texor} package's capabilities, 
+I have included a few of them.
 

### 2023-07-23 supplement/lua-filters/lua-filters.tex
New supplementary material on lua filters

% !TeX root = RJwrapper.tex
\title{Understanding pandoc lua filters}
\author{by Abhishek Ulayil}

\maketitle

\abstract{
pandoc supports intermediate modification of the Abstract Syntax Tree (AST) between
the parsing and writing phase using filters. This supplement paper will highlight
the use cases of such filters written in Lua Language on LaTeX,markdown and native AST.
}

\section{Introduction}
To understand pandoc filters, one must look at the document conversion process in
general. A document is read/parsed in one format to a common intermediate format.
In pandoc it is called an abstract syntax tree (AST). A document writer is designed 
to read the AST and write the contents in the form of desired article format.
For example converting markdown to HTML will involve these processes in pandoc.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6\textwidth]{workflow.png}
  \caption{Conversion Workflow}
  \label{fig:workflow}
\end{figure}

The filters come into picture when there are certain elements which require 
customization. For example, suppose you are converting markdown to HTML and you
require the page to have automatic numbering for figures or tables. 
If you think about it in markdown there is no system of numbering figures/tables 
automatically, to retain 1:1 conversion the HTML output will miss the numbering as well. 
At this moment you would wish that there was some option to add such a feature,
but then given the infinite possibilities of customization it is not feasible to 
add each and every customization as an option in pandoc or the writer. 
To tackle this, the idea of filters arrived, wherein you could manipulate the AST
partly or fully and modify it before the writer could read the AST. Hence generating
an output you desire \citep{pandocfilters}.

\section{Writing lua filters}

Continuing the example above lets create a dummy article, where we need to add
numbering. One way could be to keep a counter of Images, and add a prefix \verb|"Figure X :"|
to each figure caption. This will serve the purpose of numbering the Images in the end result.

\begin{verbatim}
![R logo](Rlogo-5.png){width="10%"}

![penguins](penguins.png){width="15%"}

\end{verbatim}

\begin{figure*}[htbp]
\begin{verbatim}
pandoc example.md --from markdown --to html5 --output example.html
\end{verbatim}
\caption{pandoc command without lua filter or extensions}
\label{code:2}
\end{figure*}

Now If we convert the above markdown file to HTML5 using pandoc command \ref{code:2}, we get \ref{fig:1}

\begin{figure*}[!htbp]
\centering
\includegraphics[width=0.5\linewidth]{example.png}
\caption{Vanilla HTML5 output}
\label{fig:1}
\end{figure*}

\section{Using pandoc lua filters}

As we can see, there is no figure numbering done automatically, which is generally
the expected result. However if we want to include numbering to it, we would need
to write a lua filter. This lua filter will modify the AST and make the changes we 
desire.

We call a lua filter in the pandoc command \ref{code:2} using the \verb| --lua-filter name_of_filter.lua|
option in pandoc.

Now we write a lua filter to manipulate the figures in \ref{code:3}
\subsection{TODO : explain lua filter in more detail.}
\begin{figure}[htbp]
\begin{verbatim}
figures = 0
is_fig = 0

function Figure(el)
      local label = ""
      pandoc.walk_block(el,{ Image = function(el)
                     is_fig = 1
                     end})
      if is_fig == 1 then
	        figures = figures + 1
	        label = "Figure " .. tostring(figures) .. ":"
      end
	    local caption = pandoc.utils.stringify(el.caption)
	    if not caption then
          caption = label
    	else
          caption = label .. " " .. caption
    	end
    	el.caption.long[1] = caption
    	is_fig = 0
    	return el
end
\end{verbatim}
\caption{A lua filter to add figure numbering}
\label{code:3}
\end{figure}

\section{Interpreting changes using lua filters}

The filter \ref{code:3} when included in the pandoc command will generate \ref{fig:2} 
using the \ref{code:4} command.

\begin{figure}[htbp]
\begin{verbatim}
pandoc example.md --from markdown --to html5 --output filtered-example.html
--lua-filter image_numbering_filter.lua
\end{verbatim}
\caption{pandoc command with lua filter}
\label{code:4}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=0.5\linewidth]{example-filtered.png}
\caption{HTML5 output with desired filtering}
\label{fig:2}
\end{figure}

\section{Usage of pandoc lua filters in texor package}

\section{Summary}

\section{Supplementary materials}
The example conversion files described here are included as supplementary materials
in the Rjournal article of texor.

\begin{thebibliography}{2}
    \providecommand{\natexlab}[1]{#1}
    \providecommand{\url}[1]{\texttt{#1}}
    \expandafter\ifx\csname urlstyle\endcsname\relax
      \providecommand{\doi}[1]{doi: #1}\else
      \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi

\bibitem[Krewinkel, Lucero (2023)]{pandoc}
A.~ Krewinkel and A.~ Lucero,
\newblock pandoc 3.0 Release notes,
\newblock \emph{pandoc} \penalty0 2023
\newblock URL \url{https://pandoc.org/releases.html#pandoc-3.0-2023-01-18}

\bibitem[MacFarlane and other pandoc authors (2023)]{pandocfilters}
John MacFarlane and other pandoc authors,
\newblock Pandoc Lua filters,
\newblock \emph{pandoc documentation} \penalty0 2023
\newblock URL \url{https://pandoc.org/lua-filters.html}

\end{thebibliography}

\address{%
Abhishek Ulayil\\
Student, Institute of Actuaries of India\\%
Mumbai, India\\
ORCiD: 0009-0000-6935-8690\\
}

### 2023-07-24 supplement/bibliography/rebib.tex
New supplementary material on rebib and bibliography

% !TeX root = RJwrapper.tex
\title{Working with bibliography in texor}
\author{by Abhishek Ulayil}

\maketitle

\abstract{
This is a small sample article to demonstrate usage of \CRANpkg{rebib} to manage bibliographies.
}

\section{Introduction}
The \CRANpkg{rebib} package is a spun-off from \CRANpkg{texor} package.

The decision to do this is the fact that the bibliography section 
in \CRANpkg{texor} package was expanding significantly, enough to deserve a dedicated space.

Thus \CRANpkg{rebib} was born and has improved over the past few months to include more advanced features and utilities.

\subsection{What does rebib do?}

In LaTeX, we have multiple ways to declare bibliography entries, which include embedding bibliography entries in the document itself or isolating that section to an external file commonly called a bbl file. Then we have the BibTeX format, which has its structure and is much superior in managing bibliographic entries.

LaTeX has no problem dealing with multiple formats and works well with all these bibliographic formats. However, some web publishing journals like RJ web articles from RJournal do not support embedded bibliographies. As a result, they need a BibTeX equivalent of the same.

The \CRANpkg{rebib} package converts embedded LaTeX bibliographies into a close  BibTeX equivalent. This package eases the author from the manual conversion of the documents and aids the texor project in automating the conversion of past RJournal articles to the RJ-web-article format.

\subsection{features of rebib}

\begin{itemize}
  \item Reads bib chunks to produce a very close BibTeX equivalent
  \item Title and author are usually mandatory fields
  \item URL, ISBN, publisher, pages and year are optional fields and will be enabled when relevant
  \item Rest of the data is stored in \verb|"journal"|(internally) and \verb|"publisher"|(when writing BibTeX file)
  \item Ignores commented LaTeX code 
  \item Citation tracker
  \item Bibliography Aggregation
\end{itemize}

\section{Parser}

The magic behind \CRANpkg{rebib} is a parser which parses bibliographic data and assorts them according to the matching regular expressions.

\subsubsection{Stage 1: Read the embedded bibliography block}

This stage is a minor step where it reads the Embedded Bibliography from the LaTeX document. This step also includes Filtering out the commented code to avoid un-intended entries read.

Lastly, the data is broken down based on the LaTeX macro \verb|\\bibitem| as a marker for a new entry and this assorted data is exported to a variable.

\begin{Schunk}

\textbf{Setup :}
\begin{Sinput}
dir.create(your_article_folder <- file.path(tempdir(), "exampledir"))
example_files <-  system.file("article", package = "rebib")
x <- file.copy(from = example_files,to=your_article_folder,recursive = T)
your_article_path <- paste(your_article_folder,"article",sep="/")
\end{Sinput}
\end{Schunk}

\begin{Schunk}

\textbf{Input :}
\begin{Sinput}
file_name <- rebib:::get_texfile_name(your_article_path)
bib_items <- rebib:::extract_embeded_bib_items(your_article_path,file_name)
bib_items[[1]]
\end{Sinput}

\textbf{Output :}
\begin{Soutput}
#> [1] "\\bibitem[Ihaka, Ross and Gentleman, Robert]{ihaka:1996}"                          
#> [2] "Ihaka, Ross and Gentleman, Robert"                                                 
#> [3] "\\newblock \\emph{R: A Language for Data Analysis and Graphics.}"                  
#> [4] "\\newblock \\emph{Journal of Computational and Graphical Statistics}, 3:\\penalty0"
#> [5] "299--314, 1996."                                                                   
#> [6] "\\newblock URL : \\url{https://doi.org/10.1080/10618600.1996.10474713}"
\end{Soutput}
\end{Schunk}

\subsubsection{Stage 2: Regex powered parser}

Now, with the chunks of bibliographic entries, each is passed to a parser which will break it down based on regular expressions.
The logic is to use the LaTeX macro \verb|\\newblock| as a placeholder to identify the position of text elements relative to it.

The first value to be parsed is the \verb|unique_id| also called the citation reference which is used to cite elements inside the article. Usually, this is in the first or second line of the whole entry. The position of the 
\verb|unique_id| will determine the position of the author names.
\begin{Schunk}

\textbf{Input :}
\begin{Sinput}
bib_items[[1]][1]
\end{Sinput}

\textbf{Output :}
\begin{Soutput}
#> [1] "\\bibitem[Ihaka, Ross and Gentleman, Robert]{ihaka:1996}"
\end{Soutput}
\end{Schunk}

After reading the \verb|unique_id|, the parser will attempt to read the author name(s).

\begin{Schunk}

\textbf{Input :}
\begin{Sinput}
bib_items[[1]][2]
\end{Sinput}

\textbf{Output :}
\begin{Soutput}
#> [1] "Ihaka, Ross and Gentleman, Robert"
\end{Soutput}
\end{Schunk}

Next, the title is extracted based on the position of the new blocks or the end of the bib chunk.

\begin{Schunk}

\textbf{Input :}
\begin{Sinput}
bib_items[[1]][3]
\end{Sinput}
\textbf{Output :}
\begin{Soutput}
#> [1] "\\newblock \\emph{R: A Language for Data Analysis and Graphics.}"
\end{Soutput}
\end{Schunk}

This way the crucial elements of the bibliographic entry (unique_id, author names and title ) are parsed out.
The remaining data is stored as \verb|journal| internally and \verb|publisher| when writing to a new BibTeX file.

\begin{Schunk}

\textbf{Input :}
\begin{Sinput}
bib_items[[1]][4:6]
\end{Sinput}

\textbf{Output :}
\begin{Soutput}
#> [1] "\\newblock \\emph{Journal of Computational and Graphical Statistics}, 3:\\penalty0"
#> [2] "299--314, 1996."                                                                   
#> [3] "\\newblock URL : \\url{https://doi.org/10.1080/10618600.1996.10474713}"
\end{Soutput}
\end{Schunk}

A series of filters for ISBN, URL, pages and year fields are applied to search for relevant data from the remaining data. If the data is not available then it is set as NULL and will be skipped while writing the BibTeX file. 
There is a lot of filtering of common LaTeX elements and after that, the data remaining is stored in a structured format to be written to a file.

\begin{Schunk}

\textbf{Input :}
\begin{Sinput}
bib_entry <- rebib:::bib_handler(bib_items)
bib_entry
\end{Sinput}

\textbf{Output :}
\begin{Soutput}
#> $book
#> $book[[1]]
#> $book[[1]]$unique_id
#> [1] "ihaka:1996"
#> 
#> $book[[1]]$author
#> [1] "Ihaka, Ross and Gentleman, Robert"
#> 
#> $book[[1]]$title
#> [1] "R: A Language for Data Analysis and Graphics"
#> 
#> $book[[1]]$journal
#> [1] "Journal of Computational and Graphical Statistics 3:   :"
#> 
#> $book[[1]]$year
#> [1] "1996"
#> 
#> $book[[1]]$URL
#> [1] "https://doi.org/10.1080/10618600.1996.10474713"
#> 
#> $book[[1]]$isbn
#> NULL
#> 
#> $book[[1]]$pages
#> [1] "299--314"
#> 
#> 
#> $book[[2]]
#> $book[[2]]$unique_id
#> [1] "R"
#> 
#> $book[[2]]$author
#> [1] "R Core Team"
#> 
#> $book[[2]]$title
#> [1] "R: A Language and Environment for Statistical Computing"
#> 
#> $book[[2]]$journal
#> [1] "R Foundation for Statistical Computing Vienna Austria    :"
#> 
#> $book[[2]]$year
#> [1] "2016"
#> 
#> $book[[2]]$URL
#> [1] "https://www.R-project.org/"
#> 
#> $book[[2]]$isbn
#> [1] "3-900051-07-0"
#> 
#> $book[[2]]$pages
#> NULL
\end{Soutput}
\end{Schunk}

\subsubsection{Stage 3: BibTeX writer}

After reading the bibliographic entries and splitting out meaningful values from them, we can finally write a structured file in the BibTeX format.

The writer will read the bib chunks one at a time based on the metadata extracted and will write the corresponding data fields. The default entry type is a book, which should not have any problems with the web articles. 
\begin{Schunk}

\textbf{Input :}
\begin{Sinput}
file_path <- paste0(your_article_path,"/",file_name)
bib_path <- paste0(your_article_path,"/example.bib")
\end{Sinput}
\end{Schunk}

\begin{Schunk}

\textbf{Input :}
\begin{Sinput}
rebib:::bibtex_writer(bib_entry,file_path)
cat(readLines(paste(your_article_path,"example.bib",sep="/")),sep = "\n")
\end{Sinput}

\textbf{Output :}
\begin{Soutput}
#> @book{ihaka:1996,
#> author = {{Ihaka, Ross and Gentleman, Robert}},
#> title = {{R: A Language for Data Analysis and Graphics}},
#> publisher = {Journal of Computational and Graphical Statistics 3:   :},
#> pages = {299--314},
#> year = {1996},
#> url = {https://doi.org/10.1080/10618600.1996.10474713}
#> }
#> @book{R,
#> author = {R {Core Team}},
#> title = {{R: A Language and Environment for Statistical Computing}},
#> publisher = {R Foundation for Statistical Computing Vienna Austria    :},
#> year = {2016},
#> url = {https://www.R-project.org/},
#> isbn = {3-900051-07-0}
#> }
\end{Soutput}
\end{Schunk}

I expect the authors who are converting the document to take a look and check if there are any errors or missing values.

\section{General usage}

\section{Bibliography aggregation}

\section{Summary}

In summary the \CRANpkg{rebib} package supports:
\begin{itemize}
\item Conversion of embedded natbib type bibliography used in RJournal
\item Aggregation of bibliographic entries from the article as well as BibTeX
\end{itemize}

\address{%
Abhishek Ulayil\\
Student, Institute of Actuaries of India\\%
Mumbai, India\\
ORCiD: 0009-0000-6935-8690\\
}

### 2023-07-24 supplement/bibliography/examples/aggregation/example.tex
Updated Supplementary material for rebib

% !TeX root = RJwrapper.tex
% License: Creative commons CC BY 4.0
\title{Capitalized Title Here}
\author{by Author One, Author Two and Author Three}

\maketitle

\abstract{
An abstract of less than 150 words.
}

\section{Section title in sentence case}

Introductory section which may include references in parentheses
\citep{R}, or cite a reference such as \citet{R} in the text.

\section{Another section}

There will likely be several sections, perhaps including code snippets, such as:

\begin{example}
  x <- 1:10
  result <- myFunction(x)
\end{example}

\section{Summary}

This file is only a basic article template. For full details of \emph{The R Journal} style and information on how to prepare your article for submission, see the \href{https://journal.r-project.org/share/author-guide.pdf}{Instructions for Authors}.

\begin{thebibliography}{20}
    \providecommand{\natexlab}[1]{#1}
    \providecommand{\url}[1]{\texttt{#1}}
    \expandafter\ifx\csname urlstyle\endcsname\relax
      \providecommand{\doi}[1]{doi: #1}\else
      \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi
\bibitem[Ihaka, Ross and Gentleman, Robert]{ihaka:1996}
Ihaka, Ross and Gentleman, Robert
\newblock \emph{R: A Language for Data Analysis and Graphics.}
\newblock \emph{Journal of Computational and Graphical Statistics}, 3:\penalty0
299--314, 1996.
\newblock URL : \url{https://doi.org/10.1080/10618600.1996.10474713}

\bibitem[R Core Team]{R}
R Core Team
\newblock R: A Language and Environment for Statistical Computing
\newblock \emph{R Foundation for Statistical Computing}, Vienna, Austria \penalty0 2016.
\newblock URL : \url{https://www.R-project.org/}, ISBN 3-900051-07-0

\bibitem[Tremblay(2012)]{Tremblay:2012}
A.~Tremblay.
\newblock \emph{LMERConvenienceFunctions: A suite of functions to back-fit
  fixed effects and forward-fit random effects, as well as other miscellaneous
  functions.}, 2012.
\newblock URL \url{http://CRAN.R-project.org/package=LMERConvenienceFunctions}.
\newblock R package version 1.6.8.2.
\end{thebibliography}

\address{Author One\\
  Affiliation\\
  Address\\
  Country\\
  (ORCiD if desired)\\
  \email{author1@work}
}

\address{Author Two\\
  Affiliation\\
  Address\\
  Country\\
  (ORCiD if desired)\\
  \email{author2@work}
}

\address{Author Three\\
  Affiliation\\
  Address\\
  Country\\
  (ORCiD if desired)\\
  \email{author3@work}
 }

 \bibliography{example}

### 2023-07-24 supplement/bibliography/rebib.tex
Updated Supplementary material for rebib

@@
 
-\begin{Schunk}
-
-\textbf{Setup :}
 \begin{Sinput}
@@
 your_article_path <- paste(your_article_folder,"article",sep="/")
-\end{Sinput}
-\end{Schunk}
-
-\begin{Schunk}
-
-\textbf{Input :}
-\begin{Sinput}
 file_name <- rebib:::get_texfile_name(your_article_path)
@@
 bib_items[[1]]
-\end{Sinput}
-
-\textbf{Output :}
-\begin{Soutput}
 #> [1] "\\bibitem[Ihaka, Ross and Gentleman, Robert]{ihaka:1996}"                          
@@
 #> [6] "\\newblock URL : \\url{https://doi.org/10.1080/10618600.1996.10474713}"
-\end{Soutput}
-\end{Schunk}
+\end{Sinput}
 
@@
 \verb|unique_id| will determine the position of the author names.
-\begin{Schunk}
 
-\textbf{Input :}
-\begin{Sinput}
+\begin{Soutput}
 bib_items[[1]][1]
-\end{Sinput}
-
-\textbf{Output :}
-\begin{Soutput}
 #> [1] "\\bibitem[Ihaka, Ross and Gentleman, Robert]{ihaka:1996}"
 \end{Soutput}
-\end{Schunk}
 
@@
 
-\begin{Schunk}
-
-\textbf{Input :}
 \begin{Sinput}
 bib_items[[1]][2]
+#> [1] "Ihaka, Ross and Gentleman, Robert"
 \end{Sinput}
-
-\textbf{Output :}
-\begin{Soutput}
-#> [1] "Ihaka, Ross and Gentleman, Robert"
-\end{Soutput}
-\end{Schunk}
 
@@
 
-\begin{Schunk}
-
-\textbf{Input :}
 \begin{Sinput}
 bib_items[[1]][3]
+#> [1] "\\newblock \\emph{R: A Language for Data Analysis and Graphics.}"
 \end{Sinput}
-\textbf{Output :}
-\begin{Soutput}
-#> [1] "\\newblock \\emph{R: A Language for Data Analysis and Graphics.}"
-\end{Soutput}
-\end{Schunk}
 
-This way the crucial elements of the bibliographic entry (unique_id, author names and title ) are parsed out.
+This way the crucial elements of the bibliographic entry (unique\_id, author names and title ) are parsed out.
 The remaining data is stored as \verb|journal| internally and \verb|publisher| when writing to a new BibTeX file.
 
-\begin{Schunk}
-
-\textbf{Input :}
 \begin{Sinput}
 bib_items[[1]][4:6]
-\end{Sinput}
-
-\textbf{Output :}
-\begin{Soutput}
 #> [1] "\\newblock \\emph{Journal of Computational and Graphical Statistics}, 3:\\penalty0"
@@
 #> [3] "\\newblock URL : \\url{https://doi.org/10.1080/10618600.1996.10474713}"
-\end{Soutput}
-\end{Schunk}
+\end{Sinput}
 
@@
 \begin{Schunk}
-
-\textbf{Input :}
-\begin{Sinput}
+\begin{Soutput}
 bib_entry <- rebib:::bib_handler(bib_items)
 bib_entry
-\end{Sinput}
-
-\textbf{Output :}
-\begin{Soutput}
 #> $book
@@
 
-\textbf{Input :}
 \begin{Sinput}
@@
 bib_path <- paste0(your_article_path,"/example.bib")
-\end{Sinput}
-\end{Schunk}
-
-\begin{Schunk}
-
-\textbf{Input :}
-\begin{Sinput}
 rebib:::bibtex_writer(bib_entry,file_path)
 cat(readLines(paste(your_article_path,"example.bib",sep="/")),sep = "\n")
-\end{Sinput}
-
-\textbf{Output :}
-\begin{Soutput}
 #> @book{ihaka:1996,
@@
 #> }
-\end{Soutput}
+\end{Sinput}
 \end{Schunk}
@@
 \section{General usage}
+In general to use \CRANpkg{rebib} package, it is quite easy with requiring only one function call.
+\subsubsection{Using with RJournal article}
+In bibliography construction mode, \CRANpkg{rebib} will check for existing BibTeX files and will convert the embedded bibliography only when there are no linked BibTeX files in the RJ article.
+\begin{verbatim}
+%% typically rebib will search for this line in the RJ article, 
+%% not the RJwrapper file
+\bibliography{example}
+\end{verbatim}
+If there is no linked BibTeX file, \CRANpkg{rebib} will begin its procedure of converting the embedded bibliography to BibTeX and then link it with the article file as well.
+\begin{Schunk}
+\begin{Sinput}
+# for files without BibTeX source
+rebib::handle_bibliography(your_article_path)
+#> Warning in get_bib_file(article_dir, file_name): No Bib files found !
+#> BibTeX file does not exist
+#> will parse for bibliography
+#> bibtex file created
+cat(readLines(paste(your_article_path,"example.bib",sep="/")),sep = "\n")
+#> @book{ihaka:1996,
+#> author = {{Ihaka, Ross and Gentleman, Robert}},
+#> title = {{R: A Language for Data Analysis and Graphics}},
+#> publisher = {Journal of Computational and Graphical Statistics 3:   :},
+#> pages = {299--314},
+#> year = {1996},
+#> url = {https://doi.org/10.1080/10618600.1996.10474713}
+#> }
+#> @book{R,
+#> author = {R {Core Team}},
+#> title = {{R: A Language and Environment for Statistical Computing}},
+#> publisher = {R Foundation for Statistical Computing Vienna Austria    :},
+#> year = {2016},
+#> url = {https://www.R-project.org/},
+#> isbn = {3-900051-07-0}
+#> }
+\end{Sinput}
+\end{Schunk}
 
-\section{Bibliography aggregation}
+\subsubsection{Bibliography aggregation}
+
+If you have a BibTeX file and it is missing some references then this mode can help you read the embedded bibliography and combine the two bibliographies in a single BibTeX file.
+\begin{verbatim}
+# for files with BibTeX source as well as embedded entries
+rebib::aggregate_bibliography(your_article_path)
+#> Found Bib file  example.bib
+#> bibliography aggregation possible
+#> Found Bib file  example.bib
+#> aggregation delta : 2
+#> BibTeX file aggregated
+\end{verbatim}
+
+\subsubsection{Using with any LaTeX/BBL file}
+The bibliography converter will convert any LaTex or BBL file given as a path. This way, one can generate BibTeX from the embedded bibliography or BBL file.
+\begin{verbatim}
+rebib::biblio_converter(path_to_your_file)
+#> working directory : /temp/Rtmpt6Qk0v/exampledir/standalone
+#> bib entries : 6
+#> Written BibTeX file : /temp/Rtmpt6Qk0v/exampledir/standalone/sample.bib
+\end{verbatim}
 

### 2023-07-26 texor.Rmd
documented up to pre-processing

@@
 
-The decision to convert to R Markdown format means that the output to pdf and HTML will depend on Pandoc [@pandoc]. 
-Pandoc is a versatile document conversion program written in Haskell that is core to numerous documentation systems, including Rmarkdown and Quarto.
+The decision to convert to R Markdown format means that the final output to pdf and HTML will depend on Pandoc [@pandoc].
+Pandoc is a versatile document conversion program written in Haskell that is core to numerous documentation systems, including R Markdown and Quarto.
 Pandoc first converts a document into an abstract syntax tree.
 From this, it can convert to a different format, including custom ones.
-Although creating bespoke writers/readers in `Lua` can be challenging, pandoc includes filters that allow users to customize the output generated based on matching patterns plus some logic to modify them.
+<!-- Although creating bespoke writers/readers in `Lua` can be challenging, pandoc includes filters that allow users to customize the output generated based on matching patterns plus some logic to modify them.
+-->
 
-**DC: Can you explain what a Lua filter is?**
+Pandoc can be used to do the conversion from LaTeX to R Markdown also.
+However, additional pre-processing needs to be done to handle special R Journal LaTeX styling.
+And further post-processing needs to be done to handle specific R Journal R Markdown styling.
+The \CRANpkg{texor} package contains functionality to handle this pre- and postprocessing of the document.
 
-The \CRANpkg{texor} package uses these filters heavily, even for simple tasks like choosing which supporting figures to copy.
+<!-- uses these filters heavily, even for simple tasks like choosing which supporting figures to copy.
 A `Lua` filter constructs a list of image paths, stores them momentarily and the package uses this data to copy images.
@@
 This is needed to convert some text and LaTeX commands that are not recognized or handled by Pandoc.
+-->
 
-**DC: Add a diagram here showing flow of work, just pre-process .tex -> pandoc -> lua ???**
+**DC: Add a diagram here showing flow of work, just pre-process .tex -> pandoc -> lua ; rebib operations???**
 
-**Maybe more details on the supporting packages here: LaTeX, Regex, pandoc building blocks of the packages, methods used in them [MOW]**
+<!-- **Maybe more details on the supporting packages here: LaTeX, Regex, pandoc building blocks of the packages, methods used in them [MOW]**-->
 
@@
 
-LaTeX is very descriptive and allows unfettered levels of customization to authors, which poses an issue during conversions.
-RMarkdown is quite restrictive by  syntax though you can always embed HTML to expand the feature set.
-Although
-pandoc does most of the heavy lifting, there are some instances where even  filters cannot assist. A good demonstration of the issue in hand is code environments.
+LaTeX is very descriptive language, that allows authors substantial freedom for customization.
+Markdown [@markdown], on which R Markdown is based, is more restrictive and was born to make it easier to create web pages without the distraction of a gazillion HTML tags.
+The purpose of Markdown is that it allows the author to focus on writing, without format cluttering the text.
 
-A limitation of pandoc is that it only works with a "verbatim" environment.
+While pandoc can do most of the heavy-lifting, it cannot cope with all the freedom with which LaTeX documents are written.
+An example of this is with formatting of code.
+Pandoc only handles the `verbatim` environment, but there are many ways to format code in LaTeX, and the R Journal template has a special `\code{}` command.
+If the code environment is not verbatim, then pandoc will also try to process the actual code content as LaTeX commands and will likely lose details.
+It is better to convert these synonyms into `verbatim` environments this prior to passing the document to pandoc.
+
+The functions in \CRANpkg{texor} that handle the pre-processing using regular expressions are:
+
+- `stream_editor()`: operates like the `sed` function in unix [@unix] and allows generic text pattern matching and replacing.
+- `patch_code_env()`: replaces the common code environments, `code`, `example`, `Sin`, `Sout`, `Scode`, `Sinput`, `smallverbatim`, `boxedverbatim`, `smallexample` with `verbatim`.
+- `patch_equations()`: coordinates various equation environments.
+- `patch_figure_env()`: coordinates various figure environments.
+- `patch_table_env()`: coordinates table environments.
+
+These functions are verbose and describe all the changes being made. They also create a backup of the original file before making the changes.
+
+<!-- by syntax though you can always embed HTML to expand the feature set.
+Although pandoc does most of the heavy lifting, there are some instances where even  filters cannot assist.
+A good demonstration of the issue in hand is code environments.-->
+
+<!-- A limitation of pandoc is that it only works with a "verbatim" environment.
 However, in "RJournal",  we have multiple custom code environments which are not supported or defined due to lower-level tex limitations in pandoc.
@@
 Including a Lua filter to find these elements and redefine them as verbatim environments would also be  ineffective.
-If the code environment is not verbatim, then pandoc will also try to process the actual code content as LaTeX commands and will lose out on details in some cases.
+If the code environment is not verbatim, then pandoc will also try to process the actual code content as LaTeX commands and will lose out on details in some cases.-->
 
-**Be sure to mention the hard parts**
+<!--**Be sure to mention the hard parts**-->
 
-## Post-Processing : Lua Filters
+## Post-processing using Lua filters
 

### 2023-07-26 texor.tex
documented up to pre-processing

@@
 
-The decision to convert to R Markdown format means that the output to pdf and HTML will depend on Pandoc (MacFarlane 2023).
-Pandoc is a versatile document conversion program written in Haskell that is core to numerous documentation systems, including Rmarkdown and Quarto.
+The decision to convert to R Markdown format means that the final output to pdf and HTML will depend on Pandoc (MacFarlane 2023).
+Pandoc is a versatile document conversion program written in Haskell that is core to numerous documentation systems, including R Markdown and Quarto.
 Pandoc first converts a document into an abstract syntax tree.
 From this, it can convert to a different format, including custom ones.
-Although creating bespoke writers/readers in \texttt{Lua} can be challenging, pandoc includes filters that allow users to customize the output generated based on matching patterns plus some logic to modify them.
 
-\textbf{DC: Can you explain what a Lua filter is?}
+Pandoc can be used to do the conversion from LaTeX to R Markdown also.
+However, additional pre-processing needs to be done to handle special R Journal LaTeX styling.
+And further post-processing needs to be done to handle specific R Journal R Markdown styling.
+The \CRANpkg{texor} package contains functionality to handle this pre- and postprocessing of the document.
 
-The \CRANpkg{texor} package uses these filters heavily, even for simple tasks like choosing which supporting figures to copy.
-A \texttt{Lua} filter constructs a list of image paths, stores them momentarily and the package uses this data to copy images.
-However, \CRANpkg{texor} also includes many pre-processing functions that employ regular expressions.
-This is needed to convert some text and LaTeX commands that are not recognized or handled by Pandoc.
-
-\textbf{DC: Add a diagram here showing flow of work, just pre-process .tex -\textgreater{} pandoc -\textgreater{} lua ???}
-
-\textbf{Maybe more details on the supporting packages here: LaTeX, Regex, pandoc building blocks of the packages, methods used in them {[}MOW{]}}
+\textbf{DC: Add a diagram here showing flow of work, just pre-process .tex -\textgreater{} pandoc -\textgreater{} lua ; rebib operations???}
 
@@
 
-LaTeX is very descriptive and allows unfettered levels of customization to authors, which poses an issue during conversions.
-RMarkdown is quite restrictive by syntax though you can always embed HTML to expand the feature set.
-Although
-pandoc does most of the heavy lifting, there are some instances where even filters cannot assist. A good demonstration of the issue in hand is code environments.
+LaTeX is very descriptive language, that allows authors substantial freedom for customization.
+Markdown (Gruber 2002), on which R Markdown is based, is more restrictive and was born to make it easier to create web pages without the distraction of a gazillion HTML tags.
+The purpose of Markdown is that it allows the author to focus on writing, without format cluttering the text.
 
-A limitation of pandoc is that it only works with a ``verbatim'' environment.
-However, in ``RJournal'', we have multiple custom code environments which are not supported or defined due to lower-level tex limitations in pandoc.
-So the code environments are eliminated during the conversion process.
-Including a Lua filter to find these elements and redefine them as verbatim environments would also be ineffective.
-If the code environment is not verbatim, then pandoc will also try to process the actual code content as LaTeX commands and will lose out on details in some cases.
+While pandoc can do most of the heavy-lifting, it cannot cope with all the freedom with which LaTeX documents are written.
+An example of this is with formatting of code.
+Pandoc only handles the \texttt{verbatim} environment, but there are many ways to format code in LaTeX, and the R Journal template has a special \texttt{\textbackslash{}code\{\}} command.
+If the code environment is not verbatim, then pandoc will also try to process the actual code content as LaTeX commands and will likely lose details.
+It is better to convert these synonyms into \texttt{verbatim} environments this prior to passing the document to pandoc.
 
-\textbf{Be sure to mention the hard parts}
+The functions in \CRANpkg{texor} that handle the pre-processing using regular expressions are:
 
-\hypertarget{post-processing-lua-filters}{%
-\subsection{Post-Processing : Lua Filters}\label{post-processing-lua-filters}}
+\begin{itemize}
+\tightlist
+\item
+  \texttt{stream\_editor()}: operates like the \texttt{sed} function in unix (Ritchie and Thompson 1978) and allows generic text pattern matching and replacing.
+\item
+  \texttt{patch\_code\_env()}: replaces the common code environments, \texttt{code}, \texttt{example}, \texttt{Sin}, \texttt{Sout}, \texttt{Scode}, \texttt{Sinput}, \texttt{smallverbatim}, \texttt{boxedverbatim}, \texttt{smallexample} with \texttt{verbatim}.
+\item
+  \texttt{patch\_equations()}: coordinates various equation environments.
+\item
+  \texttt{patch\_figure\_env()}: coordinates various figure environments.
+\item
+  \texttt{patch\_table\_env()}: coordinates table environments.
+\end{itemize}
+
+These functions are verbose and describe all the changes being made. They also create a backup of the original file before making the changes.
+
+\hypertarget{post-processing-using-lua-filters}{%
+\subsection{Post-processing using Lua filters}\label{post-processing-using-lua-filters}}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-markdown}{}}%
+Gruber, John. 2002. {``Markdown.''} https://daringfireball.net/projects/markdown/.
+
 \leavevmode\vadjust pre{\hypertarget{ref-pandoc}{}}%
 MacFarlane, John. 2023. {``Pandoc: A Universal Document Converter.''} https://pandoc.org.
+
+\leavevmode\vadjust pre{\hypertarget{ref-unix}{}}%
+Ritchie, O. M., and K. Thompson. 1978. {``The UNIX Time-Sharing System.''} \emph{The Bell System Technical Journal} 57 (6): 1905--29. \url{https://doi.org/10.1002/j.1538-7305.1978.tb02136.x}.
 

### 2023-07-26 texor.Rmd
Lua filters - but still needs more work

@@
 Markdown [@markdown], on which R Markdown is based, is more restrictive and was born to make it easier to create web pages without the distraction of a gazillion HTML tags.
-The purpose of Markdown is that it allows the author to focus on writing, without format cluttering the text.
+The beauty of Markdown is that it allows the author to focus on writing, without format cluttering the text.
+The drawback is that it is simple typesetting, optimized for web delivery.
 
@@
 
-This was one of the earliest and most basic Lua filters, and it was part of the \CRANpkg{texor} package:
+Lua [@lua] is a programming language, that is light-weight, fast, ideal for procedural operations.
+It is embedded in many other applications to allow custom scripting for extensibility.
+Pandoc allows users to provide custom lua filters to produce custom output formats.
+The \CRANpkg{texor} package handles post-processing of the R Markdown document into the special format for the R Journal using a suite of Lua filters.
+
+Here is an example of a Lua filter in \CRANpkg{texor}:
 
@@
 
-Above filter reads the abstract syntax tree and filters out all the Div elements.
-Then it looks for the class "thebibliography." It turns out that this Div element  contains the LaTeX bibliographic records, which usually appear at the very end
-of papers. We don't need this portion in the article with the "RJ-web-article" layout,  given its inclusion as meta-data in the footer.
+This filter reads the abstract syntax tree and filters out all the Div elements.
+Then it looks for the class "thebibliography." It turns out that this Div element  contains the LaTeX bibliographic records, which usually appear at the very end of papers.
+We don't need this portion in the article with the "RJ-web-article" layout,  given its inclusion as meta-data in the footer.
 

### 2023-07-26 texor.tex
Lua filters - but still needs more work

@@
 Markdown (Gruber 2002), on which R Markdown is based, is more restrictive and was born to make it easier to create web pages without the distraction of a gazillion HTML tags.
-The purpose of Markdown is that it allows the author to focus on writing, without format cluttering the text.
+The beauty of Markdown is that it allows the author to focus on writing, without format cluttering the text.
+The drawback is that it is simple typesetting, optimized for web delivery.
 
@@
 
-This was one of the earliest and most basic Lua filters, and it was part of the \CRANpkg{texor} package:
+Lua (Ierusalimschy, Figueiredo, and Filho 1996) is a programming language, that is light-weight, fast, ideal for procedural operations.
+It is embedded in many other applications to allow custom scripting for extensibility.
+Pandoc allows users to provide custom lua filters to produce custom output formats.
+The \CRANpkg{texor} package handles post-processing of the R Markdown document into the special format for the R Journal using a suite of Lua filters.
+
+Here is an example of a Lua filter in \CRANpkg{texor}:
 
@@
 
-Above filter reads the abstract syntax tree and filters out all the Div elements.
-Then it looks for the class ``thebibliography.'' It turns out that this Div element contains the LaTeX bibliographic records, which usually appear at the very end
-of papers. We don't need this portion in the article with the ``RJ-web-article'' layout, given its inclusion as meta-data in the footer.
+This filter reads the abstract syntax tree and filters out all the Div elements.
+Then it looks for the class ``thebibliography.'' It turns out that this Div element contains the LaTeX bibliographic records, which usually appear at the very end of papers.
+We don't need this portion in the article with the ``RJ-web-article'' layout, given its inclusion as meta-data in the footer.
 
@@
 Gruber, John. 2002. {``Markdown.''} https://daringfireball.net/projects/markdown/.
+
+\leavevmode\vadjust pre{\hypertarget{ref-lua}{}}%
+Ierusalimschy, Roberto, Luiz Henrique de Figueiredo, and Waldemar Celes Filho. 1996. {``Lua---an Extensible Extension Language.''} \emph{Software: Practice and Experience} 26 (6): 635--52.
 

### 2023-07-26 texor.Rmd
Supplementary

@@
 
-# Structure of the \CRANpkg{texor} package
+## Conversion
 
-**DC: List the main functions in an organised way, could be the pre-processing functions, the post-processing functions, utility functions.
-This can be also be summarized in a table, but there needs to be text explanations of the main functionality.**
+The only function that a user necessarily needs is `latex_to_web()`. This creates the R Journal style R Markdown file from a given R Journal style LaTeX file. 
+This is achieved in several conversion steps, `convert_to_markdown()`, `generate_rmd()` and `produce_html()`.
+
+## Handling figures
+
+## Logging the conversion
 
@@
 **Examples / getting started / usage and so on**
+
+Run through the different examples in supplementary here
+
+**From general LaTeX to generic R Markdown**
 
@@
 
- github repos for the packages, texor, rebib, rjtools
+Several supplementary example documents are provided which illustrate how different elements of LaTeX-authored papers are handled. These are:
+
+-
+
+The GitHub repos supporting this work are:
+
+- \CRANpkg{texor}: https://abhi-1u.github.io/texor
+- \CRANpkg{rebib}: https://abhi-1u.github.io/rebib/
+- This paper: https://github.com/Abhi-1U/texor-rjarticle
+- More details on \CRANpkg{rjtools} are at https://rjournal.github.io/rjtools/

### 2023-07-26 texor.tex
Supplementary

@@
 
-\hypertarget{structure-of-the-package}{%
-\section{\texorpdfstring{Structure of the \CRANpkg{texor} package}{Structure of the  package}}\label{structure-of-the-package}}
+\hypertarget{conversion}{%
+\subsection{Conversion}\label{conversion}}
 
-\textbf{DC: List the main functions in an organised way, could be the pre-processing functions, the post-processing functions, utility functions.
-This can be also be summarized in a table, but there needs to be text explanations of the main functionality.}
+The only function that a user necessarily needs is \texttt{latex\_to\_web()}. This creates the R Journal style R Markdown file from a given R Journal style LaTeX file.
+This is achieved in several conversion steps, \texttt{convert\_to\_markdown()}, \texttt{generate\_rmd()} and \texttt{produce\_html()}.
+
+\hypertarget{handling-figures}{%
+\subsection{Handling figures}\label{handling-figures}}
+
+\hypertarget{logging-the-conversion}{%
+\subsection{Logging the conversion}\label{logging-the-conversion}}
 
@@
 \textbf{Examples / getting started / usage and so on}
+
+Run through the different examples in supplementary here
+
+\textbf{From general LaTeX to generic R Markdown}
 
@@
 
-github repos for the packages, texor, rebib, rjtools
+Several supplementary example documents are provided which illustrate how different elements of LaTeX-authored papers are handled. These are:
+
+\begin{itemize}
+\tightlist
+\item
+\end{itemize}
+
+The GitHub repos supporting this work are:
+
+\begin{itemize}
+\tightlist
+\item
+  \CRANpkg{texor}: \url{https://abhi-1u.github.io/texor}
+\item
+  \CRANpkg{rebib}: \url{https://abhi-1u.github.io/rebib/}
+\item
+  This paper: \url{https://github.com/Abhi-1U/texor-rjarticle}
+\item
+  More details on \CRANpkg{rjtools} are at \url{https://rjournal.github.io/rjtools/}
+\end{itemize}
 

### 2023-07-27 supplement/figure-env/figure-example.tex
Updated Supplementary materials

@@
 
+\section{Other elements in figure objects}
+Figures can also house non-image environments like CodeBlocks and Para. This 
+also creates an opportunity to add support for CodeBlock numbering with some
+lua filter trickery.
+\begin{figure}[htbp]
+\begin{center}
+\begin{verbatim}
+code_in_figure <- function() {
+  if (pandoc_version >= 3) {
+    print("Code in Figure Supported")
+  }
+  else {
+    print("code in Figure not supported")
+  }
+}
+\end{verbatim}
+\caption{ Example Code inside Figure environment}
+\label{code:example}
+\end{center}
+\end{figure}
 \section{Summary}

### 2023-07-27 supplement/math-env/math-example.tex
Supplement: Math environment

% !TeX root = RJwrapper.tex
\title{Working with Math environments in texor}
\author{by Abhishek Ulayil}

\maketitle

\abstract{
This is a small sample article to demonstrate usage of texor to convert math environments.
}

\section{Introduction}
Math typesetting has always been LaTeX’s highlight feature, making it a de facto choice among academicians and researchers globally. However, as we proceed to our humble web interfaces, math is hard to describe traditionally. There have been advancements in JavaScript libraries to better Typeset and present math in web pages but not all LaTeX commands/math functions are available.

\subsubsection{MathJax}
The \CRANpkg{texor} package uses Mathjax version 3 to enhance the visual look of the math content in HTML.

The core of the MathJax project is the development of its state-of-the-art, open source, JavaScript platform for display of mathematics. The key design goals are \citep{mathjax}:
\begin{itemize}
    \item High-quality display of mathematics notation in all browsers.
    \item No special browser setup required.
    \item Support for LaTeX, MathML, and other equation markup directly in the HTML source.
    \item An extensible, modular design with a rich API for easy integration into web applications.
    \item Support for accessibility, copy and paste, and other rich functionality.
    \item Interoperability with other applications and math-aware search.
    \item Support for equation conversion outside a browser (e.g., preprocessing on a server).
\end{itemize}

\subsubsection{Pandoc Handling, extensions}
As \CRANpkg{texor} calls \CRANpkg{rmarkdown} to render the Rmarkdwon file into HTML,
the \verb|rjtools::rjournal_web_article| template by default uses MathJax as the math engine. 
Also we specify the Mathjax version in the metadata of the generated Rmarkdown file.

\section{Inline math}
One can define Inline Math in LaTeX using commands \verb|\(..\)| or \verb|$..$|.
It is also possible to use inline math within captions and  
\begin{table}[htbp]
  \centering
  \begin{tabular}{| c | c |}
  \hline
  Command & Math\\
    \hline
    \begin{minipage}{0.45\textwidth}
\vspace{1mm}
\begin{example}
\(\mu = (0,0,0)^\top \)
\end{example}
    \end{minipage} &
    \begin{minipage}{0.45\textwidth}
    \centering
    \( \mu = (0,0,0)^\top \)
    \end{minipage}\\
    \hline

  \begin{minipage}{0.45\textwidth}
\vspace{1mm}
\begin{example}
$\mu = (0,0,0)^\top $
\end{example}
    \end{minipage} &
    \begin{minipage}{0.45\textwidth}
    \centering
    $ \mu = (0,0,0)^\top $
    \end{minipage}\\
    \hline
  \end{tabular}
  \caption{InlineMath syntax and its output side by side}
  \label{table:1}
\end{table}

\section{Display Math}

\begin{align}
f(x) = \frac{1}{\sigma\sqrt{2\pi}} 
  \exp\left( -\frac{1}{2}\left(\frac{x-\mu}{\sigma}\right)^{\!2}\,\right)
  \label{eq:2}
\end{align}

\section{Equation Numbering}
\begin{align}
S_{T, s}(z_t) = & X^{\top} K_{b,t}^* X (Z - z_t)^s \nonumber\\
T_{T, s}(z_t) = &X^{\top} K_{b,t}^* Y (Z - z_t)^s,
\label{eq:tvpols}
\end{align}

\section{Unsupported LaTeX Commands}

like instead of \verb|\bm| \verb|\boldmath| use \verb|\mathbb|
avoid custom math commands MathJaX will never be able to parse it

\section{Summary}

In summary the \CRANpkg{texor} package with the help of pandoc and MathJax supports:
\begin{itemize}
\item some common math environments.
\item Equation Numbering
\end{itemize}

\begin{thebibliography}{2}
    \providecommand{\natexlab}[1]{#1}
    \providecommand{\url}[1]{\texttt{#1}}
    \expandafter\ifx\csname urlstyle\endcsname\relax
      \providecommand{\doi}[1]{doi: #1}\else
      \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi

\bibitem[Krewinkel, Lucero (2023)]{pandoc}
A.~ Krewinkel and A.~ Lucero
\newblock pandoc 3.0 Release notes
\newblock \emph{pandoc} \penalty0 2023
\newblock URL \url{https://pandoc.org/releases.html#pandoc-3.0-2023-01-18}

\bibitem[Mathjax authors (2021)]{mathjax}
MathJax authors
\newblock Mathjax website
\newblock  \penalty0 2021
\newblock URL \url{https://www.mathjax.org/#features}

\end{thebibliography}

\address{%
Abhishek Ulayil\\
Student, Institute of Actuaries of India\\%
Mumbai, India\\
ORCiD: 0009-0000-6935-8690\\
}

### 2023-07-28 supplement/math-env/math-example.tex
Updated math supplementary material

@@
 
+Display Math refers to equations, almost all LaTeX equations can be rendered by
+MathJax.
+
+\begin{verbatim}
 \begin{align}
@@
   \exp\left( -\frac{1}{2}\left(\frac{x-\mu}{\sigma}\right)^{\!2}\,\right)
-  \label{eq:2}
+  \label{eq:1}
+\end{align}
+\end{verbatim}
+This will render as \ref{eq:1}.
+\begin{align}
+f(x) = \frac{1}{\sigma\sqrt{2\pi}} 
+  \exp\left( -\frac{1}{2}\left(\frac{x-\mu}{\sigma}\right)^{\!2}\,\right)
+  \label{eq:1}
 \end{align}
@@
 \section{Equation Numbering}
+In LaTeX your equations get numbered automatically unless you are using \verb|\[..\]| to describe
+math. Equation numbering works a bit differently in \CRANpkg{bookdown} (the base of RJ web article format) where
+it is mandatory to have a \verb|(\#eq:xx)| which is described in more detail in \citep{bookdown}.
+
+The \CRANpkg{texor} package eases authors from manually adding \verb|(\#eq:xx)| to equations in Rmarkdown
+by using pandoc Lua filter to convert existing \verb|\label{..}| in the equations to \verb|(\#eq:xx)|
+during conversion. Equation \ref{eq:xe} shows such a use case of equation being numbered by its label in Rmakrdown
+as well as LaTeX.
+
 \begin{align}
-S_{T, s}(z_t) = & X^{\top} K_{b,t}^* X (Z - z_t)^s \nonumber\\
-T_{T, s}(z_t) = &X^{\top} K_{b,t}^* Y (Z - z_t)^s,
-\label{eq:tvpols}
+S_{T, s}(z_t) = & X^{\top} K_{b,t}^* X (Z - z_t)^s 
+\label{eq:xe}
 \end{align}
+
+Equation labels must start with the prefix \verb|eq:| in bookdown.
+All labels in bookdown must only contain alphanumeric characters, \verb|:|, \verb|-|, and/or \verb|/| as suggested in \citep{bookdown}.
 
@@
 
-like instead of \verb|\bm| \verb|\boldmath| use \verb|\mathbb|
-avoid custom math commands MathJaX will never be able to parse it
+Altough MathJax does a good job of supporting most LaTeX math functions,
+some functions do not work like \verb|\bm| \verb|\boldmath|, instead use \verb|\mathbb|.
+Also avoid custom math commands MathJaX will never be able to parse it. The reason being
+MathJax renders math case by case and custom commands will already be removed by pandoc
+during the conversion, hence MathJax will not be aware of such redefinations.
 
@@
 
+\bibitem[Yihui (2023)]{bookdown}
+Yihui Xie
+\newblock bookdown: Authoring Books and Technical Documents with R Markdown
+\newblock \penalty0 2023
+\newblock URL \url{https://bookdown.org/yihui/bookdown/markdown-extensions-by-bookdown.html#equations}
+
 \end{thebibliography}

### 2023-07-28 supplement/metadata/metadata.tex
New Supplementary article for metadata

% !TeX root = RJwrapper.tex
\title{Metadata and Other Structures in texor}
\author{by Abhishek Ulayil}

\maketitle

\abstract{
This is a small sample article to demonstrate handling of metadata and other structures during conversion in texor package.
}

\section{Introduction}
There is a difference in the structure of metadata and generally Rmarkdown has more
meta-data entries than its LaTeX counterpart. So for the conversion, main task is to 
transform the meta data and include it in the generated Rmarkdown file, matching the
LaTeX data as closely as possible.

Pandoc when used with \verb|-s| or \verb|standalone| mode reads out some of the existing
meta data. The extracted author meta-data are listed below

\subsubsection{Metadata Read from LaTeX file}

\begin{itemize}
  \item Author Names 
  \item Author address(affiliation)
  \item ORCID number
  \item Title
  \item Bibliography
  \item Abstract
  \item CRAN/BIO packages
\end{itemize}

The \verb|\address{..}| section must be typeset in this manner for the most 
effective conversion.
\begin{verbatim}
\address{
Abhishek Ulayil\\ % Your Name in the first Line
Student, Institute of Actuaries of India\\% Your affiliation which includes post and/or org
Mumbai, India\\ % Your City/Country
ORCiD: 0009-0000-6935-8690\\% ORCID number with prefix ORCiD: 
}
\end{verbatim}

\subsubsection{Metadata Read from DESCRIPTION file}

\begin{itemize}
  \item Date
  \item Slug
  \item Volume and Issue
\end{itemize}

Some details like the Volume and Issue were picked up from folder structure as well
in some legacy articles where DESCRIPTION file was not available.

\subsubsection{Other supported commands in conversion}

\begin{table}[htbp]
\centering
\begin{tabular}{| c | c |}
 \hline
 Command & Render \\
 \hline
 \verb|\acronym{A}| & \acronym{A} \\ \hline
\verb|\R| & \R  \\ \hline
 \verb|\pkg{texor}|& \pkg{texor} \\ \hline
 \verb|\CRANpkg{texor}|& \CRANpkg{texor} \\ \hline
 \verb|\BIOpkg{Biobase}|      & \BIOpkg{Biobase}  \\ \hline
 \verb|\ctv{ReproducibleResearch}| & \ctv{ReproducibleResearch}  \\ \hline
\verb|\command{help}| & \command{help} \\
\hline
\end{tabular}
\caption{Other commands supported by \CRANpkg{texor} and Pandoc}
\label{table:1}
\end{table}

\section{Miscallaneous}

\subsubsection{Citations}

As we know about citations (using \verb|\citet{..}| command here) [\citet{moreverb}]. We could also use \verb|\citep{..}| command in real \citep{pandoc}.

\begin{table}[htbp]
\centering
\begin{tabular}{| c | c |}
 \hline
 Command & Render \\
 \hline
 \verb|\citet{pandoc}| & \citet{pandoc} \\ \hline
\verb|\citep{pandoc}| & \citep{pandoc} \\ \hline
 \verb|\citet*{pandoc}| & \citet*{pandoc} \\ \hline
\verb|\citep*{pandoc}| & \citep*{pandoc} \\ \hline
 \verb|\citeyear{pandoc}| & \citeyear{pandoc} \\ \hline
\verb|\citeauthor{pandoc}| & \citeauthor{pandoc} \\ \hline
 \verb|\cite{pandoc}| & \cite{pandoc} \\ \hline
\end{tabular}
\caption{Different citations in LaTeX}
\label{table:3}
\end{table}

\subsubsection{Footnotes}
\begin{table}[htbp]
\centering
\begin{tabular}{| c | c |}
 \hline
 Command & Render \\
 \hline
 \verb|\acronym{A}\footnote{This works in Web article}| & \acronym{A}\footnote{This works in Web article} \\ \hline
\verb|\R\footnote[21]{Its 2 not 21 :(}| & \R\footnote[21]{Its 2 not 21 :(}  \\ \hline
\end{tabular}
\caption{Demo of footnotes}
\label{table:3}
\end{table}

Footnotes do not work in \verb|tabular| environment in LaTeX but it does work in Rmarkdown\footnote{This works in LaTeX and  Web article}.
Custom Footnote number works in LaTeX but does not work in web articles\footnote[21]{Forever 21 :)}.

\subsubsection{BlockQuotes}
\begin{quote}
This BlockQuote should Justify an example.
\end{quote}

\subsubsection{Links}
Most links work just fine using \verb|\href| and \verb|\url| command will work flawlessly.
\begin{table}[htbp]
\centering
\begin{tabular}{| c | c |}
 \hline
 Command & Render \\
 \hline
 \verb|\href{www.google.com}{Google}| & \href{www.google.com}{Google} \\ \hline
\verb|\url{www.google.com}| & \url{www.google.com}  \\ \hline
\verb|Table \ref{table:3}| & Table \ref{table:3}  \\ \hline
\end{tabular}
\caption{Links}
\label{table:4}
\end{table}

\section{Summary}
In summary the \CRANpkg{texor} package alongside pandoc supports:
\begin{itemize}
\item Extracting Metadata and putting it together well.
\item Handling Links, BlockQuotes, Citations and Footnotes.
\end{itemize}

\begin{thebibliography}{2}
    \providecommand{\natexlab}[1]{#1}
    \providecommand{\url}[1]{\texttt{#1}}
    \expandafter\ifx\csname urlstyle\endcsname\relax
      \providecommand{\doi}[1]{doi: #1}\else
      \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi

\bibitem[Krewinkel, Lucero (2023)]{pandoc}
A.~ Krewinkel and A.~ Lucero
\newblock pandoc 3.0 Release notes
\newblock \emph{pandoc} \penalty0 2023
\newblock URL \url{https://pandoc.org/releases.html}

\bibitem[Fairbairns, Duggan, Schöpf, Eijkhout (2011)]{moreverb}
R.~ Fairbairns, A.~ Duggan, R.~ Schöpf and V.~ Eijkhout
\newblock The moreverb package documentation
\newblock \emph{CTAN} \penalty0 2011
\newblock URL \url{https://mirror.niser.ac.in/ctan/macros/latex/contrib/moreverb/moreverb.pdf}

\end{thebibliography}

\address{%
Abhishek Ulayil\\
Student, Institute of Actuaries of India\\%
Mumbai, India\\
ORCiD: 0009-0000-6935-8690\\
}

### 2023-07-29 supplement/table-env/table-example.tex
New Supplement for tables

% !TeX root = RJwrapper.tex
\title{Working with table environments in texor}
\author{by Abhishek Ulayil}

\maketitle

\abstract{
This is a small sample article to demonstrate usage of \CRANpkg{texor} to convert table environments.
}

\section{Introduction}
Tables are commonly used in RJ articles to display data in a tabular format. However, there are differences in the way tables are handled by LaTeX and HTML.
LaTeX tables have more customization and are usually optimized for printing, whereas the web articles need tables optimized for varying sizes of media.
Pandoc converts most of the tables somewhat easily, but is unable to do well with table customization packages and complex tables.
Some pandoc extensions are used in order to tackle them, they are : \verb|simple_tables|,\verb|pipe_tables|
\section{Generic tables}
Normal LaTeX tables will be converted just fine, even if they include a bit of math and other elements.
However do note that the conversion will be done to traditional or pandoc style markdown instead of Rmarkdown
specific using \verb|knitr::kable(..)| function. Also any custom LaTeX commands or specific font charactars would
not be supported.
\begin{table}[htbp]
\centering
\begin{tabular}{l | llll }
 \hline
 Graphics Format & LaTeX & Markdown & Rmarkdown & HTML \\
 \hline
 PNG       & Yes & Yes & Yes & Yes \\
 JPG       & Yes & Yes & Yes & Yes \\
 PDF       & Yes & No & No & No \\
 SVG       & No & Yes & Yes & Yes \\
 Tikz      & Yes & No & Yes & No \\
 Algorithm & Yes & No & No & No \\
\hline
\end{tabular}
\caption{Image Format support in various Markup/Typesetting Languages}
\label{table:1}
\end{table}

\section{Multicolumn tables}
\begin{table*}[htbp]
\begin{center}
\begin{tabular}{l | llll }
 \hline
 \multicolumn{1}{c |}{EXAMPLE} & \multicolumn{2}{c}{$X$} &
\multicolumn{2}{c}{$Y$} \\
 \hline
  & 1 & 2 & 1 & 2 \\
 EX1  & X11 & X12 &  Y11  & Y12 \\
 EX2  & X21 & X22 &  Y21  & Y22 \\
 EX3  & X31 & X32 &  Y31  & Y32 \\
 EX4  & X41 & X42 &  Y41  & Y42\\
 EX5  & X51 & X52 &  Y51  & Y52 \\
\hline
\end{tabular}
\caption{An Example Table with Multicolumn}
\label{table:2}
\end{center}
\end{table*}

The above example illustrates a multicolumn table \ref{table:2}, which the \CRANpkg{texor} package can handle through subroutines and pre-processing step to transform the LaTeX code.

\begin{itemize}
    \item The Stream Editor modifies the \verb|\multicolumn{..}| to \verb|\multicolumnx{..}|
    \item Then a LaTeX macro redefines the \verb|\multicolumnx{..}| to \verb|\multicolumn{---}|, which is accepted by pandoc.
    \item Finally pandoc reads the table and transforms it to markdown.
\end{itemize}

Also note that the stream editor is used to rename \verb|table*| environment to \verb|table| environment because the markdown/HTML will not support the specific changes table* environment will bring.

\section{Complex tables}
A complex table with various other elements like figures, math , code and so on, are also supported
by \CRANpkg{texor}.  
\begin{table}[htbp]
\centering
\begin{tabular}{|l | lll |}
 \hline
 Inline Format & LaTeX Support & Web Support &  Rendering \\
 \hline
 Text       & Yes & Yes  & \emph{Hello} \\ \hline
 Image       & Yes & Yes & \begin{minipage}{0.25\textwidth}
    \centering
    \includegraphics[width=1\textwidth]{Rlogo-5.png}
    \end{minipage} \\ \hline
 CodeBlock      & Yes & Yes & \begin{minipage}{0.45\textwidth}
\vspace{1mm}
\begin{example}
x <- 1:100
y <- dbinom(x,100,prob = 0.5)
plot(x,y)
\end{example}
    \end{minipage} \\ \hline
 Math      & Yes & Yes & $e = m c^2 $ \\ \hline
 Link      & Yes & Yes &  \href{www.google.com}{Google} \\ \hline
 Nested Table & Yes & No & NaN \\
\hline
\end{tabular}
\caption{Image Format support in various Markup/Typesetting Languages}
\label{table:3}
\end{table}

\section{Wide tables}

\begin{widetable}
\centering
\begin{tabular}{lllllllll}
\toprule
index  &  A  &  B     & C   &  D  &  E  &  F  &  G  & H  \\
\midrule
1 & 359.00 & NaN & 5796.00 & 0.00 & 16.14 & 1.00 & NaN & 0.00\\
2 & 25.73 & 0.00 & 1029.20 & NaN & 40.00 & 0.68 & 0.00 & 0.00\\
2.1 & 26.26 & 0.00 & 13.40 & 0.00 & 2.14 & 0.68 & 0.00 & NaN\\
2.2 & 32.06 & 20.06 & 47.64 & 0.04 & 1.80 & 0.68 & 0.01 & NaN\\
2.3 & 51.94 & 420.27 & 21.17 & 0.20 & 1.77 & 0.74 & 0.05 & NaN\\
2.4 & 40.62 & 30.44 & \phantom{000}0.90 & 0.57 & 1.44 & 1.31 & 0.24 & NaN\\
 & \ldots & \ldots&  \ldots& \ldots &  \ldots& \ldots& \ldots & \ldots \\
\bottomrule
\end{tabular}
\begin{tabular}{llllllll}
\toprule
index &    I  &  J  &  K  &  L  &  M  &  N & O \\
\midrule
1 & 0.00 & $-$1.73 & 0.00 & 0.00 & 0.00 & 0.00 & NaN\\
2 & 0.21 & $-$33.41 & 0.00 & 0.11 & 0.00 & NaN & $-$197.85\\
2.1 & 0.24 & $-$24.00 & 0.00 & 0.15 & NaN & 0.06 & $-$70.46\\
2.2 & 0.6 & $-$19.42 & 0.00 & 0.15 & 0.00 & 0.11 & $-$16.48\\
2.3 & 0.75 & $-$ 31.77 & 0.00 & 0.18 & 0.01 & 0.37 & $-$0.82\\
2.4 & 0.26 & $-$1.89 & 0.1 & 0.55 & 0.70 & 0.22 & $-$6.55 \\
  &\ldots &  \ldots&  \ldots& \ldots & \ldots & \ldots & \ldots \\
\bottomrule
\end{tabular}
\caption{\code{A dummy research data}}
\label{table:4}

\end{widetable}

Widetable environments also gets some perks when converted by \CRANpkg{texor} package. 
They get their own block and seperate numbering (although references are shared by figures).

\begin{widetable}
\centering %
\begin{tabular}{llrr}
\toprule 
Package  & Commits  & Version  & Last Updated \tabularnewline
\toprule 
\CRANpkg{texor} & 260 & 1.1.0  & 28-Jul-2023 \tabularnewline
\CRANpkg{rebib} & 76 & 0.2.4  & 29-Jul-2023 \tabularnewline
\CRANpkg{rjtools}  & 314 & 1.0.11  & 30-Jul-2023 \tabularnewline
\CRANpkg{rmarkdown} & 3189 & 2.23  & 31-Jul-2023 \tabularnewline
\toprule 
\end{tabular}
\caption{A dummy summary of a few CRAN packages}
\label{table:5}

\end{widetable}

Here is a reference to the table \ref{table:4} and table \ref{table:5}.
\section{Summary}

In summary the \CRANpkg{texor} package supports:
\begin{itemize}
\item some common table environments.
\item wide tables.
\item some multicolumn tables.
\item inclusion of different environments like figure/code in tables.
\end{itemize}

\begin{thebibliography}{2}
    \providecommand{\natexlab}[1]{#1}
    \providecommand{\url}[1]{\texttt{#1}}
    \expandafter\ifx\csname urlstyle\endcsname\relax
      \providecommand{\doi}[1]{doi: #1}\else
      \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi

\bibitem[Krewinkel, Lucero (2023)]{pandoc}
A.~ Krewinkel and A.~ Lucero
\newblock pandoc 3.0 Release notes
\newblock \emph{pandoc} \penalty0 2023
\newblock URL \url{https://pandoc.org/releases.html}

\end{thebibliography}

\address{%
Abhishek Ulayil\\
Student, Institute of Actuaries of India\\%
Mumbai, India\\
ORCiD: 0009-0000-6935-8690\\
}

### 2023-07-31 texor.Rmd
overview of current examples

@@
 ---
-title: "Converting LaTeX R Journal Articles into Rmarkdown using texor and rebib"
+title: "Converting LaTeX Legacy R Journal Articles into R Markdown Articles using texor and rebib"
 date: "2022-11-09"
@@
   - name: Abhishek Ulayil
-    affiliation: Student, Institute of Actuaries of India
+    affiliation: Institute of Actuaries of India (student)
     address:
@@
 
-# Converting from LaTeX to R Markdown
+# The internals of converting from LaTeX to R Markdown
 
@@
 **DC: Add a diagram here showing flow of work, just pre-process .tex -> pandoc -> lua ; rebib operations???**
+
+The supplementary materials has example LaTeX documents that allow the reader to see how different common patterns in the legacy documents are handled with the conversion. These include:
+
+- `code-env`: Explains how different code environments defined by the R Journal style are handled, and additional details such as code in figure environments, and code in table environments. 
+- `math-env`: Examples of inline math, display math, how equation numbering is handled by a Lua filter to convert from LaTeX labelling to R Markdown labelling. 
+- `figure-env`: Explains how the variety of figure definitions are handled in the conversion, including different image formats, numbering, captions, labelling, multiple images, and `tikz` [@casflow] images.
+- `table-env`: Examples of how a variety of table types are converted, including multicolumn, complex and wide tables.
+- `lua-filters`: Overview and lots of small examples of Lua filters to handle the custom output needed for the R Markdown format.
+- `metadata`: This has a collection of additional format handling including extracting metadata like author names and affiliations, article identifiers used in the review process, and handling citations, footnotes and links.
+- `bibliography`: The bibliography was handled differently over the years of the journal, and this details how to use the `rebib` functionality to handle `bbl` files, embedded `bbl`, to convert into the standard `.bib` format.
+
+The explanations in the next sections are extracted from these examples.
+In each of these folders there is a `RJwrapper.tex`, and `.tex` file, with the extra template files `RJournal.sty` and `Rlogo-5.png`.
+These match the legacy template file structure, from which the `RJwrapper.pdf` file is created.
+If you follow the instructions in the `README.md` you can create the R Markdown and HTML versions of the document.
 
@@
 
-Here is an example of a Lua filter in \CRANpkg{texor}:
+Here is an example of a Lua filter available in \CRANpkg{texor}:
 
@@
 
-This filter reads the abstract syntax tree and filters out all the Div elements.
-Then it looks for the class "thebibliography." It turns out that this Div element  contains the LaTeX bibliographic records, which usually appear at the very end of papers.
-We don't need this portion in the article with the "RJ-web-article" layout,  given its inclusion as meta-data in the footer.
+This filter reads the abstract syntax tree, selecting all the Div elements.
+Then it looks for the class "thebibliography." This Div element contains the LaTeX bibliographic records, that appear at the very end of papers.
+It should not be in the document when using the "RJ-web-article" layout, because it is added from meta-data when the R Markdown is knitted.
+So the Lua filter removes this section.
 
-## Conversion
+## Handling figures
+
+## Handling equations
+
+## Handling the bibliography
+
+## Logging the conversion
+
+## One-step user-facing conversion
 
@@
 
-## Handling figures
-
-## Logging the conversion
-
 # Using \CRANpkg{texor}
 
-**Install instructions**
+The package `texor` can be installed from CRAN using:
+
+```
+install.packages("texor")
+```
+
+and the development version from https://abhi-1u.github.io/texor.
 

### 2023-07-31 texor.tex
overview of current examples

@@
 % !TeX root = RJwrapper.tex
-\title{Converting LaTeX R Journal Articles into Rmarkdown using texor and rebib}
+\title{Converting LaTeX Legacy R Journal Articles into R Markdown Articles using texor and rebib}
 
@@
 
-\hypertarget{converting-from-latex-to-r-markdown}{%
-\section{Converting from LaTeX to R Markdown}\label{converting-from-latex-to-r-markdown}}
+\hypertarget{the-internals-of-converting-from-latex-to-r-markdown}{%
+\section{The internals of converting from LaTeX to R Markdown}\label{the-internals-of-converting-from-latex-to-r-markdown}}
 
@@
 \textbf{DC: Add a diagram here showing flow of work, just pre-process .tex -\textgreater{} pandoc -\textgreater{} lua ; rebib operations???}
+
+The supplementary materials has example LaTeX documents that allow the reader to see how different common patterns in the legacy documents are handled with the conversion. These include:
+
+\begin{itemize}
+\tightlist
+\item
+  \texttt{code-env}: Explains how different code environments defined by the R Journal style are handled, and additional details such as code in figure environments, and code in table environments.
+\item
+  \texttt{math-env}: Examples of inline math, display math, how equation numbering is handled by a Lua filter to convert from LaTeX labelling to R Markdown labelling.
+\item
+  \texttt{figure-env}: Explains how the variety of figure definitions are handled in the conversion, including different image formats, numbering, captions, labelling, multiple images, and \texttt{tikz} (Josh Cassidy 2013) images.
+\item
+  \texttt{table-env}: Examples of how a variety of table types are converted, including multicolumn, complex and wide tables.
+\item
+  \texttt{lua-filters}: Overview and lots of small examples of Lua filters to handle the custom output needed for the R Markdown format.
+\item
+  \texttt{metadata}: This has a collection of additional format handling including extracting metadata like author names and affiliations, article identifiers used in the review process, and handling citations, footnotes and links.
+\item
+  \texttt{bibliography}: The bibliography was handled differently over the years of the journal, and this details how to use the \texttt{rebib} functionality to handle \texttt{bbl} files, embedded \texttt{bbl}, to convert into the standard \texttt{.bib} format.
+\end{itemize}
+
+The explanations in the next sections are extracted from these examples.
+In each of these folders there is a \texttt{RJwrapper.tex}, and \texttt{.tex} file, with the extra template files \texttt{RJournal.sty} and \texttt{Rlogo-5.png}.
+These match the legacy template file structure, from which the \texttt{RJwrapper.pdf} file is created.
+If you follow the instructions in the \texttt{README.md} you can create the R Markdown and HTML versions of the document.
 
@@
 
-This filter reads the abstract syntax tree and filters out all the Div elements.
-Then it looks for the class ``thebibliography.'' It turns out that this Div element contains the LaTeX bibliographic records, which usually appear at the very end of papers.
-We don't need this portion in the article with the ``RJ-web-article'' layout, given its inclusion as meta-data in the footer.
+This filter reads the abstract syntax tree, selecting all the Div elements.
+Then it looks for the class ``thebibliography.'' This Div element contains the LaTeX bibliographic records, that appear at the very end of papers.
+It should not be in the document when using the ``RJ-web-article'' layout, because it is added from meta-data when the R Markdown is knitted.
+So the Lua filter removes this section.
 
-\hypertarget{conversion}{%
-\subsection{Conversion}\label{conversion}}
+\hypertarget{handling-figures}{%
+\subsection{Handling figures}\label{handling-figures}}
+
+\hypertarget{handling-equations}{%
+\subsection{Handling equations}\label{handling-equations}}
+
+\hypertarget{handling-the-bibliography}{%
+\subsection{Handling the bibliography}\label{handling-the-bibliography}}
+
+\hypertarget{logging-the-conversion}{%
+\subsection{Logging the conversion}\label{logging-the-conversion}}
+
+\hypertarget{one-step-user-facing-conversion}{%
+\subsection{One-step user-facing conversion}\label{one-step-user-facing-conversion}}
 
@@
 
-\hypertarget{handling-figures}{%
-\subsection{Handling figures}\label{handling-figures}}
-
-\hypertarget{logging-the-conversion}{%
-\subsection{Logging the conversion}\label{logging-the-conversion}}
-
 \hypertarget{using}{%
@@
 
-\textbf{Install instructions}
+The package \texttt{texor} can be installed from CRAN using:
+
+\begin{verbatim}
+install.packages("texor")
+\end{verbatim}
+
+and the development version from \url{https://abhi-1u.github.io/texor}.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-casflow}{}}%
+Josh Cassidy. 2013. \emph{{LaTeX Graphics using TikZ: A Tutorial for Beginners (Part 3)---Creating Flowcharts}}. Overleaf tutorials. \url{https://www.overleaf.com/learn/latex/}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-pandoc}{}}%
@@
 Abhishek Ulayil\\
-Student, Institute of Actuaries of India\\%
+Institute of Actuaries of India (student)\\%
 Mumbai, India\\

### 2023-08-01 supplement/lua-filters/lua-filters.tex
Updated Lua filter supplement

@@
 \section{Introduction}
-To understand pandoc filters, one must look at the document conversion process in
-general. A document is read/parsed in one format to a common intermediate format.
-In pandoc it is called an abstract syntax tree (AST). A document writer is designed 
-to read the AST and write the contents in the form of desired article format.
-For example converting markdown to HTML will involve these processes in pandoc.
+To fully understand pandoc filters, one must first grasp the document conversion process.
+A document in one format is read/parsed into an intermediate form.
+An abstract syntax tree (AST) is what pandoc calls it.
+A document writer is a specialized type-setter to read an AST and produces the contents in the chosen article format.
+Converting markdown to HTML, for example, will entail these pandoc operations.
+Figure\ref{fig:workflow} shows the conversion workflow withing pandoc.
 
@@
   \centering
-  \includegraphics[width=0.6\textwidth]{workflow.png}
+  \includegraphics[width=0.6\textwidth]{figures/workflow.png}
   \caption{Conversion Workflow}
@@
 
-The filters come into picture when there are certain elements which require 
-customization. For example, suppose you are converting markdown to HTML and you
-require the page to have automatic numbering for figures or tables. 
-If you think about it in markdown there is no system of numbering figures/tables 
-automatically, to retain 1:1 conversion the HTML output will miss the numbering as well. 
-At this moment you would wish that there was some option to add such a feature,
-but then given the infinite possibilities of customization it is not feasible to 
-add each and every customization as an option in pandoc or the writer. 
-To tackle this, the idea of filters arrived, wherein you could manipulate the AST
-partly or fully and modify it before the writer could read the AST. Hence generating
-an output you desire \citep{pandocfilters}.
+When specific elements require adjustment, the filters enter the picture. Assume you're converting markdown to HTML and want the page to have automated numbering for figures or tables. 
+If you use markdown, there is no system for automatically numbering figures/tables; to retain 1:1 conversion, the HTML output will also lack numbering.
 
-\section{Writing lua filters}
+At this point, you might wish there was a way to add such a feature.
+Then there is the possibility of unlimited choices for customization, including every customization as an option in pandoc or the writer is not viable.
+To address this, the idea of filters emerged, in which you could change the AST partially and modify it before the writer could read it. As a result, you get the desired result \citep{pandocfilters}.
+
+\section{Need of lua filters}
 
@@
 \centering
-\includegraphics[width=0.5\linewidth]{example.png}
+\includegraphics[width=0.5\linewidth]{figures/example.png}
 \caption{Vanilla HTML5 output}
@@
 \end{figure*}
-
-\section{Using pandoc lua filters}
 
@@
 Now we write a lua filter to manipulate the figures in \ref{code:3}
-\subsection{TODO : explain lua filter in more detail.}
+
+\section{Writing pandoc lua filters}
+
 \begin{figure}[htbp]
@@
 
+Everything in Lua revolves around tables; for example, the pandoc AST or the document is one giant table with sub-tables.
+
+In pandoc, there are two sorts of elements: \verb|Blocks| and \verb|Inlines|.
+\verb|Blocks| are complicated constructions constructed from simple pieces (\verb|Inlines|).
+A \verb|Para|, for example, is a \verb|Block| made up of several \verb|Str| \verb|Inlines|.
+
+The first step in writing a filter is to select a target type, that is, which block or Inline this filter will target.
+Following the selection, we name the function after the target type and each element in the argument as el.
+For instance, in the case of our filter, it would be \verb|function Figure(el)|. Pandoc is smart enough to match the names of filter functions to the AST elements.
+
+Now within the filter function we can assume that \verb|el| will contain the table object of a Figure element from the document.
+we can use many functions over it or \verb|Inlines| contained in it.
+One such function to walk over all the elements inside the block is \verb|walk_block|.
+walk functions are great to check or count the presence of certain elements in a block.
+we use \verb|walk_block| to check if there are any Image inline elements within the Figure block.
+This is because after pandoc 3 \citep{pandoc} Figure blocks can now contain elements other than Images as well.
+
+If there is an Image then we append \verb|"Figure X :"| to the caption of the Figure element.
+and return the modified element.
+
 \section{Interpreting changes using lua filters}
@@
 \centering
-\includegraphics[width=0.5\linewidth]{example-filtered.png}
+\includegraphics[width=0.5\linewidth]{figures/example-filtered.png}
 \caption{HTML5 output with desired filtering}
@@
 \newblock \emph{pandoc} \penalty0 2023
-\newblock URL \url{https://pandoc.org/releases.html#pandoc-3.0-2023-01-18}
+\newblock URL \url{https://pandoc.org/releases.html}
 

### 2023-08-01 supplement/metadata/metadata.tex
Updated supplement examples
overview of current examples
updated metadata supplement

@@
 Pandoc when used with \verb|-s| or \verb|standalone| mode reads out some of the existing
-meta data. The extracted author meta-data are listed below
+meta data. The extracted author meta-data are listed below:
 
@@
 Abhishek Ulayil\\ % Your Name in the first Line
-Student, Institute of Actuaries of India\\% Your affiliation which includes post and/or org
+Institute of Actuaries of India (Student)\\% Your affiliation which includes post and/or org
 Mumbai, India\\ % Your City/Country
@@
 
-\section{Miscallaneous}
+\section{Miscellaneous}
 
@@
 
-As we know about citations (using \verb|\citet{..}| command here) [\citet{moreverb}]. We could also use \verb|\citep{..}| command in real \citep{pandoc}.
+Table \ref{table:2} represents all the ways one can use citations in LaTeX and
+expect the generated output.
 
@@
 \caption{Different citations in LaTeX}
-\label{table:3}
+\label{table:2}
 \end{table}
@@
 Abhishek Ulayil\\
-Student, Institute of Actuaries of India\\%
+Institute of Actuaries of India (Student)\\%
 Mumbai, India\\

### 2023-08-03 texor.Rmd
more details in paper

@@
 
-The supplementary materials has example LaTeX documents that allow the reader to see how different common patterns in the legacy documents are handled with the conversion. These include:
+The supplementary materials has example folders containing LaTeX documents that allow the reader to see how different common patterns in the legacy documents are handled with the conversion.
+These include:
 
@@
 
+Figure definitions in LaTeX are many and varied! The standard, single figure definition with the `figure` environment is handled by pandoc, which converts:
+
+```
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.35\textwidth]{Rlogo-5.png}
+  \caption{The logo of R.}
+  \label{figure:rlogo}
+\end{figure}
+```
+
+to
+
+```
+<figure id="figure:rlogo">
+<img src="Rlogo-5.png" style="width:35.0%" />
+<figcaption>Figure 1: The logo of R.</figcaption>
+</figure>
+```
+
+PDF figures cause a problem for HTML display, so the image is first converted to PNG, in the pre-processing of the LaTeX document.
+Then the figure environment would be post-processed using pandoc, as described above.
+
+Multiple figures are supported with the latest versions of pandoc, so definitions like:
+
+```
+\begin{figure*}[htbp]
+  \centering
+  \includegraphics[width=0.45\textwidth]{Rlogo-5.png}\includegraphics[width=0.45\textwidth]{normal}
+  \caption{Images side by side}
+  \label{fig:twoimages}
+\end{figure*}
+```
+
+will be converted to:
+
+```
+<figure id="fig:twoimages">
+<p><img src="Rlogo-5.png" style="width:45.0%" alt="image" /><img
+src="normal.png" style="width:45.0%" alt="image" /></p>
+<figcaption>Figure 3: Images side by side</figcaption>
+</figure>
+```
+
+Some legacy articles defined images using `tikz` commands, such as:
+
+```
+\begin{figure}
+
+%% Generated Image will included as a PNG above automatically
+  \centering
+\tikzstyle{process} = [rectangle, rounded corners,
+minimum width=3cm, 
+minimum height=1cm,
+text centered, 
+draw=black]
+\tikzstyle{arrow} = [thick,->,>=stealth]
+\begin{tikzpicture}[node distance=4cm]
+%Nodes
+...
+
+```
+
+This is handled by pre-rpocessing the LaTeX to create the image, as both PDF, and them PNG, for inclusion in the R Markdown document using:
+
+```
+<figure id="fig:tikz">
+<img src="tikz/figtikz.png" style="width:100.0%" />
+
+<figcaption>Figure 5: Tikz Image example</figcaption>
+</figure>
+```
+
+Interestingly, some legacy articles have algorithms are defined in the figure environment:
+
+```
+\begin{algorithm}[htbp]
+\SetAlgoLined
+\KwData{this text}
+\KwResult{how to write algorithm with \LaTeX2e }
+initialization\;
+\While{not at end of this document}{
+read current\;
+\eIf{understand}{
+go to next section\;
+current section becomes this one\;
+}{
+go back to the beginning of current section\;
+}
+}
+\caption{How to write algorithms}
+  \label{alg:how}
+\end{algorithm}
+```
+
 ## Handling equations
@@
 
+Only handling to html format, especially relevant for figures, because the legacy pdf will remain the same.
+If using for converting work from LaTeX to Rmd then this will need manual changing by author.
+See rjtools guide.
+
 Where this might be useful in the future, other applications.
+
+No shift to quarto yet.
+
+Known problems that need manual fixes.
 

### 2023-08-03 texor.tex
more details in paper

@@
 
+Figure definitions in LaTeX are many and varied! The standard, single figure definition with the \texttt{figure} environment is handled by pandoc, which converts:
+
+\begin{verbatim}
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.35\textwidth]{Rlogo-5.png}
+  \caption{The logo of R.}
+  \label{figure:rlogo}
+\end{figure}
+\end{verbatim}
+
+to
+
+\begin{verbatim}
+<figure id="figure:rlogo">
+<img src="Rlogo-5.png" style="width:35.0%" />
+<figcaption>Figure 1: The logo of R.</figcaption>
+</figure>
+\end{verbatim}
+
+PDF figures cause a problem for HTML display, so the image is first converted to PNG, in the pre-processing of the LaTeX document.
+Then the figure environment would be post-processed using pandoc, as described above.
+
+Multiple figures are supported with the latest versions of pandoc, so definitions like:
+
+\begin{verbatim}
+\begin{figure*}[htbp]
+  \centering
+  \includegraphics[width=0.45\textwidth]{Rlogo-5.png}\includegraphics[width=0.45\textwidth]{normal}
+  \caption{Images side by side}
+  \label{fig:twoimages}
+\end{figure*}
+\end{verbatim}
+
+will be converted to:
+
+\begin{verbatim}
+<figure id="fig:twoimages">
+<p><img src="Rlogo-5.png" style="width:45.0%" alt="image" /><img
+src="normal.png" style="width:45.0%" alt="image" /></p>
+<figcaption>Figure 3: Images side by side</figcaption>
+</figure>
+\end{verbatim}
+
+Some legacy articles defined images using \texttt{tikz} commands, such as:
+
+\begin{verbatim}
+\begin{figure}
+
+%% Generated Image will included as a PNG above automatically
+  \centering
+\tikzstyle{process} = [rectangle, rounded corners,
+minimum width=3cm, 
+minimum height=1cm,
+text centered, 
+draw=black]
+\tikzstyle{arrow} = [thick,->,>=stealth]
+\begin{tikzpicture}[node distance=4cm]
+%Nodes
+...
+\end{verbatim}
+
+This is handled by pre-rpocessing the LaTeX to create the image, as both PDF, and them PNG, for inclusion in the R Markdown document using:
+
+\begin{verbatim}
+<figure id="fig:tikz">
+<img src="tikz/figtikz.png" style="width:100.0%" />
+
+<figcaption>Figure 5: Tikz Image example</figcaption>
+</figure>
+\end{verbatim}
+
+Interestingly, some legacy articles have algorithms are defined in the figure environment:
+
+\begin{verbatim}
+\begin{algorithm}[htbp]
+\SetAlgoLined
+\KwData{this text}
+\KwResult{how to write algorithm with \LaTeX2e }
+initialization\;
+\While{not at end of this document}{
+read current\;
+\eIf{understand}{
+go to next section\;
+current section becomes this one\;
+}{
+go back to the beginning of current section\;
+}
+}
+\caption{How to write algorithms}
+  \label{alg:how}
+\end{algorithm}
+\end{verbatim}
+
 \hypertarget{handling-equations}{%

### 2023-08-07 supplement/math-env/math-example.tex
updates to examples sections

@@
 \subsubsection{Pandoc Handling, extensions}
-As \CRANpkg{texor} calls \CRANpkg{rmarkdown} to render the Rmarkdwon file into HTML,
+As \CRANpkg{texor} calls \CRANpkg{rmarkdown} to render the Rmarkdown file into HTML,
 the \verb|rjtools::rjournal_web_article| template by default uses MathJax as the math engine. 

### 2023-08-07 texor.Rmd
updates to examples sections

@@
 
-The format is constantly evolving.
+The format is regularly updated.
 Early articles were typeset using LateX [@latex], from a specific, but changing, template.
@@
 And further post-processing needs to be done to handle specific R Journal R Markdown styling.
-The \CRANpkg{texor} package contains functionality to handle this pre- and postprocessing of the document.
+The \CRANpkg{texor} package contains functionality to handle this pre- and post-processing of the document, in a workflow illustrated in Figure \@ref(fig:workflow).
 
@@
 
-**DC: Add a diagram here showing flow of work, just pre-process .tex -> pandoc -> lua ; rebib operations???**
+```{r workflow, fig.cap="Workflow of the document conversion conducted by texor."}
+knitr::include_graphics("workflow.png")
+```
 
@@
 
-This is handled by pre-rpocessing the LaTeX to create the image, as both PDF, and them PNG, for inclusion in the R Markdown document using:
+This is handled by pre-processing the LaTeX to create the image, as both PDF, and them PNG, for inclusion in the R Markdown document using:
 
@@
 
-Interestingly, some legacy articles have algorithms are defined in the figure environment:
+Algorithms as figures is supported, and the following description will yield the result in Algorithm \@ref(alg:how).
 
@@
 
+<figure id="alg:how">
+<img src="alghow.png" style="width:100.0%" />
+<p>initialization</p>
+<figcaption>Algorithm 1: How to write algorithms</figcaption>
+</figure>
+
+**DC: Don't really understand this example, why would html figure definition also create pdf?**
+
 ## Handling equations
+
+Math is handled primarily by pandoc. The inline math and equation descriptions are unchanged between LaTeX and R Markdown.
+
+The HTML output renders math using MathJax. This does mean that some functionality, like `\bm`, `\boldmath` and `mathbb` not supported, and special definitions cannot be handled.
+
+The numbering of equations is a bit trickier.
+LaTeX automatically numbers equations, unless specifically instructed not to.
+Equation numbering in R Markdown requires specific labelling using `(\#eq:xx)` as described in @bookdown.
+The \CRANpkg{texor} helps by adding the labelling using a Lua filter to convert the existing `\label{..}` to `(\#eq:xx)`.
 
@@
 
-Only handling to html format, especially relevant for figures, because the legacy pdf will remain the same.
+Remember only the HTML is rendered, and the original PDF is not re-generated.
+
+Only handling to HTML format, especially relevant for figures, because the legacy pdf will remain the same.
 If using for converting work from LaTeX to Rmd then this will need manual changing by author.

### 2023-08-07 texor.tex
updates to examples sections

@@
 
-The format is constantly evolving.
+The format is regularly updated.
 Early articles were typeset using LateX (The LaTeX Project 2023), from a specific, but changing, template.
@@
 And further post-processing needs to be done to handle specific R Journal R Markdown styling.
-The \CRANpkg{texor} package contains functionality to handle this pre- and postprocessing of the document.
+The \CRANpkg{texor} package contains functionality to handle this pre- and post-processing of the document, in a workflow illustrated in Figure \ref{fig:workflow}.
 
-\textbf{DC: Add a diagram here showing flow of work, just pre-process .tex -\textgreater{} pandoc -\textgreater{} lua ; rebib operations???}
+\begin{figure}
+\includegraphics[width=15.5in]{workflow} \caption{Workflow of the document conversion conducted by texor.}\label{fig:workflow}
+\end{figure}
 
@@
 
-Interestingly, some legacy articles have algorithms are defined in the figure environment:
+Algorithms as figures is supported, and the following description will yield the result in Algorithm \ref{alg:how}.
 
@@
 
+initialization
+
+Algorithm 1: How to write algorithms
+
+\textbf{DC: Don't really understand this example, why would html figure definition also create pdf?}
+
 \hypertarget{handling-equations}{%
 \subsection{Handling equations}\label{handling-equations}}
+
+Math is handled primarily by pandoc. The inline math and equation descriptions are unchanged between LaTeX and R Markdown.
+
+The HTML output renders math using MathJax.
+This does mean that some functionality, like \texttt{\textbackslash{}bm}, \texttt{\textbackslash{}boldmath} and \texttt{mathbb} not supported, and special definitions cannot be handled.
+
+The numbering of equations is a bit trickier.
+LaTeX automatically numbers equations, unless specifically instructed not to.
+Equation numbering in R Markdown requires specific labelling using \texttt{(\textbackslash{}\#eq:xx)} as described in Yihui Xie (2023).
+The \CRANpkg{texor} helps by adding the labelling using a Lua filter to convert the existing \texttt{\textbackslash{}label\{..\}} to \texttt{(\textbackslash{}\#eq:xx)}.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-bookdown}{}}%
+Yihui Xie. 2023. \emph{{bookdown: Authoring Books and Technical Documents with R Markdown}}. \#equations. \url{https://bookdown.org/yihui/bookdown/markdown-extensions-by-bookdown.html}.
+
 \end{CSLReferences}

### 2023-08-07 texor.R

@@ -6,3 +6,7 @@ knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
 library(plotly)
 library(ggplot2)
 
+
+## ----workflow, fig.cap="Workflow of the document conversion conducted by texor."----
+knitr::include_graphics("workflow.png")
+

### 2023-08-09 texor.Rmd
almost first draft

@@
 abstract: >
-  An abstract of less than 150 words.
+  In 2021 the R Journal made a change of templates for article writing to R Markdown instead of LaTeX.
+The reasons were to encourage better reproducibility of articles using dynamic documents, enable interactivity in articles, and to make the articles more accessible for vision-impaired readers.
+A resulting challenge was to explore whether legacy articles might be suitably converted into R Markdown before converting into HTML output.
+This paper describes the process to convert an R Journal LaTeX template article into an R Markdown article, and the two new R packages, \CRANpkg{texor} and \CRANpkg{rebib}, that can be used to achieve the conversion.
 draft: true
@@
 The paper is organised as follows.
-Section XXX gives an overview of the conversion process.
-Section XXX describes examples of pre-processing using regular expressions, and section XXX provides examples of post-processing using lua filters.
-Section XXX describes tools to do special handling of bibliography files.
-Section XXX runs through an example conversion process.
+Section 2 gives an overview of the conversion process, that includes pre-processing using regular expressions, post-processing using lua filters, and handling of figures, tables and equations.
+Section 3 describes the \CRANpkg{texor} that handles most of the conversion.
+Section 4 describes tools to do special handling of bibliography files.
+The supplementary materials have folders containing specific examples that can be used for understanding how the conversions are done.
 
@@
 
-# The internals of converting from LaTeX to R Markdown
+# The internals of converting from LaTeX to R Markdown {#internals}
 
@@
 
-```{r workflow, fig.cap="Workflow of the document conversion conducted by texor."}
+```{r workflow, fig.cap="Workflow of the document conversion conducted by texor.", out.width="70%", fig.align='center'}
 knitr::include_graphics("workflow.png")
 ```
-
-The supplementary materials has example folders containing LaTeX documents that allow the reader to see how different common patterns in the legacy documents are handled with the conversion.
-These include:
-
-- `code-env`: Explains how different code environments defined by the R Journal style are handled, and additional details such as code in figure environments, and code in table environments. 
-- `math-env`: Examples of inline math, display math, how equation numbering is handled by a Lua filter to convert from LaTeX labelling to R Markdown labelling. 
-- `figure-env`: Explains how the variety of figure definitions are handled in the conversion, including different image formats, numbering, captions, labelling, multiple images, and `tikz` [@casflow] images.
-- `table-env`: Examples of how a variety of table types are converted, including multicolumn, complex and wide tables.
-- `lua-filters`: Overview and lots of small examples of Lua filters to handle the custom output needed for the R Markdown format.
-- `metadata`: This has a collection of additional format handling including extracting metadata like author names and affiliations, article identifiers used in the review process, and handling citations, footnotes and links.
-- `bibliography`: The bibliography was handled differently over the years of the journal, and this details how to use the `rebib` functionality to handle `bbl` files, embedded `bbl`, to convert into the standard `.bib` format.
-
-The explanations in the next sections are extracted from these examples.
-In each of these folders there is a `RJwrapper.tex`, and `.tex` file, with the extra template files `RJournal.sty` and `Rlogo-5.png`.
-These match the legacy template file structure, from which the `RJwrapper.pdf` file is created.
-If you follow the instructions in the `README.md` you can create the R Markdown and HTML versions of the document.
-
-<!-- **Maybe more details on the supporting packages here: LaTeX, Regex, pandoc building blocks of the packages, methods used in them [MOW]**-->
 
@@
 
-While pandoc can do most of the heavy-lifting, it cannot cope with all the freedom with which LaTeX documents are written.
+While Pandoc can do most of the heavy-lifting, it cannot cope with all the freedom with which LaTeX documents are written.
 An example of this is with formatting of code.
 Pandoc only handles the `verbatim` environment, but there are many ways to format code in LaTeX, and the R Journal template has a special `\code{}` command.
-If the code environment is not verbatim, then pandoc will also try to process the actual code content as LaTeX commands and will likely lose details.
-It is better to convert these synonyms into `verbatim` environments this prior to passing the document to pandoc.
+If the code environment is not verbatim, then Pandoc will also try to process the actual code content as LaTeX commands and will likely lose details.
+It is better to convert these synonyms into `verbatim` environments this prior to passing the document to Pandoc.
 
@@
 <!-- by syntax though you can always embed HTML to expand the feature set.
-Although pandoc does most of the heavy lifting, there are some instances where even  filters cannot assist.
+Although Pandoc does most of the heavy lifting, there are some instances where even  filters cannot assist.
 A good demonstration of the issue in hand is code environments.-->
 
-<!-- A limitation of pandoc is that it only works with a "verbatim" environment.
-However, in "RJournal",  we have multiple custom code environments which are not supported or defined due to lower-level tex limitations in pandoc.
+<!-- A limitation of Pandoc is that it only works with a "verbatim" environment.
+However, in "RJournal",  we have multiple custom code environments which are not supported or defined due to lower-level tex limitations in Pandoc.
 So the code  environments are eliminated during the conversion process.
 Including a Lua filter to find these elements and redefine them as verbatim environments would also be  ineffective.
-If the code environment is not verbatim, then pandoc will also try to process the actual code content as LaTeX commands and will lose out on details in some cases.-->
+If the code environment is not verbatim, then Pandoc will also try to process the actual code content as LaTeX commands and will lose out on details in some cases.-->
 
@@
 
-## Handling figures
+## Figures
 
-Figure definitions in LaTeX are many and varied! The standard, single figure definition with the `figure` environment is handled by pandoc, which converts:
+### Standard, single figure
+
+Figure definitions in LaTeX are many and varied! The standard, single figure definition with the `figure` environment and raster image format such as PNG or JPG, is handled by Pandoc. It will convert:
 
@@
 
-PDF figures cause a problem for HTML display, so the image is first converted to PNG, in the pre-processing of the LaTeX document.
-Then the figure environment would be post-processed using pandoc, as described above.
+### PDF format images
 
-Multiple figures are supported with the latest versions of pandoc, so definitions like:
+Images in PDF format are converted to PNG, in the pre-processing of the LaTeX document, and then post-processed using Pandoc as described above.
+
+### Multiple figures
+
+Multiple figures are supported with the latest versions of Pandoc, so definitions like:
 
@@
   \centering
-  \includegraphics[width=0.45\textwidth]{Rlogo-5.png}\includegraphics[width=0.45\textwidth]{normal}
+  \includegraphics[width=0.45\textwidth]{Rlogo-5.png}
+  \includegraphics[width=0.45\textwidth]{normal}
   \caption{Images side by side}
@@
 
-Some legacy articles defined images using `tikz` commands, such as:
+### `tikz` format images
+
+Some legacy articles define images using `tikz` commands, such as:
 
@@
 
-This is handled by pre-processing the LaTeX to create the image, as both PDF, and them PNG, for inclusion in the R Markdown document using:
+This is handled by pre-processing the LaTeX to create the image, as both PDF, and then PNG, for inclusion in the R Markdown document using:
 
@@
 
-Algorithms as figures is supported, and the following description will yield the result in Algorithm \@ref(alg:how).
+### Algorithms as figures
+
+Algorithms as figures is supported, and the following description will yield the result in Figure \@ref(fig:alghow).
 
@@
 
+```
 <figure id="alg:how">
@@
 </figure>
+```
 
-**DC: Don't really understand this example, why would html figure definition also create pdf?**
+```{r alghow, fig.cap="How to write algorithms.", out.width="60%", fig.align='center'}
+knitr::include_graphics("alghow.png")
+```
 
-## Handling equations
+## Equations
 
-Math is handled primarily by pandoc. The inline math and equation descriptions are unchanged between LaTeX and R Markdown.
+Math is handled primarily by Pandoc. The inline math and equation descriptions are unchanged between LaTeX and R Markdown.
 
@@
 
-## Handling the bibliography
+## Tables
+
+Tables form one of the biggest challenges in migrating from LaTeX to R Markdown, because the sophistication is not completely replicated.
+However, there have been many improvements in table definitions for R Markdown that are increasing producing the beautifully crafted tables possible in LaTeX.
+The conversion in \CRANpkg{texor} can mostly handle the simple tables, and for producing more complex tables it may be necessary to manually edit the resulting `Rmd` file to make conditional tables, one to render specifically for HTML output using packages such as \CRANpkg{kableExtra} [@kableExtra], \CRANpkg{gt} [@gt], \CRANpkg{htmlTable} [@htmlTable], \CRANpkg{tableHTML} [@tableHTML], \CRANpkg{tables} [@tables] or \CRANpkg{DT} [@DT].
+A benefit of special HTML table output is interactivity, that would allow the table to be sorted in different ways.
+
+### Generic tables
+
+Simple LaTeX tables are converted into traditional markdown format tables, by Pandoc. So this table definition:
+
+```
+\begin{table}[htbp]
+\centering
+\begin{tabular}{l | llll }
+ \hline
+ Graphics Format & LaTeX & Markdown & Rmarkdown & HTML \\
+ \hline
+ PNG       & Yes & Yes & Yes & Yes \\
+ JPG       & Yes & Yes & Yes & Yes \\
+ PDF       & Yes & No & No & No \\
+ SVG       & No & Yes & Yes & Yes \\
+ Tikz      & Yes & No & Yes & No \\
+ Algorithm & Yes & No & No & No \\
+\hline
+\end{tabular}
+\caption{Image Format support in various Markup/Typesetting Languages}
+\label{table:1}
+\end{table}
+```
+
+will be converted to:
+
+```
+::: {#table:1}
+  -------------------------------------------------------
+  Graphics Format   LaTeX   Markdown   Rmarkdown   HTML
+  ----------------- ------- ---------- ----------- ------
+  PNG               Yes     Yes        Yes         Yes
+
+  JPG               Yes     Yes        Yes         Yes
+
+  PDF               Yes     No         No          No
+
+  SVG               No      Yes        Yes         Yes
+
+  Tikz              Yes     No         Yes         No
+
+  Algorithm         Yes     No         No          No
+  -------------------------------------------------------
+
+  : Table 1: Image Format support in various Markup/Typesetting
+  Languages
+:::
+```
+
+### Multicolumn tables
+
+A multicolumn table requires:
+
+1. The stream editor modifies the `\multicolumnx{..}` to
+    `\multicolumnxx{..}`.
+
+2. A LaTeX macro is used to redefine the `\multicolumnxx{..}` to
+    `\multicolumnx{---}`.
+
+3. Pandoc reads the table and transforms it to markdown.
+
+```{asis echo=knitr::is_latex_output()}
+\begin{table*}[htbp]
+\begin{center}
+\begin{tabular}{l | llll }
+ \hline
+ \multicolumn{1}{c |}{EXAMPLE} & \multicolumn{2}{c}{$X$} &
+\multicolumn{2}{c}{$Y$} \\
+ \hline
+  & 1 & 2 & 1 & 2 \\
+ EX1  & X11 & X12 &  Y11  & Y12 \\
+ EX2  & X21 & X22 &  Y21  & Y22 \\
+ EX3  & X31 & X32 &  Y31  & Y32 \\
+ EX4  & X41 & X42 &  Y41  & Y42\\
+ EX5  & X51 & X52 &  Y51  & Y52 \\
+\hline
+\end{tabular}
+\caption{An example multicolumn table.}
+\label{table:2}
+\end{center}
+\end{table*}
+```
+
+```{asis echo=knitr::is_html_output()}
+::: center
+::: {#table:2}
+  -------------------------------------
+  EXAMPLE   \(X\)         \(Y\)   
+  --------- ------- ----- ------- -----
+            1       2     1       2
+
+  EX1       X11     X12   Y11     Y12
+
+  EX2       X21     X22   Y21     Y22
+
+  EX3       X31     X32   Y31     Y32
+
+  EX4       X41     X42   Y41     Y42
+
+  EX5       X51     X52   Y51     Y52
+  -------------------------------------
+
+  : Table 2: An Example Table with Multicolumn
+:::
+:::
+```
+
+Also note that the stream editor is used to rename `table*` environment to `table` environment because the markdown/HTML will not support the specific changes table\* environment will bring.
+
+### Other tables
+
+Tables with images, math, code or links in the cells are generally handled. Also `widetable` tables that allow for specific width or wrapping of tables into blocks are also partially handled.
+
+## Animation
+
+XXX Shall we describe the one pdf paper that has animation and how this is handled - could be cool to include
 
@@
 
-## One-step user-facing conversion
+XXX what should go here?
 
-The only function that a user necessarily needs is `latex_to_web()`. This creates the R Journal style R Markdown file from a given R Journal style LaTeX file. 
-This is achieved in several conversion steps, `convert_to_markdown()`, `generate_rmd()` and `produce_html()`.
-
-# Using \CRANpkg{texor}
+# Using \CRANpkg{texor} {#texor}
 
@@
 
-and the development version from https://abhi-1u.github.io/texor.
+and the development version from https://github.com/Abhi-1U/texor. The website for the package, https://abhi-1u.github.io/texor, has vignettes documenting usage.
 
-**Examples / getting started / usage and so on**
+Note that you will need to use [Pandoc Version > 3.0.0 (if possible latest)]((https://pandoc.org/installing.html)) for the best results. You can check your version with:
 
-Run through the different examples in supplementary here
+```
+rmarkdown::pandoc_version()
+```
 
-**From general LaTeX to generic R Markdown**
+The only function that a user necessarily needs is `latex_to_web()`. This creates the R Journal style R Markdown file from a given R Journal style LaTeX file. 
+This is achieved by several sequential steps, `convert_to_markdown()`, `generate_rmd()` and `produce_html()`.
 
-# Managing the bibliography using \CRANpkg{rebib}
+The supplementary materials have several folders with specific examples of figures, tables, equations that can be used to understand how the conversion operates.
+
+For converting the 14 years of legacy R Journal articles, batch processing of issues was conducted. This can be achieved using:
+
+XXX
+
+For individuals who are interested in submitting their paper to the R Journal but have written their article using the legacy LaTeX format and wish to convert it to the new format, you can use the `latex_to_web()` function on your paper directory.
+
+# Managing the bibliography using \CRANpkg{rebib} {#rebib}
 
@@
 
-# Acknowledgments
-pandoc
- "TODO : Later"
+# Acknowledgments {-}
 
-# Supplementary materials
+The authors wish to thank the [Google Summer of Code](https://summerofcode.withgoogle.com) program for financially supporting Abhishek's work on this project, and the R Project Organization for their support.
 
-Several supplementary example documents are provided which illustrate how different elements of LaTeX-authored papers are handled. These are:
+# Supplementary materials {-}
 
--
+The supplementary materials has example folders containing LaTeX documents that allow the reader to see how different common patterns in the legacy documents are handled with the conversion.
+These include:
 
-The GitHub repos supporting this work are:
+- `code-env`: Explains how different code environments defined by the R Journal style are handled, and additional details such as code in figure environments, and code in table environments. 
+- `math-env`: Examples of inline math, display math, how equation numbering is handled by a Lua filter to convert from LaTeX labelling to R Markdown labelling. 
+- `figure-env`: Explains how the variety of figure definitions are handled in the conversion, including different image formats, numbering, captions, labelling, multiple images, and `tikz` [@casflow] images.
+- `table-env`: Examples of how a variety of table types are converted, including multicolumn, complex and wide tables.
+- `lua-filters`: Overview and lots of small examples of Lua filters to handle the custom output needed for the R Markdown format.
+- `metadata`: This has a collection of additional format handling including extracting metadata like author names and affiliations, article identifiers used in the review process, and handling citations, footnotes and links.
+- `bibliography`: The bibliography was handled differently over the years of the journal, and this details how to use the `rebib` functionality to handle `bbl` files, embedded `bbl`, to convert into the standard `.bib` format.
+
+In each of these folders there is a `RJwrapper.tex`, and `.tex` file, with the extra template files `RJournal.sty` and `Rlogo-5.png` and `.bib` files.
+These match the legacy template file structure, from which the `RJwrapper.pdf` file is created.
+To test the conversion for each of these examples, set the path directory to one of the folders and use the `latex_to_web()` function as follows:
+
+```
+article_dir <- "path-to-this supplementary folder"
+texor::latex_to_web(article_dir)
+```
+
+This will create an `.Rmd` and `.html` files in the same directory, that demonstrate the converted R Markdown version and the HTML output format.
+
+You'll need to ensure that you have the latest versions of \CRANpkg{texor} and \CRANpkg{rebib}, and Pandoc (at least later than version 3.0.0).
+
+## Source materials {-}
+
+The \CRANpkg{texor} and \CRANpkg{rebib} source codem and materials to reproduce this paper are available at:
 

### 2023-08-09 texor.tex
almost first draft

@@
 \abstract{%
-An abstract of less than 150 words.
+In 2021 the R Journal made a change of templates for article writing to R Markdown instead of LaTeX.
+The reasons were to encourage better reproducibility of articles using dynamic documents, enable interactivity in articles, and to make the articles more accessible for vision-impaired readers.
+A resulting challenge was to explore whether legacy articles might be suitably converted into R Markdown before converting into HTML output.
+This paper describes the process to convert an R Journal LaTeX template article into an R Markdown article, and the two new R packages, \CRANpkg{texor} and \CRANpkg{rebib}, that can be used to achieve the conversion.
 }
@@
 The paper is organised as follows.
-Section XXX gives an overview of the conversion process.
-Section XXX describes examples of pre-processing using regular expressions, and section XXX provides examples of post-processing using lua filters.
-Section XXX describes tools to do special handling of bibliography files.
-Section XXX runs through an example conversion process.
+Section 2 gives an overview of the conversion process, that includes pre-processing using regular expressions, post-processing using lua filters, and handling of figures, tables and equations.
+Section 3 describes the \CRANpkg{texor} that handles most of the conversion.
+Section 4 describes tools to do special handling of bibliography files.
+The supplementary materials have folders containing specific examples that can be used for understanding how the conversions are done.
 
-\hypertarget{the-internals-of-converting-from-latex-to-r-markdown}{%
-\section{The internals of converting from LaTeX to R Markdown}\label{the-internals-of-converting-from-latex-to-r-markdown}}
+\hypertarget{internals}{%
+\section{The internals of converting from LaTeX to R Markdown}\label{internals}}
 
@@
 \begin{figure}
-\includegraphics[width=15.5in]{workflow} \caption{Workflow of the document conversion conducted by texor.}\label{fig:workflow}
+
+{\centering \includegraphics[width=0.7\linewidth]{workflow}
+
+}
+
+\caption{Workflow of the document conversion conducted by texor.}\label{fig:workflow}
 \end{figure}
-
-The supplementary materials has example folders containing LaTeX documents that allow the reader to see how different common patterns in the legacy documents are handled with the conversion.
-These include:
-
-\begin{itemize}
-\tightlist
-\item
-  \texttt{code-env}: Explains how different code environments defined by the R Journal style are handled, and additional details such as code in figure environments, and code in table environments.
-\item
-  \texttt{math-env}: Examples of inline math, display math, how equation numbering is handled by a Lua filter to convert from LaTeX labelling to R Markdown labelling.
-\item
-  \texttt{figure-env}: Explains how the variety of figure definitions are handled in the conversion, including different image formats, numbering, captions, labelling, multiple images, and \texttt{tikz} (Josh Cassidy 2013) images.
-\item
-  \texttt{table-env}: Examples of how a variety of table types are converted, including multicolumn, complex and wide tables.
-\item
-  \texttt{lua-filters}: Overview and lots of small examples of Lua filters to handle the custom output needed for the R Markdown format.
-\item
-  \texttt{metadata}: This has a collection of additional format handling including extracting metadata like author names and affiliations, article identifiers used in the review process, and handling citations, footnotes and links.
-\item
-  \texttt{bibliography}: The bibliography was handled differently over the years of the journal, and this details how to use the \texttt{rebib} functionality to handle \texttt{bbl} files, embedded \texttt{bbl}, to convert into the standard \texttt{.bib} format.
-\end{itemize}
-
-The explanations in the next sections are extracted from these examples.
-In each of these folders there is a \texttt{RJwrapper.tex}, and \texttt{.tex} file, with the extra template files \texttt{RJournal.sty} and \texttt{Rlogo-5.png}.
-These match the legacy template file structure, from which the \texttt{RJwrapper.pdf} file is created.
-If you follow the instructions in the \texttt{README.md} you can create the R Markdown and HTML versions of the document.
 
@@
 
-While pandoc can do most of the heavy-lifting, it cannot cope with all the freedom with which LaTeX documents are written.
+While Pandoc can do most of the heavy-lifting, it cannot cope with all the freedom with which LaTeX documents are written.
 An example of this is with formatting of code.
 Pandoc only handles the \texttt{verbatim} environment, but there are many ways to format code in LaTeX, and the R Journal template has a special \texttt{\textbackslash{}code\{\}} command.
-If the code environment is not verbatim, then pandoc will also try to process the actual code content as LaTeX commands and will likely lose details.
-It is better to convert these synonyms into \texttt{verbatim} environments this prior to passing the document to pandoc.
+If the code environment is not verbatim, then Pandoc will also try to process the actual code content as LaTeX commands and will likely lose details.
+It is better to convert these synonyms into \texttt{verbatim} environments this prior to passing the document to Pandoc.
 
@@
 
-\hypertarget{handling-figures}{%
-\subsection{Handling figures}\label{handling-figures}}
+\hypertarget{figures}{%
+\subsection{Figures}\label{figures}}
 
-Figure definitions in LaTeX are many and varied! The standard, single figure definition with the \texttt{figure} environment is handled by pandoc, which converts:
+\hypertarget{standard-single-figure}{%
+\subsubsection{Standard, single figure}\label{standard-single-figure}}
+
+Figure definitions in LaTeX are many and varied!
+The standard, single figure definition with the \texttt{figure} environment and raster image format such as PNG or JPG, is handled by Pandoc.
+It will convert:
 
@@
 
-PDF figures cause a problem for HTML display, so the image is first converted to PNG, in the pre-processing of the LaTeX document.
-Then the figure environment would be post-processed using pandoc, as described above.
+\hypertarget{pdf-format-images}{%
+\subsubsection{PDF format images}\label{pdf-format-images}}
 
-Multiple figures are supported with the latest versions of pandoc, so definitions like:
+Images in PDF format are converted to PNG, in the pre-processing of the LaTeX document, and then post-processed using Pandoc as described above.
+
+\hypertarget{multiple-figures}{%
+\subsubsection{Multiple figures}\label{multiple-figures}}
+
+Multiple figures are supported with the latest versions of Pandoc, so definitions like:
 
@@
 
-Some legacy articles defined images using \texttt{tikz} commands, such as:
+\hypertarget{tikz-format-images}{%
+\subsubsection{\texorpdfstring{\texttt{tikz} format images}{tikz format images}}\label{tikz-format-images}}
+
+Some legacy articles define images using \texttt{tikz} commands, such as:
 
@@
 
-Algorithms as figures is supported, and the following description will yield the result in Algorithm \ref{alg:how}.
+\hypertarget{algorithms-as-figures}{%
+\subsubsection{Algorithms as figures}\label{algorithms-as-figures}}
+
+Algorithms as figures is supported, and the following description will yield the result in Figure \ref{fig:alghow}.
 
@@
 
-initialization
+\begin{verbatim}
+<figure id="alg:how">
+<img src="alghow.png" style="width:100.0%" />
+<p>initialization</p>
+<figcaption>Algorithm 1: How to write algorithms</figcaption>
+</figure>
+\end{verbatim}
 
-Algorithm 1: How to write algorithms
+\begin{figure}
 
-\textbf{DC: Don't really understand this example, why would html figure definition also create pdf?}
+{\centering \includegraphics[width=0.6\linewidth]{alghow}
 
-\hypertarget{handling-equations}{%
-\subsection{Handling equations}\label{handling-equations}}
+}
 
-Math is handled primarily by pandoc. The inline math and equation descriptions are unchanged between LaTeX and R Markdown.
+\caption{How to write algorithms.}\label{fig:alghow}
+\end{figure}
+
+\hypertarget{equations}{%
+\subsection{Equations}\label{equations}}
+
+Math is handled primarily by Pandoc. The inline math and equation descriptions are unchanged between LaTeX and R Markdown.
 
@@
 
-\hypertarget{handling-the-bibliography}{%
-\subsection{Handling the bibliography}\label{handling-the-bibliography}}
+\hypertarget{tables}{%
+\subsection{Tables}\label{tables}}
+
+Tables form one of the biggest challenges in migrating from LaTeX to R Markdown, because the sophistication is not completely replicated.
+However, there have been many improvements in table definitions for R Markdown that are increasing producing the beautifully crafted tables possible in LaTeX.
+The conversion in \CRANpkg{texor} can mostly handle the simple tables, and for producing more complex tables it may be necessary to manually edit the resulting \texttt{Rmd} file to make conditional tables, one to render specifically for HTML output using packages such as \CRANpkg{kableExtra} (Zhu 2021), \CRANpkg{gt} (Iannone et al.
+2023), \CRANpkg{htmlTable} (Gordon, Gragg, and Konings 2022), \CRANpkg{tableHTML} (Boutaris, Zauchner, and Jomar 2023), \CRANpkg{tables} (Murdoch 2023) or \CRANpkg{DT} (Xie, Cheng, and Tan 2023).
+A benefit of special HTML table output is interactivity, that would allow the table to be sorted in different ways.
+
+\hypertarget{generic-tables}{%
+\subsubsection{Generic tables}\label{generic-tables}}
+
+Simple LaTeX tables are converted into traditional markdown format tables, by Pandoc. So this table definition:
+
+\begin{verbatim}
+\begin{table}[htbp]
+\centering
+\begin{tabular}{l | llll }
+ \hline
+ Graphics Format & LaTeX & Markdown & Rmarkdown & HTML \\
+ \hline
+ PNG       & Yes & Yes & Yes & Yes \\
+ JPG       & Yes & Yes & Yes & Yes \\
+ PDF       & Yes & No & No & No \\
+ SVG       & No & Yes & Yes & Yes \\
+ Tikz      & Yes & No & Yes & No \\
+ Algorithm & Yes & No & No & No \\
+\hline
+\end{tabular}
+\caption{Image Format support in various Markup/Typesetting Languages}
+\label{table:1}
+\end{table}
+\end{verbatim}
+
+will be converted to:
+
+\begin{verbatim}
+::: {#table:1}
+  -------------------------------------------------------
+  Graphics Format   LaTeX   Markdown   Rmarkdown   HTML
+  ----------------- ------- ---------- ----------- ------
+  PNG               Yes     Yes        Yes         Yes
+
+  JPG               Yes     Yes        Yes         Yes
+
+  PDF               Yes     No         No          No
+
+  SVG               No      Yes        Yes         Yes
+
+  Tikz              Yes     No         Yes         No
+
+  Algorithm         Yes     No         No          No
+  -------------------------------------------------------
+
+  : Table 1: Image Format support in various Markup/Typesetting
+  Languages
+:::
+\end{verbatim}
+
+\hypertarget{multicolumn-tables}{%
+\subsubsection{Multicolumn tables}\label{multicolumn-tables}}
+
+A multicolumn table requires:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  The stream editor modifies the \texttt{\textbackslash{}multicolumnx\{..\}} to
+  \texttt{\textbackslash{}multicolumnxx\{..\}}.
+\item
+  A LaTeX macro is used to redefine the \texttt{\textbackslash{}multicolumnxx\{..\}} to
+  \texttt{\textbackslash{}multicolumnx\{-\/-\/-\}}.
+\item
+  Pandoc reads the table and transforms it to markdown.
+\end{enumerate}
+
+\begin{table*}[htbp]
+\begin{center}
+\begin{tabular}{l | llll }
+ \hline
+ \multicolumn{1}{c |}{EXAMPLE} & \multicolumn{2}{c}{$X$} &
+\multicolumn{2}{c}{$Y$} \\
+ \hline
+  & 1 & 2 & 1 & 2 \\
+ EX1  & X11 & X12 &  Y11  & Y12 \\
+ EX2  & X21 & X22 &  Y21  & Y22 \\
+ EX3  & X31 & X32 &  Y31  & Y32 \\
+ EX4  & X41 & X42 &  Y41  & Y42\\
+ EX5  & X51 & X52 &  Y51  & Y52 \\
+\hline
+\end{tabular}
+\caption{An example multicolumn table.}
+\label{table:2}
+\end{center}
+\end{table*}
+
+Also note that the stream editor is used to rename \texttt{table*} environment to \texttt{table} environment because the markdown/HTML will not support the specific changes table* environment will bring.
+
+\hypertarget{other-tables}{%
+\subsubsection{Other tables}\label{other-tables}}
+
+Tables with images, math, code or links in the cells are generally handled. Also \texttt{widetable} tables that allow for specific width or wrapping of tables into blocks are also partially handled.
+
+\hypertarget{animation}{%
+\subsection{Animation}\label{animation}}
+
+XXX Shall we describe the one pdf paper that has animation and how this is handled - could be cool to include
 
@@
 
-\hypertarget{one-step-user-facing-conversion}{%
-\subsection{One-step user-facing conversion}\label{one-step-user-facing-conversion}}
+XXX what should go here?
 
-The only function that a user necessarily needs is \texttt{latex\_to\_web()}. This creates the R Journal style R Markdown file from a given R Journal style LaTeX file.
-This is achieved in several conversion steps, \texttt{convert\_to\_markdown()}, \texttt{generate\_rmd()} and \texttt{produce\_html()}.
-
-\hypertarget{using}{%
-\section{\texorpdfstring{Using \CRANpkg{texor}}{Using }}\label{using}}
+\hypertarget{texor}{%
+\section{\texorpdfstring{Using \CRANpkg{texor}}{Using }}\label{texor}}
 
@@
 
-and the development version from \url{https://abhi-1u.github.io/texor}.
+and the development version from \url{https://github.com/Abhi-1U/texor}. The website for the package, \url{https://abhi-1u.github.io/texor}, has vignettes documenting usage.
 
-\textbf{Examples / getting started / usage and so on}
+Note that you will need to use \href{(https://pandoc.org/installing.html)}{Pandoc Version \textgreater{} 3.0.0 (if possible latest)} for the best results. You can check your version with:
 
-Run through the different examples in supplementary here
+\begin{verbatim}
+rmarkdown::pandoc_version()
+\end{verbatim}
 
-\textbf{From general LaTeX to generic R Markdown}
+The only function that a user necessarily needs is \texttt{latex\_to\_web()}. This creates the R Journal style R Markdown file from a given R Journal style LaTeX file.
+This is achieved by several sequential steps, \texttt{convert\_to\_markdown()}, \texttt{generate\_rmd()} and \texttt{produce\_html()}.
 
-\hypertarget{managing-the-bibliography-using}{%
-\section{\texorpdfstring{Managing the bibliography using \CRANpkg{rebib}}{Managing the bibliography using }}\label{managing-the-bibliography-using}}
+The supplementary materials have several folders with specific examples of figures, tables, equations that can be used to understand how the conversion operates.
+
+For converting the 14 years of legacy R Journal articles, batch processing of issues was conducted. This can be achieved using:
+
+XXX
+
+For individuals who are interested in submitting their paper to the R Journal but have written their article using the legacy LaTeX format and wish to convert it to the new format, you can use the \texttt{latex\_to\_web()} function on your paper directory.
+
+\hypertarget{rebib}{%
+\section{\texorpdfstring{Managing the bibliography using \CRANpkg{rebib}}{Managing the bibliography using }}\label{rebib}}
 
@@
 
-\hypertarget{using-1}{%
-\subsection{\texorpdfstring{Using \CRANpkg{rebib}}{Using }}\label{using-1}}
+\hypertarget{using}{%
+\subsection{\texorpdfstring{Using \CRANpkg{rebib}}{Using }}\label{using}}
 
@@
 \hypertarget{acknowledgments}{%
-\section{Acknowledgments}\label{acknowledgments}}
+\section*{Acknowledgments}\label{acknowledgments}}
+\addcontentsline{toc}{section}{Acknowledgments}
 
-pandoc
-``TODO : Later''
+The authors wish to thank the \href{https://summerofcode.withgoogle.com}{Google Summer of Code} program for financially supporting Abhishek's work on this project, and the R Project Organization for their support.
 
 \hypertarget{supplementary-materials}{%
-\section{Supplementary materials}\label{supplementary-materials}}
+\section*{Supplementary materials}\label{supplementary-materials}}
+\addcontentsline{toc}{section}{Supplementary materials}
 
-Several supplementary example documents are provided which illustrate how different elements of LaTeX-authored papers are handled. These are:
+The supplementary materials has example folders containing LaTeX documents that allow the reader to see how different common patterns in the legacy documents are handled with the conversion.
+These include:
 
@@
 \item
+  \texttt{code-env}: Explains how different code environments defined by the R Journal style are handled, and additional details such as code in figure environments, and code in table environments.
+\item
+  \texttt{math-env}: Examples of inline math, display math, how equation numbering is handled by a Lua filter to convert from LaTeX labelling to R Markdown labelling.
+\item
+  \texttt{figure-env}: Explains how the variety of figure definitions are handled in the conversion, including different image formats, numbering, captions, labelling, multiple images, and \texttt{tikz} (Josh Cassidy 2013) images.
+\item
+  \texttt{table-env}: Examples of how a variety of table types are converted, including multicolumn, complex and wide tables.
+\item
+  \texttt{lua-filters}: Overview and lots of small examples of Lua filters to handle the custom output needed for the R Markdown format.
+\item
+  \texttt{metadata}: This has a collection of additional format handling including extracting metadata like author names and affiliations, article identifiers used in the review process, and handling citations, footnotes and links.
+\item
+  \texttt{bibliography}: The bibliography was handled differently over the years of the journal, and this details how to use the \texttt{rebib} functionality to handle \texttt{bbl} files, embedded \texttt{bbl}, to convert into the standard \texttt{.bib} format.
 \end{itemize}
 
-The GitHub repos supporting this work are:
+In each of these folders there is a \texttt{RJwrapper.tex}, and \texttt{.tex} file, with the extra template files \texttt{RJournal.sty} and \texttt{Rlogo-5.png} and \texttt{.bib} files.
+These match the legacy template file structure, from which the \texttt{RJwrapper.pdf} file is created.
+To test the conversion for each of these examples, set the path directory to one of the folders and use the \texttt{latex\_to\_web()} function as follows:
+
+\begin{verbatim}
+article_dir <- "path-to-this supplementary folder"
+texor::latex_to_web(article_dir)
+\end{verbatim}
+
+This will create an \texttt{.Rmd} and \texttt{.html} files in the same directory, that demonstrate the converted R Markdown version and the HTML output format.
+
+You'll need to ensure that you have the latest versions of \CRANpkg{texor} and \CRANpkg{rebib}, and Pandoc (at least later than version 3.0.0).
+
+\hypertarget{source-materials}{%
+\subsection*{Source materials}\label{source-materials}}
+\addcontentsline{toc}{subsection}{Source materials}
+
+The \CRANpkg{texor} and \CRANpkg{rebib} source codem and materials to reproduce this paper are available at:
 
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-tableHTML}{}}%
+Boutaris, Theo, Clemens Zauchner, and Dana Jomar. 2023. \emph{tableHTML: A Tool to Create HTML Tables}. \url{https://CRAN.R-project.org/package=tableHTML}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-distill}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-htmlTable}{}}%
+Gordon, Max, Stephen Gragg, and Peter Konings. 2022. \emph{htmlTable: Advanced Tables for Markdown/HTML}. \url{https://CRAN.R-project.org/package=htmlTable}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-markdown}{}}%
 Gruber, John. 2002. {``Markdown.''} https://daringfireball.net/projects/markdown/.
+
+\leavevmode\vadjust pre{\hypertarget{ref-gt}{}}%
+Iannone, Richard, Joe Cheng, Barret Schloerke, Ellis Hughes, Alexandra Lauer, and JooYoung Seo.
+2023.
+\emph{Gt: Easily Create Presentation-Ready Display Tables}.
+\url{https://CRAN.R-project.org/package=gt}.
 
@@
 MacFarlane, John. 2023. {``Pandoc: A Universal Document Converter.''} https://pandoc.org.
+
+\leavevmode\vadjust pre{\hypertarget{ref-tables}{}}%
+Murdoch, Duncan. 2023. \emph{Tables: Formula-Driven Table Generation}. \url{https://CRAN.R-project.org/package=tables}.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-DT}{}}%
+Xie, Yihui, Joe Cheng, and Xianying Tan. 2023. \emph{DT: A Wrapper of the JavaScript Library 'DataTables'}. \url{https://CRAN.R-project.org/package=DT}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-bookdown}{}}%
 Yihui Xie. 2023. \emph{{bookdown: Authoring Books and Technical Documents with R Markdown}}. \#equations. \url{https://bookdown.org/yihui/bookdown/markdown-extensions-by-bookdown.html}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kableExtra}{}}%
+Zhu, Hao. 2021. \emph{kableExtra: Construct Complex Table with 'Kable' and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.
 

### 2023-08-09 texor.R

@@ -7,6 +7,92 @@ library(plotly)
 library(ggplot2)
 
 
-## ----workflow, fig.cap="Workflow of the document conversion conducted by texor."----
+## ----workflow, fig.cap="Workflow of the document conversion conducted by texor.", out.width="70%", fig.align='center'----
 knitr::include_graphics("workflow.png")
+
+
+## ----alghow, fig.cap="How to write algorithms.", out.width="60%", fig.align='center'----
+knitr::include_graphics("alghow.png")
+
+
+## \begin{table*}[htbp]
+
+## \begin{center}
+
+## \begin{tabular}{l | llll }
+
+##  \hline
+
+##  \multicolumn{1}{c |}{EXAMPLE} & \multicolumn{2}{c}{$X$} &
+
+## \multicolumn{2}{c}{$Y$} \\
+
+##  \hline
+
+##   & 1 & 2 & 1 & 2 \\
+
+##  EX1  & X11 & X12 &  Y11  & Y12 \\
+
+##  EX2  & X21 & X22 &  Y21  & Y22 \\
+
+##  EX3  & X31 & X32 &  Y31  & Y32 \\
+
+##  EX4  & X41 & X42 &  Y41  & Y42\\
+
+##  EX5  & X51 & X52 &  Y51  & Y52 \\
+
+## \hline
+
+## \end{tabular}
+
+## \caption{An example multicolumn table.}
+
+## \label{table:2}
+
+## \end{center}
+
+## \end{table*}
+
+
+## ::: center
+
+## ::: {#table:2}
+
+##   -------------------------------------
+
+##   EXAMPLE   \(X\)         \(Y\)
+
+##   --------- ------- ----- ------- -----
+
+##             1       2     1       2
+
+## 
+
+##   EX1       X11     X12   Y11     Y12
+
+## 
+
+##   EX2       X21     X22   Y21     Y22
+
+## 
+
+##   EX3       X31     X32   Y31     Y32
+
+## 
+
+##   EX4       X41     X42   Y41     Y42
+
+## 
+
+##   EX5       X51     X52   Y51     Y52
+
+##   -------------------------------------
+
+## 
+
+##   : Table 2: An Example Table with Multicolumn
+
+## :::
+
+## :::
 

### 2023-08-09 texor.Rmd
rebib section populated

@@
 The format is regularly updated.
-Early articles were typeset using LateX [@latex], from a specific, but changing, template.
+Early articles were typeset using LaTeX [@latex], from a specific, but changing, template.
 This requires that code is separated from the documentation, and there is a chance that code chunks in the paper don't reproduce the results reported.
@@
 The paper is organised as follows.
-Section 2 gives an overview of the conversion process, that includes pre-processing using regular expressions, post-processing using lua filters, and handling of figures, tables and equations.
+Section 2 gives an overview of the conversion process, that includes pre-processing using regular expressions, post-processing using Lua filters, and handling of figures, tables and equations.
 Section 3 describes the \CRANpkg{texor} that handles most of the conversion.
@@
 It is embedded in many other applications to allow custom scripting for extensibility.
-Pandoc allows users to provide custom lua filters to produce custom output formats.
+Pandoc allows users to provide custom Lua filters to produce custom output formats.
 The \CRANpkg{texor} package handles post-processing of the R Markdown document into the special format for the R Journal using a suite of Lua filters.
@@
 LaTeX automatically numbers equations, unless specifically instructed not to.
-Equation numbering in R Markdown requires specific labelling using `(\#eq:xx)` as described in @bookdown.
-The \CRANpkg{texor} helps by adding the labelling using a Lua filter to convert the existing `\label{..}` to `(\#eq:xx)`.
+Equation numbering in R Markdown requires specific labeling using `(\#eq:xx)` as described in @bookdown.
+The \CRANpkg{texor} helps by adding the labeling using a Lua filter to convert the existing `\label{..}` to `(\#eq:xx)`.
 
@@
 
-## Overview
+Typically bibliographies are generated during the processing of a LaTeX article using the BibTeX software [@bibtex] operating on a `.bib` list of references.
+The R Journal template requires the inclusion of the `.bib` file.
+But LaTeX actually uses a `.bbl` format for references, which is what BibTeX generates as an intermediate format during the article processing.
 
-The \CRANpkg{rebib} package addresses the issue with LaTeX articles using  built-in bibliography options with or without BibTeX files.
+During the conversion of legacy articles, especially in the earliest published articles, it was discovered that some papers were only accompanied by the intermediate `.bbl` formatted references and often these were included directly in the `.tex` file.
+Sometimes articles had both references in the `.tex` and separate `.bbl` or `.bib` files.
+While LaTeX can technically handle either `.bbl` or `.bib` formatted references R Markdown can only handle `.bib`.
+
+The \CRANpkg{rebib} package was developed to handle these tricky situations. It converts embedded LaTeX bibliographies into a close BibTeX equivalent. The features of the package are:
+
+- reading `bbl` chunks to produce a very close BibTeX equivalent.
+- title and author are usually mandatory fields.
+- URL, ISBN, publisher, pages and year are optional fields and will be enabled when available.
+- remaining data is stored in `"journal"` (internally) and `"publisher"` (when writing BibTeX file).
+- ignores commented LaTeX code.
+- provides a citation tracker
+- does bibliography aggregation XXX what is this?
+
+<!-- The \CRANpkg{rebib} package addresses the issue with LaTeX articles using built-in bibliography options with or without BibTeX files.
 While this works  well with LaTeX, it won't work with Rmarkdown.
@@
 It was initially a part of the \CRANpkg{texor} package, as those functions grew in features and became more involved; at that point, it made sense to move those functions as a separate package.
-Initially, there were some reservations about
-the usage of \CRANpkg{rebib} and its stability with various formats. However, the package has improved over time and proven to be a good performer.
+Initially, there were some reservations about the usage of \CRANpkg{rebib} and its stability with various formats.
+However, the package has improved over time and proven to be a good performer.-->
 
-## Using \CRANpkg{rebib}
+<!-- ## Using \CRANpkg{rebib}-->
 
-**Install instructions**
+The package `rebib` can be installed from CRAN using:
 
-**Examples / getting started / usage and so on**
+```
+install.packages("rebib")
+```
+
+and the development version from https://github.com/Abhi-1U/rebib. The website for the package, https://abhi-1u.github.io/rebib, has vignettes documenting usage.
 
@@
 - `code-env`: Explains how different code environments defined by the R Journal style are handled, and additional details such as code in figure environments, and code in table environments. 
-- `math-env`: Examples of inline math, display math, how equation numbering is handled by a Lua filter to convert from LaTeX labelling to R Markdown labelling. 
-- `figure-env`: Explains how the variety of figure definitions are handled in the conversion, including different image formats, numbering, captions, labelling, multiple images, and `tikz` [@casflow] images.
+- `math-env`: Examples of inline math, display math, how equation numbering is handled by a Lua filter to convert from LaTeX labeling to R Markdown labeling. 
+- `figure-env`: Explains how the variety of figure definitions are handled in the conversion, including different image formats, numbering, captions, labeling, multiple images, and `tikz` [@casflow] images.
 - `table-env`: Examples of how a variety of table types are converted, including multicolumn, complex and wide tables.
@@
 
-The \CRANpkg{texor} and \CRANpkg{rebib} source codem and materials to reproduce this paper are available at:
+The \CRANpkg{texor} and \CRANpkg{rebib} source code and materials to reproduce this paper are available at:
 

### 2023-08-09 texor.tex
rebib section populated

@@
 The format is regularly updated.
-Early articles were typeset using LateX (The LaTeX Project 2023), from a specific, but changing, template.
+Early articles were typeset using LaTeX (The LaTeX Project 2023), from a specific, but changing, template.
 This requires that code is separated from the documentation, and there is a chance that code chunks in the paper don't reproduce the results reported.
@@
 LaTeX automatically numbers equations, unless specifically instructed not to.
-Equation numbering in R Markdown requires specific labelling using \texttt{(\textbackslash{}\#eq:xx)} as described in Yihui Xie (2023).
-The \CRANpkg{texor} helps by adding the labelling using a Lua filter to convert the existing \texttt{\textbackslash{}label\{..\}} to \texttt{(\textbackslash{}\#eq:xx)}.
+Equation numbering in R Markdown requires specific labeling using \texttt{(\textbackslash{}\#eq:xx)} as described in Yihui Xie (2023).
+The \CRANpkg{texor} helps by adding the labeling using a Lua filter to convert the existing \texttt{\textbackslash{}label\{..\}} to \texttt{(\textbackslash{}\#eq:xx)}.
 
@@
 
-\hypertarget{overview}{%
-\subsection{Overview}\label{overview}}
+Typically bibliographies are generated during the processing of a LaTeX article using the BibTeX software (Feder 2006) operating on a \texttt{.bib} list of references.
+The R Journal template requires the inclusion of the \texttt{.bib} file.
+But LaTeX actually uses a \texttt{.bbl} format for references, which is what BibTeX generates as an intermediate format during the article processing.
 
-The \CRANpkg{rebib} package addresses the issue with LaTeX articles using built-in bibliography options with or without BibTeX files.
-While this works well with LaTeX, it won't work with Rmarkdown.
-Initially the goal was to use external software like Biber to convert the embedded bibliography to BibTeX.
-However the integration of external software was not viable, hence the experimental idea of a bibliography parser gained momentum.
+During the conversion of legacy articles, especially in the earliest published articles, it was discovered that some papers were only accompanied by the intermediate \texttt{.bbl} formatted references and often these were included directly in the \texttt{.tex} file.
+Sometimes articles had both references in the \texttt{.tex} and separate \texttt{.bbl} or \texttt{.bib} files.
+While LaTeX can technically handle either \texttt{.bbl} or \texttt{.bib} formatted references R Markdown can only handle \texttt{.bib}.
 
-It was initially a part of the \CRANpkg{texor} package, as those functions grew in features and became more involved; at that point, it made sense to move those functions as a separate package.
-Initially, there were some reservations about
-the usage of \CRANpkg{rebib} and its stability with various formats. However, the package has improved over time and proven to be a good performer.
+The \CRANpkg{rebib} package was developed to handle these tricky situations. It converts embedded LaTeX bibliographies into a close BibTeX equivalent. The features of the package are:
 
-\hypertarget{using}{%
-\subsection{\texorpdfstring{Using \CRANpkg{rebib}}{Using }}\label{using}}
+\begin{itemize}
+\tightlist
+\item
+  reading \texttt{bbl} chunks to produce a very close BibTeX equivalent.
+\item
+  title and author are usually mandatory fields.
+\item
+  URL, ISBN, publisher, pages and year are optional fields and will be enabled when available.
+\item
+  remaining data is stored in \texttt{"journal"} (internally) and \texttt{"publisher"} (when writing BibTeX file).
+\item
+  ignores commented LaTeX code.
+\item
+  provides a citation tracker
+\item
+  does bibliography aggregation XXX what is this?
+\end{itemize}
 
-\textbf{Install instructions}
+The package \texttt{rebib} can be installed from CRAN using:
 
-\textbf{Examples / getting started / usage and so on}
+\begin{verbatim}
+install.packages("rebib")
+\end{verbatim}
+
+and the development version from \url{https://github.com/Abhi-1U/rebib}. The website for the package, \url{https://abhi-1u.github.io/rebib}, has vignettes documenting usage.
 
@@
 \item
-  \texttt{math-env}: Examples of inline math, display math, how equation numbering is handled by a Lua filter to convert from LaTeX labelling to R Markdown labelling.
+  \texttt{math-env}: Examples of inline math, display math, how equation numbering is handled by a Lua filter to convert from LaTeX labeling to R Markdown labeling.
 \item
-  \texttt{figure-env}: Explains how the variety of figure definitions are handled in the conversion, including different image formats, numbering, captions, labelling, multiple images, and \texttt{tikz} (Josh Cassidy 2013) images.
+  \texttt{figure-env}: Explains how the variety of figure definitions are handled in the conversion, including different image formats, numbering, captions, labeling, multiple images, and \texttt{tikz} (Josh Cassidy 2013) images.
 \item
@@
 \url{https://CRAN.R-project.org/package=distill}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-bibtex}{}}%
+Feder, Alexander. 2006. {``BibTeX: Reference Management Software.''} https://www.bibtex.org.
 

### 2023-08-10 texor.Rmd
Minor Edits to the paper

@@
 
-```{r workflow, fig.cap="Workflow of the document conversion conducted by texor.", out.width="70%", fig.align='center'}
+```{r workflow, fig.cap="Workflow of the document conversion conducted by texor.", out.width="90%", fig.align='center'}
 knitr::include_graphics("workflow.png")
@@
 
-The HTML output renders math using MathJax. This does mean that some functionality, like `\bm`, `\boldmath` and `mathbb` not supported, and special definitions cannot be handled.
+The HTML output renders math using MathJax.
+This does mean that some functionality, like `\bm`, `\boldmath` and `\mathbbm` not supported, and special definitions can be handled albeit in a limited capacity.
 
@@
 
-1. The stream editor modifies the `\multicolumnx{..}` to
-    `\multicolumnxx{..}`.
+1. The stream editor modifies the `\multicolumn{..}` to
+    `\multicolumnx{..}`.
 
 2. A LaTeX macro is used to redefine the `\multicolumnxx{..}` to
-    `\multicolumnx{---}`.
+    `\multicolumn{---}` (which is accepted by pandoc).
 
@@
 For individuals who are interested in submitting their paper to the R Journal but have written their article using the legacy LaTeX format and wish to convert it to the new format, you can use the `latex_to_web()` function on your paper directory.
+
+XXX - Should we include various options/params for the function here ?
 

### 2023-08-10 texor.tex
Minor Edits to the paper

@@
 
-{\centering \includegraphics[width=0.7\linewidth]{workflow}
+{\centering \includegraphics[width=0.9\linewidth]{workflow}
 
@@
 The HTML output renders math using MathJax.
-This does mean that some functionality, like \texttt{\textbackslash{}bm}, \texttt{\textbackslash{}boldmath} and \texttt{mathbb} not supported, and special definitions cannot be handled.
+This does mean that some functionality, like \texttt{\textbackslash{}bm}, \texttt{\textbackslash{}boldmath} and \texttt{\textbackslash{}mathbbm} not supported, and special definitions can be handled albeit in a limited capacity.
 
@@
 \item
-  The stream editor modifies the \texttt{\textbackslash{}multicolumnx\{..\}} to
-  \texttt{\textbackslash{}multicolumnxx\{..\}}.
+  The stream editor modifies the \texttt{\textbackslash{}multicolumn\{..\}} to
+  \texttt{\textbackslash{}multicolumnx\{..\}}.
 \item
   A LaTeX macro is used to redefine the \texttt{\textbackslash{}multicolumnxx\{..\}} to
-  \texttt{\textbackslash{}multicolumnx\{-\/-\/-\}}.
+  \texttt{\textbackslash{}multicolumn\{-\/-\/-\}} (which is accepted by pandoc).
 \item
@@
 For individuals who are interested in submitting their paper to the R Journal but have written their article using the legacy LaTeX format and wish to convert it to the new format, you can use the \texttt{latex\_to\_web()} function on your paper directory.
+
+XXX - Should we include various options/params for the function here ?
 

### 2023-08-10 texor.R

@@ -7,7 +7,7 @@ library(plotly)
 library(ggplot2)
 
 
-## ----workflow, fig.cap="Workflow of the document conversion conducted by texor.", out.width="70%", fig.align='center'----
+## ----workflow, fig.cap="Workflow of the document conversion conducted by texor.", out.width="90%", fig.align='center'----
 knitr::include_graphics("workflow.png")
 
 

### 2023-08-10 supplement/math-env/math-example.tex
Updated math supplement with custom commands.

--- 2023-07-28 supplement/math-env/math-example.tex
@@
 
-The core of the MathJax project is the development of its state-of-the-art, open source, JavaScript platform for display of mathematics. The key design goals are \citep{mathjax}:
+The core goal of the MathJax project is the development of its state-of-the-art, open source, JavaScript platform for display of mathematics. The key design goals are \citep{mathjax}:
 \begin{itemize}
@@
 This will render as \ref{eq:1}.
+
 \begin{align}
+\label{eq:1}
 f(x) = \frac{1}{\sigma\sqrt{2\pi}} 
   \exp\left( -\frac{1}{2}\left(\frac{x-\mu}{\sigma}\right)^{\!2}\,\right)
-  \label{eq:1}
 \end{align}
@@
 \section{Equation Numbering}
-In LaTeX your equations get numbered automatically unless you are using \verb|\[..\]| to describe
-math. Equation numbering works a bit differently in \CRANpkg{bookdown} (the base of RJ web article format) where
+In LaTeX your equations get numbered automatically unless you are using \verb|\[..\]| environment or \verb|\nonumber| to suppress
+numbering. Equation numbering works a bit differently in \CRANpkg{bookdown} (the base of RJ web article format) where
 it is mandatory to have a \verb|(\#eq:xx)| which is described in more detail in \citep{bookdown}.
@@
 \begin{align}
-S_{T, s}(z_t) = & X^{\top} K_{b,t}^* X (Z - z_t)^s 
+S_{T, s}(z_t) = X^{\top} K_{b,t}^* X (Z - z_t)^s, 
 \label{eq:xe}
@@
 
+\section{Custom Math Commands}
+
+%Also avoid custom math commands MathJaX will never be able to parse it. The reason being
+%MathJax renders math case by case and custom commands will already be removed by pandoc
+%during the conversion, hence MathJax will not be aware of such redefinations.
+To ease the frequent use of the command, some custom math commands that contain LaTeX math commands. 
+However, This will not function if you use functions/environments from CTAN packages or 
+the ones that are out-of-the-box functions which do not render in HTML/markdown.
+
+\paragraph{Acceptable}
+
+\begin{verbatim}
+\newcommand{\ABS}[1]{|#1|}
+\end{verbatim}
+
+$$ \ABS{\sigma^2} = \pm 1 $$
+
+\paragraph{Not acceptable}
+
+\begin{verbatim}
+\newcommand{\rotatethis}[1]{\rotatebox[origin=c]{90}{$#1$}}
+\end{verbatim}
+
+$$ \rotatethis{\sigma^2} = \pm 1 $$
+
 \section{Unsupported LaTeX Commands}
@@
 Altough MathJax does a good job of supporting most LaTeX math functions,
-some functions do not work like \verb|\bm| \verb|\boldmath|, instead use \verb|\mathbb|.
-Also avoid custom math commands MathJaX will never be able to parse it. The reason being
-MathJax renders math case by case and custom commands will already be removed by pandoc
-during the conversion, hence MathJax will not be aware of such redefinations.
+some functions do not work like \verb|\bm| \verb|\boldmath|, instead use \verb|\mathbf|.
 
@@
 \newblock  \penalty0 2021
-\newblock URL \url{https://www.mathjax.org/#features}
+\newblock URL \url{https://www.mathjax.org/}
 
@@
 \newblock \penalty0 2023
-\newblock URL \url{https://bookdown.org/yihui/bookdown/markdown-extensions-by-bookdown.html#equations}
+\newblock URL \url{https://bookdown.org/yihui/bookdown/markdown-extensions-by-bookdown.html}
 

### 2023-08-10 texor.Rmd
Updated notes and highlight params

@@
 
-```
+```latex
 \begin{figure}[htbp]
@@
 
-```
+```html
 <figure id="figure:rlogo">
@@
 
-```
+```latex
 \begin{figure*}[htbp]
@@
 
-```
+```html
 <figure id="fig:twoimages">
@@
 
-```
+```latex
 \begin{figure}
@@
 
-```
+```html
 <figure id="fig:tikz">
@@
 
-```
+```latex
 \begin{algorithm}[htbp]
@@
 
-```
+```html
 <figure id="alg:how">
@@
 
-```
+```latex
 \begin{table}[htbp]
@@
 
-```
+```markdown
 ::: {#table:1}
@@
 
-```
+```r
 install.packages("texor")
@@
 
-```
+```r
 rmarkdown::pandoc_version()
@@
 
-XXX
+XXX ??? warn authors about the conversion not being 1:1 with rmarkdown and might need further edits in with tables and figures.
 
@@
 
-```
+```r
 install.packages("rebib")
@@
 
-```
+```r
 article_dir <- "path-to-this supplementary folder"

### 2023-08-15 supplement/figure-env/figure-example.tex
Updated supplement examples
overview of current examples
Updated figures and clarifications on differences in CodeBlocks

@@
 Images are an essential component in any article, However due to the differences in the support for various graphic formats between LaTeX and markdown/HTML we need to fallback on raster graphics. 
-It is well summarized in the table \ref{table:1}
+It is well summarized in the Table \ref{table:1}.
 
@@
 \subsubsection{Tikz Code:}
-\begin{verbatim}
-\begin{figure}
+The figure  \ref{fig:tikz} is an example of how \pkg{texor} handles tikz images,
 
-%% Generated Image will included as a PNG above automatically
-  \centering
-\tikzstyle{process} = [rectangle, rounded corners,
-minimum width=3cm, 
-minimum height=1cm,
-text centered, 
-draw=black]
-\tikzstyle{arrow} = [thick,->,>=stealth]
-\begin{tikzpicture}[node distance=4cm]
-%Nodes
-\node (start) [process] {LaTeX Source with Tikz Image};
-\node (isolate) [process, below right of=start] {Isolated tikz env};
-\node (pro1) [process, below left of=isolate] {Standalone PDF file};
-\node (pro2) [process, below right of=pro1, yshift=-0.5cm] {PNG file};
-\node (link) [process, below left of=pro2] {Include PNG in article};
-\node (stop) [process, below right of=link] {Convert To RJ-web-article};
-% arrows
-\draw [arrow] (start) -- node[anchor=east] {Extract Figure Block} (isolate);
-\draw [arrow] (isolate) -- node[anchor=east] {Compile} (pro1);
-\draw [arrow] (pro1) -- node[anchor=east] {Convert to raster} (pro2);
-\draw [arrow] (pro2) -- node[anchor=east] {Include back to source as image} (link);
-\draw [arrow] (link) -- node[anchor=east] {Using Pandoc} (stop);
-\end{tikzpicture}
-\caption{Tikz Image example}
-  \label{fig:tikz}
-\end{figure}
+\begin{itemize}
+\item First figures containing tikz images are isolated.
+\item Tikz libraries are fetched from the wrapper file,
+\item Tikz Code section is isolated into a standalone LaTeX file and compiled.
+\item The Compiled LaTeX file generates a PDF Image.
+\item This is then Converted again to PNG format as PDFs embedded in HTML wont look good.
+\item In the figure environment a line is added \verb|\includegraphics{tikz/somefigure.png}|
+\item During the conversion in pandoc, a lua filter removes extra/unncessary text from the rest of the figure
+environemnt.
+\end{itemize}
 
-\end{verbatim}
+If You use \pkg{texor} to convert your articles using \verb|texor::latex_to_web()| with \verb|temp_mode=TRUE|(it is TRUE by default).
+The resultant Rmarkdown/HTML file will not modify the contents of your LaTeX file.
+In this case you can keep reloading the article after making changes to the tikz images, without having to do the above steps manually in case you are converting the article by hand.
 
-\subsubsection{Resultant Figure :}
 \begin{figure*}
@@
 \tikzstyle{arrow} = [thick,->,>=stealth]
-\begin{tikzpicture}[node distance=4cm]
+\begin{tikzpicture}[node distance=2cm]
 %Nodes
-\node (start) [process] {LaTeX Source with Tikz Image};
-\node (isolate) [process, below right of=start] {Isolated tikz env};
-\node (pro1) [process, below left of=isolate] {Standalone PDF file};
-\node (pro2) [process, below right of=pro1, yshift=-0.5cm] {PNG file};
-\node (link) [process, below left of=pro2] {Include PNG in article};
-\node (stop) [process, below right of=link] {Convert To RJ-web-article};
+\node (start) [process] {Tikz};
+\node (isolate) [process, below right of=start] {Isolated LaTeX};
+\node (pro1) [process, below right of=isolate] {PDF};
+\node (pro2) [process, below right of=pro1, yshift=-0.5cm] {PNG};
 % arrows
-\draw [arrow] (start) -- node[anchor=east] {Extract Figure Block} (isolate);
+\draw [arrow] (start) -- node[anchor=east] {Extract} (isolate);
 \draw [arrow] (isolate) -- node[anchor=east] {Compile} (pro1);
-\draw [arrow] (pro1) -- node[anchor=east] {Convert to raster} (pro2);
-\draw [arrow] (pro2) -- node[anchor=east] {Include back to source as image} (link);
-\draw [arrow] (link) -- node[anchor=east] {Using Pandoc} (stop);
+\draw [arrow] (pro1) -- node[anchor=east] {Rasterize} (pro2);
+
 \end{tikzpicture}
@@
 Here is an example of algorithm \ref{alg:how}, referenced from \cite{algoexample}
-
-\begin{verbatim}
-\begin{algorithm}[htbp]
-\SetAlgoLined
-\KwData{this text}
-\KwResult{how to write algorithm with \LaTeX2e }
-initialization\;
-\While{not at end of this document}{
-read current\;
-\eIf{understand}{
-go to next section\;
-current section becomes this one\;
-}{
-go back to the beginning of current section\;
-}
-}
-\caption{How to write algorithms}
-  \label{alg:how}
-\end{algorithm}
-\end{verbatim}
 
@@
 \section{Other elements in figure objects}
-Figures can also house non-image environments like CodeBlocks and Para. This 
-also creates an opportunity to add support for CodeBlock numbering with some
-lua filter trickery.
+
+Figures can also house non-image environments like CodeBlocks and WideTables. This is 
+the part where things start to look a bit different in the converted articles, CodeBlocks 
+in figure environments would have numbering and caption that goes "Figure x:". However, 
+when the same CodeBlock is converted to Rmarkdown, they are numbered as "CodeBlock x:".
+This distinguishes them from normal Images and matches with authors vision more appropriately.
+
+As for Inclusion of widetables, since it houses two different tabular environments with the same caption,
+it would be treated as two different tables by pandoc sharing the same caption. This would not be the best solution hence, \pkg{texor} package uses Figure enviornment to house the widetables.
+This way they are distinguished well from normal tables, also the numbering scheme is seperated from figures "WideTable x:".
+
+However one important thing to note is, the reference links and reference numbering would be shared with ones of the figures.
+So the changes that the package makes are cosmetic to captions.
+Like CodeBlock 1 below refers to \ref{code:example}.
+The number should have been 1 but instead it is 7.
+
 \begin{figure}[htbp]

### 2023-08-16 texor.Rmd
working on conclusions

@@
 
-## Animation
-
-XXX Shall we describe the one pdf paper that has animation and how this is handled - could be cool to include
-
 ## Logging the conversion
@@
 
-The supplementary materials have several folders with specific examples of figures, tables, equations that can be used to understand how the conversion operates.
+For converting the 14 years of legacy R Journal articles, batch processing of issues was conducted.
 
-For converting the 14 years of legacy R Journal articles, batch processing of issues was conducted. This can be achieved using:
-
-XXX ??? warn authors about the conversion not being 1:1 with rmarkdown and might need further edits in with tables and figures.
+<!--XXX ??? warn authors about the conversion not being 1:1 with rmarkdown and might need further edits in with tables and figures.-->
 
 For individuals who are interested in submitting their paper to the R Journal but have written their article using the legacy LaTeX format and wish to convert it to the new format, you can use the `latex_to_web()` function on your paper directory.
+This will get you about 80% of the way to an R Markdown version of your paper.
+You will then want to
+
+- Edit the lines where figures are included. 
+- Edit the tables. 
+- Include your R code, to dynamically do the computing described in your article.
 
@@
 
-Remember only the HTML is rendered, and the original PDF is not re-generated.
+The original motivation for the \CRANpkg{texor} and \CRANpkg{rebib} packages was to convert legacy LaTeX articles into HTML format to accessibility.
+Thus it only creates the R Markdown and an HTML version leaving the original PDF as published.
+This will also be used to produce HTML versions of newly accepted papers only submitted as LaTeX.
+
+However, it could be useful to help authors convert to dynamic documents, by providing an initial R Markdown version of their LaTeX, that with some modification, such as including their R code for computations directly in the document, will produce both PDF nd HTML versions of their submitted paper.
+
+<!--Remember only the HTML is rendered, and the original PDF is not re-generated.
 
@@
 
-Where this might be useful in the future, other applications.
+Where this might be useful in the future, other applications.-->
 

### 2023-08-16 texor.tex
Updated notes and highlight params
working on conclusions

@@
 
-\hypertarget{animation}{%
-\subsection{Animation}\label{animation}}
-
-XXX Shall we describe the one pdf paper that has animation and how this is handled - could be cool to include
-
 \hypertarget{logging-the-conversion}{%
@@
 
-The supplementary materials have several folders with specific examples of figures, tables, equations that can be used to understand how the conversion operates.
-
-For converting the 14 years of legacy R Journal articles, batch processing of issues was conducted. This can be achieved using:
-
-XXX
+For converting the 14 years of legacy R Journal articles, batch processing of issues was conducted.
 
 For individuals who are interested in submitting their paper to the R Journal but have written their article using the legacy LaTeX format and wish to convert it to the new format, you can use the \texttt{latex\_to\_web()} function on your paper directory.
+This will get you about 80\% of the way to an R Markdown version of your paper.
+You will then want to
+
+\begin{itemize}
+\tightlist
+\item
+  Edit the lines where figures are included.
+\item
+  Edit the tables.
+\item
+  Include your R code, to dynamically do the computing described in your article.
+\end{itemize}
 
@@
 
-Remember only the HTML is rendered, and the original PDF is not re-generated.
+The original motivation for the \CRANpkg{texor} and \CRANpkg{rebib} packages was to convert legacy LaTeX articles into HTML format to accessibility.
+Thus it only creates the R Markdown and an HTML version leaving the original PDF as published.
+This will also be used to produce HTML versions of newly accepted papers only submitted as LaTeX.
 
-Only handling to HTML format, especially relevant for figures, because the legacy pdf will remain the same.
-If using for converting work from LaTeX to Rmd then this will need manual changing by author.
-See rjtools guide.
-
-Where this might be useful in the future, other applications.
+However, it could be useful to help authors convert to dynamic documents, by providing an initial R Markdown version of their LaTeX, that with some modification, such as including their R code for computations directly in the document, will produce both PDF nd HTML versions of their submitted paper.
 

### 2023-08-17 supplement/lua-filters/lua-filters.tex
Updated supplementary materials

@@
       end
-	    local caption = pandoc.utils.stringify(el.caption)
+	    local caption = el.caption
 	    if not caption then
-          caption = label
+          caption = {pandoc.Str(label)}
     	else
-          caption = label .. " " .. caption
+          caption = {pandoc.Str(label),pandoc.Space()}
     	end
-    	el.caption.long[1] = caption
+    	el.caption.long[1].content = caption .. el.caption.long[1].content
     	is_fig = 0
@@
 \section{Usage of pandoc lua filters in texor package}
-
-\section{Summary}
+Pandoc Lua filters are used in texor package to modify the AST for various markups.
+ Table \ref{tab:filters} summarizes the use of each filter.
+\begin{table}
+\begin{tabular}{|l|p{10cm}|}
+\hline
+File Name & Description\\
+\hline
+\verb|abs_filter.lua| & {Filters out unicode \verb|182| character.}\\
+\hline
+\verb|bib_filter.lua| & Clears out bibliography from the article itself as it will added to metadata\\
+\hline
+\verb|R_code.lua| & Adds a class component \verb|'r'| to CodeBlocks for code highlight \\
+\hline
+\verb|image_filter.lua| & Fixes extensions for image paths without extension or \verb|.pdf| extension \\
+\hline
+\verb|table_caption.lua| & Adds Table numbering in the captions similar to the ones found in Rmarkdown and kable based tables. \\
+\hline
+\verb|image_caption.lua| & Adds Figure/Algorithm/CodeBlock/WideTable numbering as well as clears out residuals of tikz/algorithm images \\
+\hline
+\verb|conversion_compat_check.lua| & This filter keeps a count all Inline and Block elements and writes it to a yaml file. \\
+\hline
+\verb|equation_filter.lua| & Adds Bookdown style equation numbering to LaTeX equation/math environments.\\
+\hline
+\verb|bookdown_ref.lua| & Changes the format of Links to bookdown (markdown) style of links\\
+\hline
+\verb|issue_checker.lua| & Searches and notifies any occourence of unrecognized math commands.\\
+\hline
+\verb|find_pdf_files.lua| & This filter creates a list for all Image PDF files included in the article.\\
+\hline
+\end{tabular}
+\caption{A summary of all Lua filters used in texor package (available in \href{https://github.com/Abhi-1U/texor/tree/master/inst}{/inst} folder of package).}
+\label{tab:filters}
+\end{table}
 

### 2023-08-18 texor.Rmd
general edits

@@
 The reasons were to encourage better reproducibility of articles using dynamic documents, enable interactivity in articles, and to make the articles more accessible for vision-impaired readers.
-A resulting challenge was to explore whether legacy articles might be suitably converted into R Markdown before converting into HTML output.
-This paper describes the process to convert an R Journal LaTeX template article into an R Markdown article, and the two new R packages, \CRANpkg{texor} and \CRANpkg{rebib}, that can be used to achieve the conversion.
+A resulting challenge was to explore whether legacy articles might be suitably converted into HTML output.
+This paper describes the process to convert an R Journal article from LaTeX to HTML format via R Markdown, and the two new R packages, \CRANpkg{texor} and \CRANpkg{rebib}, that can be used to achieve the conversion.
 draft: true
@@
     address:
-    - Newport, United Kingdom
+    - Coventry, United Kingdom
     orcid: 0000-0002-1256-3375
@@
 The R Journal is the primary open-access outlet for publications produced by the R community.
-It was born in 2008, evolving from a newsletter, that ran from 2001, into a more formal article publication to encourage documenting statistical computing research.
+It was started in 2009, evolving from a newsletter that ran from 2001, to become a more formal publication that encourages and provides credit for the documentation of statistical computing research.
 
-The format is regularly updated.
-Early articles were typeset using LaTeX [@latex], from a specific, but changing, template.
-This requires that code is separated from the documentation, and there is a chance that code chunks in the paper don't reproduce the results reported.
-With the emergence of dynamic document systems like R Markdown [@rmarkdown] a tight-coupling of code and documentation is possible.
+The tooling behind the production of the R Journal is regularly updated.
+Early articles were typeset using LaTeX [@latex], based on a specific, but changing, template.
+Using LaTeX requires that code is separated from the documentation, and there is a chance that code chunks in the paper don't reproduce the results reported.
+With the emergence of dynamic document systems such as R Markdown [@rmarkdown], a tight coupling of code and documentation is possible.
 Code chunks are dynamically executed when the document is typeset using a system like \CRANpkg{knitr} [@knitr], making reporting of computing research more reproducible.
 
-In 2019, with the help of funding from the R Consortium it was decided that it was time to update operations.
-One aspect of this was to change from LaTeX paper submissions to a more reproducible format, where code was embedded in the document, and the output could be both HTML and pdf.
+In 2019, with the help of funding from the R Consortium, work began to update operations.
+One aspect of this was to change from LaTeX article submissions to a more reproducible format, where code was embedded in the document, and the output could be both HTML and pdf.
 There are numerous benefits of HTML format:
@@
 1. Articles can include interactive graphics and tables.
-2. The format is more accessible to screen readers making the work more accessible to vision-impaired researchers.
+2. The format is more accessible to screen readers and other assistive technologies, making the work more accessible to vision-impaired researchers.
+3. HTML provides a more comfortable reading experience on mobile devices, which are increasing used as researchers work on the move and share work via social media.
+4. Search engines can easily access the full text of articles, facilitating discovery of published articles.
 
@@
 
-This latter point is a reason to consider converting all of the legacy articles into HTML.
+The latter points are motivations for converting all of the legacy R Journal articles into HTML.
 
-A key decision for creating conversion software was to decide to directly convert LaTeX to HTML, or PDF to HTML, or LaTeX to R Markdown, and then use the current journal tools to create the HTML.
+A key decision when designing the conversion software, was whether to convert the LaTeX source to HTML; the PDF output to HTML; or the LaTeX source to R Markdown, which would then be converted to HTML using the current journal tools.
 The latter approach was decided to be the most versatile and useful.
@@
 In addition to article format, changes to the web site structure were important for delivering the publication.
-Web site architectures are also constantly evolving, and the emergence of \CRANpkg{distill} [@distill] allows for the journal web site to optimally deliver R Markdown articles.
+Web site architectures are also constantly evolving, and the emergence of \CRANpkg{distill} [@distill] allows the journal web site to optimally deliver R Markdown articles.
 
-The \CRANpkg{rjtools} was developed to create articles using R Markdown for the R Journal, and to embed them into the journal web site.
-The packages described here, \CRANpkg{texor} and \CRANpkg{rebib} describes software to convert legacy LaTeX format articles into Rmarkdown, so that they can be rendered in HTML in the new web site.
+The \CRANpkg{rjtools} was developed to create articles using R Markdown for the R Journal and to embed them into the journal web site.
+The packages described here, \CRANpkg{texor} and \CRANpkg{rebib}, describe software to convert legacy LaTeX format articles into R Markdown, so that they can be rendered in HTML in the new web site.
 
@@
 Section 2 gives an overview of the conversion process, that includes pre-processing using regular expressions, post-processing using Lua filters, and handling of figures, tables and equations.
-Section 3 describes the \CRANpkg{texor} that handles most of the conversion.
+Section 3 describes the \CRANpkg{texor} package that handles most of the conversion.
 Section 4 describes tools to do special handling of bibliography files.
@@
 
-The decision to convert to R Markdown format means that the final output to pdf and HTML will depend on Pandoc [@pandoc].
+The decision to convert to R Markdown format means that the final output to PDF and HTML will depend on Pandoc [@pandoc].
 Pandoc is a versatile document conversion program written in Haskell that is core to numerous documentation systems, including R Markdown and Quarto.
@@
 Pandoc can be used to do the conversion from LaTeX to R Markdown also.
-However, additional pre-processing needs to be done to handle special R Journal LaTeX styling.
-And further post-processing needs to be done to handle specific R Journal R Markdown styling.
+However, some pre-processing is necessary to handle special R Journal LaTeX styling.
+And further post-processing is necessary to handle specific R Journal R Markdown styling.
 The \CRANpkg{texor} package contains functionality to handle this pre- and post-processing of the document, in a workflow illustrated in Figure \@ref(fig:workflow).
@@
 LaTeX is very descriptive language, that allows authors substantial freedom for customization.
-Markdown [@markdown], on which R Markdown is based, is more restrictive and was born to make it easier to create web pages without the distraction of a gazillion HTML tags.
+Markdown [@markdown], on which R Markdown is based, is more restrictive and was designed to make it easier to create web pages without the distraction of a gazillion HTML tags.
 The beauty of Markdown is that it allows the author to focus on writing, without format cluttering the text.
@@
 If the code environment is not verbatim, then Pandoc will also try to process the actual code content as LaTeX commands and will likely lose details.
-It is better to convert these synonyms into `verbatim` environments this prior to passing the document to Pandoc.
+It is better to convert these synonyms into `verbatim` environments prior to passing the document to Pandoc.
 
@@
 
-- `stream_editor()`: operates like the `sed` function in unix [@unix] and allows generic text pattern matching and replacing.
+- `stream_editor()`: operates like the `sed` function in unix [@unix] and allows generic text pattern matching and replacement.
 - `patch_code_env()`: replaces the common code environments, `code`, `example`, `Sin`, `Sout`, `Scode`, `Sinput`, `smallverbatim`, `boxedverbatim`, `smallexample` with `verbatim`.
-- `patch_equations()`: coordinates various equation environments.
-- `patch_figure_env()`: coordinates various figure environments.
-- `patch_table_env()`: coordinates table environments.
+- `patch_equations()`: handles various equation environments.
+- `patch_figure_env()`: handles various figure environments.
+- `patch_table_env()`: handles various table environments.
 
@@
 
-Lua [@lua] is a programming language, that is light-weight, fast, ideal for procedural operations.
+Lua [@lua] is a programming language, that is light-weight and fast: ideal for procedural operations.
 It is embedded in many other applications to allow custom scripting for extensibility.
@@
 
-Algorithms as figures is supported, and the following description will yield the result in Figure \@ref(fig:alghow).
+Algorithms as figures are supported, and the following description will yield the result in Figure \@ref(fig:alghow).
 
@@
 The HTML output renders math using MathJax.
-This does mean that some functionality, like `\bm`, `\boldmath` and `\mathbbm` not supported, and special definitions can be handled albeit in a limited capacity.
+This does mean that some functionality, like `\bm`, `\boldmath` and `\mathbbm` are not supported, and special definitions can only be handled in a limited capacity.
 
@@
 The conversion in \CRANpkg{texor} can mostly handle the simple tables, and for producing more complex tables it may be necessary to manually edit the resulting `Rmd` file to make conditional tables, one to render specifically for HTML output using packages such as \CRANpkg{kableExtra} [@kableExtra], \CRANpkg{gt} [@gt], \CRANpkg{htmlTable} [@htmlTable], \CRANpkg{tableHTML} [@tableHTML], \CRANpkg{tables} [@tables] or \CRANpkg{DT} [@DT].
-A benefit of special HTML table output is interactivity, that would allow the table to be sorted in different ways.
+A benefit of using conditional markup is that it can take advantage of HTML-specific features, such as sortable columns or paged tables.
 
@@
 
-Simple LaTeX tables are converted into traditional markdown format tables, by Pandoc. So this table definition:
+Simple LaTeX tables are converted into traditional markdown format tables by Pandoc. So this table definition:
 
@@
 
-Also note that the stream editor is used to rename `table*` environment to `table` environment because the markdown/HTML will not support the specific changes table\* environment will bring.
+Also note that the stream editor is used to rename `table*` environment to `table` environment because the HTML format is single column, so the asterisk indicating that the table should be drawn over the full width of the page is redundant in this case.
 
@@
 
-The only function that a user necessarily needs is `latex_to_web()`. This creates the R Journal style R Markdown file from a given R Journal style LaTeX file. 
-This is achieved by several sequential steps, `convert_to_markdown()`, `generate_rmd()` and `produce_html()`.
+The only function that a user will typically require is `latex_to_web()`. This creates the R Journal style R Markdown file from a given R Journal style LaTeX file. 
+This is achieved by several sequential steps: `convert_to_markdown()`, `generate_rmd()`, and `produce_html()`.
 
@@
 
-For individuals who are interested in submitting their paper to the R Journal but have written their article using the legacy LaTeX format and wish to convert it to the new format, you can use the `latex_to_web()` function on your paper directory.
+For individuals who are interested in submitting their paper to the R Journal but have written their article using the legacy LaTeX format and wish to convert it to the current R Markdown format, you can use the `latex_to_web()` function on your paper directory.
 This will get you about 80% of the way to an R Markdown version of your paper.
@@
 
-- Edit the lines where figures are included. 
-- Edit the tables. 
+- Edit the lines where figures are included. <!--Why? What will need changing?-->
+- Edit the tables. <!--If the simple conversion is insufficient?-->
 - Include your R code, to dynamically do the computing described in your article.
@@
 Typically bibliographies are generated during the processing of a LaTeX article using the BibTeX software [@bibtex] operating on a `.bib` list of references.
-The R Journal template requires the inclusion of the `.bib` file.
+The current R Journal template requires the inclusion of the `.bib` file.
 But LaTeX actually uses a `.bbl` format for references, which is what BibTeX generates as an intermediate format during the article processing.
 
-During the conversion of legacy articles, especially in the earliest published articles, it was discovered that some papers were only accompanied by the intermediate `.bbl` formatted references and often these were included directly in the `.tex` file.
-Sometimes articles had both references in the `.tex` and separate `.bbl` or `.bib` files.
+The instructions for R Journal authors have changed over time.
+Initially, authors were instructed to run BibTeX on their article and to include the `.bbl` formatted references directly in the `.tex` file.
+This was to avoid clashes in citation keys between multiple articles in an issue.
+Later on, authors were instructed to provide a `.bib` file instead.
+During the conversion of legacy articles, it was discovered that some papers had references in both the `.tex` file and in separate `.bbl` or `.bib` files.
+Usually, the different files were equivalent, but sometimes references were only found in one source and in some cases the files contained conflicting information!
 While LaTeX can technically handle either `.bbl` or `.bib` formatted references R Markdown can only handle `.bib`.
@@
 
-- reading `bbl` chunks to produce a very close BibTeX equivalent.
-- title and author are usually mandatory fields.
-- URL, ISBN, publisher, pages and year are optional fields and will be enabled when available.
-- remaining data is stored in `"journal"` (internally) and `"publisher"` (when writing BibTeX file).
-- ignores commented LaTeX code.
-- provides a citation tracker
-- does bibliography aggregation XXX what is this?
+- extracting `bbl` chunks from a `.tex` file;
+- creating the mandatory title and author fields;
+- creating the optional URL, ISBN, publisher, pages and year fields, when available;
+- storing remaining information in `"journal"` (internally) and `"publisher"` (when writing BibTeX file);
+- ignoring commented LaTeX code;
+- tracking citations included in the document, and
+- aggregating references from embedded `bbl` with references from supplementary `.bbl` and `.bib` files.
 
@@
 
-The authors wish to thank the [Google Summer of Code](https://summerofcode.withgoogle.com) program for financially supporting Abhishek's work on this project, and the R Project Organization for their support.
+The authors wish to thank the [Google Summer of Code](https://summerofcode.withgoogle.com) program for financially supporting Abhishek Ulayil's work on this project, and the R Project organization for their support.
+Heather Turner was supported by the Engineering and Physical Sciences Research Council [EPSRC EP/V052128/1] during this project.
 

### 2023-08-18 supplement/bibliography/rebib.tex
edits to bibliography supplement

@@
 % !TeX root = RJwrapper.tex
-\title{Working with bibliography in texor}
+\title{Handling bibliographies with texor}
 \author{by Abhishek Ulayil}
@@
 \section{Introduction}
-The \CRANpkg{rebib} package is a spun-off from \CRANpkg{texor} package.
+The \CRANpkg{rebib} package is a spin-off from the \CRANpkg{texor} package.
 
-The decision to do this is the fact that the bibliography section 
+The spin-off cam about because the bibliography section 
 in \CRANpkg{texor} package was expanding significantly, enough to deserve a dedicated space.
 
-Thus \CRANpkg{rebib} was born and has improved over the past few months to include more advanced features and utilities.
+Thus \CRANpkg{rebib} was born and it has improved over the past few months to include more advanced features and utilities.
 
@@
 
-In LaTeX, we have multiple ways to declare bibliography entries, which include embedding bibliography entries in the document itself or isolating that section to an external file commonly called a bbl file.
-Then we have the BibTeX format, which has its structure and is much superior in managing bibliographic entries.
+In LaTeX, we have multiple ways to declare bibliography entries, which include embedding entries formatted to a given style in the document itself or isolating that section to an external file commonly called a \file{.bbl} file.
+Then we have the BibTeX format, which is designed to store the full bibliographic information for each reference, making it superior for managing references.
 
 LaTeX has no problem dealing with multiple formats and works well with all these bibliographic formats.
-However, some web publishing journals like RJ web articles from RJournal do not support embedded bibliographies.
+However, some web publishing journals like R Markdown articles for the R Journal do not support embedded bibliographies.
 As a result, they need a BibTeX equivalent of the same.
@@
 The \CRANpkg{rebib} package converts embedded LaTeX bibliographies into a close  BibTeX equivalent.
-This package eases the author from the manual conversion of the documents and aids the texor project in automating the conversion of past RJournal articles to the RJ-web-article format.
+This package relieves the author from manual conversion of the documents and aids the project to convert legacy R Journal articles to the current R Markdown format.
 
-\subsection{features of rebib}
+\subsection{Features of rebib}
 
 \begin{itemize}
-  \item Reads bib chunks to produce a very close BibTeX equivalent
-  \item Title and author are usually mandatory fields
-  \item URL, ISBN, publisher, pages and year are optional fields and will be enabled when relevant
-  \item Rest of the data is stored in \verb|"journal"|(internally) and \verb|"publisher"|(when writing BibTeX file)
-  \item Ignores commented LaTeX code 
-  \item Citation tracker
-  \item Bibliography Aggregation
+  \item Extracts embedded bibliographic entries from a \file{.tex} file.
+  \item Creates the mandatory title and author fields.
+  \item Creates the optional URL, ISBN, publisher, pages and year fields, when available.
+  \item Stores the remaining information in \verb|"journal"| (internally) and \verb|"publisher"| (when writing BibTeX file).
+  \item Ignores commented LaTeX code. 
+  \item Tracks citations included in the document.
+  \item Aggregates embedded bibliographic entries with references from supplementary \file{.bbl} and \file{.bib} files.
 \end{itemize}
@@
 
-The magic behind \CRANpkg{rebib} is a parser which parses bibliographic data and assorts them according to the matching regular expressions.
+The magic behind \CRANpkg{rebib} is a parser which parses bibliographic data and sorts them according to the matching regular expressions.
 
@@
 
-This stage is a minor step where it reads the Embedded Bibliography from the LaTeX document. This step also includes Filtering out the commented code to avoid un-intended entries read.
+This stage is a minor step where it reads the Embedded Bibliography from the LaTeX document. This step also includes filtering out commented lines to avoid reading un-intended entries.
 
-Lastly, the data is broken down based on the LaTeX macro \verb|\\bibitem| as a marker for a new entry and this assorted data is exported to a variable.
+The bibliography is split into entries using the LaTeX macro \verb|\\bibitem| as a marker for a new entry and the resulting data is exported to a variable.
 
@@
 
-The first value to be parsed is the \verb|unique_id| also called the citation reference which is used to cite elements inside the article.
+The first value to be parsed is the \verb|unique_id|, also called the citation reference which is used to cite elements inside the article.
 Usually, this is in the first or second line of the whole entry.
@@
 
-This way the crucial elements of the bibliographic entry (unique\_id, author names and title ) are parsed out.
+This way the crucial elements of the bibliographic entry (unique\_id, author names and title) are parsed out.
 The remaining data is stored as \verb|journal| internally and \verb|publisher| when writing to a new BibTeX file.
@@
 A series of filters for ISBN, URL, pages and year fields are applied to search for relevant data from the remaining data.
-If the data is not available then it is set as NULL and will be skipped while writing the BibTeX file.
-There is a lot of filtering of common LaTeX elements and after that, the data remaining is stored in a structured format to be written to a file.
+If the data is not available, then it is set as \code{NULL} and will be skipped while writing the BibTeX file.
+There is a lot of filtering of common LaTeX elements and after that, the remaining data is stored in a structured format to be written to a file.
 
@@
 
-I expect the authors who are converting the document to take a look and check if there are any errors or missing values.
+Authors who are converting a bibliography from \file{bbl} should check if there are any errors or missing values.
 
 \section{General usage}
-In general to use \CRANpkg{rebib} package, it is quite easy with requiring only one function call.
+
+General use of the \CRANpkg{rebib} package is quite easy, requiring only one function call.
+
 \subsubsection{Using with RJournal article}
+
 In bibliography construction mode, \CRANpkg{rebib} will check for existing BibTeX files and will convert the embedded bibliography only when there are no linked BibTeX files in the RJ article.
@@
 
-\subsubsection{Using with any LaTeX/BBL file}
-The bibliography converter will convert any LaTex or BBL file given as a path. This way, one can generate BibTeX from the embedded bibliography or BBL file.
+\subsubsection{Using with any LaTeX/.bbl file}
+
+The bibliography converter will convert any LaTeX or \file{.bbl} file given as a path. This way, one can generate BibTeX from the embedded bibliography or \file{.bbl} file.
 \begin{verbatim}
@@
 \begin{itemize}
-\item Conversion of embedded natbib type bibliography used in RJournal
-\item Aggregation of bibliographic entries from the article as well as BibTeX
+\item Conversion of embedded natbib type bibliography entries used in the R Journal
+\item Aggregation of bibliographic entries from the article as well as separate BibTeX files
 \end{itemize}

### 2023-08-18 texor.Rmd
edits to bibliography supplement

@@
 
-- extracting `bbl` chunks from a `.tex` file;
+- extracting embedded bibliographic entries from a `.tex` file;
 - creating the mandatory title and author fields;
@@
 - tracking citations included in the document, and
-- aggregating references from embedded `bbl` with references from supplementary `.bbl` and `.bib` files.
+- aggregating references from embedded bibliographic entries with references from supplementary `.bbl` and `.bib` files.
 

### 2023-08-18 supplement/code-env/code-example.tex
Typeset code-env supplement
Updated supplementary materials
edits to code-env supplement

--- 2023-07-22 supplement/code-env/code-example.tex
@@
 \abstract{
-This is a small sample article to demonstrate usage of texor to convert code environments.
+This is a small sample article to demonstrate usage of \CRANpkg{texor} to convert code environments.
 }
@@
 
-Pandoc naturally converts verbatim environment easily, however the redefination of other commands such as example, example*, Sinput etc to verbatim does not work well in pandoc.
+Pandoc naturally converts the \code{verbatim} environment easily, however the redefinition of other commands such as \code{example}, \code{example*}, \code{Sinput} etc to \code{verbatim} does not work well in pandoc.
 
-Hence \CRANpkg{texor} package uses the stream editor to search find and replace matching code environments to verbatim before pandoc touches it.
+Hence, the \CRANpkg{texor} package uses a stream editor to find and replace matching code environments to \code{verbatim} before pandoc touches it.
 
@@
 
-Code Environment types are well summarized in the table \ref{table:1}
+Code Environment types are summarized in Table \ref{table:1}.
 
@@
  Example          & example & example* &  & & \\
- S.series         & Sin & Sout & Sinput & Soutput & Scode \\
- Special Verbatim & boxedverbatim & &  & & \\
+ S series         & Sin & Sout & Sinput & Soutput & Scode \\
+ Special verbatim & boxedverbatim & &  & & \\
 \hline
@@
 \subsection{Verbatim Series}
-While verbatim is naturally supported in pandoc,other extensions of verbatim environment
-like boxedverbatim from moreverb package \citep{moreverb}.
+While \code{verbatim} is naturally supported in pandoc, other extensions of the \code{verbatim} environment
+like \code{boxedverbatim} from the moreverb package \citep{moreverb} fall back to normal \code{verbatim}.
 % verbatim
 
-1. verbatim :
+1. \code{verbatim} :
 
@@
 
-2. boxedverbatim :
+2. \code{boxedverbatim} :
 
@@
 
-S series code environement is defined in Rjournal.sty file.
+The S series of code environments is defined in the \file{Rjournal.sty} file. Most of these are
+extensions of the \code{verbatim} environment and are represented as vanilla \code{verbatim} in the HTML output.
 
@@
 \subsection{Example series}
-Example series of code environment is defined in Rjournal.sty file.
 
-1. example :
+The Example series of code environments is defined in the \file{Rjournal.sty} file. Examples are
+extensions of \code{verbatim} environment and are represented as vanilla \code{verbatim} in the HTML output.
+
+1. \code{example} :
 
@@
 
-2. example* :
+2. \code{example*} :
 
@@
 \section{Code in Figure Environments}
-A small example of this is visble in \ref{code:example}
+A small example of code in a figure environment is demonstrated in Figure \ref{code:example} (named CodeBlock \ref{code:example} in the HTML). This is a common practice
+in R News articles as it used to add a boxed border around the code which looks 
+attractive. However, in web articles there isn't much advantage to it.
 
-\begin{figure*}[htbp]
-\begin{Sinput}
+\begin{figure}[htbp]
+\begin{center}
+\begin{verbatim}
 code_in_figure <- function() {
   if (pandoc_version >= 3) {
-    print("Code in Figure Supported")
+    print("code in figure supported")
   }
   else {
-    print("code in Figure not supported")
+    print("code in figure not supported")
   }
 }
-\end{Sinput}
+\end{verbatim}
 \caption{ Example Code inside Figure environment}
 \label{code:example}
-\end{figure*}
+\end{center}
+\end{figure}
+
 Pandoc v3 or greater \citep{pandoc} has a Figure object which allows non-image
-figures to be treated like one.
+figures to be treated like image figures. This is why the \CRANpkg{texor} package requires at least version 3 of pandoc.
 
@@
 \section{Code in Table Environments}
-We can use code environments in a table using minipage environments.
+We can use code environments in a table using minipage environments. This is not
+a common practice among LaTeX article authors, but a few articles had such complex
+structures. A few of examples are included here to demonstrate the capabilities of pandoc and the \CRANpkg{texor} package.
 
@@
     \hline
-    Language & Function Defination Syntax \\
+    Language & Function Definition Syntax \\
     \hline
@@
 \hline
-
-C & \begin{minipage}{0.75\textwidth}
-\vspace{1mm}
-\begin{example}
-int fun() {
-  printf("A function in C");
-  return 0;
-}
-
-\end{example}
-\end{minipage}\\
-\hline
 \end{tabular}
@@
 
-A similar arrangement can be had for figures/plots besides code environment.
-Table \ref{table:3} demonstrates a table with code and figure.
+A similar arrangement can be used to arrange figures/plots besides code environments.
+Table \ref{table:3} demonstrates a table with code and a figure.
 \begin{table}[htbp]
@@
   \end{tabular}
-  \caption{Code and Plot side by side}
+  \caption{Code and a plot side by side}
   \label{table:3}
@@
 
-Using inline code in LaTeX is possible using \verb|\verb| command.
-It would be repoduced similarly, as a Inline code element.
+Using inline code in LaTeX is possible using the \verb|\verb| command.
+It would be reproduced similarly, as an \code{Inline} code element.
 \begin{verbatim}
@@
 \end{verbatim}
-will be represented as \verb|x <- 1:100| in Inline format.
+will be represented as \verb|x <- 1:100| in \code{Inline} format.
 
 \section{Code chunks using Schunk}
-Code chunks within an \verb|Schunk| environment to demonstrate Input/Output
+Code chunks within an \verb|Schunk| environment to demonstrate Input/Output:
 
@@
 
-Similar arrangement can be had for Plots as well using figure environnment.
+A similar arrangement can be used for plots as well, using the figure environnment.
 
@@
 Output :\\
-\includegraphics[width=0.5\textwidth]{binom}
+\includegraphics[width=1\textwidth]{binom}
 \end{Schunk}
@@
 \begin{itemize}
-\item Almost all code environments in RJournal.
-\item Code Highlight for R language.
-\item Inline Code.
-\item Code in different environments like tables/figures.
+\item Almost all code environments in the R Journal.
+\item Code highlighting for the R language.
+\item Inline code.
+\item Code in different environments such as tables/figures.
 \end{itemize}
 
-\begin{thebibliography}{2}
+\begin{thebibliography}{99}
     \providecommand{\natexlab}[1]{#1}

### 2023-08-19 supplement/code-env/code-example.tex
edits to figure-env supplement

@@
 
-Code Environment types are summarized in Table \ref{table:1}.
+Code Environment types are summarized in Table~\ref{table:1}.
 
@@
 \section{Code in Figure Environments}
-A small example of code in a figure environment is demonstrated in Figure \ref{code:example} (named CodeBlock \ref{code:example} in the HTML). This is a common practice
+A small example of code in a figure environment is demonstrated in Figure~\ref{code:example} (named CodeBlock~\ref{code:example} in the HTML). This is a common practice
 in R News articles as it used to add a boxed border around the code which looks 
@@
 
-Table \ref{table:2} is an example of code environments within a table.
+Table~\ref{table:2} is an example of code environments within a table.
 
@@
 A similar arrangement can be used to arrange figures/plots besides code environments.
-Table \ref{table:3} demonstrates a table with code and a figure.
+Table~\ref{table:3} demonstrates a table with code and a figure.
 \begin{table}[htbp]

### 2023-08-19 supplement/lua-filters/lua-filters.tex
edits to lua-filters supplement

--- 2023-08-01 supplement/lua-filters/lua-filters.tex
@@
 pandoc supports intermediate modification of the Abstract Syntax Tree (AST) between
-the parsing and writing phase using filters. This supplement paper will highlight
-the use cases of such filters written in Lua Language on LaTeX,markdown and native AST.
+the parsing and writing phase using filters. This supplement paper highlights
+the use cases of such filters written in the Lua language on LaTeX, markdown and native AST.
 }
@@
 To fully understand pandoc filters, one must first grasp the document conversion process.
-A document in one format is read/parsed into an intermediate form.
-An abstract syntax tree (AST) is what pandoc calls it.
-A document writer is a specialized type-setter to read an AST and produces the contents in the chosen article format.
+A document in one format is read/parsed into an intermediate form, known as an abstract syntax tree (AST).
+A document writer is a specialized type-setter to read an AST and produce the contents in the chosen article format.
 Converting markdown to HTML, for example, will entail these pandoc operations.
-Figure\ref{fig:workflow} shows the conversion workflow withing pandoc.
+Figure~\ref{fig:workflow} shows the conversion workflow within pandoc.
 
@@
   \includegraphics[width=0.6\textwidth]{figures/workflow.png}
-  \caption{Conversion Workflow}
+  \caption{Conversion workflow.}
   \label{fig:workflow}
@@
 
-When specific elements require adjustment, the filters enter the picture. Assume you're converting markdown to HTML and want the page to have automated numbering for figures or tables. 
+When specific elements require adjustment, the Lua filters enter the picture. Assume you're converting markdown to HTML and want the page to have automated numbering for figures or tables. 
 If you use markdown, there is no system for automatically numbering figures/tables; to retain 1:1 conversion, the HTML output will also lack numbering.
@@
 Then there is the possibility of unlimited choices for customization, including every customization as an option in pandoc or the writer is not viable.
-To address this, the idea of filters emerged, in which you could change the AST partially and modify it before the writer could read it. As a result, you get the desired result \citep{pandocfilters}.
+To address this, the idea of filters emerged, with which you could modify the AST before the writer could read it to obtain the desired result \citep{pandocfilters}.
 
-\section{Need of lua filters}
+\section{Need of Lua filters}
 
-Continuing the example above lets create a dummy article, where we need to add
-numbering. One way could be to keep a counter of Images, and add a prefix \verb|"Figure X :"|
-to each figure caption. This will serve the purpose of numbering the Images in the end result.
+Continuing the example above, let's create a dummy article, where we need to add
+numbering. One way could be to keep a counter of images, and add a prefix \verb|"Figure X :"|
+to each figure caption. This will serve the purpose of numbering the images in the end result.
 
@@
 \end{verbatim}
-\caption{pandoc command without lua filter or extensions}
+\caption{pandoc command without lua filter or extensions.}
 \label{code:2}
@@
 
-Now If we convert the above markdown file to HTML5 using pandoc command \ref{code:2}, we get \ref{fig:1}
+Now if we convert the above markdown file to HTML5 using the pandoc command in Figure~\ref{code:2}, we get Figure~\ref{fig:1}.
 
@@
 \includegraphics[width=0.5\linewidth]{figures/example.png}
-\caption{Vanilla HTML5 output}
+\caption{Vanilla HTML5 output.}
 \label{fig:1}
@@
 As we can see, there is no figure numbering done automatically, which is generally
-the expected result. However if we want to include numbering to it, we would need
-to write a lua filter. This lua filter will modify the AST and make the changes we 
+the expected result. If we want to include numbering, we would need
+to write a Lua filter. This Lua filter will modify the AST and make the changes we 
 desire.
 
-We call a lua filter in the pandoc command \ref{code:2} using the \verb| --lua-filter name_of_filter.lua|
+We call a Lua filter in the pandoc command show in Figure~\ref{code:2} using the \verb| --lua-filter name_of_filter.lua|
 option in pandoc.
 
-Now we write a lua filter to manipulate the figures in \ref{code:3}
+In the next section we write a Lua filter to manipulate the figures in Figure~\ref{code:3}.
 
-\section{Writing pandoc lua filters}
+\section{Writing pandoc Lua filters}
 
@@
 \end{verbatim}
-\caption{A lua filter to add figure numbering}
+\caption{A Lua filter to add figure numbering.}
 \label{code:3}
@@
 
-The first step in writing a filter is to select a target type, that is, which block or Inline this filter will target.
-Following the selection, we name the function after the target type and each element in the argument as el.
+The first step in writing a filter is to select a target type, that is, which \verb|Block| or \verb|Inline| this filter will target.
+Following the selection, we name the function after the target type and each element in the argument as \verb|el|.
 For instance, in the case of our filter, it would be \verb|function Figure(el)|. Pandoc is smart enough to match the names of filter functions to the AST elements.
@@
 Now within the filter function we can assume that \verb|el| will contain the table object of a Figure element from the document.
-we can use many functions over it or \verb|Inlines| contained in it.
+We can use many functions over it or \verb|Inlines| contained in it.
 One such function to walk over all the elements inside the block is \verb|walk_block|.
-walk functions are great to check or count the presence of certain elements in a block.
-we use \verb|walk_block| to check if there are any Image inline elements within the Figure block.
-This is because after pandoc 3 \citep{pandoc} Figure blocks can now contain elements other than Images as well.
+Walk functions are a great way to check or count the presence of certain elements in a block.
+We use \verb|walk_block| to check if there are any \verb|Image| inline elements within the \verb|Figure| block.
+This is because after pandoc 3 \citep{pandoc} \verb|Figure| blocks can now contain elements other than \verb|Images| as well.
 
-If there is an Image then we append \verb|"Figure X :"| to the caption of the Figure element.
+If there is an \verb|Image| then we append \verb|"Figure X :"| to the caption of the \verb|Figure| element.
 and return the modified element.
 
-\section{Interpreting changes using lua filters}
+\section{Interpreting changes using Lua filters}
 
-The filter \ref{code:3} when included in the pandoc command will generate \ref{fig:2} 
-using the \ref{code:4} command.
+The filter in Figure~\ref{code:3} when included in the pandoc command will generate Figure~\ref{fig:2} 
+using the command in Figure~\ref{code:4}.
 
@@
 \end{verbatim}
-\caption{pandoc command with lua filter}
+\caption{pandoc command with Lua filter.}
 \label{code:4}
@@
 \includegraphics[width=0.5\linewidth]{figures/example-filtered.png}
-\caption{HTML5 output with desired filtering}
+\caption{HTML5 output with desired filtering.}
 \label{fig:2}
@@
 
-\section{Usage of pandoc lua filters in texor package}
-
-\section{Summary}
-
-\section{Supplementary materials}
-The example conversion files described here are included as supplementary materials
-in the Rjournal article of texor.
+\section{Usage of pandoc lua filters in the \CRANpkg{texor} package}
+Pandoc Lua filters are used in the \CRANpkg{texor} package to modify the AST for various markups.
+ Table~\ref{tab:filters} summarizes the use of each filter.
+\begin{table}
+\begin{tabular}{|l|p{10cm}|}
+\hline
+File Name & Description\\
+\hline
+\verb|abs_filter.lua| & {Filters out unicode \verb|182| character.}\\
+\hline
+\verb|bib_filter.lua| & Clears out the bibliography from the article itself as it will be added to the metadata.\\
+\hline
+\verb|R_code.lua| & Adds a class component \verb|'r'| to CodeBlocks for code highlighting. \\
+\hline
+\verb|image_filter.lua| & Fixes extensions for image paths without an extension or with \verb|.pdf| extension. \\
+\hline
+\verb|table_caption.lua| & Adds table numbering in the captions similar to the ones found in R Markdown and \CRANpkg{kable}-based tables. \\
+\hline
+\verb|image_caption.lua| & Adds Figure/Algorithm/CodeBlock/WideTable numbering as well as clears out residuals of tikz/algorithm images. \\
+\hline
+\verb|conversion_compat_check.lua| & This filter keeps a count of all \verb|Inline| and \verb|Block| elements and writes it to a yaml file. \\
+\hline
+\verb|equation_filter.lua| & Adds \CRANpkg{bookdown} style equation numbering to LaTeX equation/math environments.\\
+\hline
+\verb|bookdown_ref.lua| & Changes the format of links to \CRANpkg{bookdown} (markdown) style.\\
+\hline
+\verb|issue_checker.lua| & Searches and notifies any occurrence of unrecognized math commands.\\
+\hline
+\verb|find_pdf_files.lua| & This filter creates a list for all image PDF files included in the article.\\
+\hline
+\end{tabular}
+\caption{A summary of all Lua filters used in the \CRANpkg{texor} package (available in the \href{https://github.com/Abhi-1U/texor/tree/master/inst}{/inst} folder of the package).}
+\label{tab:filters}
+\end{table}
 

### 2023-08-19 supplement/metadata/metadata.tex
Updated supplementary materials
edits to metadata supplement

@@
 \section{Introduction}
-There is a difference in the structure of metadata and generally Rmarkdown has more
-meta-data entries than its LaTeX counterpart. So for the conversion, main task is to 
-transform the meta data and include it in the generated Rmarkdown file, matching the
-LaTeX data as closely as possible.
+There is a difference in the structure of metadata and generally R Markdown has more
+metadata entries than its LaTeX counterpart. So for the conversion, the main task is to 
+transform the metadata and include it in the generated R Markdown file, matching the
+LaTeX metadata as closely as possible.
 
-Pandoc when used with \verb|-s| or \verb|standalone| mode reads out some of the existing
-meta data. The extracted author meta-data are listed below:
+When used with the \verb|-s| or \verb|standalone| mode, pandoc reads out some of the existing
+meta data. The extracted metadata are listed below.
 
-\subsubsection{Metadata Read from LaTeX file}
+\subsubsection{Metadata read from LaTeX file}
 
 \begin{itemize}
-  \item Author Names 
-  \item Author address(affiliation)
+  \item Author names 
+  \item Author address (affiliation)
   \item ORCID number
@@
   \item Abstract
-  \item CRAN/BIO packages
+  \item CRAN/Bioconductor packages
 \end{itemize}
@@
 
-\subsubsection{Metadata Read from DESCRIPTION file}
+\subsubsection{Metadata read from DESCRIPTION file}
 
@@
   \item Slug
-  \item Volume and Issue
+  \item Volume and issue
 \end{itemize}
 
-Some details like the Volume and Issue were picked up from folder structure as well
-in some legacy articles where DESCRIPTION file was not available.
+Some metadata like the volume and issue were picked up from the folder structure for some legacy articles, where the DESCRIPTION file was not available.
 
@@
 \end{tabular}
-\caption{Other commands supported by \CRANpkg{texor} and Pandoc}
+\caption{Other commands supported by \CRANpkg{texor} and pandoc}
 \label{table:1}
@@
 
-Table \ref{table:2} represents all the ways one can use citations in LaTeX and
+Table~\ref{table:2} represents all the ways one can use citations in LaTeX and
 expect the generated output.
@@
 \end{tabular}
-\caption{Different citations in LaTeX}
+\caption{Different forms of citation in LaTeX.}
 \label{table:2}
@@
 \subsubsection{Footnotes}
+
+Footnotes do not work in \verb|tabular| environments in LaTeX but do work in R Markdown tables\footnote{Footnotes in paragraph text work in both LaTeX and R markdown}.
+Custom footnote numbering works in LaTeX but does not work in R Markdown\footnote[21]{Forever 21 :)}.
+These features are demonstrated in Table~\ref{table:3}.
+
 \begin{table}[htbp]
@@
  \hline
- \verb|\acronym{A}\footnote{This works in Web article}| & \acronym{A}\footnote{This works in Web article} \\ \hline
-\verb|\R\footnote[21]{Its 2 not 21 :(}| & \R\footnote[21]{Its 2 not 21 :(}  \\ \hline
+ \verb|\acronym{A}\footnote{This works in HTML}| & \acronym{A}\footnote{This works in HTML} \\ \hline
+\verb|\R\footnote[21]{HTML shows 2 rather than 21}| & \R\footnote[21]{HTML shows 2 rather than 21}  \\ \hline
 \end{tabular}
-\caption{Demo of footnotes}
+\caption{Demonstration of footnotes.}
 \label{table:3}
@@
 
-Footnotes do not work in \verb|tabular| environment in LaTeX but it does work in Rmarkdown\footnote{This works in LaTeX and  Web article}.
-Custom Footnote number works in LaTeX but does not work in web articles\footnote[21]{Forever 21 :)}.
-
-\subsubsection{BlockQuotes}
+\subsubsection{Block quotes}
 \begin{quote}
-This BlockQuote should Justify an example.
+This block quote should justify an example.
 \end{quote}
@@
 \subsubsection{Links}
-Most links work just fine using \verb|\href| and \verb|\url| command will work flawlessly.
+Most links work just fine using \verb|\href| and the \verb|\url| command will work flawlessly, as demonstrated in Table ~\ref{table:4}.
+
 \begin{table}[htbp]
@@
 
+\subsection{Colored text}
+Some commands to describe colored text also work during conversion, such as
+\verb|\textcolor{blue}{Blue Text}|  \textcolor{blue}{Blue Text}.
+
 \section{Summary}
@@
 \begin{itemize}
-\item Extracting Metadata and putting it together well.
-\item Handling Links, BlockQuotes, Citations and Footnotes.
+\item Extracting metadata and putting it together well.
+\item Handling links, block quotes, citations and footnotes.
 \end{itemize}

### 2023-08-21 supplement/figure-env/figure-example.tex

@@
 \abstract{
-This is a small sample article to demonstrate usage of texor to convert figure environments.
+This is a small sample article to demonstrate usage of \CRANpkg{texor} to convert figure environments.
 }
@@
 
-Images are an essential component in any article, However due to the differences in the support for various graphic formats between LaTeX and markdown/HTML we need to fallback on raster graphics. 
-It is well summarized in the Table \ref{table:1}.
+Images are an essential component of any article, however, due to the differences in support for various graphic formats between LaTeX and markdown/HTML we need to fall back on raster graphics. 
+The support for different image formats across markup languages is summarized in Table~\ref{table:1}.
 
@@
 \section{Image with width parameters}
-This section may contain a figure such as Figure~\ref{figure:rlogo}.
-This is the most basic example of figure.
+
+The following code includes an image with width parameters, producing the output in Figure~\ref{figure:rlogo}.
 
@@
 
+This is the most basic example of figure.
+
 \section{Images in PDF format}
 
-Image \ref{fig:normal} is a graphical representation of normal distribution.
+The following code includes an image in PDF format, shown in Figure \ref{fig:normal}.
 
@@
 \section{Multiple images}
- Pandoc v3 and above now supports a new Figure object \citep{pandoc} which supports multiple 
- images side by side or in a grid format.
+ Pandoc v3 and above now support a new \code{Figure} object \citep{pandoc} which supports multiple 
+ images side by side or in a grid format. The following subsections demonstrate these capabilities.
 %% Two images side by side
@@
 
-Here is a tikz image example in fig \ref{fig:tikz} adapted from \citep{casflow}.
+Figure \ref{fig:tikz} shows a tikz image adapted from \citep{casflow}.
 
-Another interesting aspect of including tikz image here is that you can modify the
+An interesting aspect of including a tikz image here is that you can modify the
 source and re-convert without making any other changes. Figure will get updated
@@
 \subsubsection{Tikz Code:}
-The figure  \ref{fig:tikz} is an example of how \pkg{texor} handles tikz images,
+The image in Figure  \ref{fig:tikz} is a graphical representation of how \pkg{texor} handles tikz images:
 
 \begin{itemize}
-\item First figures containing tikz images are isolated.
-\item Tikz libraries are fetched from the wrapper file,
-\item Tikz Code section is isolated into a standalone LaTeX file and compiled.
-\item The Compiled LaTeX file generates a PDF Image.
-\item This is then Converted again to PNG format as PDFs embedded in HTML wont look good.
-\item In the figure environment a line is added \verb|\includegraphics{tikz/somefigure.png}|
-\item During the conversion in pandoc, a lua filter removes extra/unncessary text from the rest of the figure
-environemnt.
+\item Figures containing tikz images are isolated.
+\item Tikz libraries are fetched from the wrapper file.
+\item The Tikz code section is isolated into a standalone LaTeX file and compiled.
+\item The compiled LaTeX file generates a PDF Image.
+\item This is converted to PNG format for embedding in the HTML output.
+\item \verb|\includegraphics{tikz/somefigure.png}| is added to the figure environment.
+\item A lua filter removes redundant text the figure environment.
 \end{itemize}
 
-If You use \pkg{texor} to convert your articles using \verb|texor::latex_to_web()| with \verb|temp_mode=TRUE|(it is TRUE by default).
+If you use \pkg{texor} to convert your articles using \verb|texor::latex_to_web()| with \verb|temp_mode=TRUE|(it is \verb|TRUE| by default).
 The resultant Rmarkdown/HTML file will not modify the contents of your LaTeX file.
@@
 
-we do support algorithm2e diagrams and images, these will be numbered differently
-and we strongly suggest to use "alg:" in labels for best results. 
-Here is an example of algorithm \ref{alg:how}, referenced from \cite{algoexample}
+Diagrams and images using the \code{algorithm2e} environment are supported, these will be numbered differently.
+We strongly suggest to use "alg:" in labels for best results. 
+Algorithm \ref{alg:how} is an example from the \CRANpkg{algorithm2e} vignette \cite{algoexample}.
 
@@
 
-Figures can also house non-image environments like CodeBlocks and WideTables. This is 
-the part where things start to look a bit different in the converted articles, CodeBlocks 
-in figure environments would have numbering and caption that goes "Figure x:". However, 
-when the same CodeBlock is converted to Rmarkdown, they are numbered as "CodeBlock x:".
-This distinguishes them from normal Images and matches with authors vision more appropriately.
+Figures can also house non-image environments like code blocks and wide tables. This is 
+one aspect where the converted articles differ between PDF and HTML formats. Code blocks 
+in figure environments would have numbering and caption of the form "Figure x:" in the PDF output. However, 
+when the code blocks are converted to R Markdown, they are numbered as "CodeBlock x:".
+This distinguishes them from normal images and matches with the authors' vision more appropriately.
 
-As for Inclusion of widetables, since it houses two different tabular environments with the same caption,
-it would be treated as two different tables by pandoc sharing the same caption. This would not be the best solution hence, \pkg{texor} package uses Figure enviornment to house the widetables.
+As for the inclusion of wide tables, since the figure houses two different tabular environments with the same caption,
+this would be treated by pandoc as two different tables sharing the same caption. This would not be the best solution hence, the \pkg{texor} package uses Figure environment to house the widetables.
 This way they are distinguished well from normal tables, also the numbering scheme is seperated from figures "WideTable x:".
@@
   if (pandoc_version >= 3) {
-    print("Code in Figure Supported")
+    print("code in figure supported")
   }
   else {
-    print("code in Figure not supported")
+    print("code in figure not supported")
   }

### 2023-08-21 supplement/math-env/math-example.tex

@@
 \abstract{
-This is a small sample article to demonstrate usage of texor to convert math environments.
+This is a small sample article to demonstrate usage of the texor package to convert math environments.
 }
@@
 \section{Introduction}
-Math typesetting has always been LaTeX’s highlight feature, making it a de facto choice among academicians and researchers globally.
-However, as we proceed to our humble web interfaces, math is hard to describe traditionally.
-There have been advancements in JavaScript libraries to better Typeset and present math in web pages but not all LaTeX commands/math functions are available.
+Math typesetting has always been a highlight of LaTeX, making it a de facto choice among academics and researchers globally.
+However, in the move to humble web interfaces, math has been traditionally hard to describe.
+There have been advances in JavaScript libraries to better typeset and present math in web pages but not all LaTeX commands/math functions are available.
 
@@
 
-\subsubsection{Pandoc Handling, extensions}
-As \CRANpkg{texor} calls \CRANpkg{rmarkdown} to render the Rmarkdwon file into HTML,
-the \verb|rjtools::rjournal_web_article| template by default uses MathJax as the math engine. 
-Also we specify the Mathjax version in the metadata of the generated Rmarkdown file.
+\subsubsection{Pandoc handling, extensions}
+As \CRANpkg{texor} calls \CRANpkg{rmarkdown} to render the R Markdown file into HTML,
+the \verb|rjtools::rjournal_web_article| template uses MathJax as the math engine by default. 
+We also specify the Mathjax version in the metadata of the generated Rmarkdown file.
 
 \section{Inline math}
-One can define Inline Math in LaTeX using commands \verb|\(..\)| or \verb|$..$|.
+One can define inline math in LaTeX using commands \verb|\(..\)| or \verb|$..$|.
 It is also possible to use inline math within captions and  
@@
   \end{tabular}
-  \caption{InlineMath syntax and its output side by side}
+  \caption{Inline math syntax and its output side by side.}
   \label{table:1}
@@
 
-\section{Display Math}
+\section{Display math}
 
-Display Math refers to equations, almost all LaTeX equations can be rendered by
+Display math refers to equations typeset on separate lines rather than in line with the text. Almost all LaTeX equations can be rendered by
 MathJax.
@@
 \end{verbatim}
-This will render as \ref{eq:1}.
+The code above will render as Figure~\ref{eq:1}.
 
@@
 
-\section{Equation Numbering}
+\section{Equation numbering}
 In LaTeX your equations get numbered automatically unless you are using \verb|\[..\]| environment or \verb|\nonumber| to suppress
-numbering. Equation numbering works a bit differently in \CRANpkg{bookdown} (the base of RJ web article format) where
-it is mandatory to have a \verb|(\#eq:xx)| which is described in more detail in \citep{bookdown}.
+numbering. Equation numbering works a bit differently in \CRANpkg{bookdown} (the base of the R Journal web article format) where
+it is mandatory to have a \verb|(\#eq:xx)| which is described in more detail in \citet{bookdown}.
 
-The \CRANpkg{texor} package eases authors from manually adding \verb|(\#eq:xx)| to equations in Rmarkdown
-by using pandoc Lua filter to convert existing \verb|\label{..}| in the equations to \verb|(\#eq:xx)|
-during conversion. Equation \ref{eq:xe} shows such a use case of equation being numbered by its label in Rmakrdown
+The \CRANpkg{texor} package relieves authors from manually adding \verb|(\#eq:xx)| to equations in R Markdown
+by using a pandoc Lua filter to convert existing \verb|\label{..}| in the equations to \verb|(\#eq:xx)|
+during conversion. Equation~\ref{eq:xe} shows such a use case of an equation being numbered by its label in R Markdown
 as well as LaTeX.
@@
 
-\section{Custom Math Commands}
+\section{Custom math commands}
 
@@
 To ease the frequent use of the command, some custom math commands that contain LaTeX math commands. 
-However, This will not function if you use functions/environments from CTAN packages or 
+However, this will not work if you use functions/environments from CTAN packages or 
 the ones that are out-of-the-box functions which do not render in HTML/markdown.
 
-\paragraph{Acceptable}
+\paragraph{Renders correctly in HTML}
 
@@
 
-\paragraph{Not acceptable}
+\paragraph{Does not render correctly in HTML}
 
@@
 
-\section{Unsupported LaTeX Commands}
+\section{Unsupported LaTeX commands}
 
 Altough MathJax does a good job of supporting most LaTeX math functions,
-some functions do not work like \verb|\bm| \verb|\boldmath|, instead use \verb|\mathbf|.
+some functions do not currently work. Common examples are \verb|\bm| and \verb|\boldmath|, for which \verb|\mathbf| can be used instead.
 
@@
 
-In summary the \CRANpkg{texor} package with the help of pandoc and MathJax supports:
+In summary the \CRANpkg{texor} package, with the help of pandoc and MathJax supports:
 \begin{itemize}
-\item some common math environments.
-\item Equation Numbering
+\item Common math environments.
+\item Equation numbering.
 \end{itemize}

### 2023-08-21 supplement/table-env/table-example.tex
almost first draft

@@
 \section{Introduction}
-Tables are commonly used in RJ articles to display data in a tabular format. However, there are differences in the way tables are handled by LaTeX and HTML.
+Tables are commonly used in R Journal articles to display data in a tabular format. However, there are differences in the way tables are handled by LaTeX and HTML.
 LaTeX tables have more customization and are usually optimized for printing, whereas the web articles need tables optimized for varying sizes of media.
 Pandoc converts most of the tables somewhat easily, but is unable to do well with table customization packages and complex tables.
-Some pandoc extensions are used in order to tackle them, they are : \verb|simple_tables|,\verb|pipe_tables|
+The \CRANpkg{texor} packages uses the pandoc extensions \verb|simple_tables| and \verb|pipe_tables| to tackle these difficult cases.
+
 \section{Generic tables}
-Normal LaTeX tables will be converted just fine, even if they include a bit of math and other elements.
-However do note that the conversion will be done to traditional or pandoc style markdown instead of Rmarkdown
-specific using \verb|knitr::kable(..)| function. Also any custom LaTeX commands or specific font charactars would
+Normal LaTeX tables will be converted just fine, even if they include a some math and other elements.
+However, note that the conversion will be done to traditional or pandoc-style markdown rather than R Markdown with the \verb|knitr::kable(..)| function.
+Also any custom LaTeX commands or specific font characters will
 not be supported.
@@
  \hline
- Graphics Format & LaTeX & Markdown & Rmarkdown & HTML \\
+ Graphics Format & LaTeX & Markdown & R Markdown & HTML \\
  \hline
@@
 \end{tabular}
-\caption{Image Format support in various Markup/Typesetting Languages}
+\caption{Image format support in various markup/typesetting languages.}
 \label{table:1}
@@
 \end{tabular}
-\caption{An Example Table with Multicolumn}
+\caption{An example table using \code{multicolumn}}
 \label{table:2}
@@
 
-The above example illustrates a multicolumn table \ref{table:2}, which the \CRANpkg{texor} package can handle through subroutines and pre-processing step to transform the LaTeX code.
+Table~\ref{table:2} illustrates a table that uses the \code{multicolumn} command, which the \CRANpkg{texor} package can handle through subroutines and pre-processing steps to transform the LaTeX code.
 
@@
 
-Also note that the stream editor is used to rename \verb|table*| environment to \verb|table| environment because the markdown/HTML will not support the specific changes table* environment will bring.
+Also note that the stream editor is used to rename \verb|table*| environment to \verb|table| environment because the HTML format is single column, so the asterisk indicating that the table should be drawn over the full width of the page is redundant in this case.
 
 \section{Complex tables}
-A complex table with various other elements like figures, math , code and so on, are also supported
+A complex table with various other elements like figures, math, code and so on, are also supported
 by \CRANpkg{texor}.  
@@
 \end{tabular}
-\caption{Image Format support in various Markup/Typesetting Languages}
+\caption{Image format support in various markup/typesetting languages}
 \label{table:3}
@@
 
-Widetable environments also gets some perks when converted by \CRANpkg{texor} package. 
-They get their own block and seperate numbering (although references are shared by figures).
+Widetable environments are treated specially when converted by the \CRANpkg{texor} package. 
+They get their own block and separate numbering (although references are shared by figures).
 
@@
 
-Here is a reference to the table \ref{table:4} and table \ref{table:5}.
+Here is a reference to Table~\ref{table:4} and Table~\ref{table:5}.
 \section{Summary}
@@
 \begin{itemize}
-\item some common table environments.
-\item wide tables.
-\item some multicolumn tables.
-\item inclusion of different environments like figure/code in tables.
+\item Some common table environments.
+\item Wide tables.
+\item Some tables with \code{multicolumn} commands.
+\item Environments such as \code{figure}/\code{code} in tables.
 \end{itemize}

### 2023-08-21 texor.Rmd

@@
 The R Journal is the primary open-access outlet for publications produced by the R community.
-It was started in 2009, evolving from a newsletter that ran from 2001, to become a more formal publication that encourages and provides credit for the documentation of statistical computing research.
+It was started in 2009, evolving from the R News newsletter that ran from 2001, to become a more formal publication that encourages and provides credit for the documentation of statistical computing research.
 

### 2023-08-21 supplement/figure-env/figure-example.tex
Updated supplementary to fix some #2 issues except math and lua filters

@@
 
-This is the most basic example of figure.
-
-\section{Images in PDF format}
-
-The following code includes an image in PDF format, shown in Figure \ref{fig:normal}.
-
-\begin{verbatim}
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5\textwidth]{normal}
-  \caption{PDF of a normal distribution}
-  \label{fig:normal}
-\end{figure}
-\end{verbatim}
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5\textwidth]{normal}
-  \caption{PDF of a normal distribution}
-  \label{fig:normal}
-\end{figure}
+This is the most basic example of figure. Similarly PDF figures will be included as rasterized images.
 
@@
 
+The \pkg{texor} package supports tikz images by rasterizing it and making it web friendly.
 Figure \ref{fig:tikz} shows a tikz image adapted from \citep{casflow}.
-
-An interesting aspect of including a tikz image here is that you can modify the
-source and re-convert without making any other changes. Figure will get updated
-in the generated html article.
 
@@
 Diagrams and images using the \code{algorithm2e} environment are supported, these will be numbered differently.
-We strongly suggest to use "alg:" in labels for best results. 
-Algorithm \ref{alg:how} is an example from the \CRANpkg{algorithm2e} vignette \cite{algoexample}.
+We strongly suggest to use "alg:" in labels for best results. As a part of the filtering requires "alg:" to number
+algorithm references seperately, otherwise they will share references with normal figures.
+
+Algorithm \ref{alg:how} is an example from the \pkg{algorithm2e} vignette \cite{algoexample}.
 
@@
 
-Figures can also house non-image environments like code blocks and wide tables. This is 
-one aspect where the converted articles differ between PDF and HTML formats. Code blocks 
-in figure environments would have numbering and caption of the form "Figure x:" in the PDF output. However, 
-when the code blocks are converted to R Markdown, they are numbered as "CodeBlock x:".
-This distinguishes them from normal images and matches with the authors' vision more appropriately.
-
-As for the inclusion of wide tables, since the figure houses two different tabular environments with the same caption,
-this would be treated by pandoc as two different tables sharing the same caption. This would not be the best solution hence, the \pkg{texor} package uses Figure environment to house the widetables.
-This way they are distinguished well from normal tables, also the numbering scheme is seperated from figures "WideTable x:".
-
-However one important thing to note is, the reference links and reference numbering would be shared with ones of the figures.
-So the changes that the package makes are cosmetic to captions.
-Like CodeBlock 1 below refers to \ref{code:example}.
-The number should have been 1 but instead it is 7.
+Figures can also house non-image environments like code blocks. Code blocks 
+in figure environments would share numbering with normal figures, similar to LaTeX.
+Here Figure \ref{code:example} refers to a code block.
 
@@
 \end{figure}
+
+\section{Limitations}
+
+The limitations of \CRANpkg{texor} package in figure handling are:
+
+\begin{itemize}
+\item Animations using the \pkg{animation} package in LaTeX wont be supported. The authors
+can include GiF equivalents of the same in the Rmarkdown/web article.
+\item Multi-page PDF files are currently not supported.
+\item Usage of subfigure,sclaebox,wrapfigure environments will not work as intended.
+\item Currently only \verb|[width=X.X\textwdith]| format is supported for defining the width
+in web article (default width would be 100\% in other cases).
+\item Other \verb|\inclduegraphics[..]| options will be ignored.
+\end{itemize}
+
 \section{Summary}
@@
 In summary the \CRANpkg{texor} package supports:
+
 \begin{itemize}
-\item Almost all image formats in LaTeX.
-\item Algorithm and tikz as well in some capacity.
 \item Multiple images in grid,side-by-side configuration.
 \item Image Captions with Numbering and Labelling.
+\item Algorithm2e and tikz figures.
 \end{itemize}

### 2023-08-21 supplement/table-env/table-example.tex
Updated supplementary to fix some #2 issues except math and lua filters

@@
 However, note that the conversion will be done to traditional or pandoc-style markdown rather than R Markdown with the \verb|knitr::kable(..)| function.
-Also any custom LaTeX commands or specific font characters will
+Also any graphic commands or specific font characters will
 not be supported.
@@
 
-\begin{itemize}
-    \item The Stream Editor modifies the \verb|\multicolumn{..}| to \verb|\multicolumnx{..}|
-    \item Then a LaTeX macro redefines the \verb|\multicolumnx{..}| to \verb|\multicolumn{---}|, which is accepted by pandoc.
-    \item Finally pandoc reads the table and transforms it to markdown.
-\end{itemize}
-
 Also note that the stream editor is used to rename \verb|table*| environment to \verb|table| environment because the HTML format is single column, so the asterisk indicating that the table should be drawn over the full width of the page is redundant in this case.
-
+-- ToDo : include more details about MultiRow and tablenotes
 \section{Complex tables}
@@
 \section{Wide tables}
+As for the inclusion of wide tables, they are represented as multiple tables with
+the first table housing the caption at the top. The numbering and references will be
+the same as LaTeX.
 
+Here is a reference to Table~\ref{table:4} and Table~\ref{table:5}.
 \begin{widetable}
@@
 
-Widetable environments are treated specially when converted by the \CRANpkg{texor} package. 
-They get their own block and separate numbering (although references are shared by figures).
-
 \begin{widetable}
@@
 
-Here is a reference to Table~\ref{table:4} and Table~\ref{table:5}.
+\section{Limitations}
+
+Limitaions of the \CRANpkg{texor} package in table handling includes:
+\begin{itemize}
+\item Usage of custom graphics/ characters like \verb|\circ| will not render in HTML properly,(using it in inline math might work).
+\item Inclusion of code blocks might not always work and it is best avoided.
+\item Currently only \verb|\\| is supported as the row end command/marker.
+\item Some table commands/environments might not work as expected.
+\item Nested tabular environments will not work as expected.
+\item Colored backgrounds in tables are not supported yet.
+\end{itemize}
+
 \section{Summary}
@@
 
-\begin{thebibliography}{2}
-    \providecommand{\natexlab}[1]{#1}
-    \providecommand{\url}[1]{\texttt{#1}}
-    \expandafter\ifx\csname urlstyle\endcsname\relax
-      \providecommand{\doi}[1]{doi: #1}\else
-      \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi
-
-\bibitem[Krewinkel, Lucero (2023)]{pandoc}
-A.~ Krewinkel and A.~ Lucero
-\newblock pandoc 3.0 Release notes
-\newblock \emph{pandoc} \penalty0 2023
-\newblock URL \url{https://pandoc.org/releases.html}
-
-\end{thebibliography}
-
 \address{%

### 2023-08-22 texor.Rmd
Compiled main paper

@@
   rjtools::rjournal_web_article:
-    web_only: yes
+    web_only: no
     self_contained: yes

### 2023-08-22 texor.tex
Compiled main paper

@@
 The reasons were to encourage better reproducibility of articles using dynamic documents, enable interactivity in articles, and to make the articles more accessible for vision-impaired readers.
-A resulting challenge was to explore whether legacy articles might be suitably converted into R Markdown before converting into HTML output.
-This paper describes the process to convert an R Journal LaTeX template article into an R Markdown article, and the two new R packages, \CRANpkg{texor} and \CRANpkg{rebib}, that can be used to achieve the conversion.
+A resulting challenge was to explore whether legacy articles might be suitably converted into HTML output.
+This paper describes the process to convert an R Journal article from LaTeX to HTML format via R Markdown, and the two new R packages, \CRANpkg{texor} and \CRANpkg{rebib}, that can be used to achieve the conversion.
 }
@@
 The R Journal is the primary open-access outlet for publications produced by the R community.
-It was born in 2008, evolving from a newsletter, that ran from 2001, into a more formal article publication to encourage documenting statistical computing research.
+It was started in 2009, evolving from the R News newsletter that ran from 2001, to become a more formal publication that encourages and provides credit for the documentation of statistical computing research.
 
-The format is regularly updated.
-Early articles were typeset using LaTeX (The LaTeX Project 2023), from a specific, but changing, template.
-This requires that code is separated from the documentation, and there is a chance that code chunks in the paper don't reproduce the results reported.
-With the emergence of dynamic document systems like R Markdown (Xie, Allaire, and Grolemund 2018) a tight-coupling of code and documentation is possible.
+The tooling behind the production of the R Journal is regularly updated.
+Early articles were typeset using LaTeX (The LaTeX Project 2023), based on a specific, but changing, template.
+Using LaTeX requires that code is separated from the documentation, and there is a chance that code chunks in the paper don't reproduce the results reported.
+With the emergence of dynamic document systems such as R Markdown (Xie, Allaire, and Grolemund 2018), a tight coupling of code and documentation is possible.
 Code chunks are dynamically executed when the document is typeset using a system like \CRANpkg{knitr} (Xie 2015), making reporting of computing research more reproducible.
 
-In 2019, with the help of funding from the R Consortium it was decided that it was time to update operations.
-One aspect of this was to change from LaTeX paper submissions to a more reproducible format, where code was embedded in the document, and the output could be both HTML and pdf.
+In 2019, with the help of funding from the R Consortium, work began to update operations.
+One aspect of this was to change from LaTeX article submissions to a more reproducible format, where code was embedded in the document, and the output could be both HTML and pdf.
 There are numerous benefits of HTML format:
@@
 \item
-  The format is more accessible to screen readers making the work more accessible to vision-impaired researchers.
+  The format is more accessible to screen readers and other assistive technologies, making the work more accessible to vision-impaired researchers.
+\item
+  HTML provides a more comfortable reading experience on mobile devices, which are increasing used as researchers work on the move and share work via social media.
+\item
+  Search engines can easily access the full text of articles, facilitating discovery of published articles.
 \end{enumerate}
 
-This latter point is a reason to consider converting all of the legacy articles into HTML.
+The latter points are motivations for converting all of the legacy R Journal articles into HTML.
 
-A key decision for creating conversion software was to decide to directly convert LaTeX to HTML, or PDF to HTML, or LaTeX to R Markdown, and then use the current journal tools to create the HTML.
+A key decision when designing the conversion software, was whether to convert the LaTeX source to HTML; the PDF output to HTML; or the LaTeX source to R Markdown, which would then be converted to HTML using the current journal tools.
 The latter approach was decided to be the most versatile and useful.
@@
 Web site architectures are also constantly evolving, and the emergence of \CRANpkg{distill} (Dervieux et al.
-2022) allows for the journal web site to optimally deliver R Markdown articles.
+2022) allows the journal web site to optimally deliver R Markdown articles.
 
-The \CRANpkg{rjtools} was developed to create articles using R Markdown for the R Journal, and to embed them into the journal web site.
-The packages described here, \CRANpkg{texor} and \CRANpkg{rebib} describes software to convert legacy LaTeX format articles into Rmarkdown, so that they can be rendered in HTML in the new web site.
+The \CRANpkg{rjtools} was developed to create articles using R Markdown for the R Journal and to embed them into the journal web site.
+The packages described here, \CRANpkg{texor} and \CRANpkg{rebib}, describe software to convert legacy LaTeX format articles into R Markdown, so that they can be rendered in HTML in the new web site.
 
@@
 
-The decision to convert to R Markdown format means that the final output to pdf and HTML will depend on Pandoc (MacFarlane 2023).
+The decision to convert to R Markdown format means that the final output to PDF and HTML will depend on Pandoc (MacFarlane 2023).
 Pandoc is a versatile document conversion program written in Haskell that is core to numerous documentation systems, including R Markdown and Quarto.
@@
 Pandoc can be used to do the conversion from LaTeX to R Markdown also.
-However, additional pre-processing needs to be done to handle special R Journal LaTeX styling.
-And further post-processing needs to be done to handle specific R Journal R Markdown styling.
+However, some pre-processing is necessary to handle special R Journal LaTeX styling.
+And further post-processing is necessary to handle specific R Journal R Markdown styling.
 The \CRANpkg{texor} package contains functionality to handle this pre- and post-processing of the document, in a workflow illustrated in Figure \ref{fig:workflow}.
@@
 LaTeX is very descriptive language, that allows authors substantial freedom for customization.
-Markdown (Gruber 2002), on which R Markdown is based, is more restrictive and was born to make it easier to create web pages without the distraction of a gazillion HTML tags.
+Markdown (Gruber 2002), on which R Markdown is based, is more restrictive and was designed to make it easier to create web pages without the distraction of a gazillion HTML tags.
 The beauty of Markdown is that it allows the author to focus on writing, without format cluttering the text.
@@
 If the code environment is not verbatim, then Pandoc will also try to process the actual code content as LaTeX commands and will likely lose details.
-It is better to convert these synonyms into \texttt{verbatim} environments this prior to passing the document to Pandoc.
+It is better to convert these synonyms into \texttt{verbatim} environments prior to passing the document to Pandoc.
 
@@
 \item
-  \texttt{stream\_editor()}: operates like the \texttt{sed} function in unix (Ritchie and Thompson 1978) and allows generic text pattern matching and replacing.
+  \texttt{stream\_editor()}: operates like the \texttt{sed} function in unix (Ritchie and Thompson 1978) and allows generic text pattern matching and replacement.
 \item
@@
 \item
-  \texttt{patch\_equations()}: coordinates various equation environments.
+  \texttt{patch\_equations()}: handles various equation environments.
 \item
-  \texttt{patch\_figure\_env()}: coordinates various figure environments.
+  \texttt{patch\_figure\_env()}: handles various figure environments.
 \item
-  \texttt{patch\_table\_env()}: coordinates table environments.
+  \texttt{patch\_table\_env()}: handles various table environments.
 \end{itemize}
@@
 
-Lua (Ierusalimschy, Figueiredo, and Filho 1996) is a programming language, that is light-weight, fast, ideal for procedural operations.
+Lua (Ierusalimschy, Figueiredo, and Filho 1996) is a programming language, that is light-weight and fast: ideal for procedural operations.
 It is embedded in many other applications to allow custom scripting for extensibility.
@@
 
-Algorithms as figures is supported, and the following description will yield the result in Figure \ref{fig:alghow}.
+Algorithms as figures are supported, and the following description will yield the result in Figure \ref{fig:alghow}.
 
@@
 The HTML output renders math using MathJax.
-This does mean that some functionality, like \texttt{\textbackslash{}bm}, \texttt{\textbackslash{}boldmath} and \texttt{\textbackslash{}mathbbm} not supported, and special definitions can be handled albeit in a limited capacity.
+This does mean that some functionality, like \texttt{\textbackslash{}bm}, \texttt{\textbackslash{}boldmath} and \texttt{\textbackslash{}mathbbm} are not supported, and special definitions can only be handled in a limited capacity.
 
@@
 2023), \CRANpkg{htmlTable} (Gordon, Gragg, and Konings 2022), \CRANpkg{tableHTML} (Boutaris, Zauchner, and Jomar 2023), \CRANpkg{tables} (Murdoch 2023) or \CRANpkg{DT} (Xie, Cheng, and Tan 2023).
-A benefit of special HTML table output is interactivity, that would allow the table to be sorted in different ways.
+A benefit of using conditional markup is that it can take advantage of HTML-specific features, such as sortable columns or paged tables.
 
@@
 
-Also note that the stream editor is used to rename \texttt{table*} environment to \texttt{table} environment because the markdown/HTML will not support the specific changes table* environment will bring.
+Also note that the stream editor is used to rename \texttt{table*} environment to \texttt{table} environment because the HTML format is single column, so the asterisk indicating that the table should be drawn over the full width of the page is redundant in this case.
 
@@
 
-The only function that a user necessarily needs is \texttt{latex\_to\_web()}. This creates the R Journal style R Markdown file from a given R Journal style LaTeX file.
-This is achieved by several sequential steps, \texttt{convert\_to\_markdown()}, \texttt{generate\_rmd()} and \texttt{produce\_html()}.
+The only function that a user will typically require is \texttt{latex\_to\_web()}. This creates the R Journal style R Markdown file from a given R Journal style LaTeX file.
+This is achieved by several sequential steps: \texttt{convert\_to\_markdown()}, \texttt{generate\_rmd()}, and \texttt{produce\_html()}.
 
@@
 
-For individuals who are interested in submitting their paper to the R Journal but have written their article using the legacy LaTeX format and wish to convert it to the new format, you can use the \texttt{latex\_to\_web()} function on your paper directory.
+For individuals who are interested in submitting their paper to the R Journal but have written their article using the legacy LaTeX format and wish to convert it to the current R Markdown format, you can use the \texttt{latex\_to\_web()} function on your paper directory.
 This will get you about 80\% of the way to an R Markdown version of your paper.
@@
 Typically bibliographies are generated during the processing of a LaTeX article using the BibTeX software (Feder 2006) operating on a \texttt{.bib} list of references.
-The R Journal template requires the inclusion of the \texttt{.bib} file.
+The current R Journal template requires the inclusion of the \texttt{.bib} file.
 But LaTeX actually uses a \texttt{.bbl} format for references, which is what BibTeX generates as an intermediate format during the article processing.
 
-During the conversion of legacy articles, especially in the earliest published articles, it was discovered that some papers were only accompanied by the intermediate \texttt{.bbl} formatted references and often these were included directly in the \texttt{.tex} file.
-Sometimes articles had both references in the \texttt{.tex} and separate \texttt{.bbl} or \texttt{.bib} files.
+The instructions for R Journal authors have changed over time.
+Initially, authors were instructed to run BibTeX on their article and to include the \texttt{.bbl} formatted references directly in the \texttt{.tex} file.
+This was to avoid clashes in citation keys between multiple articles in an issue.
+Later on, authors were instructed to provide a \texttt{.bib} file instead.
+During the conversion of legacy articles, it was discovered that some papers had references in both the \texttt{.tex} file and in separate \texttt{.bbl} or \texttt{.bib} files.
+Usually, the different files were equivalent, but sometimes references were only found in one source and in some cases the files contained conflicting information!
 While LaTeX can technically handle either \texttt{.bbl} or \texttt{.bib} formatted references R Markdown can only handle \texttt{.bib}.
@@
 \item
-  reading \texttt{bbl} chunks to produce a very close BibTeX equivalent.
+  extracting embedded bibliographic entries from a \texttt{.tex} file;
 \item
-  title and author are usually mandatory fields.
+  creating the mandatory title and author fields;
 \item
-  URL, ISBN, publisher, pages and year are optional fields and will be enabled when available.
+  creating the optional URL, ISBN, publisher, pages and year fields, when available;
 \item
-  remaining data is stored in \texttt{"journal"} (internally) and \texttt{"publisher"} (when writing BibTeX file).
+  storing remaining information in \texttt{"journal"} (internally) and \texttt{"publisher"} (when writing BibTeX file);
 \item
-  ignores commented LaTeX code.
+  ignoring commented LaTeX code;
 \item
-  provides a citation tracker
+  tracking citations included in the document, and
 \item
-  does bibliography aggregation XXX what is this?
+  aggregating references from embedded bibliographic entries with references from supplementary \texttt{.bbl} and \texttt{.bib} files.
 \end{itemize}
@@
 
-The authors wish to thank the \href{https://summerofcode.withgoogle.com}{Google Summer of Code} program for financially supporting Abhishek's work on this project, and the R Project Organization for their support.
+The authors wish to thank the \href{https://summerofcode.withgoogle.com}{Google Summer of Code} program for financially supporting Abhishek Ulayil's work on this project, and the R Project organization for their support.
+Heather Turner was supported by the Engineering and Physical Sciences Research Council {[}EPSRC EP/V052128/1{]} during this project.
 
@@
 University of Warwick\\%
-Newport, United Kingdom\\
+Coventry, United Kingdom\\
 %

### 2023-08-23 texor.Rmd
start on the check list of changes

@@
   rjtools::rjournal_web_article:
-    web_only: no
     self_contained: yes
@@
 
-## Logging the conversion
-
-XXX what should go here?
-
 # Using \CRANpkg{texor} {#texor}
 
-The package `texor` can be installed from CRAN using:
-
-```r
-install.packages("texor")
-```
-
-and the development version from https://github.com/Abhi-1U/texor. The website for the package, https://abhi-1u.github.io/texor, has vignettes documenting usage.
+The package `texor` can be installed from CRAN and the development version from https://github.com/Abhi-1U/texor.
+The website for the package, https://abhi-1u.github.io/texor, has vignettes documenting usage.
 
@@
 - Include your R code, to dynamically do the computing described in your article.
-
-XXX - Should we include various options/params for the function here ?
 
@@
 
-The package `rebib` can be installed from CRAN using:
+The package `rebib` can be installed from CRAN and the development version from https://github.com/Abhi-1U/rebib.
+The website for the package, https://abhi-1u.github.io/rebib, has vignettes documenting usage.
 
-```r
-install.packages("rebib")
-```
+# The process of converting all the legacy articles
 
-and the development version from https://github.com/Abhi-1U/rebib. The website for the package, https://abhi-1u.github.io/rebib, has vignettes documenting usage.
+# Known problems needing manual fixes
 
@@
 
-No shift to quarto yet.
-
-Known problems that need manual fixes.
+Since the shift to R Markdown, a new development for the open source community has emerged: quarto [@quarto].
+Although R Markdown would suggest a focus on R, it was always possible to include code chunks written in other languages.
+But Quarto makes this cleaner, and thus more appealing for non-R developers.
+It also provides a cleaner type-setting.
+At some point, the R Journal will likely shift to a Quarto template, which is reasonably straightforward, but for the present R Markdown is a suitable dynamic document delivery system for the R Journal.
 

### 2023-08-23 texor.tex
start on the check list of changes

@@
 LaTeX automatically numbers equations, unless specifically instructed not to.
-Equation numbering in R Markdown requires specific labeling using \texttt{(\textbackslash{}\#eq:xx)} as described in Yihui Xie (2023).
+Equation numbering in R Markdown requires specific labeling using \texttt{(\textbackslash{}\#eq:xx)} as described in Xie (2023).
 The \CRANpkg{texor} helps by adding the labeling using a Lua filter to convert the existing \texttt{\textbackslash{}label\{..\}} to \texttt{(\textbackslash{}\#eq:xx)}.
@@
 
-\hypertarget{logging-the-conversion}{%
-\subsection{Logging the conversion}\label{logging-the-conversion}}
-
-XXX what should go here?
-
 \hypertarget{texor}{%
@@
 
-The package \texttt{texor} can be installed from CRAN using:
-
-\begin{verbatim}
-install.packages("texor")
-\end{verbatim}
-
-and the development version from \url{https://github.com/Abhi-1U/texor}. The website for the package, \url{https://abhi-1u.github.io/texor}, has vignettes documenting usage.
+The package \texttt{texor} can be installed from CRAN and the development version from \url{https://github.com/Abhi-1U/texor}.
+The website for the package, \url{https://abhi-1u.github.io/texor}, has vignettes documenting usage.
 
@@
 \end{itemize}
-
-XXX - Should we include various options/params for the function here ?
 
@@
 
-The package \texttt{rebib} can be installed from CRAN using:
+The package \texttt{rebib} can be installed from CRAN and the development version from \url{https://github.com/Abhi-1U/rebib}.
+The website for the package, \url{https://abhi-1u.github.io/rebib}, has vignettes documenting usage.
 
-\begin{verbatim}
-install.packages("rebib")
-\end{verbatim}
+\hypertarget{the-process-of-converting-all-the-legacy-articles}{%
+\section{The process of converting all the legacy articles}\label{the-process-of-converting-all-the-legacy-articles}}
 
-and the development version from \url{https://github.com/Abhi-1U/rebib}. The website for the package, \url{https://abhi-1u.github.io/rebib}, has vignettes documenting usage.
+\hypertarget{known-problems-needing-manual-fixes}{%
+\section{Known problems needing manual fixes}\label{known-problems-needing-manual-fixes}}
 
@@
 
-No shift to quarto yet.
-
-Known problems that need manual fixes.
+Since the shift to R Markdown, a new development for the open source community has emerged: quarto (Allaire et al.
+2022).
+Although R Markdown would suggest a focus on R, it was always possible to include code chunks written in other languages.
+But Quarto makes this cleaner, and thus more appealing for non-R developers.
+It also provides a cleaner type-setting.
+At some point, the R Journal will likely shift to a Quarto template, which is reasonably straightforward, but for the present R Markdown is a suitable dynamic document delivery system for the R Journal.
 
@@
 \item
-  \texttt{figure-env}: Explains how the variety of figure definitions are handled in the conversion, including different image formats, numbering, captions, labeling, multiple images, and \texttt{tikz} (Josh Cassidy 2013) images.
+  \texttt{figure-env}: Explains how the variety of figure definitions are handled in the conversion, including different image formats, numbering, captions, labeling, multiple images, and \texttt{tikz} (Cassidy 2013) images.
 \item
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-quarto}{}}%
+Allaire, J. J., Charles Teague, Yihui Xie, and Christophe Dervieux. 2022. {``Quarto.''} Zenodo. \url{https://doi.org/10.5281/zenodo.5960048}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-tableHTML}{}}%
 Boutaris, Theo, Clemens Zauchner, and Dana Jomar. 2023. \emph{tableHTML: A Tool to Create HTML Tables}. \url{https://CRAN.R-project.org/package=tableHTML}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-casflow}{}}%
+Cassidy, Josh. 2013. \emph{{LaTeX Graphics using TikZ: A Tutorial for Beginners (Part 3)---Creating Flowcharts}}. Overleaf tutorials. \url{https://www.overleaf.com/learn/latex/}.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-casflow}{}}%
-Josh Cassidy. 2013. \emph{{LaTeX Graphics using TikZ: A Tutorial for Beginners (Part 3)---Creating Flowcharts}}. Overleaf tutorials. \url{https://www.overleaf.com/learn/latex/}.
-
 \leavevmode\vadjust pre{\hypertarget{ref-pandoc}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-bookdown}{}}%
+---------. 2023. \emph{{bookdown: Authoring Books and Technical Documents with R Markdown}}. \#equations. \url{https://bookdown.org/yihui/bookdown/markdown-extensions-by-bookdown.html}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-rmarkdown}{}}%
@@
 Xie, Yihui, Joe Cheng, and Xianying Tan. 2023. \emph{DT: A Wrapper of the JavaScript Library 'DataTables'}. \url{https://CRAN.R-project.org/package=DT}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-bookdown}{}}%
-Yihui Xie. 2023. \emph{{bookdown: Authoring Books and Technical Documents with R Markdown}}. \#equations. \url{https://bookdown.org/yihui/bookdown/markdown-extensions-by-bookdown.html}.
 

### 2023-08-25 texor.tex
pandoc citation

@@
 
-The decision to convert to R Markdown format means that the final output to PDF and HTML will depend on Pandoc (MacFarlane 2023).
+The decision to convert to R Markdown format means that the final output to PDF and HTML will depend on Pandoc (MacFarlane, Krewinkel, and Rosenthal 2023).
 Pandoc is a versatile document conversion program written in Haskell that is core to numerous documentation systems, including R Markdown and Quarto.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-pandoc}{}}%
-MacFarlane, John. 2023. {``Pandoc: A Universal Document Converter.''} https://pandoc.org.
+MacFarlane, John, Albert Krewinkel, and Jesse Rosenthal. 2023. {``Pandoc.''} https://github.com/jgm/pandoc.
 

### 2023-08-25 texor.Rmd
software version numbers

@@
 
-- \CRANpkg{texor}: https://abhi-1u.github.io/texor
-- \CRANpkg{rebib}: https://abhi-1u.github.io/rebib/
+- \CRANpkg{texor}: Version 1.2.0 https://abhi-1u.github.io/texor
+- \CRANpkg{rebib}: Version 0.3.0 https://abhi-1u.github.io/rebib/
 - This paper: https://github.com/Abhi-1U/texor-rjarticle

### 2023-08-25 texor.Rmd
manual changes for new users given more detail

@@
 
-- Edit the lines where figures are included. <!--Why? What will need changing?-->
-- Edit the tables. <!--If the simple conversion is insufficient?-->
-- Include your R code, to dynamically do the computing described in your article.
+- Edit the lines where figures are included.
+The conversion will create HTML code to define the image.
+This should be changed into Markdown description `![Caption](Image)` or using `knitr::include_graphic(Image)` in an R code chunk, in order for both HTML and PDF versions of the paper to be created when the `rjtools` article template is knitted.
+<!--Why?
+What will need changing?-->
+- Edit the tables for aesthetics.
+The conversion will create Markdown tables, which will convert appropriately to HTML and PDF.
+However, to have more control and to create more elegant tables using `knitr::kable` provides finer level control.
+<!--If the simple conversion is insufficient?-->
+- Include your R code, to dynamically do the computing described in your article. If any code block is time-consuming to complete, saving intermediate output, or caching would be recommended.
 

### 2023-08-25 texor.tex
software version numbers
manual changes for new users given more detail

@@
   Edit the lines where figures are included.
+The conversion will create HTML code to define the image.
+This should be changed into Markdown description \texttt{!{[}Caption{]}(Image)} or using \texttt{knitr::include\_graphic(Image)} in an R code chunk, in order for both HTML and PDF versions of the paper to be created when the \texttt{rjtools} article template is knitted.
 \item
-  Edit the tables.
+  Edit the tables for aesthetics.
+The conversion will create Markdown tables, which will convert appropriately to HTML and PDF.
+However, to have more control and to create more elegant tables using \texttt{knitr::kable} provides finer level control.
 \item
-  Include your R code, to dynamically do the computing described in your article.
+  Include your R code, to dynamically do the computing described in your article. If any code block is time-consuming to complete, saving intermediate output, or caching would be recommended.
 \end{itemize}
@@
 \item
-  \CRANpkg{texor}: \url{https://abhi-1u.github.io/texor}
+  \CRANpkg{texor}: Version 1.2.0 \url{https://abhi-1u.github.io/texor}
 \item
-  \CRANpkg{rebib}: \url{https://abhi-1u.github.io/rebib/}
+  \CRANpkg{rebib}: Version 0.3.0 \url{https://abhi-1u.github.io/rebib/}
 \item

### 2023-08-25 texor.Rmd
partially completed the processing of legacy articles section

@@
 
-# Known problems needing manual fixes
+The R Journal has been operating since 2009, and has published 682 articles through 2022 in addition to numerous editorials and news items.
+These LaTeX articles are the main focus of the conversions made with \CRANpkg{texor} and \CRANpkg{rebib}.
+
+## Batch processing
+
+The code is designed to process an issue at a time, by operating on all the files in the directory housing the issue.
+Each processing function, from pre-processing the text in the LaTeX file to post-processing accepts a directory name as the argument.
+The vignette [Introduction to texor](https://abhi-1u.github.io/texor/articles/introduction-to-texor.html) provides more details.
+
+## Known problems requiring manual fixes
+
+## Plan for release
+
+The HTML versions of all successfully converted legacy articles will be on the [R Journal web site](https://journal.r-project.org) late in 2023.
+The final step in the integration requires the R Markdown file to be knitted in place to capture the meta data for each article.
+
+## Reporting problems
+
+As a consequence of this project, the journal web site has been furnished with an additional link titled "CONTRIBUTE".
+Visiting this link will give you details on how to report a problem with an article that you are readin, or even to make a suggestion on improving operations.
 

### 2023-08-25 texor.tex
partially completed the processing of legacy articles section

@@
 
-\hypertarget{known-problems-needing-manual-fixes}{%
-\section{Known problems needing manual fixes}\label{known-problems-needing-manual-fixes}}
+The R Journal has been operating since 2009, and has published 682 articles through 2022 in addition to numerous editorials and news items.
+These LaTeX articles are the main focus of the conversions made with \CRANpkg{texor} and \CRANpkg{rebib}.
+
+\hypertarget{batch-processing}{%
+\subsection{Batch processing}\label{batch-processing}}
+
+The code is designed to process an issue at a time, by operating on all the files in the directory housing the issue.
+Each processing function, from pre-processing the text in the LaTeX file to post-processing accepts a directory name as the argument.
+The vignette \href{https://abhi-1u.github.io/texor/articles/introduction-to-texor.html}{Introduction to texor} provides more details.
+
+\hypertarget{known-problems-requiring-manual-fixes}{%
+\subsection{Known problems requiring manual fixes}\label{known-problems-requiring-manual-fixes}}
+
+\hypertarget{plan-for-release}{%
+\subsection{Plan for release}\label{plan-for-release}}
+
+The HTML versions of all successfully converted legacy articles will be on the \href{https://journal.r-project.org}{R Journal web site} late in 2023.
+The final step in the integration requires the R Markdown file to be knitted in place to capture the meta data for each article.
+
+\hypertarget{reporting-problems}{%
+\subsection{Reporting problems}\label{reporting-problems}}
+
+As a consequence of this project, the journal web site has been furnished with an additional link titled ``CONTRIBUTE''.
+Visiting this link will give you details on how to report a problem with an article that you are readin, or even to make a suggestion on improving operations.
 

### 2023-08-28 texor.Rmd
may still need to add more problems to the list

@@
 
-The code is designed to process an issue at a time, by operating on all the files in the directory housing the issue.
+The code is designed to process all the articles in an issue at a time, by operating on all the files in the directory housing the issue.
 Each processing function, from pre-processing the text in the LaTeX file to post-processing accepts a directory name as the argument.
+Generally, if a problem is encountered with one article, the processing will skip on to the next.
 The vignette [Introduction to texor](https://abhi-1u.github.io/texor/articles/introduction-to-texor.html) provides more details.
@@
 ## Known problems requiring manual fixes
+
+The problems encountered with some articles include:
+
+- The table format is often not as neat in the HTML as in the LaTeX. It can be made close to the original with extra effort in the R Markdown description.
+- Some of the figure files are missing - so the figure may need to be extracted from the pdf file and included with `knitr::include_graphics()`.
 

### 2023-08-28 texor.tex
may still need to add more problems to the list

@@
 
-The code is designed to process an issue at a time, by operating on all the files in the directory housing the issue.
+The code is designed to process all the articles in an issue at a time, by operating on all the files in the directory housing the issue.
 Each processing function, from pre-processing the text in the LaTeX file to post-processing accepts a directory name as the argument.
+Generally, if a problem is encountered with one article, the processing will skip on to the next.
 The vignette \href{https://abhi-1u.github.io/texor/articles/introduction-to-texor.html}{Introduction to texor} provides more details.
@@
 \subsection{Known problems requiring manual fixes}\label{known-problems-requiring-manual-fixes}}
+
+The problems encountered with some articles include:
+
+\begin{itemize}
+\tightlist
+\item
+  The table format is often not as neat in the HTML as in the LaTeX. It can be made close to the original with extra effort in the R Markdown description.
+\item
+  Some of the figure files are missing - so the figure may need to be extracted from the pdf file and included with \texttt{knitr::include\_graphics()}.
+\end{itemize}
 

### 2023-08-28 texor.Rmd
added a paragraph about alt-text

@@
 
+One of the useful elements when an article is in HTML format is that figures can be made richer with an `alt-text` element.
+In an R Markdown code chunk generating a figure, this can be set using `fig.alt`.
+Alternative text provides a textual alternative to non-text content in HTML documents, serves various purposes: read by screen readers are accessible to people with visual disabilities, is displayed in place of the image if it fails to load, and can serve to assist search engines.
+It is a work in progress to have these added by authors of new article submissions, and at some potin would be useful to add to the converted legacy articles.
+
 <!--Remember only the HTML is rendered, and the original PDF is not re-generated.

### 2023-08-28 texor.tex
added a paragraph about alt-text

@@
 
+One of the useful elements when an article is in HTML format is that figures can be made richer with an \texttt{alt-text} element.
+In an R Markdown code chunk generating a figure, this can be set using \texttt{fig.alt}.
+Alternative text provides a textual alternative to non-text content in HTML documents, serves various purposes: read by screen readers are accessible to people with visual disabilities, is displayed in place of the image if it fails to load, and can serve to assist search engines.
+It is a work in progress to have these added by authors of new article submissions, and at some potin would be useful to add to the converted legacy articles.
+
 Since the shift to R Markdown, a new development for the open source community has emerged: quarto (Allaire et al.

### 2023-08-28 texor.Rmd
list of manual fixes

@@
 
-The problems encountered with some articles include:
+The problems encountered with some articles are many and varied, and include:
 
@@
 - Some of the figure files are missing - so the figure may need to be extracted from the pdf file and included with `knitr::include_graphics()`.
+- Figures and equations may need some resizing.
+Sometimes the figure or equation might extend outside the text region in the HTML, which can be fixed with specifying preferred size, or breaks in the equation lines.
+- Alignment in multi-line equations may need adjusting.
+- Alignment of columns in a table may not be the same as the original, for example text might be center-aligned, instead of the original left-aligned.
+- Specialty LaTeX definitions included by an author for their article may not be recognized by Pandoc, leaving the original command as is, in the article.
 

### 2023-08-28 texor.tex
list of manual fixes

@@
   Some of the figure files are missing - so the figure may need to be extracted from the pdf file and included with \texttt{knitr::include\_graphics()}.
+\item
+  Figures and equations may need some resizing.
+Sometimes the figure or equation might extend outside the text region in the HTML, which can be fixed with specifying preferred size, or breaks in the equation lines.
+\item
+  Alignment in multi-line equations may need adjusting.
+\item
+  Alignment of columns in a table may not be the same as the original, for example text might be center-aligned, instead of the original left-aligned.
+\item
+  Specialty LaTeX definitions included by an author for their article may not be recognized by Pandoc, leaving the original command as is, in the article.
 \end{itemize}

### 2023-09-05 supplement/bibliography/rebib.tex
More up-to date info about the texor package. 1. code: Including R code chunks. 2. rebib: Limitations of rebib 3. figure: updated the pdf and html 4. luafilters: corrected some mistakes.

@@
 
+\section{Limitations}
+The limitations of \CRANpkg{rebib} package are:
+\begin{itemize}
+\item The \CRANpkg{rebib} package only supports embedded bibliography format suggested/ commonly
+used in R Journal.
+\item Encoding and character set issues might not work with rebib well. Try to use UTF-8 encoding 
+if possible.
+\item Bibliography aggregation won't fix existing issues like missing commas/brackets.
+\end{itemize}
 \section{Summary}

### 2023-09-05 supplement/code-env/code-example.tex
Updated supplementary to fix some #2 issues except math and lua filters
More up-to date info about the texor package. 1. code: Including R code chunks. 2. rebib: Limitations of rebib 3. figure: updated the pdf and html 4. luafilters: corrected some mistakes.

--- 2023-07-23 supplement/code-env/code-example.tex
@@
 Output :\\
-\includegraphics[width=0.5\textwidth]{binom}
+\includegraphics[width=1\textwidth]{binom}
 \end{Schunk}
+
+\section{Including 'R code chunks'}
+
+Including R code chunks in an article that is to be converted to R markdown
+can be a bit tricky as the \CRANpkg{knitr} package will interpret and execute the R code chunks as
+actual code bits.
+A workaround for this issue has been described in this \href{https://support.posit.co/hc/en-us/articles/360018181633-Including-verbatim-R-code-chunks-inside-R-Markdown}{posit support article}.
+Below is an example with the workaround adapted from \cite{rcodechunks}.
+\begin{example}
+`r ''````{r}
+plot(cars)
+```
+\end{example}
 
@@
 
-\begin{thebibliography}{2}
+\begin{thebibliography}{99}
     \providecommand{\natexlab}[1]{#1}
@@
 
+\bibitem[Stephens (2022)]{rcodechunks}
+Nathan Stephens
+\newblock Including verbatim R code chunks inside R Markdown
+\newblock \emph{Posit support forum}, 2022 
+\newblock URL \url{https://support.posit.co/hc/en-us/articles/360018181633-Including-verbatim-R-code-chunks-inside-R-Markdown}
 \end{thebibliography}

### 2023-09-05 supplement/lua-filters/lua-filters.tex
More up-to date info about the texor package. 1. code: Including R code chunks. 2. rebib: Limitations of rebib 3. figure: updated the pdf and html 4. luafilters: corrected some mistakes.

@@
 \abstract{
-pandoc supports intermediate modification of the Abstract Syntax Tree (AST) between
+Pandoc supports intermediate modification of the Abstract Syntax Tree (AST) between
 the parsing and writing phase using filters. This supplement paper highlights
@@
 To fully understand pandoc filters, one must first grasp the document conversion process.
-A document in one format is read/parsed into an intermediate form, known as an abstract syntax tree (AST).
-A document writer is a specialized type-setter to read an AST and produce the contents in the chosen article format.
+Pandoc will read/parse an article and then store it internally as an intermediate form, known as an abstract syntax tree (AST). 
+Then, A document writer(a specialized type-setter) will read the AST and produce the contents in the chosen article format. 
 Converting markdown to HTML, for example, will entail these pandoc operations. 
@@
 
-At this point, you might wish there was a way to add such a feature.
-Then there is the possibility of unlimited choices for customization, including every customization as an option in pandoc or the writer is not viable.
+At this point, you might wish there was a way to add such a feature. But including every customization as an option in pandoc or the writer is not viable. 
 To address this, the idea of filters emerged, with which you could modify the AST before the writer could read it to obtain the desired result \citep{pandocfilters}.
 
+Pandoc filters can technically be written in any language (called JSON filters) \footnote{More on this \href{https://pandoc.org/filters.html}{here}.}.
 \section{Need of Lua filters}
@@
 
-In pandoc, there are two sorts of elements: \verb|Blocks| and \verb|Inlines|.
-\verb|Blocks| are complicated constructions constructed from simple pieces (\verb|Inlines|).
-A \verb|Para|, for example, is a \verb|Block| made up of several \verb|Str| \verb|Inlines|.
+In pandoc, there are two sorts of elements: '\verb|Blocks|' and '\verb|Inlines|'.
+'\verb|Blocks|' are complicated constructions constructed from simpler pieces ('\verb|Inlines|').
+A '\verb|Para|', for example, is a '\verb|Block|' made up of several '\verb|Str|' '\verb|Inlines|'.
 
-The first step in writing a filter is to select a target type, that is, which \verb|Block| or \verb|Inline| this filter will target.
-Following the selection, we name the function after the target type and each element in the argument as \verb|el|.
+The first step in writing a filter is to select a target type, that is, which '\verb|Block|' or '\verb|Inline|' this filter will target.
+Following the selection, we name the function after the target type and each element in the argument as '\verb|el|'.
 For instance, in the case of our filter, it would be \verb|function Figure(el)|. Pandoc is smart enough to match the names of filter functions to the AST elements.
@@
 Now within the filter function we can assume that \verb|el| will contain the table object of a Figure element from the document.
-We can use many functions over it or \verb|Inlines| contained in it.
+We can use many functions over it or '\verb|Inlines|' contained in it.
 One such function to walk over all the elements inside the block is \verb|walk_block|.
 Walk functions are a great way to check or count the presence of certain elements in a block.
-We use \verb|walk_block| to check if there are any \verb|Image| inline elements within the \verb|Figure| block.
-This is because after pandoc 3 \citep{pandoc} \verb|Figure| blocks can now contain elements other than \verb|Images| as well.
+We use \verb|walk_block| to check if there are any '\verb|Image|' inline elements within the '\verb|Figure|' block.
+This is because after pandoc 3 \citep{pandoc} '\verb|Figure|' blocks can now contain elements other than '\verb|Images|' as well.
 
@@
 \hline
-\verb|image_caption.lua| & Adds Figure/Algorithm/CodeBlock/WideTable numbering as well as clears out residuals of tikz/algorithm images. \\
+\verb|image_caption.lua| & Adds Figure/Algorithm numbering as well as clears out residuals of tikz/algorithm images. \\
 \hline
@@
 \hline
-\verb|bookdown_ref.lua| & Changes the format of links to \CRANpkg{bookdown} (markdown) style.\\
+\verb|bookdown_ref.lua| & Corrects numbering for various elements.\\
 \hline
@@
 \hline
+\verb|widetable_patcher.lua| & This filter sets the representation for widetables.\\
+\hline
 \end{tabular}
+
 \caption{A summary of all Lua filters used in the \CRANpkg{texor} package (available in the \href{https://github.com/Abhi-1U/texor/tree/master/inst}{/inst} folder of the package).}

### 2023-09-06 supplement/math-env/math-example.tex
More Improvements to Supplementary materials 1. Tables: Multirow,longtables 2. Metadata: lists 3. Math: Equation reference correction

@@
 
-The core goal of the MathJax project is the development of its state-of-the-art, open source, JavaScript platform for display of mathematics. The key design goals are \citep{mathjax}:
+The core goal of the MathJax project is the development of its state-of-the-art, open source, JavaScript platform for display of mathematics. 
+The key advantages are \citep{mathjax}:
 \begin{itemize}
@@
     \item Support for LaTeX, MathML, and other equation markup directly in the HTML source.
-    \item An extensible, modular design with a rich API for easy integration into web applications.
     \item Support for accessibility, copy and paste, and other rich functionality.
     \item Interoperability with other applications and math-aware search.
-    \item Support for equation conversion outside a browser (e.g., preprocessing on a server).
 \end{itemize}
 
-\subsubsection{Pandoc handling, extensions}
 As \CRANpkg{texor} calls \CRANpkg{rmarkdown} to render the R Markdown file into HTML,
@@
 One can define inline math in LaTeX using commands \verb|\(..\)| or \verb|$..$|.
-It is also possible to use inline math within captions and  
 \begin{table}[htbp]
@@
 \end{verbatim}
-The code above will render as Figure~\ref{eq:1}.
+The code above will render as Equation~\ref{eq:1}.
 
@@
 All labels in bookdown must only contain alphanumeric characters, \verb|:|, \verb|-|, and/or \verb|/| as suggested in \citep{bookdown}.
+To accomadate this, the \pkg{texor} package has a lua filter implementation to correct and modify equation labels and references to the
+bookdown accepted format. For example lets say this equation has a non standard label
+\paragraph{LaTeX source code}
+\begin{example}
+\begin{equation}\label{binomial pdf}
+  f\left(k\right) = \binom{n}{k} p^k\left(1-p\right)^{n-k}
+
+\end{equation}
+The reference to above equation \eqref{binomial pdf}
+\end{example}
+
+\paragraph{Modified R markdown using \pkg{texor}}
+\begin{verbatim}
+$$\label{binomial pdf}
+  f\left(k\right) = \binom{n}{k} p^k\left(1-p\right)^{n-k}   (\#eq:binomial-pdf)$$
+
+The reference to above equation \@ref(eq:binomial-pdf)
+\end{verbatim}
+
+\paragraph{Rendering of the above example}
+\begin{equation}\label{binomial pdf}
+  f\left(k\right) = \binom{n}{k} p^k\left(1-p\right)^{n-k}
+\end{equation}
+
+The reference to above equation \eqref{binomial pdf}
 
@@
 
-%Also avoid custom math commands MathJaX will never be able to parse it. The reason being
-%MathJax renders math case by case and custom commands will already be removed by pandoc
-%during the conversion, hence MathJax will not be aware of such redefinations.
-To ease the frequent use of the command, some custom math commands that contain LaTeX math commands. 
-However, this will not work if you use functions/environments from CTAN packages or 
-the ones that are out-of-the-box functions which do not render in HTML/markdown.
+The existence of user-defined LaTeX commands intended for the Math environments will work as long as they do not contain non-math LaTeX commands or commands from other CTAN math packages.
 

### 2023-09-06 supplement/metadata/metadata.tex
More Improvements to Supplementary materials 1. Tables: Multirow,longtables 2. Metadata: lists 3. Math: Equation reference correction

@@
 
-\section{Miscellaneous}
+\section{Miscellaneous}\label{misc}
 
@@
 \verb|Table \ref{table:3}| & Table \ref{table:3}  \\ \hline
+\verb|Section \nameref{misc}| & Section \nameref{misc}\\ \hline
+\verb|\autoref{table:3}| & \autoref{table:3} \\ \hline
 \end{tabular}
@@
 
-\subsection{Colored text}
+\paragraph{Note}
+Certain reference commands like \verb|\nameref{}| and \verb|\autoref{}| will not work as intended in HTML.
+
+\subsubsection{Lists}
+
+An example of usage of lists in LaTeX and \pkg{texor} generated HTML.
+
+\paragraph{itemize}
+\begin{itemize}
+\item A
+\item B
+\end{itemize}
+\paragraph{enumerate}
+\begin{enumerate}
+\item A
+\item B
+\end{enumerate}
+\paragraph{description}
+\begin{description}
+\item[\pkg{rebib}] Convert and Aggregate Bibliographies
+\item[\pkg{texor}] Converting 'LaTeX' 'R Journal' Articles into 'RJ-web-articles'
+\end{description}
+
+\paragraph{Note:}
+Some options for lists such as special numbering schemes or markups to description
+item points might not work as intended and the generated HTML might be simple.
+
+\subsubsection{Colored text}
 Some commands to describe colored text also work during conversion, such as

### 2023-09-06 supplement/table-env/table-example.tex
pandoc citation
More Improvements to Supplementary materials 1. Tables: Multirow,longtables 2. Metadata: lists 3. Math: Equation reference correction

@@
 
-\section{Multicolumn tables}
+\section{Multicolumn and Multirow tables}
 \begin{table*}[htbp]
 \begin{center}
-\begin{tabular}{l | llll }
+\begin{tabular}{|l | llll |}
 \hline
- \multicolumn{1}{c |}{EXAMPLE} & \multicolumn{2}{c}{$X$} &
-\multicolumn{2}{c}{$Y$} \\
- \hline
-  & 1 & 2 & 1 & 2 \\
- EX1  & X11 & X12 &  Y11  & Y12 \\
- EX2  & X21 & X22 &  Y21  & Y22 \\
- EX3  & X31 & X32 &  Y31  & Y32 \\
- EX4  & X41 & X42 &  Y41  & Y42\\
- EX5  & X51 & X52 &  Y51  & Y52 \\
-\hline
+ \multicolumn{1}{|c |}{EXAMPLE} & \multicolumn{2}{c|}{$X$} &
+\multicolumn{2}{c|}{$Y$} \\\hline
+  & 1 & 2 & 1 & 2 \\\hline
+ EX1  & X11 & X12 &  Y11  & Y12 \\\hline
+ EX2  & X21 & X22 &  Y21  & Y22 \\\hline
+\multirow{ 2}{*}{EX3} & X31 & X32 &  Y31  & Y32 \\
+& X311 & X322 & Y311 & Y322\\\hline
+ \multirow{ 2}{*}{EX4}  & X41 & X42 &  Y41  & Y42\\
+ & X411 & X422 &  Y411  & Y422\\\hline
+ EX5  & X51 & X52 &  Y51  & Y52 \\\hline
 \end{tabular}
-\caption{An example table using \code{multicolumn}}
+\caption{An example table using \code{multicolumn} and \code{multirow}}
 \label{table:2}
@@
 
-Table~\ref{table:2} illustrates a table that uses the \code{multicolumn} command, which the \CRANpkg{texor} package can handle through subroutines and pre-processing steps to transform the LaTeX code.
+Table~\ref{table:2} illustrates a table that uses the \code{multicolumn} and \code{multirow} command, which the \CRANpkg{texor} package can handle through subroutines and pre-processing steps to transform the LaTeX code.
 
 Also note that the stream editor is used to rename \verb|table*| environment to \verb|table| environment because the HTML format is single column, so the asterisk indicating that the table should be drawn over the full width of the page is redundant in this case.
--- ToDo : include more details about MultiRow and tablenotes
+
 \section{Complex tables}
@@
 
+\section{Long Tables}
+
+Pandoc supports long tables from \pkg{longtable} CTAN package as well, here is an example as table \ref{table:6}.
+\begin{longtable}{|p{2cm}|p{2.5cm}|p{1cm}|p{2cm}|p{2cm}|p{2cm}|p{1cm}|}
+\caption{Table of Car data from mtcars dataset}
+\label{table:6}\\
+\hline
+\textbf{Manafacturer} & \textbf{Model and Make} & \textbf{MPG} & \textbf{Engine Cylinders} & \textbf{Engine Displacement (cu.in.)} & \textbf{Weight (1000 lbs)} & \textbf{Number of forward gears} \\
+\hline
+Mazda &RX4         &  21.0  & 6 & 160.0 & 2.620 &   4\\ \hline
+Mazda &RX4 Wag      & 21.0  & 6 & 160.0 & 2.875  &  4\\ \hline
+Datsun &710         & 22.8  & 4 &108.0 &2.320   & 4\\ \hline
+Hornet &4 Drive     & 21.4  & 6 &258.0 &3.215   & 3\\ \hline
+Hornet &Sportabout  & 18.7  & 8 &360.0 &3.440   & 3\\ \hline
+Valiant     &      &  18.1  & 6 &225.0 &3.460   & 3\\ \hline
+Duster &360        &  14.3  & 8 &360.0 &3.570   & 3\\ \hline
+Merc &240D         &  24.4  & 4 &146.7 &3.190   & 4\\ \hline
+Merc &230         &   22.8  & 4 &140.8 &3.150   & 4\\ \hline
+Merc &280          &  19.2  & 6 &167.6 &3.440   & 4\\ \hline
+Merc &280C        &   17.8  & 6 &167.6 &3.440   & 4\\ \hline
+Merc &450SE        &  16.4  & 8 &275.8 &4.070   & 3\\ \hline
+Merc &450SL        &  17.3  & 8 &275.8 &3.730   & 3\\ \hline
+Merc &450SLC       &  15.2  & 8 &275.8 &3.780   & 3\\ \hline
+Cadillac &Fleetwood & 10.4  & 8 &472.0 &5.250   & 3\\ \hline
+Lincoln &Continental& 10.4  & 8 &460.0 &5.424   & 3\\ \hline
+Chrysler &Imperial  & 14.7  & 8 &440.0 &5.345   & 3\\ \hline
+Fiat & 128          &  32.4 &  4 & 78.7 &2.200  &  4\\ \hline
+Honda &Civic        & 30.4  & 4 & 75.7 &1.615   & 4\\ \hline
+Toyota &Corolla     & 33.9  & 4 & 71.1 &1.835   & 4\\ \hline
+Toyota &Corona      & 21.5  & 4 &120.1 &2.465   & 3\\ \hline
+Dodge &Challenger   & 15.5  & 8 &318.0 &3.520   & 3\\ \hline
+AMC &Javelin        & 15.2  & 8 &304.0 &3.435   & 3\\ \hline
+Cheverolet &Camaro Z28 & 13.3 & 8& 350.0 &3.840   & 3\\ \hline
+Pontiac &Firebird   & 19.2  & 8 &400.0 &3.845   & 3\\ \hline
+Fiat &X1-9          & 27.3  & 4 & 79.0 &1.935   & 4\\ \hline
+Porsche& 914-2      & 26.0  & 4 &120.3 &2.140   & 5\\ \hline
+Lotus& Europa       & 30.4  & 4 & 95.1 &1.513   & 5\\ \hline
+Ford &Pantera L     & 15.8  & 8 &351.0 73.170   & 5\\ \hline
+Ferrari &Dino       & 19.7  & 6 &145.0 &2.770   & 5\\ \hline
+Maserati& Bora      & 15.0  & 8 &301.0 &3.570   & 5\\ \hline
+Volvo &142E         & 21.4  & 4 &121.0 &2.780   & 4\\ 
+\hline
+\end{longtable}
+
 \section{Limitations}
 
-Limitaions of the \CRANpkg{texor} package in table handling includes:
+Limitations of the \CRANpkg{texor} package in table handling includes:
 \begin{itemize}
@@
 \item Some common table environments.
-\item Wide tables.
-\item Some tables with \code{multicolumn} commands.
+\item Long and Wide Tables.
+\item Some tables with \code{multicolumn} and \code{multirow} commands.
 \item Environments such as \code{figure}/\code{code} in tables.
 \end{itemize}
+
+\section{Acknowledgements}
+
+Special thanks to Albert Krewinkel for his contributions towards earlier versions of Lua filters to add table numbering.
 

### 2023-09-06 texor.Rmd
whitespace change to test github action
try to sort dependency issue
testing workflow
using `\pkg{}` instead of `\CRANpkg{}`

@@
 A resulting challenge was to explore whether legacy articles might be suitably converted into HTML output.
-This paper describes the process to convert an R Journal article from LaTeX to HTML format via R Markdown, and the two new R packages, \CRANpkg{texor} and \CRANpkg{rebib}, that can be used to achieve the conversion.
+This paper describes the process to convert an R Journal article from LaTeX to HTML format via R Markdown, and the two new R packages, \pkg{texor} and \pkg{rebib}, that can be used to achieve the conversion.
 draft: true
@@
 Heather Turner was supported by the Engineering and Physical Sciences Research Council [EPSRC EP/V052128/1] during this project.
+Also in the initial stages of the development, Zola Batsaikhan carefully checked the conversion of articles, listing and fixing the problems after the conversion.
+This helped to refine the software to catch these problems and fix in the automatic conversion.
 

### 2023-09-06 texor.tex
using `\pkg{}` instead of `\CRANpkg{}`

@@
 A resulting challenge was to explore whether legacy articles might be suitably converted into HTML output.
-This paper describes the process to convert an R Journal article from LaTeX to HTML format via R Markdown, and the two new R packages, \CRANpkg{texor} and \CRANpkg{rebib}, that can be used to achieve the conversion.
+This paper describes the process to convert an R Journal article from LaTeX to HTML format via R Markdown, and the two new R packages, \pkg{texor} and \pkg{rebib}, that can be used to achieve the conversion.
 }
@@
 Heather Turner was supported by the Engineering and Physical Sciences Research Council {[}EPSRC EP/V052128/1{]} during this project.
+Also in the initial stages of the development, Zola Batsaikhan carefully checked the conversion of articles, listing and fixing the problems after the conversion.
+This helped to refine the software to catch these problems and fix in the automatic conversion.
 

### 2023-09-07 texor.Rmd
try to commit rendered output from GitHub action
do git add before commit
fix some typos

--- 2023-08-28 texor.Rmd
@@
 As a consequence of this project, the journal web site has been furnished with an additional link titled "CONTRIBUTE".
-Visiting this link will give you details on how to report a problem with an article that you are readin, or even to make a suggestion on improving operations.
+Visiting this link will give you details on how to report a problem with an article that you are reading, or even to make a suggestion on improving operations.
 
@@
 Alternative text provides a textual alternative to non-text content in HTML documents, serves various purposes: read by screen readers are accessible to people with visual disabilities, is displayed in place of the image if it fails to load, and can serve to assist search engines.
-It is a work in progress to have these added by authors of new article submissions, and at some potin would be useful to add to the converted legacy articles.
+It is a work in progress to have these added by authors of new article submissions, and at some point would be useful to add to the converted legacy articles.
 

### 2023-09-07 texor.Rmd
don't hard code date in Rmd

@@
 title: "Converting LaTeX Legacy R Journal Articles into R Markdown Articles using texor and rebib"
-date: "2022-11-09"
+date: "`r Sys.Date()`"
 abstract: >

### 2023-09-08 supplement/table-env/table-example.tex
Using DiagrammeR for mermaid diagram (not working) fixes to supplementary material

@@
 Hornet &Sportabout  & 18.7  & 8 &360.0 &3.440   & 3\\ \hline
-Valiant     &      &  18.1  & 6 &225.0 &3.460   & 3\\ \hline
+Plymouth      &   Valiant   &  18.1  & 6 &225.0 &3.460   & 3\\ \hline
 Duster &360        &  14.3  & 8 &360.0 &3.570   & 3\\ \hline
@@
 Lotus& Europa       & 30.4  & 4 & 95.1 &1.513   & 5\\ \hline
-Ford &Pantera L     & 15.8  & 8 &351.0 73.170   & 5\\ \hline
+Ford &Pantera L     & 15.8  & 8 &351.0 &73.170   & 5\\ \hline
 Ferrari &Dino       & 19.7  & 6 &145.0 &2.770   & 5\\ \hline

### 2023-09-08 texor.Rmd
Using DiagrammeR for mermaid diagram (not working) fixes to supplementary material

@@
 
-```{r workflow, fig.cap="Workflow of the document conversion conducted by texor.", out.width="90%", fig.align='center'}
+```{r workflow1, fig.cap="Workflow of the document conversion conducted by texor.", out.width="90%", fig.align='center', eval=knitr::is_latex_output()}
 knitr::include_graphics("workflow.png")
+```
+
+```{r workflow, fig.cap="Workflow of the document conversion conducted by texor.", out.width="90%", fig.align='center', eval=knitr::is_html_output()}
+#knitr::include_graphics("workflow.png")
+library(DiagrammeR)
+DiagrammeR::mermaid(diagram = "flowchart TD
+    A[Original Article] --> |Pre processing in R| B[Pre-Processed Article]
+    B --> |pandoc reader|C{AST} 
+    C -->|pandoc writer| D[Converted Article]
+    C -->|modify AST using lua filter| C{AST}")
 ```

### 2023-09-08 texor.tex
Revert "using `\pkg{}` instead of `\CRANpkg{}`"
This reverts commit 093d359e1e9cd1d49cdca625fb0dc3f668b9808a.
Using DiagrammeR for mermaid diagram (not working) fixes to supplementary material

--- 2023-08-28 texor.tex
@@
 
-\caption{Workflow of the document conversion conducted by texor.}\label{fig:workflow}
+\caption{Workflow of the document conversion conducted by texor.}\label{fig:workflow1}
 \end{figure}
@@
 As a consequence of this project, the journal web site has been furnished with an additional link titled ``CONTRIBUTE''.
-Visiting this link will give you details on how to report a problem with an article that you are readin, or even to make a suggestion on improving operations.
+Visiting this link will give you details on how to report a problem with an article that you are reading, or even to make a suggestion on improving operations.
 

### 2023-09-08 texor.R

@@ -7,8 +7,18 @@ library(plotly)
 library(ggplot2)
 
 
-## ----workflow, fig.cap="Workflow of the document conversion conducted by texor.", out.width="90%", fig.align='center'----
+## ----workflow1, fig.cap="Workflow of the document conversion conducted by texor.", out.width="90%", fig.align='center', eval=knitr::is_latex_output()----
 knitr::include_graphics("workflow.png")
+
+
+## ----workflow, fig.cap="Workflow of the document conversion conducted by texor.", out.width="90%", fig.align='center', eval=knitr::is_html_output()----
+#> #knitr::include_graphics("workflow.png")
+#> library(DiagrammeR)
+#> DiagrammeR::mermaid(diagram = "flowchart TD
+#>     A[Original Article] --> |Pre processing in R| B[Pre-Processed Article]
+#>     B --> |pandoc reader|C{AST}
+#>     C -->|pandoc writer| D[Converted Article]
+#>     C -->|modify AST using lua filter| C{AST}")
 
 
 ## ----alghow, fig.cap="How to write algorithms.", out.width="60%", fig.align='center'----

### 2023-09-08 texor.Rmd
using one r chunk for figure

--- 2023-09-07 texor.Rmd
@@
 ```{r workflow, fig.cap="Workflow of the document conversion conducted by texor.", out.width="90%", fig.align='center'}
+#
+if (knitr::is_html_output()) {
+  library(DiagrammeR)
+  DiagrammeR::mermaid(diagram = "flowchart TD
+    A[Original Article] --> |Pre processing in R| B[Pre-Processed Article]
+    B --> |pandoc reader|C{AST} 
+    C -->|pandoc writer| D[Converted Article]
+    C -->|modify AST using lua filter| C{AST}")
+}
+if (knitr::is_latex_output()) {
   knitr::include_graphics("workflow.png")
+}
 ```

### 2023-09-08 texor.R

@@ -7,18 +7,19 @@ library(plotly)
 library(ggplot2)
 
 
-## ----workflow1, fig.cap="Workflow of the document conversion conducted by texor.", out.width="90%", fig.align='center', eval=knitr::is_latex_output()----
-knitr::include_graphics("workflow.png")
-
-
-## ----workflow, fig.cap="Workflow of the document conversion conducted by texor.", out.width="90%", fig.align='center', eval=knitr::is_html_output()----
-#> #knitr::include_graphics("workflow.png")
-#> library(DiagrammeR)
-#> DiagrammeR::mermaid(diagram = "flowchart TD
-#>     A[Original Article] --> |Pre processing in R| B[Pre-Processed Article]
-#>     B --> |pandoc reader|C{AST}
-#>     C -->|pandoc writer| D[Converted Article]
-#>     C -->|modify AST using lua filter| C{AST}")
+## ----workflow, fig.cap="Workflow of the document conversion conducted by texor.", out.width="90%", fig.align='center'----
+#
+if (knitr::is_html_output()) {
+  library(DiagrammeR)
+  DiagrammeR::mermaid(diagram = "flowchart TD
+    A[Original Article] --> |Pre processing in R| B[Pre-Processed Article]
+    B --> |pandoc reader|C{AST} 
+    C -->|pandoc writer| D[Converted Article]
+    C -->|modify AST using lua filter| C{AST}")
+}
+if (knitr::is_latex_output()) {
+  knitr::include_graphics("workflow.png")
+}
 
 
 ## ----alghow, fig.cap="How to write algorithms.", out.width="60%", fig.align='center'----

### 2023-09-09 texor.Rmd
More specific mention of algorithm2e graphics instead of algorithm in general.

@@
 
-### Algorithms as figures
+### Algorithm2e graphics as figures
 
-Algorithms as figures are supported, and the following description will yield the result in Figure \@ref(fig:alghow).
+Algorithm2e graphics using the `algorithm2e` package in LaTeX  are supported, and the following description will yield the result in Figure \@ref(fig:alghow).
 
@@
 <img src="alghow.png" style="width:100.0%" />
-<p>initialization</p>
+
 <figcaption>Algorithm 1: How to write algorithms</figcaption>

### 2023-09-09 texor.tex
using one r chunk for figure
More specific mention of algorithm2e graphics instead of algorithm in general.

@@
 
-\caption{Workflow of the document conversion conducted by texor.}\label{fig:workflow1}
+\caption{Workflow of the document conversion conducted by texor.}\label{fig:workflow}
 \end{figure}
@@
 
-\hypertarget{algorithms-as-figures}{%
-\subsubsection{Algorithms as figures}\label{algorithms-as-figures}}
+\hypertarget{algorithm2e-graphics-as-figures}{%
+\subsubsection{Algorithm2e graphics as figures}\label{algorithm2e-graphics-as-figures}}
 
-Algorithms as figures are supported, and the following description will yield the result in Figure \ref{fig:alghow}.
+Algorithm2e graphics using the \texttt{algorithm2e} package in LaTeX are supported, and the following description will yield the result in Figure \ref{fig:alghow}.
 

### 2023-09-09 texor.Rmd
Bumped the rebib version with the CRAN version.

@@
 - \CRANpkg{texor}: Version 1.2.0 https://abhi-1u.github.io/texor
-- \CRANpkg{rebib}: Version 0.3.0 https://abhi-1u.github.io/rebib/
+- \CRANpkg{rebib}: Version 0.3.2 https://abhi-1u.github.io/rebib/
 - This paper: https://github.com/Abhi-1U/texor-rjarticle

### 2023-09-20 texor.Rmd
handled two of the issues

@@
 
-In 2019, with the help of funding from the R Consortium, work began to update operations.
-One aspect of this was to change from LaTeX article submissions to a more reproducible format, where code was embedded in the document, and the output could be both HTML and pdf.
+In 2019, with the help of funding from the R Consortium, work began to update operations to adopt an R Markdown template, that would produce both HTML and PDF versions of the paper.
+In mind, was also to develop, adopt and encourage good standards for R Journal articles.
+
 There are numerous benefits of HTML format:
@@
 The original motivation for the \CRANpkg{texor} and \CRANpkg{rebib} packages was to convert legacy LaTeX articles into HTML format to accessibility.
-Thus it only creates the R Markdown and an HTML version leaving the original PDF as published.
-This will also be used to produce HTML versions of newly accepted papers only submitted as LaTeX.
+Thus it only creates an `.Rmd` (R Markdown) file which is used to produce only an HTML version.
+The original PDF remains as published.
+The goals is use use the packages for future articles where the authors have chosen to submit only in the legacy LaTeX template.
 
-However, it could be useful to help authors convert to dynamic documents, by providing an initial R Markdown version of their LaTeX, that with some modification, such as including their R code for computations directly in the document, will produce both PDF nd HTML versions of their submitted paper.
+However, the software can also be used by authors themselves to convert to dynamic documents, by providing an initial R Markdown version of their LaTeX, that with some modification, such as including their R code for computations directly in the document, will produce both PDF nd HTML versions of their submitted paper.
 
-One of the useful elements when an article is in HTML format is that figures can be made richer with an `alt-text` element.
+Adopting the new R Markdown template also provides better standards for R Journal articles.
+It encourages authors to be more conscious that their work is reproducible.
+There is more standardization of file names and file structure, and this helps the editors and reviewers check papers.
+This in turn makes it easier to build the issues eliminating some common errors that authors make that cause compilation failure.
+
+There are many benefits of providing good standards, though, and an important one is that this can have positive spillover effects for the R community.
+Authors can become more aware of good practice.
+For example, one of the useful elements when an article is in HTML format is that figures can be made richer with an `alt-text` element.
 In an R Markdown code chunk generating a figure, this can be set using `fig.alt`.
@@
 It is a work in progress to have these added by authors of new article submissions, and at some point would be useful to add to the converted legacy articles.
+
+There are many other journals that have a similar backlog of legacy articles.
+A current and ongoing project is being conducted by @arxiv to convert their collection of articles from LaTeX (shared as PDF) to HTML, utilizing LaTeXML [@latexML].
+The work conducted for the R Journal might of interest to similar small journals that operate without paid staff.
+Converting legacy articles provides accessibility to a wealth of educational material.
 

### 2023-09-20 texor.tex
Bumped the rebib version with the CRAN version.
handled two of the issues

@@
 The original motivation for the \CRANpkg{texor} and \CRANpkg{rebib} packages was to convert legacy LaTeX articles into HTML format to accessibility.
-Thus it only creates the R Markdown and an HTML version leaving the original PDF as published.
-This will also be used to produce HTML versions of newly accepted papers only submitted as LaTeX.
+Thus it only creates an \texttt{.Rmd} (R Markdown) file which is used to produce only an HTML version.
+The original PDF remains as published.
+The goals is use use the packages for future articles where the authors have chosen to submit only in the legacy LaTeX template.
 
-However, it could be useful to help authors convert to dynamic documents, by providing an initial R Markdown version of their LaTeX, that with some modification, such as including their R code for computations directly in the document, will produce both PDF nd HTML versions of their submitted paper.
+However, the software can also be used by authors themselves to convert to dynamic documents, by providing an initial R Markdown version of their LaTeX, that with some modification, such as including their R code for computations directly in the document, will produce both PDF nd HTML versions of their submitted paper.
 
-One of the useful elements when an article is in HTML format is that figures can be made richer with an \texttt{alt-text} element.
+Adopting the new R Markdown template also provides better standards for R Journal articles.
+It encourages authors to be more conscious that their work is reproducible.
+There is more standardization of file names and file structure, and this helps the editors and reviewers check papers.
+This in turn makes it easier to build the issues eliminating some common errors that authors make that cause compilation failure.
+
+There are many benefits of providing good standards, though, and an important one is that this can have positive spillover effects for the R community.
+Authors can become more aware of good practice.
+For example, one of the useful elements when an article is in HTML format is that figures can be made richer with an \texttt{alt-text} element.
 In an R Markdown code chunk generating a figure, this can be set using \texttt{fig.alt}.
@@
 It is a work in progress to have these added by authors of new article submissions, and at some point would be useful to add to the converted legacy articles.
+
+There are many other journals that have a similar backlog of legacy articles.
+A current and ongoing project is being conducted by {``HTML as an Accessible Format for Papers''} (2023) to convert their collection of articles from LaTeX (shared as PDF) to HTML, utilizing LaTeXML ({``{LaTeXML}: A {LaTeX} to {XML/HTML/MathML} Converter''} 2023).
+The work conducted for the R Journal might of interest to similar small journals that operate without paid staff.
+Converting legacy articles provides accessibility to a wealth of educational material.
 
@@
 \item
-  \CRANpkg{rebib}: Version 0.3.0 \url{https://abhi-1u.github.io/rebib/}
+  \CRANpkg{rebib}: Version 0.3.2 \url{https://abhi-1u.github.io/rebib/}
 \item
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-arxiv}{}}%
+{``HTML as an Accessible Format for Papers.''} 2023. https://info.arxiv.org/about/accessible\_HTML.html.
+
 \leavevmode\vadjust pre{\hypertarget{ref-gt}{}}%
@@
 Ierusalimschy, Roberto, Luiz Henrique de Figueiredo, and Waldemar Celes Filho. 1996. {``Lua---an Extensible Extension Language.''} \emph{Software: Practice and Experience} 26 (6): 635--52.
+
+\leavevmode\vadjust pre{\hypertarget{ref-latexML}{}}%
+{``{LaTeXML}: A {LaTeX} to {XML/HTML/MathML} Converter.''} 2023. https://math.nist.gov/~BMiller/LaTeXML/.
 

### 2023-09-20 texor.Rmd
two more issues completed

@@
   In 2021 the R Journal made a change of templates for article writing to R Markdown instead of LaTeX.
-The reasons were to encourage better reproducibility of articles using dynamic documents, enable interactivity in articles, and to make the articles more accessible for vision-impaired readers.
+The reasons were to encourage better reproducibility of articles using dynamic documents, enable interactivity in articles, and to make the articles more accessible for print disabled readers.
 A resulting challenge was to explore whether legacy articles might be suitably converted into HTML output.
@@
 1. Articles can include interactive graphics and tables. 
-2. The format is more accessible to screen readers and other assistive technologies, making the work more accessible to vision-impaired researchers.
+2. The format is more accessible to screen readers and other assistive technologies, making the work more accessible to print disabled researchers. (See @W3C for current guidelines.)
 3. HTML provides a more comfortable reading experience on mobile devices, which are increasing used as researchers work on the move and share work via social media.
@@
 In an R Markdown code chunk generating a figure, this can be set using `fig.alt`.
-Alternative text provides a textual alternative to non-text content in HTML documents, serves various purposes: read by screen readers are accessible to people with visual disabilities, is displayed in place of the image if it fails to load, and can serve to assist search engines.
+Alternative text provides a textual alternative to non-text content in HTML documents, serves various purposes: read by screen readers are accessible to print disabled people, is displayed in place of the image if it fails to load, and can serve to assist search engines.
 It is a work in progress to have these added by authors of new article submissions, and at some point would be useful to add to the converted legacy articles.

### 2023-09-20 texor.tex
two more issues completed

@@
 In 2021 the R Journal made a change of templates for article writing to R Markdown instead of LaTeX.
-The reasons were to encourage better reproducibility of articles using dynamic documents, enable interactivity in articles, and to make the articles more accessible for vision-impaired readers.
+The reasons were to encourage better reproducibility of articles using dynamic documents, enable interactivity in articles, and to make the articles more accessible for print disabled readers.
 A resulting challenge was to explore whether legacy articles might be suitably converted into HTML output.
@@
 \item
-  The format is more accessible to screen readers and other assistive technologies, making the work more accessible to vision-impaired researchers.
+  The format is more accessible to screen readers and other assistive technologies, making the work more accessible to print disabled researchers.
+(See Sheri Byrne-Haber (2023) for current guidelines.)
 \item
@@
 In an R Markdown code chunk generating a figure, this can be set using \texttt{fig.alt}.
-Alternative text provides a textual alternative to non-text content in HTML documents, serves various purposes: read by screen readers are accessible to people with visual disabilities, is displayed in place of the image if it fails to load, and can serve to assist search engines.
+Alternative text provides a textual alternative to non-text content in HTML documents, serves various purposes: read by screen readers are accessible to print disabled people, is displayed in place of the image if it fails to load, and can serve to assist search engines.
 It is a work in progress to have these added by authors of new article submissions, and at some point would be useful to add to the converted legacy articles.
@@
 There are many other journals that have a similar backlog of legacy articles.
-A current and ongoing project is being conducted by {``HTML as an Accessible Format for Papers''} (2023) to convert their collection of articles from LaTeX (shared as PDF) to HTML, utilizing LaTeXML ({``{LaTeXML}: A {LaTeX} to {XML/HTML/MathML} Converter''} 2023).
+A current and ongoing project is being conducted by arXiv Accessibility Project (2023) to convert their collection of articles from LaTeX (shared as PDF) to HTML, utilizing LaTeXML (Miller and Ginev 2023).
 The work conducted for the R Journal might of interest to similar small journals that operate without paid staff.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-arxiv}{}}%
+arXiv Accessibility Project. 2023. {``HTML as an Accessible Format for Papers.''} \url{https://info.arxiv.org/about/accessible_HTML.html}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-tableHTML}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-arxiv}{}}%
-{``HTML as an Accessible Format for Papers.''} 2023. https://info.arxiv.org/about/accessible\_HTML.html.
-
 \leavevmode\vadjust pre{\hypertarget{ref-gt}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-latexML}{}}%
-{``{LaTeXML}: A {LaTeX} to {XML/HTML/MathML} Converter.''} 2023. https://math.nist.gov/~BMiller/LaTeXML/.
-
 \leavevmode\vadjust pre{\hypertarget{ref-pandoc}{}}%
 MacFarlane, John, Albert Krewinkel, and Jesse Rosenthal. 2023. {``Pandoc.''} https://github.com/jgm/pandoc.
+
+\leavevmode\vadjust pre{\hypertarget{ref-latexML}{}}%
+Miller, Bruce, and Deyan Ginev. 2023. {``{LaTeXML}: A {LaTeX} to {XML/HTML/MathML} Converter.''} \url{https://math.nist.gov/~BMiller/LaTeXML/}.
 
@@
 Ritchie, O. M., and K. Thompson. 1978. {``The UNIX Time-Sharing System.''} \emph{The Bell System Technical Journal} 57 (6): 1905--29. \url{https://doi.org/10.1002/j.1538-7305.1978.tb02136.x}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-W3C}{}}%
+Sheri Byrne-Haber, Charles LaPierre, David Fazio. 2023. {``W3C Accessibility Maturity Model.''} \url{https://www.w3.org/TR/maturity-model/}.
 

### 2023-09-20 texor.Rmd
phew! finally have a flow diagram. Layout with diagrammer is an absolute mess and takes a lot of work to get it looking ok

@@
 
-1. Articles can include interactive graphics and tables. 
-2. The format is more accessible to screen readers and other assistive technologies, making the work more accessible to print disabled researchers. (See @W3C for current guidelines.)
+1.
+Articles can include interactive graphics (to allow readers to query, re-focus, watch an animation, for example) and interactive tables (enabling re-sorting in different ways to better digest the information).
+2. It is the best format for accessibility [@Daisy]. Screen readers and other assistive technologies can deliver the content to print disabled readers. (See @W3C for current guidelines.)
 3. HTML provides a more comfortable reading experience on mobile devices, which are increasing used as researchers work on the move and share work via social media.
@@
 
-<!-- traditional HTML is the most universally accessible format - Australian Human Rights Commission https://humanrights.gov.au/our-work/disability-rights/world-wide-web-access-disability-discrimination-act-advisory-notes-ver and https://printdisability.org -->
-
-The latter points are motivations for converting all of the legacy R Journal articles into HTML.
+These provide the motivations for converting all of the legacy R Journal articles into HTML.
 
@@
 
-```{r workflow, fig.cap="Workflow of the document conversion conducted by texor.", out.width="90%", fig.align='center'}
-#
-if (knitr::is_html_output()) {
+```{r workflow, fig.cap="Workflow of the document conversion conducted by texor.
+(Note: AST is pandoc's abstract syntax tree.)", out.width="100%", fig.align='center', fig.width=4, fig.height=3, layout="l-body", fig.alt="Flow diagram showing original article pre-processed in R passed to pandoc which creates the AST and modifies using many lua filters then passed to pacdoc writer to produce the converted article."}
 library(DiagrammeR)
-  DiagrammeR::mermaid(diagram = "flowchart TD
-    A[Original Article] --> |Pre processing in R| B[Pre-Processed Article]
-    B --> |pandoc reader|C{AST} 
-    C -->|pandoc writer| D[Converted Article]
-    C -->|modify AST using lua filter| C{AST}")
-}
-if (knitr::is_latex_output()) {
-  knitr::include_graphics("workflow.png")
-}
+wf <- create_node_df(
+        n=8,
+        type=c("a", "b", "a", "b", 
+               "a", "b", "b", "a"),
+        label=c("original article", "pre-processing in R",
+                "pre-processed article", "pandoc reader",
+                "AST", 
+                "modify AST using lua filter",
+                "pandoc writer", 
+                "converted article"
+                ),
+        shape="rectangle",
+        width=c(1.3, 1.7, 1.9, 1.3, 0.5, 2.3, 1.2, 1.5),
+        fillcolor = c("darkseagreen2", "grey85", "darkseagreen2", "grey85", "darkseagreen2", "grey85", "grey85", "darkseagreen2"),
+        color = c("black", "grey85", "black", "grey85", "black", "grey85", "grey85", "black"),
+        fontcolor="black",
+        x=c(1,1,1,1,1,2,1,1),
+        y=c(1,2,3,4,5,5,6,6),
+        fontsize=14
+        )
+wf_arrows <- create_edge_df(from = c(1, 2, 3, 4, 5, 6, 5, 7),
+                            to = c(2, 3, 4, 5, 6, 5, 7, 8),
+                            color="black")
+wf_graph <- create_graph(wf, wf_arrows, 
+                   directed = TRUE, attr_theme = "tb")
+set.seed(7)
+render_graph(wf_graph, layout="fr", height=500)
 ```
@@
 In an R Markdown code chunk generating a figure, this can be set using `fig.alt`.
-Alternative text provides a textual alternative to non-text content in HTML documents, serves various purposes: read by screen readers are accessible to print disabled people, is displayed in place of the image if it fails to load, and can serve to assist search engines.
-It is a work in progress to have these added by authors of new article submissions, and at some point would be useful to add to the converted legacy articles.
+Alt-text provides a textual alternative to non-text content in HTML documents, and serves various purposes: screen readers can deliver the content to print disabled people, it is displayed in place of the image if it fails to load, and can serve to assist search engines.
+It is a work in progress: to encourage adoption by authors of new article submissions, and it would be useful to edit into the converted legacy articles at a later date.
 

### 2023-09-20 texor.tex
phew! finally have a flow diagram. Layout with diagrammer is an absolute mess and takes a lot of work to get it looking ok

@@
 \item
-  Articles can include interactive graphics and tables.
+  Articles can include interactive graphics (to allow readers to query, re-focus, watch an animation, for example) and interactive tables (enabling re-sorting in different ways to better digest the information).
 \item
-  The format is more accessible to screen readers and other assistive technologies, making the work more accessible to print disabled researchers.
+  It is the best format for accessibility (The Daisy Consortium 2023).
+Screen readers and other assistive technologies can deliver the content to print disabled readers.
 (See Sheri Byrne-Haber (2023) for current guidelines.)
@@
 
-The latter points are motivations for converting all of the legacy R Journal articles into HTML.
+These provide the motivations for converting all of the legacy R Journal articles into HTML.
 
@@
 
-{\centering \includegraphics[width=0.9\linewidth]{workflow}
+{\centering \includegraphics[width=1\linewidth]{texor_files/figure-latex/workflow-1}
 
@@
 
-\caption{Workflow of the document conversion conducted by texor.}\label{fig:workflow}
+\caption{Workflow of the document conversion conducted by texor. (Note: AST is pandoc's abstract syntax tree.)}\label{fig:workflow}
 \end{figure}
@@
 In an R Markdown code chunk generating a figure, this can be set using \texttt{fig.alt}.
-Alternative text provides a textual alternative to non-text content in HTML documents, serves various purposes: read by screen readers are accessible to print disabled people, is displayed in place of the image if it fails to load, and can serve to assist search engines.
-It is a work in progress to have these added by authors of new article submissions, and at some point would be useful to add to the converted legacy articles.
+Alt-text provides a textual alternative to non-text content in HTML documents, and serves various purposes: screen readers can deliver the content to print disabled people, it is displayed in place of the image if it fails to load, and can serve to assist search engines.
+It is a work in progress: to encourage adoption by authors of new article submissions, and it would be useful to edit into the converted legacy articles at a later date.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Daisy}{}}%
+The Daisy Consortium. 2023. {``Creating the Best Way to Read and Publish.''} \url{https://daisy.org}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-latex}{}}%

### 2023-09-20 texor.R

@@ -7,19 +7,35 @@ library(plotly)
 library(ggplot2)
 
 
-## ----workflow, fig.cap="Workflow of the document conversion conducted by texor.", out.width="90%", fig.align='center'----
-#
-if (knitr::is_html_output()) {
-  library(DiagrammeR)
-  DiagrammeR::mermaid(diagram = "flowchart TD
-    A[Original Article] --> |Pre processing in R| B[Pre-Processed Article]
-    B --> |pandoc reader|C{AST} 
-    C -->|pandoc writer| D[Converted Article]
-    C -->|modify AST using lua filter| C{AST}")
-}
-if (knitr::is_latex_output()) {
-  knitr::include_graphics("workflow.png")
-}
+## ----workflow, fig.cap="Workflow of the document conversion conducted by texor. (Note: AST is pandoc's abstract syntax tree.)", out.width="100%", fig.align='center', fig.width=4, fig.height=3, layout="l-body", fig.alt="Flow diagram showing original article pre-processed in R passed to pandoc which creates the AST and modifies using many lua filters then passed to pacdoc writer to produce the converted article."----
+library(DiagrammeR)
+wf <- create_node_df(
+        n=8,
+        type=c("a", "b", "a", "b", 
+               "a", "b", "b", "a"),
+        label=c("original article", "pre-processing in R",
+                "pre-processed article", "pandoc reader",
+                "AST", 
+                "modify AST using lua filter",
+                "pandoc writer", 
+                "converted article"
+                ),
+        shape="rectangle",
+        width=c(1.3, 1.7, 1.9, 1.3, 0.5, 2.3, 1.2, 1.5),
+        fillcolor = c("darkseagreen2", "grey85", "darkseagreen2", "grey85", "darkseagreen2", "grey85", "grey85", "darkseagreen2"),
+        color = c("black", "grey85", "black", "grey85", "black", "grey85", "grey85", "black"),
+        fontcolor="black",
+        x=c(1,1,1,1,1,2,1,1),
+        y=c(1,2,3,4,5,5,6,6),
+        fontsize=14
+        )
+wf_arrows <- create_edge_df(from = c(1, 2, 3, 4, 5, 6, 5, 7),
+                            to = c(2, 3, 4, 5, 6, 5, 7, 8),
+                            color="black")
+wf_graph <- create_graph(wf, wf_arrows, 
+                   directed = TRUE, attr_theme = "tb")
+set.seed(7)
[…]

### 2023-09-22 texor.Rmd
added the alt-text, used png for figure 1 bc its easier for now and put figures into subdirectory

@@
 And further post-processing is necessary to handle specific R Journal R Markdown styling.
-The \CRANpkg{texor} package contains functionality to handle this pre- and post-processing of the document, in a workflow illustrated in Figure \@ref(fig:workflow).
+The \CRANpkg{texor} package contains functionality to handle this pre- and post-processing of the document, in a workflow illustrated in Figure \@ref(fig:wf-fig).
 
-<!-- uses these filters heavily, even for simple tasks like choosing which supporting figures to copy.
-A `Lua` filter constructs a list of image paths, stores them momentarily and the package uses this data to copy images.
-However, \CRANpkg{texor} also includes many pre-processing functions that employ regular expressions.
-This is needed to convert some text and LaTeX commands that are not recognized or handled by Pandoc.
--->
-
-```{r workflow, fig.cap="Workflow of the document conversion conducted by texor.
-(Note: AST is pandoc's abstract syntax tree.)", out.width="100%", fig.align='center', fig.width=4, fig.height=3, layout="l-body", fig.alt="Flow diagram showing original article pre-processed in R passed to pandoc which creates the AST and modifies using many lua filters then passed to pacdoc writer to produce the converted article."}
+```{r workflow, eval=FALSE}
 library(DiagrammeR)
@@
 render_graph(wf_graph, layout="fr", height=500)
+```
+
+```{r wf-fig, fig.cap="Workflow of the document conversion conducted by texor.
+(Note: AST is pandoc's abstract syntax tree.)", out.width="100%", fig.align='center', fig.width=3, fig.height=5, layout="l-body", fig.alt="Flow diagram showing original article pre-processed in R passed to pandoc which creates the AST and modifies using many lua filters then passed to pacdoc writer to produce the converted article."}
+knitr::include_graphics("figures/wf.png")
 ```
@@
 
-```{r alghow, fig.cap="How to write algorithms.", out.width="60%", fig.align='center'}
-knitr::include_graphics("alghow.png")
+```{r alghow, fig.cap="How to write algorithms.", out.width="60%", fig.align='center', fig.alt="Illustration of an algorithm starting with Data and Result, followed by a while, and an if/else block."}
+knitr::include_graphics("figures/alghow.png")
 ```

### 2023-09-22 texor.tex
added the alt-text, used png for figure 1 bc its easier for now and put figures into subdirectory

@@
 And further post-processing is necessary to handle specific R Journal R Markdown styling.
-The \CRANpkg{texor} package contains functionality to handle this pre- and post-processing of the document, in a workflow illustrated in Figure \ref{fig:workflow}.
+The \CRANpkg{texor} package contains functionality to handle this pre- and post-processing of the document, in a workflow illustrated in Figure \ref{fig:wf-fig}.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{texor_files/figure-latex/workflow-1}
+{\centering \includegraphics[width=1\linewidth]{figures/wf}
 
@@
 
-\caption{Workflow of the document conversion conducted by texor. (Note: AST is pandoc's abstract syntax tree.)}\label{fig:workflow}
+\caption{Workflow of the document conversion conducted by texor. (Note: AST is pandoc's abstract syntax tree.)}\label{fig:wf-fig}
 \end{figure}
@@
 
-{\centering \includegraphics[width=0.6\linewidth]{alghow}
+{\centering \includegraphics[width=0.6\linewidth]{figures/alghow}
 

### 2023-09-22 texor.R

@@ -7,39 +7,43 @@ library(plotly)
 library(ggplot2)
 
 
-## ----workflow, fig.cap="Workflow of the document conversion conducted by texor. (Note: AST is pandoc's abstract syntax tree.)", out.width="100%", fig.align='center', fig.width=4, fig.height=3, layout="l-body", fig.alt="Flow diagram showing original article pre-processed in R passed to pandoc which creates the AST and modifies using many lua filters then passed to pacdoc writer to produce the converted article."----
-library(DiagrammeR)
-wf <- create_node_df(
-        n=8,
-        type=c("a", "b", "a", "b", 
-               "a", "b", "b", "a"),
-        label=c("original article", "pre-processing in R",
-                "pre-processed article", "pandoc reader",
-                "AST", 
-                "modify AST using lua filter",
-                "pandoc writer", 
-                "converted article"
-                ),
-        shape="rectangle",
-        width=c(1.3, 1.7, 1.9, 1.3, 0.5, 2.3, 1.2, 1.5),
-        fillcolor = c("darkseagreen2", "grey85", "darkseagreen2", "grey85", "darkseagreen2", "grey85", "grey85", "darkseagreen2"),
-        color = c("black", "grey85", "black", "grey85", "black", "grey85", "grey85", "black"),
-        fontcolor="black",
-        x=c(1,1,1,1,1,2,1,1),
-        y=c(1,2,3,4,5,5,6,6),
-        fontsize=14
-        )
-wf_arrows <- create_edge_df(from = c(1, 2, 3, 4, 5, 6, 5, 7),
-                            to = c(2, 3, 4, 5, 6, 5, 7, 8),
-                            color="black")
-wf_graph <- create_graph(wf, wf_arrows, 
-                   directed = TRUE, attr_theme = "tb")
-set.seed(7)
-render_graph(wf_graph, layout="fr", height=500)
+## ----workflow, eval=FALSE-----------------------------------------------------
+#> library(DiagrammeR)
+#> wf <- create_node_df(
+#>         n=8,
+#>         type=c("a", "b", "a", "b",
+#>                "a", "b", "b", "a"),
+#>         label=c("original article", "pre-processing in R",
+#>                 "pre-processed article", "pandoc reader",
+#>                 "AST",
+#>                 "modify AST using lua filter",
+#>                 "pandoc writer",
+#>                 "converted article"
[…]

### 2023-09-25 texor.tex
grammarly corrections

@@
 In 2021 the R Journal made a change of templates for article writing to R Markdown instead of LaTeX.
-The reasons were to encourage better reproducibility of articles using dynamic documents, enable interactivity in articles, and to make the articles more accessible for print disabled readers.
+The reasons were to encourage better reproducibility of articles using dynamic documents, enable interactivity in articles, and to make the articles more accessible for print-disabled readers.
 A resulting challenge was to explore whether legacy articles might be suitably converted into HTML output.
-This paper describes the process to convert an R Journal article from LaTeX to HTML format via R Markdown, and the two new R packages, \CRANpkg{texor} and \CRANpkg{rebib}, that can be used to achieve the conversion.
+This paper describes the process of converting an R Journal article from LaTeX to HTML format via R Markdown, and the two new R packages, \CRANpkg{texor} and \CRANpkg{rebib}, that can be used to achieve the conversion.
 }
@@
 \item
-  Articles can include interactive graphics (to allow readers to query, re-focus, watch an animation, for example) and interactive tables (enabling re-sorting in different ways to better digest the information).
+  Articles can include interactive graphics (to allow readers to query, re-focus, and watch an animation, for example) and interactive tables (enabling re-sorting in different ways to better digest the information).
 \item
   It is the best format for accessibility (The Daisy Consortium 2023).
-Screen readers and other assistive technologies can deliver the content to print disabled readers.
+Screen readers and other assistive technologies can deliver the content to print-disabled readers.
 (See Sheri Byrne-Haber (2023) for current guidelines.)
 \item
-  HTML provides a more comfortable reading experience on mobile devices, which are increasing used as researchers work on the move and share work via social media.
+  HTML provides a more comfortable reading experience on mobile devices, which are increasingly used as researchers work on the move and share work via social media.
 \item
-  Search engines can easily access the full text of articles, facilitating discovery of published articles.
+  Search engines can easily access the full text of articles, facilitating the discovery of published articles.
 \end{enumerate}
@@
 
-A key decision when designing the conversion software, was whether to convert the LaTeX source to HTML; the PDF output to HTML; or the LaTeX source to R Markdown, which would then be converted to HTML using the current journal tools.
+A key decision when designing the conversion software was whether to convert the LaTeX source to HTML; the PDF output to HTML; or the LaTeX source to R Markdown, which would then be converted to HTML using the current journal tools.
 The latter approach was decided to be the most versatile and useful.
@@
 
-In addition to article format, changes to the web site structure were important for delivering the publication.
-Web site architectures are also constantly evolving, and the emergence of \CRANpkg{distill} (Dervieux et al.
-2022) allows the journal web site to optimally deliver R Markdown articles.
+In addition to the article format, changes to the website structure were important for delivering the publication.
+Website architectures are also constantly evolving, and the emergence of \CRANpkg{distill} (Dervieux et al.
+2022) allows the journal website to optimally deliver R Markdown articles.
 
-The \CRANpkg{rjtools} was developed to create articles using R Markdown for the R Journal and to embed them into the journal web site.
-The packages described here, \CRANpkg{texor} and \CRANpkg{rebib}, describe software to convert legacy LaTeX format articles into R Markdown, so that they can be rendered in HTML in the new web site.
+The \CRANpkg{rjtools} was developed to create articles using R Markdown for the R Journal and to embed them into the journal website.
+The packages described here, \CRANpkg{texor} and \CRANpkg{rebib}, describe software to convert legacy LaTeX format articles into R Markdown, so that they can be rendered in HTML in the new website.
 
 The paper is organised as follows.
-Section 2 gives an overview of the conversion process, that includes pre-processing using regular expressions, post-processing using Lua filters, and handling of figures, tables and equations.
+Section 2 gives an overview of the conversion process, which includes pre-processing using regular expressions, post-processing using Lua filters, and handling of figures, tables and equations.
 Section 3 describes the \CRANpkg{texor} package that handles most of the conversion.
-Section 4 describes tools to do special handling of bibliography files.
+Section 4 describes tools for special handling of bibliography files.
 The supplementary materials have folders containing specific examples that can be used for understanding how the conversions are done.
@@
 
-While Pandoc can do most of the heavy-lifting, it cannot cope with all the freedom with which LaTeX documents are written.
+While Pandoc can do most of the heavy lifting, it cannot cope with all the freedom with which LaTeX documents are written.
 An example of this is with formatting of code.
@@
 Pandoc allows users to provide custom Lua filters to produce custom output formats.
-The \CRANpkg{texor} package handles post-processing of the R Markdown document into the special format for the R Journal using a suite of Lua filters.
+The \CRANpkg{texor} package handles the post-processing of the R Markdown document into the special format for the R Journal using a suite of Lua filters.
 
@@
 Figure definitions in LaTeX are many and varied!
-The standard, single figure definition with the \texttt{figure} environment and raster image format such as PNG or JPG, is handled by Pandoc.
+The standard, single-figure definition with the \texttt{figure} environment and raster image format such as PNG or JPG, is handled by Pandoc.
 It will convert:
@@
 The conversion will create HTML code to define the image.
-This should be changed into Markdown description \texttt{!{[}Caption{]}(Image)} or using \texttt{knitr::include\_graphic(Image)} in an R code chunk, in order for both HTML and PDF versions of the paper to be created when the \texttt{rjtools} article template is knitted.
+This should be changed into Markdown description \texttt{!{[}Caption{]}(Image)} or using \texttt{knitr::include\_graphic(Image)} in an R code chunk, in order for both HTML and PDF versions of the paper to be created when the \texttt{rjtools} article template is knitted.\\
 \item
@@
 
-The R Journal has been operating since 2009, and has published 682 articles through 2022 in addition to numerous editorials and news items.
+The R Journal has been operating since 2009 and has published 682 articles through 2022 in addition to numerous editorials and news items.
 These LaTeX articles are the main focus of the conversions made with \CRANpkg{texor} and \CRANpkg{rebib}.
@@
 Each processing function, from pre-processing the text in the LaTeX file to post-processing accepts a directory name as the argument.
-Generally, if a problem is encountered with one article, the processing will skip on to the next.
+Generally, if a problem is encountered with one article, the processing will skip to the next.
 The vignette \href{https://abhi-1u.github.io/texor/articles/introduction-to-texor.html}{Introduction to texor} provides more details.
@@
   Figures and equations may need some resizing.
-Sometimes the figure or equation might extend outside the text region in the HTML, which can be fixed with specifying preferred size, or breaks in the equation lines.
+Sometimes the figure or equation might extend outside the text region in the HTML, which can be fixed by specifying the preferred size, or breaks in the equation lines.
 \item
@@
 \item
-  Alignment of columns in a table may not be the same as the original, for example text might be center-aligned, instead of the original left-aligned.
+  Alignment of columns in a table may not be the same as the original, for example, the text might be center-aligned, instead of the original left-aligned.
 \item
@@
 
-The HTML versions of all successfully converted legacy articles will be on the \href{https://journal.r-project.org}{R Journal web site} late in 2023.
-The final step in the integration requires the R Markdown file to be knitted in place to capture the meta data for each article.
+The HTML versions of all successfully converted legacy articles will be on the \href{https://journal.r-project.org}{R Journal website} late in 2023.
+The final step in the integration requires the R Markdown file to be knitted in place to capture the metadata for each article.
 
@@
 
-As a consequence of this project, the journal web site has been furnished with an additional link titled ``CONTRIBUTE''.
+As a consequence of this project, the journal website has been furnished with an additional link titled ``CONTRIBUTE''.
 Visiting this link will give you details on how to report a problem with an article that you are reading, or even to make a suggestion on improving operations.
@@
 
-The original motivation for the \CRANpkg{texor} and \CRANpkg{rebib} packages was to convert legacy LaTeX articles into HTML format to accessibility.
+The original motivation for the \CRANpkg{texor} and \CRANpkg{rebib} packages was to convert legacy LaTeX articles into HTML format for accessibility.
 Thus it only creates an \texttt{.Rmd} (R Markdown) file which is used to produce only an HTML version.
 The original PDF remains as published.
-The goals is use use the packages for future articles where the authors have chosen to submit only in the legacy LaTeX template.
+The goal is to use the packages for future articles where the authors have chosen to submit only in the legacy LaTeX template.
 
-However, the software can also be used by authors themselves to convert to dynamic documents, by providing an initial R Markdown version of their LaTeX, that with some modification, such as including their R code for computations directly in the document, will produce both PDF nd HTML versions of their submitted paper.
+However, the software can also be used by authors themselves to convert to dynamic documents, by providing an initial R Markdown version of their LaTeX, that with some modification, such as including their R code for computations directly in the document, will produce both PDF and HTML versions of their submitted paper.
 
@@
 In an R Markdown code chunk generating a figure, this can be set using \texttt{fig.alt}.
-Alt-text provides a textual alternative to non-text content in HTML documents, and serves various purposes: screen readers can deliver the content to print disabled people, it is displayed in place of the image if it fails to load, and can serve to assist search engines.
+Alt-text provides a textual alternative to non-text content in HTML documents, and serves various purposes: screen readers can deliver the content to print-disabled people, it is displayed in place of the image if it fails to load, and can serve to assist search engines.
 It is a work in progress: to encourage adoption by authors of new article submissions, and it would be useful to edit into the converted legacy articles at a later date.
@@
 
-Since the shift to R Markdown, a new development for the open source community has emerged: quarto (Allaire et al.
+Since the shift to R Markdown, a new development for the open-source community has emerged: quarto (Allaire et al.
 2022).
@@
 \item
-  \texttt{math-env}: Examples of inline math, display math, how equation numbering is handled by a Lua filter to convert from LaTeX labeling to R Markdown labeling.
+  \texttt{math-env}: Examples of inline math, display math, and how equation numbering is handled by a Lua filter to convert from LaTeX labeling to R Markdown labeling.
 \item

### 2023-09-28 texor.Rmd

@@
   In 2021 the R Journal made a change of templates for article writing to R Markdown instead of LaTeX.
-The reasons were to encourage better reproducibility of articles using dynamic documents, enable interactivity in articles, and to make the articles more accessible for print disabled readers.
+The reasons were to encourage better reproducibility of articles using dynamic documents, enable interactivity in articles, and to make the articles more accessible for print-disabled readers.
 A resulting challenge was to explore whether legacy articles might be suitably converted into HTML output.
-This paper describes the process to convert an R Journal article from LaTeX to HTML format via R Markdown, and the two new R packages, \CRANpkg{texor} and \CRANpkg{rebib}, that can be used to achieve the conversion.
+This paper describes the process of converting an R Journal article from LaTeX to HTML format via R Markdown, and the two new R packages, \CRANpkg{texor} and \CRANpkg{rebib}, that can be used to achieve the conversion.
 draft: true
@@
 1.
-Articles can include interactive graphics (to allow readers to query, re-focus, watch an animation, for example) and interactive tables (enabling re-sorting in different ways to better digest the information).
-2. It is the best format for accessibility [@Daisy]. Screen readers and other assistive technologies can deliver the content to print disabled readers. (See @W3C for current guidelines.)
-3. HTML provides a more comfortable reading experience on mobile devices, which are increasing used as researchers work on the move and share work via social media.
-4. Search engines can easily access the full text of articles, facilitating discovery of published articles.
+Articles can include interactive graphics (to allow readers to query, re-focus, and watch an animation, for example) and interactive tables (enabling re-sorting in different ways to better digest the information).
+2. It is the best format for accessibility [@Daisy]. Screen readers and other assistive technologies can deliver the content to print-disabled readers. (See @W3C for current guidelines.)
+3. HTML provides a more comfortable reading experience on mobile devices, which are increasingly used as researchers work on the move and share work via social media.
+4. Search engines can easily access the full text of articles, facilitating the discovery of published articles.
 
@@
 
-In addition to article format, changes to the web site structure were important for delivering the publication.
-Web site architectures are also constantly evolving, and the emergence of \CRANpkg{distill} [@distill] allows the journal web site to optimally deliver R Markdown articles.
+In addition to the article format, changes to the website structure were important for delivering the publication.
+Website architectures are also constantly evolving, and the emergence of \CRANpkg{distill} [@distill] allows the journal website to optimally deliver R Markdown articles.
 
-The \CRANpkg{rjtools} was developed to create articles using R Markdown for the R Journal and to embed them into the journal web site.
-The packages described here, \CRANpkg{texor} and \CRANpkg{rebib}, describe software to convert legacy LaTeX format articles into R Markdown, so that they can be rendered in HTML in the new web site.
+The \CRANpkg{rjtools} was developed to create articles using R Markdown for the R Journal and to embed them into the journal website.
+The packages described here, \CRANpkg{texor} and \CRANpkg{rebib}, describe software to convert legacy LaTeX format articles into R Markdown, so that they can be rendered in HTML in the new website.
 
 The paper is organised as follows.
-Section 2 gives an overview of the conversion process, that includes pre-processing using regular expressions, post-processing using Lua filters, and handling of figures, tables and equations.
+Section 2 gives an overview of the conversion process, which includes pre-processing using regular expressions, post-processing using Lua filters, and handling of figures, tables and equations.
 Section 3 describes the \CRANpkg{texor} package that handles most of the conversion.
-Section 4 describes tools to do special handling of bibliography files.
+Section 4 describes tools for special handling of bibliography files.
 The supplementary materials have folders containing specific examples that can be used for understanding how the conversions are done.
@@
 
-<!-- by syntax though you can always embed HTML to expand the feature set.
-Although Pandoc does most of the heavy lifting, there are some instances where even  filters cannot assist.
-A good demonstration of the issue in hand is code environments.-->
-
-<!-- A limitation of Pandoc is that it only works with a "verbatim" environment.
-However, in "RJournal",  we have multiple custom code environments which are not supported or defined due to lower-level tex limitations in Pandoc.
-So the code  environments are eliminated during the conversion process.
-Including a Lua filter to find these elements and redefine them as verbatim environments would also be  ineffective.
-If the code environment is not verbatim, then Pandoc will also try to process the actual code content as LaTeX commands and will lose out on details in some cases.-->
-
-<!--**Be sure to mention the hard parts**-->
-
 ## Post-processing using Lua filters
@@
 
-Figure definitions in LaTeX are many and varied! The standard, single figure definition with the `figure` environment and raster image format such as PNG or JPG, is handled by Pandoc. It will convert:
+Figure definitions in LaTeX are many and varied! The standard, single-figure definition with the `figure` environment and raster image format such as PNG or JPG, is handled by Pandoc. It will convert:
 
@@
 
-<!--XXX ??? warn authors about the conversion not being 1:1 with rmarkdown and might need further edits in with tables and figures.-->
-
 For individuals who are interested in submitting their paper to the R Journal but have written their article using the legacy LaTeX format and wish to convert it to the current R Markdown format, you can use the `latex_to_web()` function on your paper directory.
@@
 This should be changed into Markdown description `![Caption](Image)` or using `knitr::include_graphic(Image)` in an R code chunk, in order for both HTML and PDF versions of the paper to be created when the `rjtools` article template is knitted.
-<!--Why?
-What will need changing?-->
 - Edit the tables for aesthetics.
@@
 However, to have more control and to create more elegant tables using `knitr::kable` provides finer level control.
-<!--If the simple conversion is insufficient?-->
 - Include your R code, to dynamically do the computing described in your article. If any code block is time-consuming to complete, saving intermediate output, or caching would be recommended.
@@
 While this works well with LaTeX, it won't work with Rmarkdown.
-Initially the goal was to use external software like Biber to convert the embedded bibliography to BibTeX.
+Initially, the goal was to use external software like Biber to convert the embedded bibliography to BibTeX.
 However the integration of external software was not viable, hence the experimental idea of a bibliography parser gained momentum.
@@
 Each processing function, from pre-processing the text in the LaTeX file to post-processing accepts a directory name as the argument.
-Generally, if a problem is encountered with one article, the processing will skip on to the next.
+Generally, if a problem is encountered with one article, the processing will skip to the next.
 The vignette [Introduction to texor](https://abhi-1u.github.io/texor/articles/introduction-to-texor.html) provides more details.
@@
 - Figures and equations may need some resizing.
-Sometimes the figure or equation might extend outside the text region in the HTML, which can be fixed with specifying preferred size, or breaks in the equation lines.
+Sometimes the figure or equation might extend outside the text region in the HTML, which can be fixed by specifying the preferred size, or breaks in the equation lines.
 - Alignment in multi-line equations may need adjusting.
-- Alignment of columns in a table may not be the same as the original, for example text might be center-aligned, instead of the original left-aligned.
+- Alignment of columns in a table may not be the same as the original, for example, the text might be center-aligned, instead of the original left-aligned.
 - Specialty LaTeX definitions included by an author for their article may not be recognized by Pandoc, leaving the original command as is, in the article.
@@
 
-The HTML versions of all successfully converted legacy articles will be on the [R Journal web site](https://journal.r-project.org) late in 2023.
-The final step in the integration requires the R Markdown file to be knitted in place to capture the meta data for each article.
+The HTML versions of all successfully converted legacy articles will be on the [R Journal website](https://journal.r-project.org) late in 2023.
+The final step in the integration requires the R Markdown file to be knitted in place to capture the metadata for each article.
 
@@
 
-As a consequence of this project, the journal web site has been furnished with an additional link titled "CONTRIBUTE".
+As a consequence of this project, the journal website has been furnished with an additional link titled "CONTRIBUTE".
 Visiting this link will give you details on how to report a problem with an article that you are reading, or even to make a suggestion on improving operations.
@@
 
-The original motivation for the \CRANpkg{texor} and \CRANpkg{rebib} packages was to convert legacy LaTeX articles into HTML format to accessibility.
+The original motivation for the \CRANpkg{texor} and \CRANpkg{rebib} packages was to convert legacy LaTeX articles into HTML format for accessibility.
 Thus it only creates an `.Rmd` (R Markdown) file which is used to produce only an HTML version.
 The original PDF remains as published.
-The goals is use use the packages for future articles where the authors have chosen to submit only in the legacy LaTeX template.
+The goal is to use the packages for future articles where the authors have chosen to submit only in the legacy LaTeX template.
 
-However, the software can also be used by authors themselves to convert to dynamic documents, by providing an initial R Markdown version of their LaTeX, that with some modification, such as including their R code for computations directly in the document, will produce both PDF nd HTML versions of their submitted paper.
+However, the software can also be used by authors themselves to convert to dynamic documents, by providing an initial R Markdown version of their LaTeX, that with some modification, such as including their R code for computations directly in the document, will produce both PDF and HTML versions of their submitted paper.
 
@@
 In an R Markdown code chunk generating a figure, this can be set using `fig.alt`.
-Alt-text provides a textual alternative to non-text content in HTML documents, and serves various purposes: screen readers can deliver the content to print disabled people, it is displayed in place of the image if it fails to load, and can serve to assist search engines.
+Alt-text provides a textual alternative to non-text content in HTML documents, and serves various purposes: screen readers can deliver the content to print-disabled people, it is displayed in place of the image if it fails to load, and can serve to assist search engines.
 It is a work in progress: to encourage adoption by authors of new article submissions, and it would be useful to edit into the converted legacy articles at a later date.
@@
 
-<!--Remember only the HTML is rendered, and the original PDF is not re-generated.
-
-Only handling to HTML format, especially relevant for figures, because the legacy pdf will remain the same.
-If using for converting work from LaTeX to Rmd then this will need manual changing by author.
-See rjtools guide.
-
-Where this might be useful in the future, other applications.-->
-
-Since the shift to R Markdown, a new development for the open source community has emerged: quarto [@quarto].
+Since the shift to R Markdown, a new development for the open-source community has emerged: quarto [@quarto].
 Although R Markdown would suggest a focus on R, it was always possible to include code chunks written in other languages.
@@
 - `code-env`: Explains how different code environments defined by the R Journal style are handled, and additional details such as code in figure environments, and code in table environments. 
-- `math-env`: Examples of inline math, display math, how equation numbering is handled by a Lua filter to convert from LaTeX labeling to R Markdown labeling. 
+- `math-env`: Examples of inline math, display math, and how equation numbering is handled by a Lua filter to convert from LaTeX labeling to R Markdown labeling. 
 - `figure-env`: Explains how the variety of figure definitions are handled in the conversion, including different image formats, numbering, captions, labeling, multiple images, and `tikz` [@casflow] images.

### 2023-09-28 supplement/metadata/metadata.tex
Added data about theorem environments.

@@
 
+\section{Theorem environments}
+
+\begin{theorem}\label{theorem:1}
+This is a sample theorem..
+\end{theorem}
+
+\begin{proposition}\label{prop:1}
+As with $X \in \mathbb{R}$ this proposition ...  
+\end{proposition}
+
+\begin{remark}\label{rm:1}
+\CRANpkg{texor} makes the article conversion easier.
+\end{remark}
+
+\pkg{Pandoc} supports default LaTeX theorem environments. One can refer to these environments
+like Theorem~\ref{theorem:1},Proposition~\ref{prop:1} and Remark~\ref{rm:1} made here.
+However custom theorem environments are not supported yet.
+
 \section{Summary}
@@
 \item Handling links, block quotes, citations and footnotes.
+\item Default theorem environments.
 \end{itemize}

### 2023-09-28 texor.Rmd
bumped texor version and minor change to figure-env

@@
 - `math-env`: Examples of inline math, display math, and how equation numbering is handled by a Lua filter to convert from LaTeX labeling to R Markdown labeling. 
-- `figure-env`: Explains how the variety of figure definitions are handled in the conversion, including different image formats, numbering, captions, labeling, multiple images, and `tikz` [@casflow] images.
+- `figure-env`: Explains how the variety of figure definitions are handled in the conversion, including different image formats, numbering, captions, labeling, multiple images, and `tikz` images with an example adapted from [@casflow].
 - `table-env`: Examples of how a variety of table types are converted, including multicolumn, complex and wide tables.
@@
 
-- \CRANpkg{texor}: Version 1.2.0 https://abhi-1u.github.io/texor
+- \CRANpkg{texor}: Version 1.3.0 https://abhi-1u.github.io/texor
 - \CRANpkg{rebib}: Version 0.3.2 https://abhi-1u.github.io/rebib/

### 2023-09-28 texor.tex
bumped texor version and minor change to figure-env

@@
 \item
-  \texttt{figure-env}: Explains how the variety of figure definitions are handled in the conversion, including different image formats, numbering, captions, labeling, multiple images, and \texttt{tikz} (Cassidy 2013) images.
+  \texttt{figure-env}: Explains how the variety of figure definitions are handled in the conversion, including different image formats, numbering, captions, labeling, multiple images, and \texttt{tikz} images with an example adapted from (Cassidy 2013).
 \item
@@
 \item
-  \CRANpkg{texor}: Version 1.2.0 \url{https://abhi-1u.github.io/texor}
+  \CRANpkg{texor}: Version 1.3.0 \url{https://abhi-1u.github.io/texor}
 \item

### 2023-09-28 supplement/code-env/code-example.tex
Fixed #19

@@
 S series code environement is defined in Rjournal.sty file. Most of these are
-extensions of verbatim environment and are represented as vanilla verbatim in HTML.
+extensions of verbatim environment and retain the original formatting of the content
+in HTML.
 
@@
 Example series of code environment is defined in Rjournal.sty file. Examples are
-extensions of verbatim environment and are represented as vanilla verbatim in HTML.
+extensions of verbatim environment and retain the original formatting of the content
+in HTML.
 

### 2025-03-31 supplement/table-env/table-example.tex
improve definition of simple tables
Updating article to work with newer pandoc/texor versions

@@
 
-\abstract{
+\begin{abstract}
 This is a small sample article to demonstrate usage of \CRANpkg{texor} to convert table environments.
-}
+\end{abstract}
 
@@
 
-\section{Generic tables}
-Normal LaTeX tables will be converted just fine, even if they include a some math and other elements.
+\section{Simple tables}
+
+A simple table has $n$ rows and $p$ columns leading to $np$ cells.
+Simple LaTeX tables will be converted just fine, even if they include a some math and other elements.
 However, note that the conversion will be done to traditional or pandoc-style markdown rather than R Markdown with the \verb|knitr::kable(..)| function.
@@
     \centering
-    \includegraphics[width=1\textwidth]{Rlogo-5.png}
+    \includegraphics[width=1\textwidth]{R.png}
     \end{minipage} \\ \hline

### 2025-03-31 texor.Rmd
Updating article to work with newer pandoc/texor versions

@@
     orcid: 0009-0000-6935-8690
-    email: perricoq@outlook.com
+    email: abhishek.ulayil.m@gmail.com
   - name: Mitch O'Hara-Wild
@@
 output: 
+  rjtools::rjournal_pdf_article:
+    toc: no
   rjtools::rjournal_web_article:
@@
 
-Since the shift to R Markdown, a new development for the open-source community has emerged: quarto [@quarto].
+Since the shift to R Markdown, a new development for the open-source community has emerged: quarto [@Allaire_Quarto_2024].
 Although R Markdown would suggest a focus on R, it was always possible to include code chunks written in other languages.

### 2025-03-31 texor.tex
Updating article to work with newer pandoc/texor versions

@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}
 
@@
 The tooling behind the production of the R Journal is regularly updated.
-Early articles were typeset using LaTeX (The LaTeX Project 2023), based on a specific, but changing, template.
+Early articles were typeset using LaTeX \citep{latex}, based on a specific, but changing, template.
 Using LaTeX requires that code is separated from the documentation, and there is a chance that code chunks in the paper don't reproduce the results reported.
-With the emergence of dynamic document systems such as R Markdown (Xie, Allaire, and Grolemund 2018), a tight coupling of code and documentation is possible.
-Code chunks are dynamically executed when the document is typeset using a system like \CRANpkg{knitr} (Xie 2015), making reporting of computing research more reproducible.
+With the emergence of dynamic document systems such as R Markdown \citep{rmarkdown}, a tight coupling of code and documentation is possible.
+Code chunks are dynamically executed when the document is typeset using a system like \CRANpkg{knitr} \citep{knitr}, making reporting of computing research more reproducible.
 
@@
 \item
-  It is the best format for accessibility (The Daisy Consortium 2023).
-Screen readers and other assistive technologies can deliver the content to print-disabled readers.
-(See Sheri Byrne-Haber (2023) for current guidelines.)
+  It is the best format for accessibility \citep{Daisy}. Screen readers and other assistive technologies can deliver the content to print-disabled readers. (See \citet{W3C} for current guidelines.)
 \item
@@
 In addition to the article format, changes to the website structure were important for delivering the publication.
-Website architectures are also constantly evolving, and the emergence of \CRANpkg{distill} (Dervieux et al.
-2022) allows the journal website to optimally deliver R Markdown articles.
+Website architectures are also constantly evolving, and the emergence of \CRANpkg{distill} \citep{distill} allows the journal website to optimally deliver R Markdown articles.
 
@@
 
-\hypertarget{internals}{%
-\section{The internals of converting from LaTeX to R Markdown}\label{internals}}
+\section{The internals of converting from LaTeX to R Markdown}\label{internals}
 
-The decision to convert to R Markdown format means that the final output to PDF and HTML will depend on Pandoc (MacFarlane, Krewinkel, and Rosenthal 2023).
+The decision to convert to R Markdown format means that the final output to PDF and HTML will depend on Pandoc \citep{pandoc}.
 Pandoc is a versatile document conversion program written in Haskell that is core to numerous documentation systems, including R Markdown and Quarto.
@@
 
-{\centering \includegraphics[width=1\linewidth]{figures/wf}
+{\centering \includegraphics[width=1\linewidth,alt={Flow diagram showing original article pre-processed in R passed to pandoc which creates the AST and modifies using many lua filters then passed to pacdoc writer to produce the converted article.}]{figures/wf}
 
@@
 
-\hypertarget{pre-processing-using-regular-expressions}{%
-\subsection{Pre-processing using regular expressions}\label{pre-processing-using-regular-expressions}}
+\subsection{Pre-processing using regular expressions}\label{pre-processing-using-regular-expressions}
 
 LaTeX is very descriptive language, that allows authors substantial freedom for customization.
-Markdown (Gruber 2002), on which R Markdown is based, is more restrictive and was designed to make it easier to create web pages without the distraction of a gazillion HTML tags.
+Markdown \citep{markdown}, on which R Markdown is based, is more restrictive and was designed to make it easier to create web pages without the distraction of a gazillion HTML tags.
 The beauty of Markdown is that it allows the author to focus on writing, without format cluttering the text.
@@
 \item
-  \texttt{stream\_editor()}: operates like the \texttt{sed} function in unix (Ritchie and Thompson 1978) and allows generic text pattern matching and replacement.
+  \texttt{stream\_editor()}: operates like the \texttt{sed} function in unix \citep{unix} and allows generic text pattern matching and replacement.
 \item
@@
 
-\hypertarget{post-processing-using-lua-filters}{%
-\subsection{Post-processing using Lua filters}\label{post-processing-using-lua-filters}}
+\subsection{Post-processing using Lua filters}\label{post-processing-using-lua-filters}
 
-Lua (Ierusalimschy, Figueiredo, and Filho 1996) is a programming language, that is light-weight and fast: ideal for procedural operations.
+Lua \citep{lua} is a programming language, that is light-weight and fast: ideal for procedural operations.
 It is embedded in many other applications to allow custom scripting for extensibility.
@@
 
-\hypertarget{figures}{%
-\subsection{Figures}\label{figures}}
+\subsection{Figures}\label{figures}
 
-\hypertarget{standard-single-figure}{%
-\subsubsection{Standard, single figure}\label{standard-single-figure}}
+\subsubsection{Standard, single figure}\label{standard-single-figure}
 
@@
 
-\hypertarget{pdf-format-images}{%
-\subsubsection{PDF format images}\label{pdf-format-images}}
+\subsubsection{PDF format images}\label{pdf-format-images}
 
@@
 
-\hypertarget{multiple-figures}{%
-\subsubsection{Multiple figures}\label{multiple-figures}}
+\subsubsection{Multiple figures}\label{multiple-figures}
 
@@
 
-\hypertarget{tikz-format-images}{%
-\subsubsection{\texorpdfstring{\texttt{tikz} format images}{tikz format images}}\label{tikz-format-images}}
+\subsubsection{\texorpdfstring{\texttt{tikz} format images}{tikz format images}}\label{tikz-format-images}
 
@@
 
-\hypertarget{algorithm2e-graphics-as-figures}{%
-\subsubsection{Algorithm2e graphics as figures}\label{algorithm2e-graphics-as-figures}}
+\subsubsection{Algorithm2e graphics as figures}\label{algorithm2e-graphics-as-figures}
 
@@
 
-{\centering \includegraphics[width=0.6\linewidth]{figures/alghow}
+{\centering \includegraphics[width=0.6\linewidth,alt={Illustration of an algorithm starting with Data and Result, followed by a while, and an if/else block.}]{figures/alghow}
 
@@
 
-\hypertarget{equations}{%
-\subsection{Equations}\label{equations}}
+\subsection{Equations}\label{equations}
 
@@
 LaTeX automatically numbers equations, unless specifically instructed not to.
-Equation numbering in R Markdown requires specific labeling using \texttt{(\textbackslash{}\#eq:xx)} as described in Xie (2023).
+Equation numbering in R Markdown requires specific labeling using \texttt{(\textbackslash{}\#eq:xx)} as described in \citet{bookdown}.
 The \CRANpkg{texor} helps by adding the labeling using a Lua filter to convert the existing \texttt{\textbackslash{}label\{..\}} to \texttt{(\textbackslash{}\#eq:xx)}.
 
-\hypertarget{tables}{%
-\subsection{Tables}\label{tables}}
+\subsection{Tables}\label{tables}
 
@@
 However, there have been many improvements in table definitions for R Markdown that are increasing producing the beautifully crafted tables possible in LaTeX.
-The conversion in \CRANpkg{texor} can mostly handle the simple tables, and for producing more complex tables it may be necessary to manually edit the resulting \texttt{Rmd} file to make conditional tables, one to render specifically for HTML output using packages such as \CRANpkg{kableExtra} (Zhu 2021), \CRANpkg{gt} (Iannone et al.
-2023), \CRANpkg{htmlTable} (Gordon, Gragg, and Konings 2022), \CRANpkg{tableHTML} (Boutaris, Zauchner, and Jomar 2023), \CRANpkg{tables} (Murdoch 2023) or \CRANpkg{DT} (Xie, Cheng, and Tan 2023).
+The conversion in \CRANpkg{texor} can mostly handle the simple tables, and for producing more complex tables it may be necessary to manually edit the resulting \texttt{Rmd} file to make conditional tables, one to render specifically for HTML output using packages such as \CRANpkg{kableExtra} \citep{kableExtra}, \CRANpkg{gt} \citep{gt}, \CRANpkg{htmlTable} \citep{htmlTable}, \CRANpkg{tableHTML} \citep{tableHTML}, \CRANpkg{tables} \citep{tables} or \CRANpkg{DT} \citep{DT}.
 A benefit of using conditional markup is that it can take advantage of HTML-specific features, such as sortable columns or paged tables.
 
-\hypertarget{generic-tables}{%
-\subsubsection{Generic tables}\label{generic-tables}}
+\subsubsection{Generic tables}\label{generic-tables}
 
@@
 
-\hypertarget{multicolumn-tables}{%
-\subsubsection{Multicolumn tables}\label{multicolumn-tables}}
+\subsubsection{Multicolumn tables}\label{multicolumn-tables}
 
@@
 
-\hypertarget{other-tables}{%
-\subsubsection{Other tables}\label{other-tables}}
+\subsubsection{Other tables}\label{other-tables}
 
@@
 
-\hypertarget{texor}{%
-\section{\texorpdfstring{Using \CRANpkg{texor}}{Using }}\label{texor}}
+\section{\texorpdfstring{Using \CRANpkg{texor}}{Using }}\label{texor}
 
@@
 
-\hypertarget{rebib}{%
-\section{\texorpdfstring{Managing the bibliography using \CRANpkg{rebib}}{Managing the bibliography using }}\label{rebib}}
+\section{\texorpdfstring{Managing the bibliography using \CRANpkg{rebib}}{Managing the bibliography using }}\label{rebib}
 
-Typically bibliographies are generated during the processing of a LaTeX article using the BibTeX software (Feder 2006) operating on a \texttt{.bib} list of references.
+Typically bibliographies are generated during the processing of a LaTeX article using the BibTeX software \citep{bibtex} operating on a \texttt{.bib} list of references.
 The current R Journal template requires the inclusion of the \texttt{.bib} file.
@@
 
-\hypertarget{the-process-of-converting-all-the-legacy-articles}{%
-\section{The process of converting all the legacy articles}\label{the-process-of-converting-all-the-legacy-articles}}
+\section{The process of converting all the legacy articles}\label{the-process-of-converting-all-the-legacy-articles}
 
@@
 
-\hypertarget{batch-processing}{%
-\subsection{Batch processing}\label{batch-processing}}
+\subsection{Batch processing}\label{batch-processing}
 
@@
 
-\hypertarget{known-problems-requiring-manual-fixes}{%
-\subsection{Known problems requiring manual fixes}\label{known-problems-requiring-manual-fixes}}
+\subsection{Known problems requiring manual fixes}\label{known-problems-requiring-manual-fixes}
 
@@
 
-\hypertarget{plan-for-release}{%
-\subsection{Plan for release}\label{plan-for-release}}
+\subsection{Plan for release}\label{plan-for-release}
 
@@
 
-\hypertarget{reporting-problems}{%
-\subsection{Reporting problems}\label{reporting-problems}}
+\subsection{Reporting problems}\label{reporting-problems}
 
@@
 
-\hypertarget{summary}{%
-\section{Summary}\label{summary}}
+\section{Summary}\label{summary}
 
@@
 There are many other journals that have a similar backlog of legacy articles.
-A current and ongoing project is being conducted by arXiv Accessibility Project (2023) to convert their collection of articles from LaTeX (shared as PDF) to HTML, utilizing LaTeXML (Miller and Ginev 2023).
+A current and ongoing project is being conducted by \citet{arxiv} to convert their collection of articles from LaTeX (shared as PDF) to HTML, utilizing LaTeXML \citep{latexML}.
 The work conducted for the R Journal might of interest to similar small journals that operate without paid staff.
@@
 
-Since the shift to R Markdown, a new development for the open-source community has emerged: quarto (Allaire et al.
-2022).
+Since the shift to R Markdown, a new development for the open-source community has emerged: quarto \citep{Allaire_Quarto_2024}.
 Although R Markdown would suggest a focus on R, it was always possible to include code chunks written in other languages.
@@
 
-\hypertarget{acknowledgments}{%
-\section*{Acknowledgments}\label{acknowledgments}}
+\section*{Acknowledgments}\label{acknowledgments}
 \addcontentsline{toc}{section}{Acknowledgments}
@@
 
-\hypertarget{supplementary-materials}{%
-\section*{Supplementary materials}\label{supplementary-materials}}
+\section*{Supplementary materials}\label{supplementary-materials}
 \addcontentsline{toc}{section}{Supplementary materials}
@@
 \item
-  \texttt{figure-env}: Explains how the variety of figure definitions are handled in the conversion, including different image formats, numbering, captions, labeling, multiple images, and \texttt{tikz} images with an example adapted from (Cassidy 2013).
+  \texttt{figure-env}: Explains how the variety of figure definitions are handled in the conversion, including different image formats, numbering, captions, labeling, multiple images, and \texttt{tikz} images with an example adapted from \citep{casflow}.
 \item
@@
 
-\hypertarget{source-materials}{%
-\subsection*{Source materials}\label{source-materials}}
+\subsection*{Source materials}\label{source-materials}
 \addcontentsline{toc}{subsection}{Source materials}
@@
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-quarto}{}}%
-Allaire, J. J., Charles Teague, Yihui Xie, and Christophe Dervieux. 2022. {``Quarto.''} Zenodo. \url{https://doi.org/10.5281/zenodo.5960048}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-arxiv}{}}%
-arXiv Accessibility Project. 2023. {``HTML as an Accessible Format for Papers.''} \url{https://info.arxiv.org/about/accessible_HTML.html}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-tableHTML}{}}%
-Boutaris, Theo, Clemens Zauchner, and Dana Jomar. 2023. \emph{tableHTML: A Tool to Create HTML Tables}. \url{https://CRAN.R-project.org/package=tableHTML}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-casflow}{}}%
-Cassidy, Josh. 2013. \emph{{LaTeX Graphics using TikZ: A Tutorial for Beginners (Part 3)---Creating Flowcharts}}. Overleaf tutorials. \url{https://www.overleaf.com/learn/latex/}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-distill}{}}%
-Dervieux, Christophe, JJ Allaire, Rich Iannone, Alison Presmanes Hill, and Yihui Xie.
-2022.
-\emph{Distill: 'R Markdown' Format for Scientific and Technical Writing}.
-\url{https://CRAN.R-project.org/package=distill}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-bibtex}{}}%
-Feder, Alexander. 2006. {``BibTeX: Reference Management Software.''} https://www.bibtex.org.
-
-\leavevmode\vadjust pre{\hypertarget{ref-htmlTable}{}}%
-Gordon, Max, Stephen Gragg, and Peter Konings. 2022. \emph{htmlTable: Advanced Tables for Markdown/HTML}. \url{https://CRAN.R-project.org/package=htmlTable}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-markdown}{}}%
-Gruber, John. 2002. {``Markdown.''} https://daringfireball.net/projects/markdown/.
-
-\leavevmode\vadjust pre{\hypertarget{ref-gt}{}}%
-Iannone, Richard, Joe Cheng, Barret Schloerke, Ellis Hughes, Alexandra Lauer, and JooYoung Seo.
-2023.
-\emph{Gt: Easily Create Presentation-Ready Display Tables}.
-\url{https://CRAN.R-project.org/package=gt}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-lua}{}}%
-Ierusalimschy, Roberto, Luiz Henrique de Figueiredo, and Waldemar Celes Filho. 1996. {``Lua---an Extensible Extension Language.''} \emph{Software: Practice and Experience} 26 (6): 635--52.
-
-\leavevmode\vadjust pre{\hypertarget{ref-pandoc}{}}%
-MacFarlane, John, Albert Krewinkel, and Jesse Rosenthal. 2023. {``Pandoc.''} https://github.com/jgm/pandoc.
-
-\leavevmode\vadjust pre{\hypertarget{ref-latexML}{}}%
-Miller, Bruce, and Deyan Ginev. 2023. {``{LaTeXML}: A {LaTeX} to {XML/HTML/MathML} Converter.''} \url{https://math.nist.gov/~BMiller/LaTeXML/}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-tables}{}}%
-Murdoch, Duncan. 2023. \emph{Tables: Formula-Driven Table Generation}. \url{https://CRAN.R-project.org/package=tables}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-unix}{}}%
-Ritchie, O. M., and K. Thompson. 1978. {``The UNIX Time-Sharing System.''} \emph{The Bell System Technical Journal} 57 (6): 1905--29. \url{https://doi.org/10.1002/j.1538-7305.1978.tb02136.x}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-W3C}{}}%
-Sheri Byrne-Haber, Charles LaPierre, David Fazio. 2023. {``W3C Accessibility Maturity Model.''} \url{https://www.w3.org/TR/maturity-model/}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Daisy}{}}%
-The Daisy Consortium. 2023. {``Creating the Best Way to Read and Publish.''} \url{https://daisy.org}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-latex}{}}%
-The LaTeX Project. 2023. {``LaTeX -- a Document Preparation System.''} https://www.latex-project.org.
-
-\leavevmode\vadjust pre{\hypertarget{ref-knitr}{}}%
-Xie, Yihui. 2015. \emph{Dynamic Documents with {R} and Knitr}. 2nd ed. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://yihui.org/knitr/}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-bookdown}{}}%
----------. 2023. \emph{{bookdown: Authoring Books and Technical Documents with R Markdown}}. \#equations. \url{https://bookdown.org/yihui/bookdown/markdown-extensions-by-bookdown.html}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-rmarkdown}{}}%
-Xie, Yihui, J. J. Allaire, and Garrett Grolemund. 2018. \emph{R Markdown: The Definitive Guide}. Boca Raton, Florida: Chapman; Hall/CRC. \url{https://bookdown.org/yihui/rmarkdown}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-DT}{}}%
-Xie, Yihui, Joe Cheng, and Xianying Tan. 2023. \emph{DT: A Wrapper of the JavaScript Library 'DataTables'}. \url{https://CRAN.R-project.org/package=DT}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-kableExtra}{}}%
-Zhu, Hao. 2021. \emph{kableExtra: Construct Complex Table with 'Kable' and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.
-
-\end{CSLReferences}
-
 \bibliography{RJreferences.bib}
@@
 \textit{ORCiD: \href{https://orcid.org/0009-0000-6935-8690}{0009-0000-6935-8690}}\\%
-\href{mailto:perricoq@outlook.com}{\nolinkurl{perricoq@outlook.com}}%
+\href{mailto:abhishek.ulayil.m@gmail.com}{\nolinkurl{abhishek.ulayil.m@gmail.com}}%
 }

### 2025-03-31 texor.R

@@ -8,34 +8,34 @@ library(ggplot2)
 
 
 ## ----workflow, eval=FALSE-----------------------------------------------------
-#> library(DiagrammeR)
-#> wf <- create_node_df(
-#>         n=8,
-#>         type=c("a", "b", "a", "b",
-#>                "a", "b", "b", "a"),
-#>         label=c("original article", "pre-processing in R",
-#>                 "pre-processed article", "pandoc reader",
-#>                 "AST",
-#>                 "modify AST using lua filter",
-#>                 "pandoc writer",
-#>                 "converted article"
-#>                 ),
-#>         shape="rectangle",
-#>         width=c(1.3, 1.7, 1.9, 1.3, 0.5, 2.3, 1.2, 1.5),
-#>         fillcolor = c("darkseagreen2", "grey85", "darkseagreen2", "grey85", "darkseagreen2", "grey85", "grey85", "darkseagreen2"),
-#>         color = c("black", "grey85", "black", "grey85", "black", "grey85", "grey85", "black"),
-#>         fontcolor="black",
-#>         x=c(1,1,1,1,1,2,1,1),
-#>         y=c(1,2,3,4,5,5,6,6),
-#>         fontsize=14
-#>         )
-#> wf_arrows <- create_edge_df(from = c(1, 2, 3, 4, 5, 6, 5, 7),
-#>                             to = c(2, 3, 4, 5, 6, 5, 7, 8),
-#>                             color="black")
-#> wf_graph <- create_graph(wf, wf_arrows,
-#>                    directed = TRUE, attr_theme = "tb")
-#> set.seed(7)
-#> render_graph(wf_graph, layout="fr", height=500)
+# library(DiagrammeR)
+# wf <- create_node_df(
+#         n=8,
+#         type=c("a", "b", "a", "b",
+#                "a", "b", "b", "a"),
+#         label=c("original article", "pre-processing in R",
+#                 "pre-processed article", "pandoc reader",
+#                 "AST",
+#                 "modify AST using lua filter",
+#                 "pandoc writer",
+#                 "converted article"
+#                 ),
+#         shape="rectangle",
[…]

### 2025-03-31 texor.Rmd
pdf rendering ok now

--- 2023-09-28 texor.Rmd
@@
 output: 
-  rjtools::rjournal_web_article:
-    self_contained: yes
-    toc: no
+  rjtools::rjournal_article
 bibliography: RJreferences.bib
-
 ---
@@
 A current and ongoing project is being conducted by @arxiv to convert their collection of articles from LaTeX (shared as PDF) to HTML, utilizing LaTeXML [@latexML].
-The work conducted for the R Journal might of interest to similar small journals that operate without paid staff.
+The work conducted for the R Journal might be of interest to similarly small journals that operate without paid staff.
 Converting legacy articles provides accessibility to a wealth of educational material.
 
-Since the shift to R Markdown, a new development for the open-source community has emerged: quarto [@quarto].
+After the change was made to use R Markdown, a new development for the open-source scientific and technical publishing community has emerged: quarto [@quarto].
 Although R Markdown would suggest a focus on R, it was always possible to include code chunks written in other languages.

### 2025-03-31 texor.tex
pdf rendering ok now

@@
 A current and ongoing project is being conducted by \citet{arxiv} to convert their collection of articles from LaTeX (shared as PDF) to HTML, utilizing LaTeXML \citep{latexML}.
-The work conducted for the R Journal might of interest to similar small journals that operate without paid staff.
+The work conducted for the R Journal might be of interest to similarly small journals that operate without paid staff.
 Converting legacy articles provides accessibility to a wealth of educational material.
 
-Since the shift to R Markdown, a new development for the open-source community has emerged: quarto \citep{Allaire_Quarto_2024}.
+After the change was made to use R Markdown, a new development for the open-source scientific and technical publishing community has emerged: quarto \citep{quarto}.
 Although R Markdown would suggest a focus on R, it was always possible to include code chunks written in other languages.

### 2025-03-31 texor.Rmd
self-contained html now

@@
 output: 
-  rjtools::rjournal_article
-bibliography: RJreferences.bib
+  rjtools::rjournal_article:
+    self_contained: yes
+    toc: no
+bibliography: texor.bib
 ---

### 2025-04-01 texor.Rmd
Re-Submission Article.

@@
 A resulting challenge was to explore whether legacy articles might be suitably converted into HTML output.
-This paper describes the process of converting an R Journal article from LaTeX to HTML format via R Markdown, and the two new R packages, \CRANpkg{texor} and \CRANpkg{rebib}, that can be used to achieve the conversion.
+This paper describes the process of converting an R Journal article from LaTeX to HTML format via R Markdown, and the two new R packages, texor and rebib, that can be used to achieve the conversion.
 draft: true

### 2025-04-01 texor.tex
self-contained html now
Re-Submission Article.

@@
 A resulting challenge was to explore whether legacy articles might be suitably converted into HTML output.
-This paper describes the process of converting an R Journal article from LaTeX to HTML format via R Markdown, and the two new R packages, \CRANpkg{texor} and \CRANpkg{rebib}, that can be used to achieve the conversion.
+This paper describes the process of converting an R Journal article from LaTeX to HTML format via R Markdown, and the two new R packages, texor and rebib, that can be used to achieve the conversion.
 }
@@
 
-\bibliography{RJreferences.bib}
+\bibliography{texor.bib}
 

### 2025-07-07 review/response_to_review.Rmd
changes in response to the review

---
title: "Response: Converting LaTeX Legacy R Journal Articles into R Markdown Articles using texor and rebib"
output: html_document
---

We thank the editor and reviewers for providing helpful comments on our paper. This is how we have addressed them. *Our response is in italics.*

## General

This article describes how legacy articles of the R Journal between 2009 and 2022 were converted from their original LaTeX format to R Markdown. The process was clearly documented and the results are readily available on the R Journal website today. I'd like to congratulate the authors on completing this seemingly impossible task (at least it was unimaginable to me, given the potential complexity of LaTeX). I wholeheartedly agree with the authors on the benefits of the HTML format, and have also been keeping an eye on the similar effort by arXiv (which was cited at the end of the article). I think the contribution of this article is tremendous, and would like to recommend that we accept it for publication.

*Thank you*

## Detailed comments

Some minor comments:

P6: Is there an extra `x` in `\multicolumnxx{..}`?

*Fixed*

P7: The link on "Pandoc Version > 3.0.0 (if possible latest)" appears to be missing and doesn't work for me.

*Fixed*

P7: The bullet list in Section 3 seems to have a superfluous blank line after the first list item.

*It's not clear how ot fix this. There is nothing special added to the list. It has to be something that latex is doing that we do not control. The html version doesn't have the problem.*

P8: I don't see the link titled "CONTRIBUTE" on the R Journal website as mentioned in Section 5.4.

*This is coming. We expect the current editor will add this, before the paper appears.*

P9: I'd add "(alternate text)" after "alt-text" in case readers are not familiar with the `alt` property of images, or add a link to the text, e.g., https://developer.mozilla.org/en-US/docs/Web/API/HTMLImageElement/alt.

*Done*

## Code

The packages texor and rebib have clear workflows, and their documentation are well-written.

I didn't fully review the source code since it was rather lengthy, but I could see room for improvement, e.g.,

- The stringr wrappers in texor are not necessary: https://github.com/Abhi-1U/texor/blob/e669ec3f9bf7ad5fb5c4b929f13d18ad7944f735/R/string-tools.R#L3-L28 What we need is `@importFrom stringr str_trim str_detect ...`, and the stringr functions can be used without the `stringr::` qualifier.

- The `stream_editor()` function (https://github.com/Abhi-1U/texor/blob/main/R/stream-editor.R) plays a central role in texor but its implementation didn't take advantage of the built-in vectorization of `gsub()`, i.e., there is absolutely no need to use a `for` loop to modify the input line by line (to mimic `sed`), which is very inefficient, and we can just do `gsub()` on all matches:

    i <- grep(pattern, raw_lines) 
    raw_lines[i] <- gsub(target, replacement, raw_lines[i])

- The supplementary materials (https://github.com/Abhi-1U/texor-rjarticle/tree/master/supplement) are quite helpful, and I really like the examples. It will be nice to also include the intermediate `.Rmd` or `.md` files when converting `.tex` to `.html`, so readers can see how various LaTeX environments are converted to their Markdown counterparts, although the article has already provided some examples in Section 2. The function `texor::latex_to_web()` doesn't seem to have an argument for preserving the intermediate Markdown output, and it may be worth adding one.

### 2025-07-07 texor.Rmd
changes in response to the review

@@
 
-2. A LaTeX macro is used to redefine the `\multicolumnxx{..}` to
+2. A LaTeX macro is used to redefine the `\multicolumnx{..}` to
     `\multicolumn{---}` (which is accepted by pandoc).
@@
 
-Note that you will need to use [Pandoc Version > 3.0.0 (if possible latest)]((https://pandoc.org/installing.html)) for the best results. You can check your version with:
+Note that you will need to use [Pandoc Version > 3.0.0 (if possible latest)](https://pandoc.org/installing.html) for the best results. You can check your version with:
 
@@
 This will get you about 80% of the way to an R Markdown version of your paper.
-You will then want to
+You will then want to:
 
@@
 Authors can become more aware of good practice.
-For example, one of the useful elements when an article is in HTML format is that figures can be made richer with an `alt-text` element.
+For example, one of the useful elements when an article is in HTML format is that figures can be made richer with an `alt-text` (alternative text) element.
 In an R Markdown code chunk generating a figure, this can be set using `fig.alt`.

### 2025-07-07 texor.tex
changes in response to the review

@@
 \item
-  A LaTeX macro is used to redefine the \texttt{\textbackslash{}multicolumnxx\{..\}} to
+  A LaTeX macro is used to redefine the \texttt{\textbackslash{}multicolumnx\{..\}} to
   \texttt{\textbackslash{}multicolumn\{-\/-\/-\}} (which is accepted by pandoc).
@@
 
-Note that you will need to use \href{(https://pandoc.org/installing.html)}{Pandoc Version \textgreater{} 3.0.0 (if possible latest)} for the best results. You can check your version with:
+Note that you will need to use \href{https://pandoc.org/installing.html}{Pandoc Version \textgreater{} 3.0.0 (if possible latest)} for the best results. You can check your version with:
 
@@
 This will get you about 80\% of the way to an R Markdown version of your paper.
-You will then want to
+You will then want to:
 

### 2025-07-20 texor.tex
responses to code section

@@
 The conversion will create HTML code to define the image.
-This should be changed into Markdown description \texttt{!{[}Caption{]}(Image)} or using \texttt{knitr::include\_graphic(Image)} in an R code chunk, in order for both HTML and PDF versions of the paper to be created when the \texttt{rjtools} article template is knitted.\\
+This should be changed into Markdown description \texttt{!{[}Caption{]}(Image)} or using \texttt{knitr::include\_graphic(Image)} in an R code chunk, in order for both HTML and PDF versions of the paper to be created when the \texttt{rjtools} article template is knitted.
 \item
@@
 Authors can become more aware of good practice.
-For example, one of the useful elements when an article is in HTML format is that figures can be made richer with an \texttt{alt-text} element.
+For example, one of the useful elements when an article is in HTML format is that figures can be made richer with an \texttt{alt-text} (alternative text) element.
 In an R Markdown code chunk generating a figure, this can be set using \texttt{fig.alt}.

### 2025-09-07 review/response_to_review.Rmd
responses to code section
Final response for inclusion of Rmd files.

@@
 
-*It's not clear how ot fix this. There is nothing special added to the list. It has to be something that latex is doing that we do not control. The html version doesn't have the problem.*
+*Fixed*
 
@@
 
+*Done, will be included in the latest CRAN release of the package.*
+
 - The `stream_editor()` function (https://github.com/Abhi-1U/texor/blob/main/R/stream-editor.R) plays a central role in texor but its implementation didn't take advantage of the built-in vectorization of `gsub()`, i.e., there is absolutely no need to use a `for` loop to modify the input line by line (to mimic `sed`), which is very inefficient, and we can just do `gsub()` on all matches:
@@
 
+*Done, will be included in the latest CRAN release of the package.*
+
 - The supplementary materials (https://github.com/Abhi-1U/texor-rjarticle/tree/master/supplement) are quite helpful, and I really like the examples.
@@
 The function `texor::latex_to_web()` doesn't seem to have an argument for preserving the intermediate Markdown output, and it may be worth adding one.
+
+*Done, intermediate `.Rmd` and `.md` files would be included in the next iteration of article.
+They are generated via the `texor::latex_to_web()` function and, by default, are retained automatically after a successful conversion without requiring explicit arguments.*

### 2025-09-07 supplement/bibliography/RJwrapper.Rmd
Final response for inclusion of Rmd files.

---
title: Handling bibliographies with texor
abstract: |
  This is a small sample article to demonstrate usage of
  [**rebib**](https://CRAN.R-project.org/package=rebib) to manage
  bibliographies.
author:
- name: Abhishek Ulayil
  affiliation: Student, Institute of Actuaries of India
  orcid: 0009-0000-6935-8690
  address: Mumbai, India
date: '2025-09-01'
date_received: ~
journal:
  firstpage: ~
  lastpage: ~
volume: 17
issue: 3
slug: RJ-2025-000
citation_url: https://rjournal.github.io/
packages:
  cran:
  - rebib
  - texor
  bioc: []
preview: preview.png
bibliography: ~
CTV: ~
legacy_pdf: yes
legacy_converted: yes
output:
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
    mathjax: https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-chtml-full.js
    md_extension: -tex_math_single_backslash

---

::: article
## Introduction

The [**rebib**](https://CRAN.R-project.org/package=rebib) package is a
spin-off from the [**texor**](https://CRAN.R-project.org/package=texor)
package.

The spin-off cam about because the bibliography section in
[**texor**](https://CRAN.R-project.org/package=texor) package was
expanding significantly, enough to deserve a dedicated space.

Thus [**rebib**](https://CRAN.R-project.org/package=rebib) was born and
it has improved over the past few months to include more advanced
features and utilities.

### What does rebib do?

In LaTeX, we have multiple ways to declare bibliography entries, which
include embedding entries formatted to a given style in the document
itself or isolating that section to an external file commonly called a
`.bbl` file. Then we have the BibTeX format, which is designed to store
the full bibliographic information for each reference, making it
superior for managing references.

LaTeX has no problem dealing with multiple formats and works well with
all these bibliographic formats. However, some web publishing journals
like R Markdown articles for the R Journal do not support embedded
bibliographies. As a result, they need a BibTeX equivalent of the same.

The [**rebib**](https://CRAN.R-project.org/package=rebib) package
converts embedded LaTeX bibliographies into a close BibTeX equivalent.
This package relieves the author from manual conversion of the documents
and aids the project to convert legacy R Journal articles to the current
R Markdown format.

### Features of rebib

-   Extracts embedded bibliographic entries from a `.tex` file.

-   Creates the mandatory title and author fields.

-   Creates the optional URL, ISBN, publisher, pages and year fields,
    when available.

-   Stores the remaining information in `"journal"` (internally) and
    `"publisher"` (when writing BibTeX file).

-   Ignores commented LaTeX code.

-   Tracks citations included in the document.

-   Aggregates embedded bibliographic entries with references from
    supplementary `.bbl` and `.bib` files.

## Parser

The magic behind [**rebib**](https://CRAN.R-project.org/package=rebib)
is a parser which parses bibliographic data and sorts them according to
the matching regular expressions.

#### Stage 1: Read the embedded bibliography block

This stage is a minor step where it reads the Embedded Bibliography from
the LaTeX document. This step also includes filtering out commented
lines to avoid reading un-intended entries.

The bibliography is split into entries using the LaTeX macro `\\bibitem`
as a marker for a new entry and the resulting data is exported to a
variable.

``` r
dir.create(your_article_folder <- file.path(tempdir(), "exampledir"))
example_files <-  system.file("article", package = "rebib")
x <- file.copy(from = example_files,to=your_article_folder,recursive = T)
your_article_path <- paste(your_article_folder,"article",sep="/")
file_name <- rebib:::get_texfile_name(your_article_path)
bib_items <- rebib:::extract_embeded_bib_items(your_article_path,file_name)
bib_items[[1]]
#> [1] "\\bibitem[Ihaka, Ross and Gentleman, Robert]{ihaka:1996}"                          
#> [2] "Ihaka, Ross and Gentleman, Robert"                                                 
#> [3] "\\newblock \\emph{R: A Language for Data Analysis and Graphics.}"                  
#> [4] "\\newblock \\emph{Journal of Computational and Graphical Statistics}, 3:\\penalty0"
#> [5] "299--314, 1996."                                                                   
#> [6] "\\newblock URL : \\url{https://doi.org/10.1080/10618600.1996.10474713}"
```

#### Stage 2: Regex powered parser

Now, with the chunks of bibliographic entries, each is passed to a
parser which will break it down based on regular expressions. The logic
is to use the LaTeX macro `\\newblock` as a placeholder to identify the
position of text elements relative to it.

The first value to be parsed is the `unique_id`, also called the
citation reference which is used to cite elements inside the article.
Usually, this is in the first or second line of the whole entry. The
position of the `unique_id` will determine the position of the author
names.

``` r
bib_items[[1]][1]
#> [1] "\\bibitem[Ihaka, Ross and Gentleman, Robert]{ihaka:1996}"
```

After reading the `unique_id`, the parser will attempt to read the
author name(s).

``` r
bib_items[[1]][2]
#> [1] "Ihaka, Ross and Gentleman, Robert"
```

Next, the title is extracted based on the position of the new blocks or
the end of the bib chunk.

``` r
bib_items[[1]][3]
#> [1] "\\newblock \\emph{R: A Language for Data Analysis and Graphics.}"
```

This way the crucial elements of the bibliographic entry (unique_id,
author names and title) are parsed out. The remaining data is stored as
`journal` internally and `publisher` when writing to a new BibTeX file.

``` r
bib_items[[1]][4:6]
#> [1] "\\newblock \\emph{Journal of Computational and Graphical Statistics}, 3:\\penalty0"
#> [2] "299--314, 1996."                                                                   
#> [3] "\\newblock URL : \\url{https://doi.org/10.1080/10618600.1996.10474713}"
```

A series of filters for ISBN, URL, pages and year fields are applied to
search for relevant data from the remaining data. If the data is not
available, then it is set as `NULL` and will be skipped while writing
the BibTeX file. There is a lot of filtering of common LaTeX elements
and after that, the remaining data is stored in a structured format to
be written to a file.

``` r
bib_entry <- rebib:::bib_handler(bib_items)
bib_entry
#> $book
#> $book[[1]]
#> $book[[1]]$unique_id
#> [1] "ihaka:1996"
#> 
#> $book[[1]]$author
#> [1] "Ihaka, Ross and Gentleman, Robert"
#> 
#> $book[[1]]$title
#> [1] "R: A Language for Data Analysis and Graphics"
#> 
#> $book[[1]]$journal
#> [1] "Journal of Computational and Graphical Statistics 3:   :"
#> 
#> $book[[1]]$year
#> [1] "1996"
#> 
#> $book[[1]]$URL
#> [1] "https://doi.org/10.1080/10618600.1996.10474713"
#> 
#> $book[[1]]$isbn
#> NULL
#> 
#> $book[[1]]$pages
#> [1] "299--314"
#> 
#> 
#> $book[[2]]
#> $book[[2]]$unique_id
#> [1] "R"
#> 
#> $book[[2]]$author
#> [1] "R Core Team"
#> 
#> $book[[2]]$title
#> [1] "R: A Language and Environment for Statistical Computing"
#> 
#> $book[[2]]$journal
#> [1] "R Foundation for Statistical Computing Vienna Austria    :"
#> 
#> $book[[2]]$year
#> [1] "2016"
#> 
#> $book[[2]]$URL
#> [1] "https://www.R-project.org/"
#> 
#> $book[[2]]$isbn
#> [1] "3-900051-07-0"
#> 
#> $book[[2]]$pages
#> NULL
```

#### Stage 3: BibTeX writer

After reading the bibliographic entries and splitting out meaningful
values from them, we can finally write a structured file in the BibTeX
format.

The writer will read the bib chunks one at a time based on the metadata
extracted and will write the corresponding data fields. The default
entry type is a book, which should not have any problems with the web
articles.

``` r
file_path <- paste0(your_article_path,"/",file_name)
bib_path <- paste0(your_article_path,"/example.bib")
rebib:::bibtex_writer(bib_entry,file_path)
cat(readLines(paste(your_article_path,"example.bib",sep="/")),sep = "\n")
#> @book{ihaka:1996,
#> author = {{Ihaka, Ross and Gentleman, Robert}},
#> title = {{R: A Language for Data Analysis and Graphics}},
#> publisher = {Journal of Computational and Graphical Statistics 3:   :},
#> pages = {299--314},
#> year = {1996},
#> url = {https://doi.org/10.1080/10618600.1996.10474713}
#> }
#> @book{R,
#> author = {R {Core Team}},
#> title = {{R: A Language and Environment for Statistical Computing}},
#> publisher = {R Foundation for Statistical Computing Vienna Austria    :},
#> year = {2016},
#> url = {https://www.R-project.org/},
#> isbn = {3-900051-07-0}
#> }
```

Authors who are converting a bibliography from `bbl` should check if
there are any errors or missing values.

## General usage

General use of the [**rebib**](https://CRAN.R-project.org/package=rebib)
package is quite easy, requiring only one function call.

#### Using with RJournal article

In bibliography construction mode,
[**rebib**](https://CRAN.R-project.org/package=rebib) will check for
existing BibTeX files and will convert the embedded bibliography only
when there are no linked BibTeX files in the RJ article.

``` r
%% typically rebib will search for this line in the RJ article, 
%% not the RJwrapper file
\bibliography{example}
```

If there is no linked BibTeX file,
[**rebib**](https://CRAN.R-project.org/package=rebib) will begin its
procedure of converting the embedded bibliography to BibTeX and then
link it with the article file as well.

``` r
# for files without BibTeX source
rebib::handle_bibliography(your_article_path)
#> Warning in get_bib_file(article_dir, file_name): No Bib files found !
#> BibTeX file does not exist
#> will parse for bibliography
#> bibtex file created
cat(readLines(paste(your_article_path,"example.bib",sep="/")),sep = "\n")
#> @book{ihaka:1996,
#> author = {{Ihaka, Ross and Gentleman, Robert}},
#> title = {{R: A Language for Data Analysis and Graphics}},
#> publisher = {Journal of Computational and Graphical Statistics 3:   :},
#> pages = {299--314},
#> year = {1996},
#> url = {https://doi.org/10.1080/10618600.1996.10474713}
#> }
#> @book{R,
#> author = {R {Core Team}},
#> title = {{R: A Language and Environment for Statistical Computing}},
#> publisher = {R Foundation for Statistical Computing Vienna Austria    :},
#> year = {2016},
#> url = {https://www.R-project.org/},
#> isbn = {3-900051-07-0}
#> }
```

#### Bibliography aggregation

If you have a BibTeX file and it is missing some references then this
mode can help you read the embedded bibliography and combine the two
bibliographies in a single BibTeX file.

``` r
# for files with BibTeX source as well as embedded entries
rebib::aggregate_bibliography(your_article_path)
#> Found Bib file  example.bib
#> bibliography aggregation possible
#> Found Bib file  example.bib
#> aggregation delta : 2
#> BibTeX file aggregated
```

#### Using with any LaTeX/.bbl file

The bibliography converter will convert any LaTeX or `.bbl` file given
as a path. This way, one can generate BibTeX from the embedded
bibliography or `.bbl` file.

``` r
rebib::biblio_converter(path_to_your_file)
#> working directory : /temp/Rtmpt6Qk0v/exampledir/standalone
#> bib entries : 6
#> Written BibTeX file : /temp/Rtmpt6Qk0v/exampledir/standalone/sample.bib
```

## Limitations

The limitations of [**rebib**](https://CRAN.R-project.org/package=rebib)
package are:

-   The [**rebib**](https://CRAN.R-project.org/package=rebib) package
    only supports embedded bibliography format suggested/ commonly used
    in R Journal.

-   Encoding and character set issues might not work with rebib well.
    Try to use UTF-8 encoding if possible.

-   Bibliography aggregation won't fix existing issues like missing
    commas/brackets.

## Summary

In summary the [**rebib**](https://CRAN.R-project.org/package=rebib)
package supports:

-   Conversion of embedded natbib type bibliography entries used in the
    R Journal

-   Aggregation of bibliographic entries from the article as well as
    separate BibTeX files
:::

### 2025-09-07 supplement/code-env/RJwrapper.Rmd
Final response for inclusion of Rmd files.

---
title: Working with Code environments in texor
abstract: |
  This is a small sample article to demonstrate usage of texor to
  convert code environments.
author:
- name: Abhishek Ulayil
  affiliation: Student, Institute of Actuaries of India
  orcid: 0009-0000-6935-8690
  address: Mumbai, India
date: '2025-09-01'
date_received: ~
journal:
  firstpage: ~
  lastpage: ~
volume: 17
issue: 3
slug: RJ-2025-000
citation_url: https://rjournal.github.io/
packages:
  cran:
  - texor
  - knitr
  bioc: []
preview: preview.png
bibliography: code-example.bib
CTV: ~
legacy_pdf: yes
legacy_converted: yes
output:
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
    mathjax: https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-chtml-full.js
    md_extension: -tex_math_single_backslash

---

:::::: article
## Introduction

Pandoc naturally converts verbatim environment easily, however the
redefination of other commands such as example, example\*, Sinput etc to
verbatim does not work well in pandoc.

Hence, the [**texor**](https://CRAN.R-project.org/package=texor) package
uses a stream editor to search find and replace matching code
environments to verbatim before pandoc touches it.

This way the the code is not lost in conversion, also a pandoc extension
is used to add attributes to the markdown code using
`fenced_code_attributes`

Code Environment types are well summarized in the table
[1](#tab:T1){reference-type="ref" reference="table:1"}

::: {#table:1}
  ------------------------------------------------------------------------------
  Code Environment Type                                                  
  ----------------------- --------------- ----------- -------- --------- -------
  Example                 example         example\*

  S.series                Sin             Sout        Sinput   Soutput   Scode

  Special Verbatim        boxedverbatim                                  
  ------------------------------------------------------------------------------

  : (#tab:T1) Code Environment support in texor
:::

## Environments

### Verbatim Series

While verbatim is naturally supported in pandoc,other extensions of
verbatim environment like boxedverbatim from moreverb package
[@moreverb] falls back to normal verbatim.

1\. verbatim :

``` r
print("Hello world")
```

2\. boxedverbatim :

``` r
print("Hello world")
```

### S series

S series code environement is defined in Rjournal.sty file. Most of
these are extensions of verbatim environment and retain the original
formatting of the content in HTML.

1\. Sinput :

``` r
print("Hello world")
```

2\. Soutput :

``` r
[1] "hello world"
```

3\. Sin :

``` r
print("Hello world")
```

4\. Sout :

``` r
[1] "hello world"
```

### Example series

Example series of code environment is defined in Rjournal.sty file.
Examples are extensions of verbatim environment and retain the original
formatting of the content in HTML.

1\. example :

``` r
print("Hello world")
```

2\. example\* :

``` r
print("Hello world")
```

## Code in Figure Environments

A small example of this is visble in
[1](#code:example){reference-type="ref" reference="code:example"}. This
is a common practice in Rnews articles as it used to add a boxed border
around the code which looks attractive. However, in web articles there
isn't much advantage to it.

<figure id="code:example">
<div class="center">
<div class="sourceCode" id="cb1"><pre class="sourceCode r"><code class="sourceCode r"><span id="cb1-1"><a href="#cb1-1" aria-hidden="true" tabindex="-1"></a>code_in_figure <span class="ot">&lt;-</span> <span class="cf">function</span>() {</span>
<span id="cb1-2"><a href="#cb1-2" aria-hidden="true" tabindex="-1"></a>  <span class="cf">if</span> (pandoc_version <span class="sc">&gt;=</span> <span class="dv">3</span>) {</span>
<span id="cb1-3"><a href="#cb1-3" aria-hidden="true" tabindex="-1"></a>    <span class="fu">print</span>(<span class="st">&quot;Code in Figure Supported&quot;</span>)</span>
<span id="cb1-4"><a href="#cb1-4" aria-hidden="true" tabindex="-1"></a>  }</span>
<span id="cb1-5"><a href="#cb1-5" aria-hidden="true" tabindex="-1"></a>  <span class="cf">else</span> {</span>
<span id="cb1-6"><a href="#cb1-6" aria-hidden="true" tabindex="-1"></a>    <span class="fu">print</span>(<span class="st">&quot;code in Figure not supported&quot;</span>)</span>
<span id="cb1-7"><a href="#cb1-7" aria-hidden="true" tabindex="-1"></a>  }</span>
<span id="cb1-8"><a href="#cb1-8" aria-hidden="true" tabindex="-1"></a>}</span></code></pre></div>
</div>
<figcaption>Figure 1: Example Code inside Figure
environment</figcaption>
</figure>

Pandoc v3 or greater [@pandoc] has a Figure object which allows
non-image figures to be treated like one. This is why
[**texor**](https://CRAN.R-project.org/package=texor) package requires
atleast version 3 of pandoc.

## Code in Table Environments

We can use code environments in a table using minipage environments.
This is not a common practice among LaTeX article authors, but a few
articles had such complex structures. So, as a example to demonstrate
pandoc and [**texor**](https://CRAN.R-project.org/package=texor)
package's capabilities, I have included a few of them.

Table [2](#tab:T2){reference-type="ref" reference="table:2"} is an
example of code environments within a table.

::: {#table:2}
+----------+---------------------------------+
| Language | Function Defination Syntax      |
+:========:+:===============================:+
| R        | ::: minipage                    |
|          | ``` r                           |
|          | fun <- function(){              |
|          |   print("A function in R")      |
|          |   return(0)                     |
|          | }                               |
|          | ```                             |
|          | :::                             |
+----------+---------------------------------+
| Python   | ::: minipage                    |
|          | ``` r                           |
|          | def fun():                      |
|          |   print("A function in Python") |
|          | ```                             |
|          | :::                             |
+----------+---------------------------------+
| Lua      | ::: minipage                    |
|          | ``` r                           |
|          | function fun()                  |
|          |   print("A function in Lua")    |
|          | end                             |
|          | ```                             |
|          | :::                             |
+----------+---------------------------------+

: (#tab:T2) Code in a table
:::

A similar arrangement can be had for figures/plots besides code
environment. Table [3](#tab:T3){reference-type="ref"
reference="table:3"} demonstrates a table with code and figure.

::: {#table:3}
+-------------------------------+-------------------------------------+
| Code                          | Plot                                |
+:=============================:+:===================================:+
| ::: minipage                  | ::: minipage                        |
| ``` r                         | ![image](binom.png){width="100%"    |
| x <- 1:100                    | alt="graphic without alt text"}     |
| y <- dbinom(x,100,prob = 0.5) | :::                                 |
| plot(x,y)                     |                                     |
| ```                           |                                     |
| :::                           |                                     |
+-------------------------------+-------------------------------------+

: (#tab:T3) Code and Plot side by side
:::

## Inline Code usage

Using inline code in LaTeX is possible using `\verb` command. It would
be repoduced similarly, as a Inline code element.

``` r
\verb|x <- 1:100|
```

will be represented as `x <- 1:100` in Inline format.

## Code chunks using Schunk

Code chunks within an `Schunk` environment to demonstrate Input/Output

Input :

``` r
print("Hello world")
```

Output :

``` r
[1] "hello world"
```

Similar arrangement can be had for Plots as well using figure
environnment.

Input :

``` r
x <- 1:100
y <- dbinom(x,100,prob = 0.5)
plot(x,y)
```

Output :\
![image](binom.png){width="100%" alt="graphic without alt text"}

## Including 'R code chunks'

Including R code chunks in an article that is to be converted to R
markdown can be a bit tricky as the
[**knitr**](https://CRAN.R-project.org/package=knitr) package will
interpret and execute the R code chunks as actual code bits. A
workaround for this issue has been described in this [posit support
article](https://support.posit.co/hc/en-us/articles/360018181633-Including-verbatim-R-code-chunks-inside-R-Markdown).
Below is an example with the workaround adapted from [@rcodechunks].

```` r
`r ''````{r}
plot(cars)
```
````

## Summary

In summary the [**texor**](https://CRAN.R-project.org/package=texor)
package supports:

-   Almost all code environments in RJournal.

-   Code Highlight for R language.

-   Inline Code.

-   Code in different environments like tables/figures.
::::::

### 2025-09-07 supplement/figure-env/RJwrapper.Rmd
Final response for inclusion of Rmd files.

@@
 title: Working with figure environments in texor
-abstract: This is a small sample article to demonstrate usage of texor to convert
+abstract: |
+  This is a small sample article to demonstrate usage of
+  [**texor**](https://CRAN.R-project.org/package=texor) to convert
   figure environments.
@@
   address: Mumbai, India
-date: '2009-06-01'
+date: '2025-09-01'
 date_received: ~
@@
   lastpage: ~
-volume: 1
-issue: 1
-slug: figure-env
+volume: 17
+issue: 3
+slug: RJ-2025-000
+citation_url: https://rjournal.github.io/
 packages:
@@
 CTV: ~
+legacy_pdf: yes
+legacy_converted: yes
 output:
@@
     toc: no
-    legacy_pdf: yes
-    web_only: yes
     mathjax: https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-chtml-full.js
@@
 
-::: article
-# Introduction
+:::: article
+## Introduction
 
-Images are an essential component in any article, However due to the
-differences in the support for various graphic formats between LaTeX and
-markdown/HTML we need to fallback on raster graphics. It is well
-summarized in the table [1](#table:1)
+Images are an essential component of any article, however, due to the
+differences in support for various graphic formats between LaTeX and
+markdown/HTML we need to fall back on raster graphics. The support for
+different image formats across markup languages is summarized in
+Table [1](#tab:T1){reference-type="ref" reference="table:1"}.
 
@@
 
-  : Table 1: Image Format support in various Markup/Typesetting
+  : (#tab:T1) Image Format support in various Markup/Typesetting
   Languages
@@
 
-# Image with width parameters
+## Image with width parameters
 
-This section may contain a figure such as Figure [1](#figure:rlogo).
-This is the most basic example of figure.
+The following code includes an image with width parameters, producing
+the output in Figure [1](#figure:rlogo){reference-type="ref"
+reference="figure:rlogo"}.
 
@@
 
-![Figure 1: The logo of R.](Rlogo-5.png){#figure:rlogo width="35.0%"}
+![Figure 1: The logo of R.](Rlogo-5.png){#figure:rlogo width="35.0%"
+alt="graphic without alt text"}
 
-# Images in PDF format
+This is the most basic example of figure. Similarly PDF figures will be
+included as rasterized images.
 
-Image [2](#fig:normal) is a graphical representation of normal
-distribution.
+## Multiple images
 
-``` r
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5\textwidth]{normal}
-  \caption{PDF of a normal distribution}
-  \label{fig:normal}
-\end{figure}
-```
+Pandoc v3 and above now support a new `Figure` object [@pandoc] which
+supports multiple images side by side or in a grid format. The following
+subsections demonstrate these capabilities.
 
-![Figure 2: PDF of a normal distribution](normal.png){#fig:normal
-width="50.0%"}
-
-# Multiple images
-
-Pandoc v3 and above now supports a new Figure object [@pandoc] which
-supports multiple images side by side or in a grid format.
-
-### Two or more Images side by side
+#### Two or more Images side by side
 
@@
 <figure id="fig:twoimages">
-<p><img src="Rlogo-5.png" style="width:45.0%" alt="image" /><img
-src="normal.png" style="width:45.0%" alt="image" /></p>
-<figcaption>Figure 3: Images side by side</figcaption>
+<p><img src="Rlogo-5.png" style="width:45.0%"
+alt="graphic without alt text" /><img src="normal.png"
+style="width:45.0%" alt="graphic without alt text" /></p>
+<figcaption>Figure 2: Images side by side</figcaption>
 </figure>
 
-### Four Images in a grid
+#### Four Images in a grid
 
@@
 <figure id="fig:fourimages">
-<p><img src="Rlogo-5.png" style="width:45.0%" alt="image" /><img
-src="normal.png" style="width:45.0%" alt="image" /></p>
-<p><img src="normal.png" style="width:45.0%" alt="image" /><img
-src="Rlogo-5.png" style="width:45.0%" alt="image" /></p>
-<figcaption>Figure 4: Multiple images in a grid</figcaption>
+<p><img src="Rlogo-5.png" style="width:45.0%"
+alt="graphic without alt text" /><img src="normal.png"
+style="width:45.0%" alt="graphic without alt text" /></p>
+<p><img src="normal.png" style="width:45.0%"
+alt="graphic without alt text" /><img src="Rlogo-5.png"
+style="width:45.0%" alt="graphic without alt text" /></p>
+<figcaption>Figure 3: Multiple images in a grid</figcaption>
 </figure>
 
-# Tikz images
+## Tikz images
 
-Here is a tikz image example in fig [5](#fig:tikz) adapted from
-[@casflow].
+The **texor** package supports tikz images by rasterizing it and making
+it web friendly. Figure [4](#fig:tikz){reference-type="ref"
+reference="fig:tikz"} shows a tikz image adapted from [@casflow].
 
-Another interesting aspect of including tikz image here is that you can
-modify the source and re-convert without making any other changes.
-Figure will get updated in the generated html article.
+#### Tikz Code:
 
-### Tikz Code:
+The image in Figure [4](#fig:tikz){reference-type="ref"
+reference="fig:tikz"} is a graphical representation of how **texor**
+handles tikz images:
 
-``` r
-\begin{figure}
-\includegraphics{tikz/figtikz.png}
+-   Figures containing tikz images are isolated.
 
-%% Generated Image will included as a PNG above automatically
-  \centering
-\tikzstyle{process} = [rectangle, rounded corners,
-minimum width=3cm, 
-minimum height=1cm,
-text centered, 
-draw=black]
-\tikzstyle{arrow} = [thick,->,>=stealth]
-\begin{tikzpicture}[node distance=4cm]
-%Nodes
-\node (start) [process] {LaTeX Source with Tikz Image};
-\node (isolate) [process, below right of=start] {Isolated tikz env};
-\node (pro1) [process, below left of=isolate] {Standalone PDF file};
-\node (pro2) [process, below right of=pro1, yshift=-0.5cm] {PNG file};
-\node (link) [process, below left of=pro2] {Include PNG in article};
-\node (stop) [process, below right of=link] {Convert To RJ-web-article};
-% arrows
-\draw [arrow] (start) -- node[anchor=east] {Extract Figure Block} (isolate);
-\draw [arrow] (isolate) -- node[anchor=east] {Compile} (pro1);
-\draw [arrow] (pro1) -- node[anchor=east] {Convert to raster} (pro2);
-\draw [arrow] (pro2) -- node[anchor=east] {Include back to source as image} (link);
-\draw [arrow] (link) -- node[anchor=east] {Using Pandoc} (stop);
-\end{tikzpicture}
-\caption{Tikz Image example}
-  \label{fig:tikz}
-\end{figure}
-```
+-   Tikz libraries are fetched from the wrapper file.
 
-### Resultant Figure :
+-   The Tikz code section is isolated into a standalone LaTeX file and
+    compiled.
+
+-   The compiled LaTeX file generates a PDF Image.
+
+-   This is converted to PNG format for embedding in the HTML output.
+
+-   `\includegraphics{tikz/somefigure.png}` is added to the figure
+    environment.
+
+-   A lua filter removes redundant text the figure environment.
+
+If you use **texor** to convert your articles using
+`texor::latex_to_web()` with `temp_mode=TRUE`(it is `TRUE` by default).
+The resultant Rmarkdown/HTML file will not modify the contents of your
+LaTeX file. In this case you can keep reloading the article after making
+changes to the tikz images, without having to do the above steps
+manually in case you are converting the article by hand.
 
 <figure id="fig:tikz">
-<img src="tikz/figtikz.png" style="width:100.0%" />
+<img src="tikz/figtikz.png" style="width:100.0%"
+alt="graphic without alt text" />
 
-<figcaption>Figure 5: Tikz Image example</figcaption>
+<figcaption>Figure 4: Tikz Image example</figcaption>
 </figure>
 
-# Algorithm2e diagrams
+## Algorithm2e diagrams
 
-we do support algorithm2e diagrams and images, these will be numbered
-differently and we strongly suggest to use \"alg:\" in labels for best
-results. Here is an example of algorithm [6](#alg:how), referenced from
-[@algoexample]
+Diagrams and images using the `algorithm2e` environment are supported,
+these will be numbered differently. We strongly suggest to use \"alg:\"
+in labels for best results. As a part of the filtering requires \"alg:\"
+to number algorithm references seperately, otherwise they will share
+references with normal figures.
 
-``` r
-\begin{figure}[htbp]
-\includegraphics{alg/alghow.png}
-
-\SetAlgoLined
-\KwData{this text}
-\KwResult{how to write algorithm with \LaTeX2e }
-initialization\;
-\While{not at end of this document}{
-read current\;
-\eIf{understand}{
-go to next section\;
-current section becomes this one\;
-}{
-go back to the beginning of current section\;
-}
-}
-\caption{How to write algorithms}
-  \label{alg:how}
-\end{figure}
-```
+Algorithm [1](#alg:how){reference-type="ref" reference="alg:how"} is an
+example from the **algorithm2e** vignette [@algoexample].
 
 <figure id="alg:how">
-<img src="alg/alghow.png" style="width:100.0%" />
-<p>initialization</p>
+<img src="alg/alghow.png" style="width:100.0%"
+alt="graphic without alt text" />
+
 <figcaption>Algorithm 1: How to write algorithms</figcaption>
@@
 
-# Summary
+## Other elements in figure objects
 
-In summary the [*texor*](https://CRAN.R-project.org/package=texor)
+Figures can also house non-image environments like code blocks. Code
+blocks in figure environments would share numbering with normal figures,
+similar to LaTeX. Here Figure [5](#code:example){reference-type="ref"
+reference="code:example"} refers to a code block.
+
+<figure id="code:example">
+<div class="center">
+<div class="sourceCode" id="cb1"><pre class="sourceCode r"><code class="sourceCode r"><span id="cb1-1"><a href="#cb1-1" aria-hidden="true" tabindex="-1"></a>code_in_figure <span class="ot">&lt;-</span> <span class="cf">function</span>() {</span>
+<span id="cb1-2"><a href="#cb1-2" aria-hidden="true" tabindex="-1"></a>  <span class="cf">if</span> (pandoc_version <span class="sc">&gt;=</span> <span class="dv">3</span>) {</span>
+<span id="cb1-3"><a href="#cb1-3" aria-hidden="true" tabindex="-1"></a>    <span class="fu">print</span>(<span class="st">&quot;code in figure supported&quot;</span>)</span>
+<span id="cb1-4"><a href="#cb1-4" aria-hidden="true" tabindex="-1"></a>  }</span>
+<span id="cb1-5"><a href="#cb1-5" aria-hidden="true" tabindex="-1"></a>  <span class="cf">else</span> {</span>
+<span id="cb1-6"><a href="#cb1-6" aria-hidden="true" tabindex="-1"></a>    <span class="fu">print</span>(<span class="st">&quot;code in figure not supported&quot;</span>)</span>
+<span id="cb1-7"><a href="#cb1-7" aria-hidden="true" tabindex="-1"></a>  }</span>
+<span id="cb1-8"><a href="#cb1-8" aria-hidden="true" tabindex="-1"></a>}</span></code></pre></div>
+</div>
+<figcaption>Figure 5: Example Code inside Figure
+environment</figcaption>
+</figure>
+
+## Limitations
+
+The limitations of [**texor**](https://CRAN.R-project.org/package=texor)
+package in figure handling are:
+
+-   Animations using the **animation** package in LaTeX wont be
+    supported. The authors can include GiF equivalents of the same in
+    the Rmarkdown/web article.
+
+-   Multi-page PDF files are currently not supported.
+
+-   Usage of subfigure,sclaebox,wrapfigure environments will not work as
+    intended.
+
+-   Currently only `[width=X.X\textwdith]` format is supported for
+    defining the width in web article (default width would be 100% in
+    other cases).
+
+-   Other `\inclduegraphics[..]` options will be ignored.
+
+## Summary
+
+In summary the [**texor**](https://CRAN.R-project.org/package=texor)
 package supports:
-
--   Almost all image formats in LaTeX.
-
--   Algorithm and tikz as well in some capacity.
 
@@
 -   Image Captions with Numbering and Labelling.
-:::
+
+-   Algorithm2e and tikz figures.
+::::

### 2025-09-07 supplement/lua-filters/RJwrapper.Rmd
Final response for inclusion of Rmd files.

---
title: Understanding pandoc lua filters
abstract: |
  Pandoc supports intermediate modification of the Abstract Syntax Tree
  (AST) between the parsing and writing phase using filters. This
  supplement paper highlights the use cases of such filters written in
  the Lua language on LaTeX, markdown and native AST.
author:
- name: Abhishek Ulayil
  affiliation: Student, Institute of Actuaries of India
  orcid: 0009-0000-6935-8690
  address: Mumbai, India
date: '2025-09-01'
date_received: ~
journal:
  firstpage: ~
  lastpage: ~
volume: 17
issue: 3
slug: RJ-2025-000
citation_url: https://rjournal.github.io/
packages:
  cran:
  - texor
  - kable
  - bookdown
  bioc: []
preview: preview.png
bibliography: lua-filters.bib
CTV: ~
legacy_pdf: yes
legacy_converted: yes
output:
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
    mathjax: https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-chtml-full.js
    md_extension: -tex_math_single_backslash

---

:::: article
## Introduction

To fully understand pandoc filters, one must first grasp the document
conversion process. Pandoc will read/parse an article and then store it
internally as an intermediate form, known as an abstract syntax tree
(AST). Then, A document writer(a specialized type-setter) will read the
AST and produce the contents in the chosen article format. Converting
markdown to HTML, for example, will entail these pandoc operations.
Figure [1](#fig:workflow){reference-type="ref" reference="fig:workflow"}
shows the conversion workflow within pandoc.

![Figure 1: Conversion workflow.](figures/workflow.png){#fig:workflow
width="60.0%" alt="graphic without alt text"}

When specific elements require adjustment, the Lua filters enter the
picture. Assume you're converting markdown to HTML and want the page to
have automated numbering for figures or tables. If you use markdown,
there is no system for automatically numbering figures/tables; to retain
1:1 conversion, the HTML output will also lack numbering.

At this point, you might wish there was a way to add such a feature. But
including every customization as an option in pandoc or the writer is
not viable. To address this, the idea of filters emerged, with which you
could modify the AST before the writer could read it to obtain the
desired result [@pandocfilters].

Pandoc filters can technically be written in any language (called JSON
filters) [^1].

## Need of Lua filters

Continuing the example above, let's create a dummy article, where we
need to add numbering. One way could be to keep a counter of images, and
add a prefix `"Figure X :"` to each figure caption. This will serve the
purpose of numbering the images in the end result.

``` r
![R logo](Rlogo-5.png){width="10%"}

![penguins](penguins.png){width="15%"}
```

<figure id="code:2">
<div class="sourceCode" id="cb1"><pre class="sourceCode r"><code class="sourceCode r"><span id="cb1-1"><a href="#cb1-1" aria-hidden="true" tabindex="-1"></a>pandoc example.md <span class="sc">--</span>from markdown <span class="sc">--</span>to html5 <span class="sc">--</span>output example.html</span></code></pre></div>
<figcaption>Figure 2: pandoc command without lua filter or
extensions.</figcaption>
</figure>

Now if we convert the above markdown file to HTML5 using the pandoc
command in Figure [2](#code:2){reference-type="ref" reference="code:2"},
we get Figure [3](#fig:1){reference-type="ref" reference="fig:1"}.

![Figure 3: Vanilla HTML5 output.](figures/example.png){#fig:1
width="50%"}

As we can see, there is no figure numbering done automatically, which is
generally the expected result. If we want to include numbering, we would
need to write a Lua filter. This Lua filter will modify the AST and make
the changes we desire.

We call a Lua filter in the pandoc command show in
Figure [2](#code:2){reference-type="ref" reference="code:2"} using the
` --lua-filter name_of_filter.lua` option in pandoc.

In the next section we write a Lua filter to manipulate the figures in
Figure [4](#code:3){reference-type="ref" reference="code:3"}.

## Writing pandoc Lua filters

<figure id="code:3">
<div class="sourceCode" id="cb1"><pre class="sourceCode r"><code class="sourceCode r"><span id="cb1-1"><a href="#cb1-1" aria-hidden="true" tabindex="-1"></a>figures <span class="ot">=</span> <span class="dv">0</span></span>
<span id="cb1-2"><a href="#cb1-2" aria-hidden="true" tabindex="-1"></a>is_fig <span class="ot">=</span> <span class="dv">0</span></span>
<span id="cb1-3"><a href="#cb1-3" aria-hidden="true" tabindex="-1"></a></span>
<span id="cb1-4"><a href="#cb1-4" aria-hidden="true" tabindex="-1"></a><span class="cf">function</span> <span class="fu">Figure</span>(el)</span>
<span id="cb1-5"><a href="#cb1-5" aria-hidden="true" tabindex="-1"></a>      local label <span class="ot">=</span> <span class="st">&quot;&quot;</span></span>
<span id="cb1-6"><a href="#cb1-6" aria-hidden="true" tabindex="-1"></a>      <span class="fu">pandoc.walk_block</span>(el,{ Image <span class="ot">=</span> <span class="cf">function</span>(el)</span>
<span id="cb1-7"><a href="#cb1-7" aria-hidden="true" tabindex="-1"></a>                     is_fig <span class="ot">=</span> <span class="dv">1</span></span>
<span id="cb1-8"><a href="#cb1-8" aria-hidden="true" tabindex="-1"></a>                     end})</span>
<span id="cb1-9"><a href="#cb1-9" aria-hidden="true" tabindex="-1"></a>      <span class="cf">if</span> is_fig <span class="sc">==</span> <span class="dv">1</span> then</span>
<span id="cb1-10"><a href="#cb1-10" aria-hidden="true" tabindex="-1"></a>	        figures <span class="ot">=</span> figures <span class="sc">+</span> <span class="dv">1</span></span>
<span id="cb1-11"><a href="#cb1-11" aria-hidden="true" tabindex="-1"></a>	        label <span class="ot">=</span> <span class="st">&quot;Figure &quot;</span> .. <span class="fu">tostring</span>(figures) .. <span class="st">&quot;:&quot;</span></span>
<span id="cb1-12"><a href="#cb1-12" aria-hidden="true" tabindex="-1"></a>      end</span>
<span id="cb1-13"><a href="#cb1-13" aria-hidden="true" tabindex="-1"></a>	    local caption <span class="ot">=</span> el.caption</span>
<span id="cb1-14"><a href="#cb1-14" aria-hidden="true" tabindex="-1"></a>	    <span class="cf">if</span> not caption then</span>
<span id="cb1-15"><a href="#cb1-15" aria-hidden="true" tabindex="-1"></a>          caption <span class="ot">=</span> {<span class="fu">pandoc.Str</span>(label)}</span>
<span id="cb1-16"><a href="#cb1-16" aria-hidden="true" tabindex="-1"></a>    	<span class="cf">else</span></span>
<span id="cb1-17"><a href="#cb1-17" aria-hidden="true" tabindex="-1"></a>          caption <span class="ot">=</span> {<span class="fu">pandoc.Str</span>(label),<span class="fu">pandoc.Space</span>()}</span>
<span id="cb1-18"><a href="#cb1-18" aria-hidden="true" tabindex="-1"></a>    	end</span>
<span id="cb1-19"><a href="#cb1-19" aria-hidden="true" tabindex="-1"></a>    	el.caption.long[<span class="dv">1</span>].content <span class="ot">=</span> caption .. el.caption.long[<span class="dv">1</span>].content</span>
<span id="cb1-20"><a href="#cb1-20" aria-hidden="true" tabindex="-1"></a>    	is_fig <span class="ot">=</span> <span class="dv">0</span></span>
<span id="cb1-21"><a href="#cb1-21" aria-hidden="true" tabindex="-1"></a>    	return el</span>
<span id="cb1-22"><a href="#cb1-22" aria-hidden="true" tabindex="-1"></a>end</span></code></pre></div>
<figcaption>Figure 4: A Lua filter to add figure numbering.</figcaption>
</figure>

Everything in Lua revolves around tables; for example, the pandoc AST or
the document is one giant table with sub-tables.

In pandoc, there are two sorts of elements: '`Blocks`' and '`Inlines`'.
'`Blocks`' are complicated constructions constructed from simpler pieces
('`Inlines`'). A '`Para`', for example, is a '`Block`' made up of
several '`Str`' '`Inlines`'.

The first step in writing a filter is to select a target type, that is,
which '`Block`' or '`Inline`' this filter will target. Following the
selection, we name the function after the target type and each element
in the argument as '`el`'. For instance, in the case of our filter, it
would be `function Figure(el)`. Pandoc is smart enough to match the
names of filter functions to the AST elements.

Now within the filter function we can assume that `el` will contain the
table object of a Figure element from the document. We can use many
functions over it or '`Inlines`' contained in it. One such function to
walk over all the elements inside the block is `walk_block`. Walk
functions are a great way to check or count the presence of certain
elements in a block. We use `walk_block` to check if there are any
'`Image`' inline elements within the '`Figure`' block. This is because
after pandoc 3 [@pandoc] '`Figure`' blocks can now contain elements
other than '`Images`' as well.

If there is an `Image` then we append `"Figure X :"` to the caption of
the `Figure` element. and return the modified element.

## Interpreting changes using Lua filters

The filter in Figure [4](#code:3){reference-type="ref"
reference="code:3"} when included in the pandoc command will generate
Figure [6](#fig:2){reference-type="ref" reference="fig:2"} using the
command in Figure [5](#code:4){reference-type="ref" reference="code:4"}.

<figure id="code:4">
<div class="sourceCode" id="cb1"><pre class="sourceCode r"><code class="sourceCode r"><span id="cb1-1"><a href="#cb1-1" aria-hidden="true" tabindex="-1"></a>pandoc example.md <span class="sc">--</span>from markdown <span class="sc">--</span>to html5 <span class="sc">--</span>output filtered<span class="sc">-</span>example.html</span>
<span id="cb1-2"><a href="#cb1-2" aria-hidden="true" tabindex="-1"></a><span class="sc">--</span>lua<span class="sc">-</span>filter image_numbering_filter.lua</span></code></pre></div>
<figcaption>Figure 5: pandoc command with Lua filter.</figcaption>
</figure>

![Figure 6: HTML5 output with desired
filtering.](figures/example-filtered.png){#fig:2 width="50%"}

## Usage of pandoc lua filters in the [**texor**](https://CRAN.R-project.org/package=texor) package

Pandoc Lua filters are used in the
[**texor**](https://CRAN.R-project.org/package=texor) package to modify
the AST for various markups. Table [1](#tab:T1){reference-type="ref"
reference="tab:filters"} summarizes the use of each filter.

::: {#tab:filters}
  --------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
  File Name                       Description
  ------------------------------- ------------------------------------------------------------------------------------------------------------------------------------------------------
  `abs_filter.lua`                Filters out unicode `182` character.

  `bib_filter.lua`                Clears out the bibliography from the article itself as it will be added to the metadata.

  `R_code.lua`                    Adds a class component `'r'` to CodeBlocks for code highlighting.

  `image_filter.lua`              Fixes extensions for image paths without an extension or with `.pdf` extension.

  `table_caption.lua`             Adds table numbering in the captions similar to the ones found in R Markdown and [**kable**](https://CRAN.R-project.org/package=kable)-based tables.

  `image_caption.lua`             Adds Figure/Algorithm numbering as well as clears out residuals of tikz/algorithm images.

  `conversion_compat_check.lua`   This filter keeps a count of all `Inline` and `Block` elements and writes it to a yaml file.

  `equation_filter.lua`           Adds [**bookdown**](https://CRAN.R-project.org/package=bookdown) style equation numbering to LaTeX equation/math environments.

  `bookdown_ref.lua`              Corrects numbering for various elements.

  `issue_checker.lua`             Searches and notifies any occurrence of unrecognized math commands.

  `find_pdf_files.lua`            This filter creates a list for all image PDF files included in the article.

  `widetable_patcher.lua`         This filter sets the representation for widetables.
  --------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

  : (#tab:T1) A summary of all Lua filters used in the
  [**texor**](https://CRAN.R-project.org/package=texor) package
  (available in the
  [/inst](https://github.com/Abhi-1U/texor/tree/master/inst) folder of
  the package).
:::
::::

[^1]: More on this [here](https://pandoc.org/filters.html).

### 2025-09-07 supplement/math-env/RJwrapper.Rmd
Final response for inclusion of Rmd files.

---
title: Working with Math environments in texor
abstract: |
  This is a small sample article to demonstrate usage of the texor
  package to convert math environments.
author:
- name: Abhishek Ulayil
  affiliation: Student, Institute of Actuaries of India
  orcid: 0009-0000-6935-8690
  address: Mumbai, India
date: '2025-09-01'
date_received: ~
journal:
  firstpage: ~
  lastpage: ~
volume: 17
issue: 3
slug: RJ-2025-000
citation_url: https://rjournal.github.io/
packages:
  cran:
  - texor
  - rmarkdown
  - bookdown
  bioc: []
preview: preview.png
bibliography: math-example.bib
CTV: ~
legacy_pdf: yes
legacy_converted: yes
output:
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
    mathjax: https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-chtml-full.js
    md_extension: -tex_math_single_backslash

---

:::: article
## Introduction

Math typesetting has always been a highlight of LaTeX, making it a de
facto choice among academics and researchers globally. However, in the
move to humble web interfaces, math has been traditionally hard to
describe. There have been advances in JavaScript libraries to better
typeset and present math in web pages but not all LaTeX commands/math
functions are available.

#### MathJax

The [**texor**](https://CRAN.R-project.org/package=texor) package uses
Mathjax version 3 to enhance the visual look of the math content in
HTML.

The core goal of the MathJax project is the development of its
state-of-the-art, open source, JavaScript platform for display of
mathematics. The key advantages are [@mathjax]:

-   High-quality display of mathematics notation in all browsers.

-   No special browser setup required.

-   Support for LaTeX, MathML, and other equation markup directly in the
    HTML source.

-   Support for accessibility, copy and paste, and other rich
    functionality.

-   Interoperability with other applications and math-aware search.

As [**texor**](https://CRAN.R-project.org/package=texor) calls
[**rmarkdown**](https://CRAN.R-project.org/package=rmarkdown) to render
the R Markdown file into HTML, the `rjtools::rjournal_web_article`
template uses MathJax as the math engine by default. We also specify the
Mathjax version in the metadata of the generated Rmarkdown file.

## Inline math

One can define inline math in LaTeX using commands `\(..\)` or `$..$`.

::: {#table:1}
+-------------------------+----------------------+
| Command                 | Math                 |
+:=======================:+:====================:+
| ::: minipage            | ::: minipage         |
| ``` r                   | $\mu = (0,0,0)^\top$ |
| \(\mu = (0,0,0)^\top \) | :::                  |
| ```                     |                      |
| :::                     |                      |
+-------------------------+----------------------+
| ::: minipage            | ::: minipage         |
| ``` r                   | $\mu = (0,0,0)^\top$ |
| $\mu = (0,0,0)^\top $   | :::                  |
| ```                     |                      |
| :::                     |                      |
+-------------------------+----------------------+

: (#tab:T1) Inline math syntax and its output side by side.
:::

## Display math

Display math refers to equations typeset on separate lines rather than
in line with the text. Almost all LaTeX equations can be rendered by
MathJax.

``` r
\begin{align}
f(x) = \frac{1}{\sigma\sqrt{2\pi}} 
  \exp\left( -\frac{1}{2}\left(\frac{x-\mu}{\sigma}\right)^{\!2}\,\right)
  \label{eq:1}
\end{align}
```

The code above will render as Equation \@ref(eq:1).

$$\begin{aligned}
\label{eq:1}
f(x) = \frac{1}{\sigma\sqrt{2\pi}} 
  \exp\left( -\frac{1}{2}\left(\frac{x-\mu}{\sigma}\right)^{\!2}\,\right)
\end{aligned}   (\#eq:1)$$

## Equation numbering

In LaTeX your equations get numbered automatically unless you are using
a `displaymath` environment (equivalently `\[..\]`), a starred math
environment such as `equation*`, or `\nonumber` to suppress numbering.
Equation numbering works a bit differently in
[**bookdown**](https://CRAN.R-project.org/package=bookdown) (the base of
the R Journal web article format) where it is mandatory to have a
`(\#eq:xx)` which is described in more detail in @bookdown.

The [**texor**](https://CRAN.R-project.org/package=texor) package
relieves authors from manually adding `(\#eq:xx)` to equations in R
Markdown by using a pandoc Lua filter to convert existing `\label{..}`
in the equations to `(\#eq:xx)` during conversion. Equation \@ref(eq:xe)
shows such a use case of an equation being numbered by its label in R
Markdown as well as LaTeX.

$$\begin{aligned}
S_{T, s}(z_t) = X^{\top} K_{b,t}^* X (Z - z_t)^s, 
\label{eq:xe}
\end{aligned}   (\#eq:xe)$$

Equation labels must start with the prefix `eq:` in bookdown. All labels
in bookdown must only contain alphanumeric characters, `:`, `-`, and/or
`/` as suggested in [@bookdown]. To accomadate this, the **texor**
package has a lua filter implementation to correct and modify equation
labels and references to the bookdown accepted format.

##### LaTeX source code

``` r
\begin{equation}\label{binomial pdf}
  f\left(k\right) = \binom{n}{k} p^k\left(1-p\right)^{n-k}

\end{equation}
The reference to above equation \eqref{binomial pdf}
```

##### Generated R markdown using **texor**

``` r
$$\label{binomial pdf}
  f\left(k\right) = \binom{n}{k} p^k\left(1-p\right)^{n-k}   (\#eq:binomial-pdf)$$

The reference to above equation \@ref(eq:binomial-pdf)
```

##### Rendering of the above example

$$\label{binomial pdf}
  f\left(k\right) = \binom{n}{k} p^k\left(1-p\right)^{n-k}   (\#eq:binomial-pdf)$$

The reference to above equation \@ref(eq:binomial-pdf)

## Custom math commands

The existence of user-defined LaTeX commands intended for the Math
environments will work as long as they do not contain non-math LaTeX
commands or commands from other CTAN math packages.

##### Renders correctly in HTML

``` r
\newcommand{\ABS}[1]{\left|#1\right|}
```

$$|\sigma^2| = \pm 1$$

##### Does not render correctly in HTML

``` r
\newcommand{\rotatethis}[1]{\rotatebox[origin=c]{90}{$#1$}}
```

$$\rotatebox[origin=c]{90}{\sigma^2} = \pm 1$$

## Unsupported LaTeX commands

Altough MathJax does a good job of supporting most LaTeX math functions,
some functions do not currently work. Common examples are `\bm` and
`\boldmath`, for which `\mathbf` can be used instead.

## Summary

In summary the [**texor**](https://CRAN.R-project.org/package=texor)
package, with the help of pandoc and MathJax supports:

-   Common math environments.

-   Equation numbering.
::::

### 2025-09-07 supplement/metadata/RJwrapper.Rmd
Final response for inclusion of Rmd files.

---
title: Metadata and Other Structures in texor
abstract: |
  This is a small sample article to demonstrate handling of metadata and
  other structures during conversion in texor package.
author:
- name: Abhishek Ulayil
  affiliation: Institute of Actuaries of India (Student)
  orcid: 0009-0000-6935-8690
  address: Mumbai, India
date: '2025-09-01'
date_received: ~
journal:
  firstpage: ~
  lastpage: ~
volume: 17
issue: 3
slug: RJ-2025-000
citation_url: https://rjournal.github.io/
packages:
  cran: texor
  bioc: Biobase
preview: preview.png
bibliography: metadata.bib
CTV: ~
legacy_pdf: yes
legacy_converted: yes
output:
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
    mathjax: https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-chtml-full.js
    md_extension: -tex_math_single_backslash

---

:::::::::: article
## Introduction

There is a difference in the structure of metadata and generally R
Markdown has more metadata entries than its LaTeX counterpart. So for
the conversion, the main task is to transform the metadata and include
it in the generated R Markdown file, matching the LaTeX metadata as
closely as possible.

When used with the `-s` or `standalone` mode, pandoc reads out some of
the existing meta data. The extracted metadata are listed below.

#### Metadata read from LaTeX file

-   Author names

-   Author address (affiliation)

-   ORCID number

-   Title

-   Bibliography

-   Abstract

-   CRAN/Bioconductor packages

The `\address{..}` section must be typeset in this manner for the most
effective conversion.

``` r
\address{
Abhishek Ulayil\\ % Your Name in the first Line
Institute of Actuaries of India (Student)\\% Your affiliation which includes post and/or org
Mumbai, India\\ % Your City/Country
ORCiD: 0009-0000-6935-8690\\% ORCID number with prefix ORCiD: 
}
```

#### Metadata read from DESCRIPTION file

-   Date

-   Slug

-   Volume and issue

Some metadata like the volume and issue were picked up from the folder
structure for some legacy articles, where the DESCRIPTION file was not
available.

#### Other supported commands in conversion

::: {#table:1}
  --------------------------------------------------------------------------------------------------------------------
             Command                                                    Render
  ------------------------------ -------------------------------------------------------------------------------------
          `\acronym{A}`                                                    A

               `\R`                                                        R

          `\pkg{texor}`                                                **texor**

        `\CRANpkg{texor}`                        [**texor**](https://CRAN.R-project.org/package=texor)

        `\BIOpkg{Biobase}`        [**Biobase**](https://www.bioconductor.org/packages/release/bioc/html/Biobase.html)

   `\ctv{ReproducibleResearch}`     [*ReproducibleResearch*](https://CRAN.R-project.org/view=ReproducibleResearch)

         `\command{help}`                                               `help`
  --------------------------------------------------------------------------------------------------------------------

  : (#tab:T1) Other commands supported by
  [**texor**](https://CRAN.R-project.org/package=texor) and pandoc
:::

## Miscellaneous {#misc}

#### Citations

Table [2](#tab:T2){reference-type="ref" reference="table:2"} represents
all the ways one can use citations in LaTeX and expect the generated
output.

::: {#table:2}
  ------------------------------------
          Command            Render
  ----------------------- ------------
     `\citet{pandoc}`       @pandoc

     `\citep{pandoc}`      [@pandoc]

     `\citet*{pandoc}`      @pandoc

     `\citep*{pandoc}`     [@pandoc]

    `\citeyear{pandoc}`    [-@pandoc]

   `\citeauthor{pandoc}`    @pandoc

      `\cite{pandoc}`      [@pandoc]
  ------------------------------------

  : (#tab:T2) Different forms of citation in LaTeX.
:::

#### Footnotes

Footnotes do not work in `tabular` environments in LaTeX but do work in
R Markdown tables[^1]. Custom footnote numbering works in LaTeX but does
not work in R Markdown[^2]. These features are demonstrated in
Table [3](#tab:T3){reference-type="ref" reference="table:3"}.

::: {#table:3}
  -----------------------------------------------------------------------
                             Command                              Render
  -------------------------------------------------------------- --------
   `\acronym{A}\footnote{This works in HTML, but not in LaTeX}`   A[^3]

          `\R\footnote[25]{HTML shows 4 rather than 25}`          R[^4]
  -----------------------------------------------------------------------

  : (#tab:T3) Demonstration of footnotes.
:::

#### Block quotes

> This block quote should justify an example.

#### Links

Most links work just fine using `\href` and the `\url` command will work
flawlessly, as demonstrated in Table  [4](#tab:T4){reference-type="ref"
reference="table:4"}.

::: {#table:4}
  ------------------------------------------------------------------------------------------------
               Command                                          Render
  --------------------------------- --------------------------------------------------------------
   `\href{www.google.com}{Google}`                 [Google](http://www.google.com)

       `\url{www.google.com}`               [www.google.com](http://www.google.com){.uri}

        `Table \ref{table:3}`        Table [3](#tab:T3){reference-type="ref" reference="table:3"}

      `Section \nameref{misc}`        Section [2](#misc){reference-type="ref" reference="misc"}

         `\autoref{table:3}`            [3](#tab:T3){reference-type="ref" reference="table:3"}
  ------------------------------------------------------------------------------------------------

  : (#tab:T4) Links
:::

##### Note

Certain reference commands like `\nameref{}` and `\autoref{}` may not
work as intended in HTML.

#### Lists

An example of usage of lists in LaTeX and **texor** generated HTML.

##### itemize

-   A

-   B

##### enumerate

1.  A

2.  B

##### description

**rebib**

:   Convert and Aggregate Bibliographies

**texor**

:   Converting 'LaTeX' 'R Journal' Articles into 'RJ-web-articles'

##### Note:

Some options for lists such as special numbering schemes or markups to
description item points might not work as intended and the generated
HTML might be simple.

#### Colored text

Some commands to describe colored text also work during conversion, such
as `\textcolor{blue}{Blue Text}` [Blue Text]{style="color: blue"}.

## Theorem environments

::: {#theorem:1 .theorem}
**Theorem 1**. *This is a sample theorem..*
:::

::: {#prop:1 .proposition}
**Proposition 1**. *As with $X \in \mathbb{R}$ this proposition \...*
:::

::: {#rm:1 .remark}
**Remark 1**. *[**texor**](https://CRAN.R-project.org/package=texor)
makes the article conversion easier.*
:::

**Pandoc** supports default LaTeX theorem environments. One can refer to
these environments like Theorem [1](#theorem:1){reference-type="ref"
reference="theorem:1"},Proposition [1](#prop:1){reference-type="ref"
reference="prop:1"} and Remark [1](#rm:1){reference-type="ref"
reference="rm:1"} made here. However custom theorem environments are not
supported yet.

## Summary

In summary the [**texor**](https://CRAN.R-project.org/package=texor)
package alongside pandoc supports:

-   Extracting metadata and putting it together well.

-   Handling links, block quotes, citations and footnotes.

-   Default theorem environments.
::::::::::

[^1]: Footnotes in paragraph text work in both LaTeX and R markdown

[^2]: Footnote 25

[^3]: This works in HTML, but not in LaTeX

[^4]: HTML shows 4 rather than 25

### 2025-09-07 supplement/metadata/metadata.tex
Final response for inclusion of Rmd files.

@@
 Footnotes do not work in \verb|tabular| environments in LaTeX but do work in R Markdown tables\footnote{Footnotes in paragraph text work in both LaTeX and R markdown}.
-Custom footnote numbering works in LaTeX but does not work in R Markdown\footnote[21]{Forever 21 :)}.
+Custom footnote numbering works in LaTeX but does not work in R Markdown\footnote[25]{Footnote 25 }.
 These features are demonstrated in Table~\ref{table:3}.
@@
  \hline
- \verb|\acronym{A}\footnote{This works in HTML}| & \acronym{A}\footnote{This works in HTML} \\ \hline
-\verb|\R\footnote[21]{HTML shows 2 rather than 21}| & \R\footnote[21]{HTML shows 2 rather than 21}  \\ \hline
+ \verb|\acronym{A}\footnote{This works in HTML, but not in LaTeX}| & \acronym{A}\footnote{This works in HTML, but not in LaTeX}} \\ \hline
+\verb|\R\footnote[25]{HTML shows 4 rather than 25}| & \R\footnote[25]{HTML shows 4 rather than 25}  \\ \hline
 \end{tabular}
@@
 \paragraph{Note}
-Certain reference commands like \verb|\nameref{}| and \verb|\autoref{}| will not work as intended in HTML.
+Certain reference commands like \verb|\nameref{}| and \verb|\autoref{}| may not work as intended in HTML.
 

### 2025-09-07 supplement/table-env/RJwrapper.Rmd
Final response for inclusion of Rmd files.

---
title: Working with table environments in texor
abstract: |
  This is a small sample article to demonstrate usage of
  [**texor**](https://CRAN.R-project.org/package=texor) to convert table
  environments.
author:
- name: Abhishek Ulayil
  affiliation: Student, Institute of Actuaries of India
  orcid: 0009-0000-6935-8690
  address: Mumbai, India
date: '2025-09-01'
date_received: ~
journal:
  firstpage: ~
  lastpage: ~
volume: 17
issue: 3
slug: RJ-2025-000
citation_url: https://rjournal.github.io/
packages:
  cran:
  - texor
  - rebib
  - rjtools
  - rmarkdown
  bioc: []
preview: preview.png
bibliography: ~
CTV: ~
legacy_pdf: yes
legacy_converted: yes
output:
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
    mathjax: https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-chtml-full.js
    md_extension: -tex_math_single_backslash

---

::::::::::: article
## Introduction

Tables are commonly used in R Journal articles to display data in a
tabular format. However, there are differences in the way tables are
handled by LaTeX and HTML. LaTeX tables have more customization and are
usually optimized for printing, whereas the web articles need tables
optimized for varying sizes of media. Pandoc converts most of the tables
somewhat easily, but is unable to do well with table customization
packages and complex tables. The
[**texor**](https://CRAN.R-project.org/package=texor) packages uses the
pandoc extensions `simple_tables` and `pipe_tables` to tackle these
difficult cases.

## Simple tables

A simple table has $n$ rows and $p$ columns leading to $np$ cells.
Simple LaTeX tables will be converted just fine, even if they include a
some math and other elements. However, note that the conversion will be
done to traditional or pandoc-style markdown rather than R Markdown with
the `knitr::kable(..)` function. Also any graphic commands or specific
font characters will not be supported.

::: {#table:1}
  --------------------------------------------------------
  Graphics Format   LaTeX   Markdown   R Markdown   HTML
  ----------------- ------- ---------- ------------ ------
  PNG               Yes     Yes        Yes          Yes

  JPG               Yes     Yes        Yes          Yes

  PDF               Yes     No         No           No

  SVG               No      Yes        Yes          Yes

  Tikz              Yes     No         Yes          No

  Algorithm         Yes     No         No           No
  --------------------------------------------------------

  : (#tab:T1) Image format support in various markup/typesetting
  languages.
:::

## Multicolumn and Multirow tables

:::: center
::: {#table:2}
  -------------------------------------
  EXAMPLE   $X$           $Y$    
  --------- ------ ------ ------ ------
            1      2      1      2

  EX1       X11    X12    Y11    Y12

  EX2       X21    X22    Y21    Y22

  EX3       X31    X32    Y31    Y32

            X311   X322   Y311   Y322

  EX4       X41    X42    Y41    Y42

            X411   X422   Y411   Y422

  EX5       X51    X52    Y51    Y52
  -------------------------------------

  : (#tab:T2) An example table using `multicolumn` and `multirow`
:::
::::

Table [2](#tab:T2){reference-type="ref" reference="table:2"} illustrates
a table that uses the `multicolumn` and `multirow` command, which the
[**texor**](https://CRAN.R-project.org/package=texor) package can handle
through subroutines and pre-processing steps to transform the LaTeX
code.

Also note that the stream editor is used to rename `table*` environment
to `table` environment because the HTML format is single column, so the
asterisk indicating that the table should be drawn over the full width
of the page is redundant in this case.

## Complex tables

A complex table with various other elements like figures, math, code and
so on, are also supported by
[**texor**](https://CRAN.R-project.org/package=texor).

::: {#table:3}
+---------------+---------------+-------------+---------------------+
| Inline Format | LaTeX Support | Web Support | Rendering           |
+:==============+:==============+:============+:====================+
| Text          | Yes           | Yes         | *Hello*             |
+---------------+---------------+-------------+---------------------+
| Image         | Yes           | Yes         | ::: minipage        |
|               |               |             | ![image](           |
|               |               |             | R.png){width="100%" |
|               |               |             | alt="graphic        |
|               |               |             |  without alt text"} |
|               |               |             | :::                 |
+---------------+---------------+-------------+---------------------+
| CodeBlock     | Yes           | Yes         | ::: minipage        |
|               |               |             | ``` r               |
|               |               |             | x <- 1:100          |
|               |               |             | y <- dbino          |
|               |               |             | m(x,100,prob = 0.5) |
|               |               |             | plot(x,y)           |
|               |               |             | ```                 |
|               |               |             | :::                 |
+---------------+---------------+-------------+---------------------+
| Math          | Yes           | Yes         | $e = m c^2$         |
+---------------+---------------+-------------+---------------------+
| Link          | Yes           | Yes         | [Google](htt        |
|               |               |             | p://www.google.com) |
+---------------+---------------+-------------+---------------------+
| Nested Table  | Yes           | No          | NaN                 |
+---------------+---------------+-------------+---------------------+

: (#tab:T3) Image format support in various markup/typesetting
languages
:::

## Wide tables

As for the inclusion of wide tables, they are represented as multiple
tables with the first table housing the caption at the top. The
numbering and references will be the same as LaTeX.

Here is a reference to Table [4](#tab:T4){reference-type="ref"
reference="table:4"} and Table [5](#tab:T5){reference-type="ref"
reference="table:5"}.

::: {#table:4}
  -----------------------------------------------------------------------
  index   A        B        C         D      E       F      G      H
  ------- -------- -------- --------- ------ ------- ------ ------ ------
  1       359.00   NaN      5796.00   0.00   16.14   1.00   NaN    0.00

  2       25.73    0.00     1029.20   NaN    40.00   0.68   0.00   0.00

  2.1     26.26    0.00     13.40     0.00   2.14    0.68   0.00   NaN

  2.2     32.06    20.06    47.64     0.04   1.80    0.68   0.01   NaN

  2.3     51.94    420.27   21.17     0.20   1.77    0.74   0.05   NaN

  2.4     40.62    30.44    0.90      0.57   1.44    1.31   0.24   NaN

          ...      ...      ...       ...    ...     ...    ...    ...
  -----------------------------------------------------------------------

  : (#tab:T4) `A dummy research data`
:::

::: {#table:4}
  ---------------------------------------------------------------------------------------------------
  index   I      J                           K      L      M      N      O
  ------- ------ --------------------------- ------ ------ ------ ------ ----------------------------
  1       0.00   $-$`<!-- -->`{=html}1.73    0.00   0.00   0.00   0.00   NaN

  2       0.21   $-$`<!-- -->`{=html}33.41   0.00   0.11   0.00   NaN    $-$`<!-- -->`{=html}197.85

  2.1     0.24   $-$`<!-- -->`{=html}24.00   0.00   0.15   NaN    0.06   $-$`<!-- -->`{=html}70.46

  2.2     0.6    $-$`<!-- -->`{=html}19.42   0.00   0.15   0.00   0.11   $-$`<!-- -->`{=html}16.48

  2.3     0.75   $-$ 31.77                   0.00   0.18   0.01   0.37   $-$`<!-- -->`{=html}0.82

  2.4     0.26   $-$`<!-- -->`{=html}1.89    0.1    0.55   0.70   0.22   $-$`<!-- -->`{=html}6.55

          ...    ...                         ...    ...    ...    ...    ...
  ---------------------------------------------------------------------------------------------------

  : 
:::

::: {#table:5}
  --------------------------------------------------------------------------------------------------
  Package                                                         Commits     Version   Last Updated
  --------------------------------------------------------------- --------- --------- --------------
  [**texor**](https://CRAN.R-project.org/package=texor)           260           1.1.0    28-Jul-2023

  [**rebib**](https://CRAN.R-project.org/package=rebib)           76            0.2.4    29-Jul-2023

  [**rjtools**](https://CRAN.R-project.org/package=rjtools)       314          1.0.11    30-Jul-2023

  [**rmarkdown**](https://CRAN.R-project.org/package=rmarkdown)   3189           2.23    31-Jul-2023
  --------------------------------------------------------------------------------------------------

  : (#tab:T5) A dummy summary of a few CRAN packages
:::

## Long Tables

Pandoc supports long tables from **longtable** CTAN package as well,
here is an example as table [6](#tab:T6){reference-type="ref"
reference="table:6"}.

::: {#table:6}
  -----------------------------------------------------------------------------------------------------------------------------------------------------------------
  **Manafacturer**   **Model and Make**   **MPG**   **Engine Cylinders**   **Engine Displacement (cu.in.)**   **Weight (1000 lbs)**   **Number of forward gears**
  ------------------ -------------------- --------- ---------------------- ---------------------------------- ----------------------- -----------------------------
  Mazda              RX4                  21.0      6                      160.0                              2.620                   4

  Mazda              RX4 Wag              21.0      6                      160.0                              2.875                   4

  Datsun             710                  22.8      4                      108.0                              2.320                   4

  Hornet             4 Drive              21.4      6                      258.0                              3.215                   3

  Hornet             Sportabout           18.7      8                      360.0                              3.440                   3

  Plymouth           Valiant              18.1      6                      225.0                              3.460                   3

  Duster             360                  14.3      8                      360.0                              3.570                   3

  Merc               240D                 24.4      4                      146.7                              3.190                   4

  Merc               230                  22.8      4                      140.8                              3.150                   4

  Merc               280                  19.2      6                      167.6                              3.440                   4

  Merc               280C                 17.8      6                      167.6                              3.440                   4

  Merc               450SE                16.4      8                      275.8                              4.070                   3

  Merc               450SL                17.3      8                      275.8                              3.730                   3

  Merc               450SLC               15.2      8                      275.8                              3.780                   3

  Cadillac           Fleetwood            10.4      8                      472.0                              5.250                   3

  Lincoln            Continental          10.4      8                      460.0                              5.424                   3

  Chrysler           Imperial             14.7      8                      440.0                              5.345                   3

  Fiat               128                  32.4      4                      78.7                               2.200                   4

  Honda              Civic                30.4      4                      75.7                               1.615                   4

  Toyota             Corolla              33.9      4                      71.1                               1.835                   4

  Toyota             Corona               21.5      4                      120.1                              2.465                   3

  Dodge              Challenger           15.5      8                      318.0                              3.520                   3

  AMC                Javelin              15.2      8                      304.0                              3.435                   3

  Cheverolet         Camaro Z28           13.3      8                      350.0                              3.840                   3

  Pontiac            Firebird             19.2      8                      400.0                              3.845                   3

  Fiat               X1-9                 27.3      4                      79.0                               1.935                   4

  Porsche            914-2                26.0      4                      120.3                              2.140                   5

  Lotus              Europa               30.4      4                      95.1                               1.513                   5

  Ford               Pantera L            15.8      8                      351.0                              73.170                  5

  Ferrari            Dino                 19.7      6                      145.0                              2.770                   5

  Maserati           Bora                 15.0      8                      301.0                              3.570                   5

  Volvo              142E                 21.4      4                      121.0                              2.780                   4
  -----------------------------------------------------------------------------------------------------------------------------------------------------------------

  : (#tab:T6) Table of Car data from mtcars dataset
:::

## Limitations

Limitations of the [**texor**](https://CRAN.R-project.org/package=texor)
package in table handling includes:

-   Usage of custom graphics/ characters like `\circ` will not render in
    HTML properly,(using it in inline math might work).

-   Inclusion of code blocks might not always work and it is best
    avoided.

-   Currently only `\\` is supported as the row end command/marker.

-   Some table commands/environments might not work as expected.

-   Nested tabular environments will not work as expected.

-   Colored backgrounds in tables are not supported yet.

## Summary

In summary the [**texor**](https://CRAN.R-project.org/package=texor)
package supports:

-   Some common table environments.

-   Long and Wide Tables.

-   Some tables with `multicolumn` and `multirow` commands.

-   Environments such as `figure`/`code` in tables.

## Acknowledgements

Special thanks to Albert Krewinkel for his contributions towards earlier
versions of Lua filters to add table numbering.
:::::::::::
